---
title: "02_compound_target_prediction"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

# Load packages
```{r}
if (!require("tidyverse", quietly = TRUE)) install.packages("tidyverse", dependencies = TRUE)
if (!require("data.table", quietly = TRUE)) install.packages("data.table", dependencies = TRUE)
if (!require("openxlsx", quietly = TRUE)) install.packages("openxlsx", dependencies = TRUE)

if (!require("future", quietly = TRUE)) install.packages("future", dependencies = TRUE)
if (!require("furrr", quietly = TRUE)) install.packages("furrr", dependencies = TRUE)
if (!require("progressr", quietly = TRUE)) install.packages("progressr", dependencies = TRUE)
if (!require("httr", quietly = TRUE)) install.packages("httr", dependencies = TRUE)


library(tidyverse)
library(data.table)
library(openxlsx)

library(future)     
library(furrr)      
library(progressr) 
library(httr)      
```

# Compounds of ZWHQD
## Step 1: Load Raw Compound Data from DCABM-TCM Database
```{r dcabm}
rm(list = ls())

# Import raw compound data
compounds_raw <- read.xlsx("../data/raw/tcm/zwhqd_compound_dcabm.xlsx")

# Save raw data backup for reproducibility
save(
  compounds_raw,
  file = "../data/processed/tcm/01_zwhqd_compounds_raw_dcabm.RData"
)
```

## Step 2: Structured of raw data

```{r}
source("utils.R")

# Structured processing of raw data
compounds_structured <- compounds_raw %>%
  # Trim leading/trailing whitespace from all character columns
  mutate(across(where(is.character), trimws)) %>%
  
  # Split multiple compounds separated by "|" into individual rows
  separate_rows(Compounds, sep = "\\|") %>%
  
  # Extract core metadata from the unstructured Compounds string
  mutate(
    CID = str_extract(Compounds, "(?<=Pubchem CID:)[^)]+") %>% str_trim(),
    DCABM.ID = str_extract(Compounds, "(?<=DCABM ID:)[^)]+") %>% str_trim(),
    temp = str_remove(Compounds, "\\s+\\(Pubchem CID:[^)]+\\)")
  ) %>%
  
  # Extract clean compound name and molecular formula
  mutate(
    temp2 = str_remove(temp, "\\([^)]+\\)$") %>% str_trim(),
    Compound = str_remove(temp2, "\\([^)]+\\)$") %>% str_trim(),
    Formula = str_match(temp2, "\\(([^)]+)\\)$")[, 2] %>% str_trim()
  ) %>%
  
  # Remove redundant intermediate columns
  dplyr::select(-Compounds, -temp, -temp2, -DCABM.ID) %>%
  
  # Final whitespace standardization for all character columns
  mutate(across(where(is.character), trimws)) %>%
  
  dplyr::filter(!is.na(CID)) %>%
  
  dplyr::select(-Type) %>%
  
  # deduplication
  distinct(Herb.Name.Chinese,CID,Compound,Formula,.keep_all = TRUE)

# Save cleaned dataset
save(
  compounds_structured,
  file = "../data/processed/tcm/02_zwhqd_compounds_raw_structured.RData"
)

openxlsx::write.xlsx(compounds_structured,"../data/processed/tcm/02_zwhqd_compounds_raw_structured.xlsx")
```

---

## Step 3: Load Manually zwhqd-compounds data

- Manually Add Cycloastragenol Entry

- Manually annotated compound

Input:

- Manually annotated Excel file: "../data/processed/tcm/03_zwhqd_compounds_manual_annotation.xlsx"

```{r}
# Load manually curated compound properties
compounds_manual <- read.xlsx("../data/processed/tcm/03_zwhqd_compounds_manual_annotation.xlsx")

# Merge manual annotations with herb metadata, filter, and deduplicate
compounds_cleaned <- compounds_manual %>%
  dplyr::filter(Is.Keep == "Yes") %>%
  dplyr::select(
    -c(CID, Compound, Formula,`Q-Marker`,Revision.Note,Is.Keep,Remark)
  ) %>% 
  dplyr::rename(CID = CID.Modified,
                Compound = Compound.Name.Modified,
                Formula = Formula.Corrected)

table(compounds_cleaned$Herb.Name.Pinyin)

# Save
save(
  compounds_cleaned,
  file = "../data/processed/tcm/04_compounds_cleaned.RData"
)
```

---

## Step 4: Batch Retrieve PubChem Compound Annotations

Description: Use parallel processing to batch-retrieve standardized compound annotations from PubChem, including IUPAC name, molecular formula, molecular weight, and SMILES structure.

```{r}
# load("../data/processed/tcm/04_compounds_cleaned.RData")
source("utils.R")

# Standardize CID format and isolate unique CIDs for efficient PubChem queries
compounds_pubchem <- compounds_cleaned %>%
  mutate(`CID` = clean_cids(CID)) %>%
  filter(!is.na(CID)) %>%
  distinct(CID, .keep_all = TRUE)

# Define PubChem properties to retrieve
properties <- c("IUPACName", "MolecularFormula", "MolecularWeight",
                "SMILES")

# Split unique CIDs into batches of 20 to comply with PubChem API rate limits
unique_cids <- unique(compounds_pubchem$CID)
cid_batches <- split(unique_cids, ceiling(seq_along(unique_cids)/20))

# Set up parallel processing: use all available cores minus 2 to avoid system overload
plan(multisession, workers = availableCores() - 2)

# Execute parallel batch queries with real-time progress tracking
with_progress({
  p <- progressor(steps = length(cid_batches))
  chem_info <- future_map_dfr(cid_batches, function(b) {
    p()
    fetch_batch(properties, b)
  })
})

# Reset to sequential processing mode
plan(sequential)

# Merge PubChem annotations with the curated compound list and standardize final output
compounds_final <- compounds_cleaned %>%
  left_join(chem_info, by = "CID") %>%
  dplyr::select(
    Herb.Name.Chinese,
    Herb.Name.Pinyin, 
    Herb.Name.Latin,
    Herb.Name.English,
    Compound, 
    IUPACName, 
    MolecularFormula, 
    MolecularWeight, 
    CID, 
    SMILES
  ) %>%
  mutate(across(where(is.character), trimws))
```

---

## Step 5: Calculate Cross-Herb Compound Sharing Statistics

Description: Identify compounds shared across multiple herbs, count the number of herbs each compound appears in, and generate a sorted list of shared compounds for downstream analysis.

```{r}
compound_sharing <- compounds_final %>%
  distinct(Herb.Name.Pinyin, CID,.keep_all = TRUE) %>%
  group_by(CID) %>%
  summarise(
    Herb_Count = n(),
    Herbs = paste(unique(Herb.Name.Pinyin), collapse = ", ")
  ) %>%
  filter(Herb_Count >= 2) %>%
  arrange(desc(Herb_Count))  %>%
  left_join(compounds_final[,c("CID","Compound")],by="CID") %>%
  distinct(CID,.keep_all = TRUE)

# Save all final results
save(
  compounds_final, compound_sharing,
  file = "../data/processed/tcm/05_compounds_final.RData"
)
```

---

## Step 6: Batch Download PubChem SDF Structure Files

Description: Download 3D or 2D structure SDF files for all unique compounds from PubChem, with automatic directory creation and download summary reporting.

```{r}
unique_cids <- unique(compounds_pubchem$CID)
sdf_files <- "../data/raw/tcm/sdf/"

results <- download_all_sdf(unique_cids, sdf_files)
```

## Table S1. Bioactive Compounds of ZWHQD.

Description: Rename columns to publication-ready format and export the final compound list as an Excel file.

```{r}
# load("../data/processed/tcm/05_compounds_final.RData")

table_data <- compounds_final %>%
  dplyr::rename(
   ` Herb(Chinese)` = Herb.Name.Chinese,
    `Herb(Pinyin)` = Herb.Name.Pinyin,
    `Herb(Latin)` = Herb.Name.Latin,
   `Herb(English)` = Herb.Name.English,
   `Compound Name` = Compound,
   `Compound Name (IUPAC)` = IUPACName,
   `Molecular Formula (g/mol)` = MolecularFormula,
   `Pubchme CID` = CID,
   `SMILES` = SMILES
  )

# Export to publication-ready Excel file
write.xlsx(
  list(zwhqd_compounds_raw_dcabm = compounds_raw,
       zwhqd_compounds_raw_structured = compounds_structured,
       zwhqd_compounds_cleaned = table_data), 
  "../supplementary/tables/Table_S1.xlsx"
)

cat("Table S1 has been successfully generated and saved to `../supplementary/tables/Table_S1.xlsx`\n")
```

## Report
```{r}
# load("../data/processed/tcm/05_compounds_final.RData")

print(table(compounds_final$Herb.Name.Pinyin))
```


# Potential targets of compounds

## Step 1: Extract Targets from BATMAN-TCM
```{r}
# Clear environment to avoid variable conflicts
rm(list = ls())

load("../data/raw/uniprot/uniprot_human_reviewed.RData") 
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry,Gene.Names.primary)

# Load input datasets
load("../data/processed/tcm/05_compounds_final.RData")
load("../data/raw/tcm/batman/batman_target_known.RData")
load("../data/raw/tcm/batman/batman_target_pred.RData")

# Known targets from BATMAN-TCM
compound_target_known_batman <- batman_target_known %>%
  dplyr::filter(CID %in% unique(compounds_final$CID)) %>%
  dplyr::select(CID, Uniprot.ID) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "BATMAN-Known")

# Predicted targets from BATMAN-TCM (Score ≥ 0.80,LR=48,p_adj<0.5)
compound_target_pred_batman <- batman_target_pred %>%
  dplyr::filter(CID %in% unique(compounds_final$CID)) %>%
  dplyr::select(CID, Uniprot.ID, Score) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol), Score >= 0.80) %>%
  dplyr::select(-Score) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "BATMAN-Pred")

length(unique(compound_target_pred_batman$CID))
length(unique(compound_target_pred_batman$Symbol))
length(unique(compound_target_pred_batman$Symbol)) / length(unique(compound_target_pred_batman$CID))

# Save processed BATMAN-TCM targets
save(
  compound_target_known_batman,
  compound_target_pred_batman,
  file = "../data/processed/tcm/06_batman_targets.RData"
)
```

## Step 2: Extract Targets from Super-Pred
```{r}
rm(list = ls())

source("utils.R")


load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry,Gene.Names.primary)

# known targets
compound_target_known_super <- read_compound_target("../data/raw/tcm/super/known") %>%
  dplyr::select(CID, Uniprot.ID = `UniProt ID`) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "SuperPred-Known")

# Predicted targets (Prob >= 60, Model Accuracy >= 85) (modify,70/90)
compound_target_pred_super <- read_compound_target("../data/raw/tcm/super") %>%
  dplyr::filter(Probability >= 60, `Model accuracy` >= 80) %>%
  dplyr::select(CID, Uniprot.ID = `UniProt ID`) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  distinct(CID, Symbol, .keep_all = TRUE)%>%
  mutate(Source = "SuperPred-Pred")

length(unique(compound_target_pred_super$CID))
length(unique(compound_target_pred_super$Symbol))
length(unique(compound_target_pred_super$Symbol)) / length(unique(compound_target_pred_super$CID))

# Save processed Super-Pred targets
save(
  compound_target_known_super,
  compound_target_pred_super,
  file = "../data/processed/tcm/07_super_targets.RData"
)
```

## Step 3: Extract Targets from SwissTargetPrediction
```{r}
rm(list = ls())

source("utils.R")

load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry,Gene.Names.primary)

# Process SwissTargetPrediction results (Probability > 0)

compound_target_pred_swiss <- read_compound_target("../data/raw/tcm/swiss") %>%
  filter(`Probability*` > 0) %>%
  separate_rows(`Common name`, `Uniprot ID`, sep = " ") %>%
  mutate(
    CID = clean_cids(CID),
    Uniprot.ID = gsub("[[:space:]]", "", `Uniprot ID`)
  ) %>%
  dplyr::select(CID, Uniprot.ID) %>%
  distinct(CID, Uniprot.ID, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  distinct(CID, Symbol, .keep_all = TRUE)%>%
  mutate(Source = "SwissTargetPrediction")

length(unique(compound_target_pred_swiss$Symbol))

length(unique(compound_target_pred_swiss$CID))

length(unique(compound_target_pred_swiss$Symbol)) / length(unique(compound_target_pred_swiss$CID))


# Save processed SwissTargetPrediction targets
save(
  compound_target_pred_swiss,
  file = "../data/processed/tcm/08_swiss_targets.RData"
)
```

---

## Step 4: Extract Targets from PharmMapper
```{r}
rm(list = ls())

source("utils.R")

load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry,Entry.Name, Gene.Names.primary)

# Process PharmMapper results (Human only, high-confidence filters) (modify,0.5)
compound_target_pred_pharm <- read_compound_target2("../data/raw/tcm/pharm",skip_row = 1) %>%
  dplyr::filter(grepl("_HUMAN", Uniplot),zscore>0, `Norm Fit`>0.5) %>%
  dplyr::select(CID, Uniplot) %>%
  distinct(CID, Uniplot, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Uniplot" = "Entry.Name")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  dplyr::select(CID,Uniprot.ID = Entry,Symbol) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "PharmMapper")

length(unique(compound_target_pred_pharm$Symbol))

length(unique(compound_target_pred_pharm$CID))

length(unique(compound_target_pred_pharm$Symbol)) / length(unique(compound_target_pred_pharm$CID))

# Save processed PharmMapper targets
save(
  compound_target_pred_pharm,
  file = "../data/processed/tcm/09_pharm_targets.RData"
)
```

---

## Step 5: Extract Targets from SEA
```{r}
rm(list = ls())

source("utils.R")

load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry, Entry.Name, Gene.Names.primary)


# Process SEA results (confidence filters)
# P < 1e-5、MaxTc >= 0.35(modify)
compound_target_pred_sea <- read.csv("../data/raw/tcm/sea/sea-results.xls") %>%
  mutate(CID = as.numeric(gsub("compound","",Query.ID)), .before = 1) %>%
  arrange(CID) %>%
  mutate(CID = as.character(CID)) %>%
  dplyr::select(-Query.ID) %>%
  dplyr::filter(
    grepl("_HUMAN", Target.ID), 
    P.Value < 1e-5, 
    Max.Tc >= 0.35
  ) %>%
  distinct(CID, Target.ID, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Target.ID" = "Entry.Name")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  dplyr::select(CID,Uniprot.ID = Entry,Symbol) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "SEA")

length(unique(compound_target_pred_sea$Symbol))

length(unique(compound_target_pred_sea$CID))

length(unique(compound_target_pred_sea$Symbol)) / length(unique(compound_target_pred_sea$CID))


# Save processed SEA targets
save(
  compound_target_pred_sea,
  file = "../data/processed/tcm/10_sea_targets.RData"
)
```

---

## Step 6: Integrate All Targets and Deduplicate
```{r}
rm(list = ls())

# Load all processed target datasets
load("../data/processed/tcm/06_batman_targets.RData")
load("../data/processed/tcm/07_super_targets.RData")
load("../data/processed/tcm/08_swiss_targets.RData")
load("../data/processed/tcm/09_pharm_targets.RData")
load("../data/processed/tcm/10_sea_targets.RData")

# Merge all targets 
compound_target <- rbind(
  compound_target_known_batman,
  compound_target_known_super,
  compound_target_pred_batman, 
  compound_target_pred_super, 
  compound_target_pred_swiss,
  compound_target_pred_pharm,
  compound_target_pred_sea
) %>%
  group_by(CID, Symbol) %>%  
  summarise(
    Source = paste(unique(na.omit(Source)), collapse = "; "),
    across(everything(), ~ dplyr::first(.)),  
    .groups = "drop"
  )

# Filter high-confidence targets (known or predicted by ≥1  databases)
compound_target_final <- compound_target %>%
  mutate(
    is_known = str_detect(Source, "Known"),
    platform_count = str_split(Source, "; ") %>%
      map_dbl(~ length(unique(.)))
  ) %>%
  filter(
    is_known | platform_count >= 1
  ) %>%
  dplyr::select(CID,Symbol,Uniprot.ID,Source)

length(unique(compound_target_final$Symbol))

length(unique(compound_target_final$CID))

length(unique(compound_target_final$Symbol)) / length(unique(compound_target_final$CID))

# Save final target list
save(
  compound_target_final,
  file = "../data/processed/tcm/11_compound_target_final.RData"
)
```

---

## Step 7: Target Statistics Across Prediction Platforms.
```{r}
# Generate summary table for manuscript
summary_table <- tibble(
  Database = c(
    "BATMAN-TCM (Known)",
    "SuperPred (Known)",
    "BATMAN-TCM (Predicted)",
    "SuperPred (Predicted)",
    "SwissTargetPrediction",
    "PharmMapper",
    "SEA",
    "Final Integrated Results"
  ),
  
  # Number of unique compounds with targets (CID)
  `Number of compounds with targets` = c(
    n_distinct(compound_target_known_batman$CID),
    n_distinct(compound_target_known_super$CID),
    n_distinct(compound_target_pred_batman$CID),
    n_distinct(compound_target_pred_super$CID),
    n_distinct(compound_target_pred_swiss$CID),
    n_distinct(compound_target_pred_pharm$CID),
    n_distinct(compound_target_pred_sea$CID),
    n_distinct(compound_target_final$CID)
  ),
  
  # Number of unique targets based on UniProt ID
  `Unique Targets (UniProt ID)` = c(
    n_distinct(compound_target_known_batman$Uniprot.ID),
    n_distinct(compound_target_known_super$Uniprot.ID),
    n_distinct(compound_target_pred_batman$Uniprot.ID),
    n_distinct(compound_target_pred_super$Uniprot.ID),
    n_distinct(compound_target_pred_swiss$Uniprot.ID),
    n_distinct(compound_target_pred_pharm$Uniprot.ID),
    n_distinct(compound_target_pred_sea$Uniprot.ID),
    n_distinct(compound_target_final$Uniprot.ID)
  )
)

# Save statistics

save(summary_table, file =  "../data/processed/tcm/12_target_statistics.RData")
```

## Step 8: herb-compound-target
```{r}
rm(list = ls())

load("../data/processed/tcm/05_compounds_final.RData")
load("../data/processed/tcm/11_compound_target_final.RData")

herb_compound_target <- compound_target_final %>%
  left_join(
    compounds_final,
    by = "CID",
    relationship = "many-to-many"
  ) %>%
  dplyr::select(
    Herb.Name.Chinese, Herb.Name.Pinyin, Herb.Name.Latin, Herb.Name.English,
    Compound, CID, MolecularFormula, MolecularWeight, SMILES, IUPACName,
    Symbol, Uniprot.ID
  ) %>%
  distinct(Herb.Name.Chinese, CID, Symbol, .keep_all = TRUE)

length(unique(herb_compound_target$Herb.Name.Chinese))
length(unique(herb_compound_target$CID))
length(unique(herb_compound_target$Symbol))

save(herb_compound_target, file = "../data/processed/tcm/13_herb_compound_target.RData")
```

## Table S2: Potential Targets of ZWHQD.
```{r}
load("../data/processed/tcm/06_batman_targets.RData")
load("../data/processed/tcm/07_super_targets.RData")
load("../data/processed/tcm/08_swiss_targets.RData")
load("../data/processed/tcm/09_pharm_targets.RData")
load("../data/processed/tcm/10_sea_targets.RData")

# Export Table S2
write.xlsx(
  list(`Potential_Targets_ZWHQD` = herb_compound_target,
       `Compound_Target_Known_Batman` = compound_target_known_batman,
    `Compound_Target_Known_Super` = compound_target_known_super,
    `Compound_Target_Pred_Batman` = compound_target_pred_batman,
    `Compound_Target_Pred_Super` = compound_target_pred_super,
    `Compound_Target_Pred_Swiss` = compound_target_pred_swiss,
    `Compound_Target_Pred_Pharm` = compound_target_pred_pharm,
    `Compound_Target_Pred_SEA` = compound_target_pred_sea,
    `Compound_Target_Final` = compound_target_final),
  "../supplementary/tables/Table_S2.xlsx"
)
```

## Report
```{r}
num_compound_known <- length(unique(compound_target_known_batman$Symbol,compound_target_known_super$Symbol))

num_compound_pred <- length(unique(herb_compound_target$Symbol)) - num_compound_known

cat("Number of Known targets: ", num_compound_known,"\n")

cat("Number of Predicted targets: ", num_compound_pred,"\n")
```

