---
title: "03_disease_target_identification"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

# Load packages
```{r}
if (!require("tidyverse", quietly = TRUE)) install.packages("tidyverse", dependencies = TRUE)
if (!require("data.table", quietly = TRUE)) install.packages("data.table", dependencies = TRUE)
if (!require("openxlsx", quietly = TRUE)) install.packages("openxlsx", dependencies = TRUE)

if (!require("RColorBrewer", quietly = TRUE)) install.packages("RColorBrewer", dependencies = TRUE)
if (!require("patchwork", quietly = TRUE)) install.packages("patchwork", dependencies = TRUE)
if (!require("pheatmap", quietly = TRUE)) install.packages("pheatmap", dependencies = TRUE)
if (!require("ggsci", quietly = TRUE)) install.packages("ggsci", dependencies = TRUE)
if (!require("ggvenn", quietly = TRUE)) install.packages("ggvenn", dependencies = TRUE)
if (!require("corrplot", quietly = TRUE)) install.packages("corrplot", dependencies = TRUE)
if (!require("magick", quietly = TRUE)) install.packages("magick", dependencies = TRUE)
if (!require("CMplot", quietly = TRUE)) install.packages("CMplot", dependencies = TRUE)

if (!require("AnnoProbe", quietly = TRUE)) install.packages("AnnoProbe", dependencies = TRUE)
if (!require("genefilter", quietly = TRUE)) BiocManager::install("genefilter", dependencies = TRUE)

if (!require("limma", quietly = TRUE)) install.packages("limma", dependencies = TRUE)
if (!require("WGCNA", quietly = TRUE)) install.packages("WGCNA", dependencies = TRUE)
if (!require("flashClust", quietly = TRUE)) install.packages("flashClust", dependencies = TRUE)
if (!require("robustbase", quietly = TRUE)) install.packages("robustbase", dependencies = TRUE)
if (!require("dynamicTreeCut", quietly = TRUE)) install.packages("dynamicTreeCut", dependencies = TRUE)
if (!require("GenomicRanges")) BiocManager::install("GenomicRanges")
if (!require("org.Hs.eg.db", quietly = TRUE)) BiocManager::install("org.Hs.eg.db", dependencies = TRUE)
if (!require("TxDb.Hsapiens.UCSC.hg19.knownGene", quietly = TRUE)) BiocManager::install("TxDb.Hsapiens.UCSC.hg19.knownGene", dependencies = TRUE)

library(tidyverse)
library(data.table)
library(openxlsx)

library(AnnoProbe)
library(genefilter)
library(limma)
library(org.Hs.eg.db)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)

library(WGCNA)
library(flashClust)
library(robustbase) 
library(dynamicTreeCut)

library(GenomicRanges)

library(RColorBrewer)
library(patchwork)
library(pheatmap)
library(ggsci)
library(ggvenn)
library(corrplot)
library(magick)
library(CMplot) 
```

# DEG
## Step 1: Loading Data
```{r}
rm(list = ls())
load("../data/raw/geo/GSE57338_raw.RData")

HC <- gse57338_pheno_raw$`disease status:ch1` == "non-failing"
DCM_HF <- gse57338_pheno_raw$`disease status:ch1` == "idiopathic dilated CMP"

keep_samples <- rownames(gse57338_pheno_raw)[HC | DCM_HF]
pheno <- gse57338_pheno_raw[keep_samples, ]

pheno$group <- ifelse(pheno$`disease status:ch1` == "non-failing", "HC", "DCM_HF") 
pheno$group <- factor(pheno$group, levels = c("HC", "DCM_HF"))

expr_keep <- gse57338_exprs_raw[,keep_samples]
```

---

## Step 2: Preprocessing

- Figure S1A: Distribution of expression density across samples.

- Figure S1B: Boxplot of gene expression levels across samples.

- Figure S1C: Principal component analysis (PCA) of sample expression profiles.

```{r}
# log2 transformation
ex <- expr_keep
qx <- as.numeric(quantile(ex, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))
LogC <- (qx[5] > 100) ||
          (qx[6]-qx[1] > 50 && qx[2] > 0)
if (LogC) { ex[which(ex <= 0)] <- NaN
  expr_keep <- log2(ex) }

# check GPL and annotate the probes to genes.

probe2gene=idmap(gse57338_gpl) 
expr_annotated <- filterEM(expr_keep, probe2gene)

# Processing of duplicated genes
expr <- as.matrix(expr_annotated)
if (any(duplicated(rownames(expr)))) {
  expr <- aggregate(expr, by = list(rownames(expr)), FUN = mean)
  rownames(expr) <- expr[,1]
  expr <- expr[,-1]
  expr <- as.matrix(expr)
}

# Express the overall distribution of values and observe where the "tail" of background noise falls
df <- reshape2::melt(expr)
low_expr_plot <- ggplot(df, aes(x = value, color = Var2)) + 
  geom_density(show.legend = FALSE) + 
  # labs(title = "Expression density across samples") +
  xlab("log2 expression") + 
  theme_bw() +
  scale_x_continuous(breaks = seq(from = floor(min(df$value)), to = ceiling(max(df$value)), by = 1)) +
  geom_vline(xintercept = 3, linetype = "dashed", color = "blue", linewidth = 1)

print(low_expr_plot)

ggsave("../supplementary/figures/Figure_S1A.pdf",low_expr_plot, width = 14,height = 8,units = "cm", dpi = 300)

# Filtering low expression genes
threshold_log2 <- 3
ffun <- filterfun(pOverA(p = 0.5, A = threshold_log2))
keep <- genefilter(expr, ffun)

expr_filt <- expr[keep, ]
dim(expr_filt)


# Check normalization
pdf("../supplementary/figures/Figure_S1B.pdf", width = 14, height = 8)
boxplot(expr_filt, 
        las = 2,  
        # main = "Check normalization", 
        ylab = "log2 expression", 
        col = "lightblue")

abline(h = median(expr_filt, na.rm = TRUE), col = "red", lty = 2)
dev.off()

# Check batch effect
# PCA
pca <- prcomp(t(expr_filt), scale. = TRUE)
pca_df <- data.frame(pca$x[,1:2], Group = pheno$group)
pca_plot <- ggplot(pca_df, aes(PC1, PC2, color = Group)) + 
  geom_point(size = 1) + 
  stat_ellipse() + 
  theme_minimal()

print(pca_plot)
ggsave("../supplementary/figures/Figure_S1C.pdf",pca_plot, width = 7,height = 5,units = "cm", dpi = 300)

# Correlation within the group
cor_mat <- cor(expr_filt)

within_cor <- sapply(colnames(expr_filt), function(s) {
  idx <- which(pheno$geo_accession == s)
  gr <- pheno$group[idx]
  same_group <- pheno$geo_accession[pheno$group == gr]
  same_group <- setdiff(same_group, s)
  
  if (length(same_group) == 0) return(NA)
  mean(cor_mat[s, same_group], na.rm = TRUE)
})

between_cor <- sapply(colnames(expr_filt), function(s) {
  idx <- which(pheno$geo_accession == s)
  gr <- pheno$group[idx]
  diff_group <- pheno$geo_accession[pheno$group != gr]
  if(length(diff_group)==0) return(NA)
  mean(cor_mat[s, diff_group], na.rm = TRUE)
})

mean_within = mean(within_cor, na.rm = TRUE)
mean_between = mean(between_cor, na.rm = TRUE)

cat("Average correlation coefficient within the group：", round(mean_within, 4), "\n")
cat("Average correlation coefficient between groups：", round(mean_between, 4), "\n")

if (mean_within > mean_between) {
  cat("✅ No significant batch effect\n")
} else {
  cat("⚠️ Have significant batch effect, suggest batch correction(ComBat/SVA)\n")
}

save(pheno, expr_filt, file = "../data/processed/geo/gse57338.RData")
```


## Step 3: DEG
```{r}
load("../data/processed/geo/gse57338.RData")
group <- factor(pheno$group)  
design <- model.matrix(~ 0 + group)
colnames(design) <- levels(group)
contr <- paste(levels(group)[2], levels(group)[1], sep = "-")
contr.matrix <- makeContrasts(contrasts = contr, levels = design)

fit <- lmFit(expr_filt, design)
fit2 <- contrasts.fit(fit, contr.matrix)
fit2 <- eBayes(fit2)

res <- topTable(fit2, adjust = "fdr", number = Inf)

save(res, file = "../data/processed/geo/gse57338_res.RData")
```

## Step 4: Significant DEGs
```{r}
# lfc <- log2(1.5)
lfc <- 0.585

deg_all <- as.data.frame(res) %>%
  dplyr::rename(log2FC = logFC)

deg_up <- deg_all %>% filter(adj.P.Val < 0.05, log2FC > lfc)
deg_down <- deg_all %>% filter(adj.P.Val < 0.05, log2FC < -lfc)

cat("The number of all DEGs:", nrow(deg_all), "\n")
cat("The number of up DEGs:", nrow(deg_up), "\n")
cat("The number of down DEGs:", nrow(deg_down), "\n")
```

---

##  Figure 2A: Volcano Plot

```{r}
deg_all$significance <- ifelse(
  deg_all$`adj.P.Val` < 0.05 & abs(deg_all$log2FC) > lfc,
  ifelse(deg_all$log2FC > lfc, paste0("Up(",nrow(deg_up),")"), paste0("Down(",nrow(deg_down),")")), "NS")

volcano_plot <- ggplot(deg_all, aes(x = log2FC, y = -log10(adj.P.Val), color = significance)) +
  geom_point(alpha = 0.7, size = 1) +
  scale_color_manual(values = c("blue", "gray", "red")) +
  geom_vline(xintercept = c(-lfc, lfc), linetype = "dashed", color = "gray40", linewidth = 0.4) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "gray40", linewidth = 0.4) +
  theme_bw(base_size = 10) +
  labs(
    # tag = "A",
    x = "log2(Fold Change)", 
    y = "-log10(Adjusted P-value)") +
  theme_bw() + theme(
    plot.title = element_text(hjust = 0.5)
  )

ggsave("../main/figures/Figure_2A.pdf",volcano_plot, width = 7,height = 5,dpi = 300)
```

---

## Figure 2B. Heatmap
```{r}
deg_sig <- deg_all %>% filter(adj.P.Val < 0.05, abs(log2FC) > lfc)

N_top <- 50

up_top <- deg_sig %>%
  filter(log2FC > 0) %>%
  arrange(desc(log2FC)) %>%
  head(N_top)

down_top <- deg_sig %>%
  filter(log2FC < 0) %>%
  arrange(log2FC) %>%
  head(N_top)

heat_genes <- c(rownames(up_top), rownames(down_top))

sample_order <- pheno %>%
  arrange(group) %>%   
  pull(geo_accession)  

mat <- expr_filt[heat_genes, sample_order]

mat_z <- t(scale(t(mat)))

breaks <- seq(-2, 2, length.out = 100)
palette <- colorRampPalette(c("blue", "white", "red"))(99)

# Group annotation
anno_col <- data.frame(
  Group = pheno$group[match(sample_order, pheno$geo_accession)]
)
rownames(anno_col) <- sample_order

p_heat <- pheatmap(
  mat_z,
  annotation_col = anno_col,
  annotation_names_col = F,
  annotation_legend = T,
  annotation_height = 0.2,
  color = palette,
  breaks = breaks,
  scale = "none",      
  show_rownames = FALSE,
  show_colnames = FALSE,
  treeheight_row = 20,
  treeheight_col = 0,
  cluster_cols = FALSE,    
  cluster_rows = TRUE,   
  border_color = NA,
  fontsize = 10
)

pdf("../main/figures/Figure_2B.pdf", width = 7, height = 5)
print(p_heat)
dev.off()
```

## Step 7: Saving DEG Results
```{r}
save(deg_all,deg_sig,deg_up, deg_down, lfc, file = "../data/processed/geo/gse57338_degs_results.RData")
```

## Table S2: DEG Result
```{r}
load("../data/processed/geo/gse57338_degs_results.RData")

deg_all <- deg_all %>%
  arrange(significance)

write.xlsx(
  list(DEG_Result = deg_all),
  "../tables/Table_S3.xlsx"
)
```

# WGCNA
## Step 1: Loading Data
```{r}
rm(list = ls())

load("../data/processed/geo/gse57338.RData")

# expression matrix
# WGCNA required: Sample (row) x  Gene(column)

dat <- expr_filt

common_samples <- intersect(colnames(dat), rownames(pheno))

dat <- dat[, common_samples, drop = FALSE]
pheno <- pheno[match(common_samples, rownames(pheno)), , drop = FALSE]

if (ncol(dat) == nrow(pheno)) {
  datExpr <- as.data.frame(t(dat))
} else if (nrow(dat) == nrow(pheno)) {
  datExpr <- as.data.frame(dat)
} else {
  stop("Error: Sample dimension mismatch between expression matrix and phenotype data.")
} 
rownames(datExpr) <- common_samples

dim(datExpr)

# trait: group (HC=0, DCM_HF=1)
datTraits <- data.frame(
  Group = factor(pheno$group)
)
rownames(datTraits) <- colnames(expr_filt)

datTraits$Disease <- ifelse(pheno$group == "DCM_HF", 1, 0)
```


## Step 2: Sample clustering (checking for outlier samples, quality control) 

- Figure S2A. 

- 
```{r}
# goodSamplesGenes
# Quality control: Remove bad genes/samples
gsg <- goodSamplesGenes(datExpr, verbose = 3)
if (!gsg$allOK) {
  if (sum(!gsg$goodGenes) > 0) {
    printFlush(paste("Removing genes:", paste(names(datExpr)[!gsg$goodGenes], collapse = ",")))
  }
  if (sum(!gsg$goodSamples) > 0) {
    printFlush(paste("Removing samples:", paste(rownames(datExpr)[!gsg$goodSamples], collapse = ",")))
  }
  datExpr <- datExpr[gsg$goodSamples, gsg$goodGenes]
}

# Screening for highly variable genes: MAD. Retain the first 5000
gene_mad <- apply(datExpr, 2, mad, na.rm = TRUE) 
N_top <- 5000
n_select <- min(N_top, ncol(datExpr))   
selected_genes <- names(sort(gene_mad, decreasing = TRUE)[1:n_select])
datExpr <- datExpr[, selected_genes]
cat("Retain the number of highly variable genes：", ncol(datExpr), "\n")

# Sample clustering: sampleTree
sampleTree <- hclust(dist(datExpr), method = "average")
group_color = ifelse(pheno$group=="DCM_HF","tomato","cyan")
color_mat <- cbind(Group = group_color)

pdf("../supplementary/figures/Figure_S2A.pdf", width=10, height=5)
plotDendroAndColors(sampleTree, color_mat, 
                    dendroLabels = FALSE, 
                    addGuide = TRUE,       
                    guideCol = "grey"
                    # ,   
                    # main = "Sample clustering"
                    )
dev.off()


save(datTraits,datExpr,pheno,file = "../data/processed/geo/gse57338_datExpr.RData")
```

---

## Step 3: Select soft threshold

- Figure S2B. 

```{r}
# load("../data/processed/geo/gse57338_datExpr.RData")

enableWGCNAThreads()
powers <- c(c(1:10), seq(from = 12, to=30, by=2))

R2_cut <- 0.8

sft <- pickSoftThreshold(
  datExpr,
  powerVector = powers,
  RsquaredCut = R2_cut, 
  verbose = 5,
  networkType = "unsigned"
)
disableWGCNAThreads()

power <- sft$powerEstimate
if(is.na(power)) power <- 8  
cat("power = ", power, "\n")

pdf("../supplementary/figures/Figure_S2B.pdf", width=14, height=5)
par(mfrow = c(1,2))

# Left: R2
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Power",ylab=expression(Scale ~ Free ~ R^2),type="n")
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers, col="red")
abline(h=0.8, col="blue")

# Right: Connectivity
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Power",ylab="Mean Connectivity",type="n")
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, col="red")
dev.off()

save(sft,file="../data/processed/geo/gse57338_sft.RData")

```

## Step 4: Construct a co-expression network

---

```{r}
enableWGCNAThreads()
set.seed(123)

net <- blockwiseModules(
  datExpr,
  power = power,
  TOMType = "unsigned",
  minModuleSize = 30,      # Minimum number of genes in the module
  mergeCutHeight = 0.25,   # Merge similar modules
  numericLabels = TRUE,
  pamRespectsDendro = FALSE,
  verbose = 3
)
disableWGCNAThreads()

save(net,file = "../data/processed/geo/gse57338_net.RData")

```

## Step 5: Clustering

---

- Figure_2C

```{r}
load("../data/processed/geo/gse57338_net.RData")

# Convert to Color
moduleColors <- labels2colors(net$colors)
table(moduleColors) # Check the number of genes in each module

# Genetree & module color
pdf("../main/figures/Figure_2C.pdf", width=14, height=5)

plotDendroAndColors(
  net$dendrograms[[1]],
  moduleColors[net$blockGenes[[1]]],
  "Module colors",
  dendroLabels = FALSE,
  marAll = c(0.5, 3, 1, 0),
  addGuide = TRUE,
  guideHang = 0.2
)
dev.off()
```

## Step 6: Module correlation with diseases

---

- Figure_2D.

```{r}
load("../data/processed/geo/gse57338_datExpr.RData")

MEs <- net$MEs
MEs0 <- moduleEigengenes(datExpr, moduleColors)$eigengenes
MEs <- orderMEs(MEs0)

datTraits <- model.matrix(~0 + group, data = data.frame(group = pheno$group))

colnames(datTraits) <- levels(factor(pheno$group))  
rownames(datTraits) <- rownames(pheno)

moduleTraitCor <- cor(MEs, datTraits, use = "p")
moduleTraitPvalue <- corPvalueStudent(moduleTraitCor, nrow(datExpr))

pdf("../main/figures/Figure_2D.pdf", width = 7, height = 6)

textMatrix <- paste0(signif(moduleTraitCor, 2), "\n(", signif(moduleTraitPvalue, 1), ")")
dim(textMatrix) <- dim(moduleTraitCor)

par(mar = c(1.5, 2, 1, 0.5))

labeledHeatmap(
  Matrix = moduleTraitCor,
  xLabels = colnames(datTraits),        # HC + HF
  yLabels = names(MEs),
  colors = blueWhiteRed(50),
  textMatrix = textMatrix,
  setStdMargins = FALSE,
  cex.text = 0.6,
  zlim = c(-1, 1),
  main = "Module-Trait Correlation",
  xLabelsAngle = 0,        
  xLabelsAdj = 0.5          
)
dev.off()
```


## Step 7: Extract DCM_HF‑related modules and candidate hub genes

---

- Figure_2E.

- Figure_2F.

```{r}
key_modules <- names(MEs)[abs(moduleTraitCor) > 0.5 & moduleTraitPvalue < 0.05]
key_modules <- gsub("ME","",key_modules)

print(key_modules)

mod_cor_HF = moduleTraitCor[,"DCM_HF"]
mod_p_HF  = moduleTraitPvalue[,"DCM_HF"]
mod_name = sub("ME","",rownames(moduleTraitCor))

mod_df = data.frame(
  Module = mod_name,
  Cor_HF = mod_cor_HF,
  Pval_HF = mod_p_HF,
  AbsCor = abs(mod_cor_HF)
) 

mod_sig <- mod_df %>%
  dplyr::filter(Pval_HF < 0.05, AbsCor >= 0.5) %>%
  arrange(desc(AbsCor))

# Calculate kME values for all genes, Module membership degree, screening the internal core of the module
kME <- signedKME(datExpr, MEs)
colnames(kME) <- paste0("kME.", gsub("kME", "", colnames(kME)))

# Calculate GS (Gene Disease Correlation)
colnames(datTraits) <- levels(factor(pheno$group))
HF <- datTraits[, "DCM_HF"] 

GS <- apply(datExpr, 2, function(x) cor(x, HF, use = "p"))
GS_p <- apply(datExpr, 2, function(x) corPvalueStudent(cor(x, HF, use = "p"), nSamples = nrow(datExpr)))

geneInfo <- data.frame(
  Gene = colnames(datExpr),
  Module = moduleColors,
  GS_HF = GS,
  GS_pval = GS_p
)
geneInfo <- cbind(geneInfo, kME)

# Screening Hub genes (Key Modules + |kME| > 0.6 + |GS| > 0.2)

mm_cut <- 0.6
gs_cut <- 0.2

# wgcna genes
wgcna_Genes <- geneInfo %>%
  subset(Module %in% mod_sig$Module) %>%
  rowwise() %>%
  mutate(
    own_kME = get(paste0("kME.", Module)), 
    own_abs_kME = abs(own_kME)
  ) %>%
  subset(own_abs_kME > mm_cut & abs(GS_HF) > gs_cut)

cat("Number of screened wgcna genes: ", nrow(wgcna_Genes), "\n")
table(wgcna_Genes$Module)


# M~M
cor_ME <- cor(MEs)
mod_names <- gsub("ME","",colnames(MEs))
colnames(cor_ME) <- mod_names
rownames(cor_ME) <- mod_names
wgcnacol <- blueWhiteRed(100)

pdf("../supplementary/figures/Figure_S2C.pdf", width=7, height=5)
par(mar = c(0, 0, 0, 0), oma = c(0, 0, 0, 0))
corrplot(
  cor_ME,
  method = "circle",
  type = "upper",
  diag = FALSE,
  col = wgcnacol,
  tl.col = "black", 
  # tl.col = mod_names,
  tl.cex = 0.90,
  tl.srt = 45,
  mar = c(0, 0, 0, 0),   
  cl.pos = "r",
  cl.ratio = 0.15,
  addgrid.col = NA           
)
dev.off()


# |GS|~|MM|
library(cowplot)

select_mod <- mod_sig$Module

plot_list <- lapply(select_mod, function(mod) {
  df <- geneInfo %>% filter(Module == mod)
  df$MM <- abs(df[[paste0("kME.", mod)]])
  df$GS_abs <- abs(df$GS_HF)
  df$group <- ifelse(df$MM > mm_cut & df$GS_abs > gs_cut, "selected", "unselected")
  
  ggplot(df, aes(x = MM, y = GS_abs)) +
    geom_point(aes(color = group), size = 1.3, alpha = 0.6) +
    scale_color_manual(values = c(selected = "#E6550D", unselected = "gray70")) +
    geom_vline(xintercept = mm_cut, linetype = "dashed", col = "red") +
    geom_hline(yintercept = gs_cut, linetype = "dashed", col = "red")  + xlab("|MM|") + ylab("|GS|") +
    theme_bw() + theme(legend.position = "none")
})

pdf("../main/figures/Figure_2E.pdf", width=7, height=6)
plot_grid(plotlist = plot_list, ncol=3)
dev.off()

wgcna_genes_df <- wgcna_Genes[,"Gene"]
save(wgcna_genes_df, file =  "../data/processed/geo/gse57338_wgcna_genes.RData")
```

## Step 8: geo_genes: intersecting DEGs and WGCNA

- Figure 2G.

```{r}
rm(list = ls())

load("../data/processed/geo/gse57338_degs_results.RData")
load("../data/processed/geo/gse57338_wgcna_genes.RData")

deg_genes <- rownames(deg_sig) 
wgcna_genes <- wgcna_genes_df$Gene 

# geo_genes: intersecting DEGs and WGCNA
geo_genes <- intersect(deg_genes,wgcna_genes)

deg_genes_df <- data.frame(Gene = deg_genes)

geo_genes_df <- data.frame(Gene = geo_genes) %>%
  arrange(Gene)

gene_list <- list(
  DEGs = deg_genes,
  `WGCNA Genes` = wgcna_genes
)

p <- ggvenn(
  gene_list,
  fill_color = pal_aaas()(2),
  stroke_color = "white",
  stroke_size = 1,
  set_name_size = 5,
  text_size = 5,
  text_color = "white",
  show_percentage = FALSE
) +
  theme(plot.title = element_text(hjust = 0.5, size=10))

# save
ggsave("../main/figures/Figure_2F.pdf", p, width=6, height=6, dpi=300)


save(deg_genes_df, wgcna_genes_df,geo_genes_df, file = "../data/processed/geo/geo_genes.RData")

write.xlsx(
  list(
    DEGs = deg_genes_df,
    WGCNAGenes = wgcna_genes_df
  ),
  "../supplementary/tables/Table_S3.xlsx"
)

deg_candidate <- deg_sig[rownames(deg_sig) %in% geo_genes, ]

up_genes <- rownames(deg_candidate[ deg_candidate$significance =="Up(151)",])

down_genes <- rownames(deg_candidate[ deg_candidate$significance =="Down(162)",])

cat("Up genes: ", length(up_genes), "\n")
print(up_genes)

cat("\nDown genes: ", length(down_genes), "\n")
print(down_genes)
```

## Figure 2
```{r}
files <- c(
  "../main/figures/Figure_2A.pdf",
  "../main/figures/Figure_2B.pdf",
  "../main/figures/Figure_2C.pdf",
  "../main/figures/Figure_2D.pdf",
  "../main/figures/Figure_2E.pdf"
)

labels <- c("A", "B", "C", "D", "E")

pdfs <- list()
for (i in seq_along(files)) {
  img <- image_read_pdf(files[i], density = 300)
  img <- image_annotate(img, labels[i], 
                        gravity = "northwest",  
                        location = "+10+10",    
                        size = 12,                                 weight = 700,    
                        color = "black")
  pdfs[[i]] <- img
}

row1 <- image_append(c(pdfs[[1]],pdfs[[2]]), stack = FALSE)
row2 <- image_append(c(pdfs[[3]]), stack = FALSE)
row3 <- image_append(c(pdfs[[4]],pdfs[[5]]), stack = FALSE)

combined <- image_append(c(row1, row2, row3), stack = TRUE)

image_write(combined, "../main/figures/Figure_2.pdf", format = "pdf")
```


# Genetic genes extracted from HERMES2.0 GWAS
## Step 1: Load HERMES2.0 GWAS summary 
```{r}
rm(list = ls())
gc()

gwas_file <- "../data/raw/opengwas/HERMES2_GWAS_DCM_EUR/FORMAT-METAL_Pheno5-DCM_EUR.tsv.gz"

gwas <- fread(gwas_file, header = TRUE, sep = "\t", data.table = FALSE)
```

## Step2: Screening significant SNPs （P < 5e-8）
```{r}
sig_snps <- gwas %>% 
  filter(pval < 5e-8) %>%
  distinct(chr, pos_b37, .keep_all = TRUE)
cat(sprintf("Number of significant SNPs: %d\n", nrow(sig_snps)))
```

## Step 3: Gene annotation
```{r}
txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene  
genes_gr <- genes(txdb, single.strand.genes.only = FALSE) 

gene_ids <- names(genes_gr)

gene_sym <- AnnotationDbi::select(org.Hs.eg.db, keys = gene_ids, columns = "SYMBOL", keytype = "ENTREZID")

gene_sym_vec <- setNames(gene_sym$SYMBOL, gene_sym$ENTREZID)
```

## 5. Constructing GRanges objects with significant SNPs
```{r}
sig_gr <- GRanges(
  seqnames = paste0("chr", sig_snps$chr),
  ranges = IRanges(start = sig_snps$pos_b37 - 50000, end = sig_snps$pos_b37 + 50000),
  rsID = sig_snps$rsID,
  pval = sig_snps$pval,
  pos = sig_snps$pos_b37
)

overlaps <- findOverlaps(sig_gr, genes_gr)
neighbor_genes <- data.frame(
  SNP_rsID = sig_gr$rsID[queryHits(overlaps)],
  SNP_chr = as.character(seqnames(sig_gr)[queryHits(overlaps)]),
  SNP_pos = sig_gr$pos[queryHits(overlaps)],
  Gene_ENTREZID = names(genes_gr)[subjectHits(overlaps)],
  stringsAsFactors = FALSE
) %>% distinct()

# Add gene symbol
neighbor_genes$Gene_SYMBOL <- gene_sym_vec[neighbor_genes$Gene_ENTREZID]

gwas_genes <- unique(neighbor_genes$Gene_SYMBOL[!is.na(neighbor_genes$Gene_SYMBOL)])

gwas_genes_df <- data.frame(Gene = gwas_genes)

save(
  sig_snps,
  gwas_genes_df,
  file = "../data/processed/opengwas/GWAS_neighbor_genes.RData"
)
write.xlsx(gwas_genes, "../data/processed/opengwas/GWAS_neighbor_genes.xlsx")
```

# Known pathogenic genes of DCM

```{r}
clinGen <- c(
  "ACTC1", "ACTN2", "BAG3", "BAG5", "DES", "DSP", "FLII", "FLNC",
  "JPH2", "LDB3", "LMNA", "LMOD2", "MYH7", "MYLK3", "MYZAP", "NKX2-5",
  "NEXN", "NRAP", "PLN", "PPA2", "PLEKHM2", "PPP1R13L", "PRDM16", "RBM20",
  "RPL3L", "SCN5A", "TBX20", "TNNI3", "TNNI3K", "TNNC1", "TNNT2", "TPM1",
  "TTN", "VCL"
)

review_genes <- c(
  "TTN", "MYH7", "TNNT2",
  "LMNA", "EMD",
  "SCN5A", "CACNA1C", "RYR2",
  "DSC2", "DSP",
  "RBM20", "KLF13", "ETS1", "BMP10"
)

pathogenic_genes <- union(clinGen, review_genes)

pathogenic_genes_df <- data.frame(Gene = pathogenic_genes)

save(clinGen, review_genes, pathogenic_genes_df, file = "../data/processed/opengwas/pathogenic_genes.Rdata")
```

# DCM_HF-related genes
```{r}
rm(list = ls())

load("../data/processed/geo/geo_genes.RData")

load("../data/processed/opengwas/GWAS_neighbor_genes.RData")
load("../data/processed/opengwas/pathogenic_genes.Rdata")

disease_genes <- Reduce(union, list(deg_genes_df$Gene, wgcna_genes_df$Gene, gwas_genes_df$Gene, pathogenic_genes_df$Gene))

disease_genes_df <- data.frame(Gene = disease_genes)

write.xlsx(
  list(
    DEGs = deg_genes_df,
    WGCNA_Genes = wgcna_genes_df,
    GWAS_Genes = gwas_genes_df,
    High_Evidence_Pathogenic_Genes = pathogenic_genes_df,
    DCM_HF_Related_Genes = disease_genes_df
  ),
  "../supplementary/tables/Table_S3.xlsx"
)

save(disease_genes_df, 
     file = "../data/processed/disease_genes_df.RData")
```




