---
title: "01. Collection of bioactive compounds in ZWHQD"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
# Global chunk settings for reproducible output
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

# Extended timeout for large file download
options(timeout = 36000)

# Disable automatic string-to-factor conversion
options(stringsAsFactors = FALSE)

# Download settings for stable HTTPS requests
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

# Set Bioconductor & CRAN mirrors (Tsinghua)
options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

---

# Load Packages
```{r}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("data.table")) install.packages("data.table")
if (!require("openxlsx")) install.packages("openxlsx")

if (!require("future")) install.packages("future")
if (!require("furrr")) install.packages("furrr")
if (!require("progressr")) install.packages("progressr")
if (!require("httr")) install.packages("httr")

library(tidyverse)  #Core toolkit for data manipulation, pipe operations, and data cleaning
library(data.table) # Fast, memory-efficient handling of large datasets
library(openxlsx)   # Read and write Excel files
library(future)     # Parallel processing framework for batch API queries
library(furrr)      # Parallel processing framework for batch API queries
library(progressr)  # Real-time progress tracking for parallel operations
library(httr)       # HTTP request handling for file downloads
```

# Step 1: Load Raw Compound Data from DCABM-TCM Database
```{r dcabm}
# Clear all objects from the working environment
rm(list = ls())

# Import raw compound data
compounds_raw <- read.xlsx("../data/raw/tcm/zwhqd_compound_dcabm.xlsx")

# Save raw data backup for reproducibility
save(
  compounds_raw,
  file = "../data/processed/tcm/01_zwhqd_compounds_raw_dcabm.RData"
)
```

---

# Step 2: Structured processing of raw data

Description: Parse compound metadata.

```{r}
source("utils.R")

# Structured processing of raw data
compounds_structured <- compounds_raw %>%
  # Trim leading/trailing whitespace from all character columns
  mutate(across(where(is.character), trimws)) %>%
  
  # Split multiple compounds separated by "|" into individual rows
  separate_rows(Compounds, sep = "\\|") %>%
  
  # Extract core metadata from the unstructured Compounds string
  mutate(
    CID = str_extract(Compounds, "(?<=Pubchem CID:)[^)]+") %>% str_trim(),
    DCABM.ID = str_extract(Compounds, "(?<=DCABM ID:)[^)]+") %>% str_trim(),
    temp = str_remove(Compounds, "\\s+\\(Pubchem CID:[^)]+\\)")
  ) %>%
  
  # Extract clean compound name and molecular formula
  mutate(
    temp2 = str_remove(temp, "\\([^)]+\\)$") %>% str_trim(),
    Compound = str_remove(temp2, "\\([^)]+\\)$") %>% str_trim(),
    Formula = str_match(temp2, "\\(([^)]+)\\)$")[, 2] %>% str_trim()
  ) %>%
  
  # Remove redundant intermediate columns
  dplyr::select(-Compounds, -temp, -temp2, -DCABM.ID) %>%
  
  # Final whitespace standardization for all character columns
  mutate(across(where(is.character), trimws)) %>%
  
  dplyr::filter(!is.na(CID)) %>%
  
  dplyr::select(-Type) %>%
  
  # deduplication
  distinct(Herb.Name.Chinese,CID,Compound,Formula,.keep_all = TRUE)

# Save cleaned dataset
save(
  compounds_structured,
  file = "../data/processed/tcm/02_zwhqd_compounds_raw_structured.RData"
)

openxlsx::write.xlsx(compounds_structured,"../data/processed/tcm/02_zwhqd_compounds_raw_structured.xlsx")
```

---

# Step 3: Load Manually zwhqd-compounds data

- Manually Add Cycloastragenol Entry

- Manually annotated compound

Input:

- Manually annotated Excel file: "../data/processed/tcm/03_zwhqd_compounds_manual_annotation.xlsx"

```{r}
# Load manually curated compound properties
compounds_manual <- read.xlsx("../data/processed/tcm/03_zwhqd_compounds_manual_annotation.xlsx")

# Merge manual annotations with herb metadata, filter, and deduplicate
compounds_cleaned <- compounds_manual %>%
  dplyr::filter(Is.Keep == "Yes") %>%
  dplyr::select(
    -c(CID, Compound, Formula,`Q-Marker`,Revision.Note,Is.Keep,Remark)
  ) %>% 
  dplyr::rename(CID = CID.Modified,
                Compound = Compound.Name.Modified,
                Formula = Formula.Corrected)

table(compounds_cleaned$Herb.Name.Pinyin)

# Save
save(
  compounds_cleaned,
  file = "../data/processed/tcm/04_compounds_cleaned.RData"
)
```

---

# Step 4: Batch Retrieve PubChem Compound Annotations

Description: Use parallel processing to batch-retrieve standardized compound annotations from PubChem, including IUPAC name, molecular formula, molecular weight, and SMILES structure.

```{r}
# load("../data/processed/tcm/04_compounds_cleaned.RData")
source("utils.R")

# Standardize CID format and isolate unique CIDs for efficient PubChem queries
compounds_pubchem <- compounds_cleaned %>%
  mutate(`CID` = clean_cids(CID)) %>%
  filter(!is.na(CID)) %>%
  distinct(CID, .keep_all = TRUE)

# Define PubChem properties to retrieve
properties <- c("IUPACName", "MolecularFormula", "MolecularWeight",
                "SMILES")

# Split unique CIDs into batches of 20 to comply with PubChem API rate limits
unique_cids <- unique(compounds_pubchem$CID)
cid_batches <- split(unique_cids, ceiling(seq_along(unique_cids)/20))

# Set up parallel processing: use all available cores minus 2 to avoid system overload
plan(multisession, workers = availableCores() - 2)

# Execute parallel batch queries with real-time progress tracking
with_progress({
  p <- progressor(steps = length(cid_batches))
  chem_info <- future_map_dfr(cid_batches, function(b) {
    p()
    fetch_batch(properties, b)
  })
})

# Reset to sequential processing mode
plan(sequential)

# Merge PubChem annotations with the curated compound list and standardize final output
compounds_final <- compounds_cleaned %>%
  left_join(chem_info, by = "CID") %>%
  dplyr::select(
    Herb.Name.Chinese,
    Herb.Name.Pinyin, 
    Herb.Name.Latin,
    Herb.Name.English,
    Compound, 
    IUPACName, 
    MolecularFormula, 
    MolecularWeight, 
    CID, 
    SMILES
  ) %>%
  mutate(across(where(is.character), trimws))
```

---

# Step 5: Calculate Cross-Herb Compound Sharing Statistics

Description: Identify compounds shared across multiple herbs, count the number of herbs each compound appears in, and generate a sorted list of shared compounds for downstream analysis.

```{r}
compound_sharing <- compounds_final %>%
  distinct(Herb.Name.Pinyin, CID,.keep_all = TRUE) %>%
  group_by(CID) %>%
  summarise(
    Herb_Count = n(),
    Herbs = paste(unique(Herb.Name.Pinyin), collapse = ", ")
  ) %>%
  filter(Herb_Count >= 2) %>%
  arrange(desc(Herb_Count))  %>%
  left_join(compounds_final[,c("CID","Compound")],by="CID") %>%
  distinct(CID,.keep_all = TRUE)

# Save all final results
save(
  compounds_final, compound_sharing,
  file = "../data/processed/tcm/05_compounds_final.RData"
)
```

---

# Step 6: Batch Download PubChem SDF Structure Files

Description: Download 3D or 2D structure SDF files for all unique compounds from PubChem, with automatic directory creation and download summary reporting.

```{r}
unique_cids <- unique(compounds_pubchem$CID)
sdf_files <- "../data/raw/tcm/sdf/"

results <- download_all_sdf(unique_cids, sdf_files)
```

## Table S1. Bioactive Constituents of ZWHQD.

Description: Rename columns to publication-ready format and export the final compound list as an Excel file.

```{r}
# load("../data/processed/tcm/05_compounds_final.RData")

table_data <- compounds_final %>%
  dplyr::rename(
   ` Herb(Chinese)` = Herb.Name.Chinese,
    `Herb(Pinyin)` = Herb.Name.Pinyin,
    `Herb(Latin)` = Herb.Name.Latin,
   `Herb(English)` = Herb.Name.English,
   `Compound Name` = Compound,
   `Compound Name (IUPAC)` = IUPACName,
   `Molecular Formula (g/mol)` = MolecularFormula,
   `Pubchme CID` = CID,
   `SMILES` = SMILES
  )

# Export to publication-ready Excel file
write.xlsx(
  list(zwhqd_compounds_raw_dcabm = compounds_raw,
       zwhqd_compounds_raw_structured = compounds_structured,
       zwhqd_compounds_cleaned = table_data), 
  "../tables/Table_S1.xlsx"
)

cat("Table S1 has been successfully generated and saved to `../tables/Table_S1.xlsx`\n")
```