---
title: "02. Prediction and Identification of Potential Targets of ZWHQD"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
# Global chunk settings for reproducible output
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

# Extended timeout for large file download
options(timeout = 36000)

# Disable automatic string-to-factor conversion
options(stringsAsFactors = FALSE)

# Download settings for stable HTTPS requests
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

# Set Bioconductor & CRAN mirrors (Tsinghua)
options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

---

# Load Packages
```{r}
if (!require("tidyverse")) install.packages("tidyverse", dependencies = TRUE)
if (!require("data.table")) install.packages("data.table", dependencies = TRUE)
if (!require("openxlsx")) install.packages("openxlsx", dependencies = TRUE)

library(tidyverse)  #Core toolkit for data manipulation, pipe operations, and data cleaning
library(data.table) # Fast, memory-efficient handling of large datasets
library(openxlsx)   # Read and write Excel files
```

---

# Step 1: Extract Targets from BATMAN-TCM
```{r}
# Clear environment to avoid variable conflicts
rm(list = ls())

load("../data/raw/uniprot/uniprot_human_reviewed.RData") 
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry,Gene.Names.primary)

# Load input datasets
load("../data/processed/tcm/05_compounds_final.RData")
load("../data/raw/tcm/batman/batman_target_known.RData")
load("../data/raw/tcm/batman/batman_target_pred.RData")

# Known targets from BATMAN-TCM
compound_target_known_batman <- batman_target_known %>%
  dplyr::filter(CID %in% unique(compounds_final$CID)) %>%
  dplyr::select(CID, Uniprot.ID) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "BATMAN-Known")

# Predicted targets from BATMAN-TCM (Score ≥ 0.80,LR=48,p_adj<0.5)
compound_target_pred_batman <- batman_target_pred %>%
  dplyr::filter(CID %in% unique(compounds_final$CID)) %>%
  dplyr::select(CID, Uniprot.ID, Score) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol), Score >= 0.80) %>%
  dplyr::select(-Score) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "BATMAN-Pred")

length(unique(compound_target_pred_batman$CID))
length(unique(compound_target_pred_batman$Symbol))
length(unique(compound_target_pred_batman$Symbol)) / length(unique(compound_target_pred_batman$CID))

# Save processed BATMAN-TCM targets
save(
  compound_target_known_batman,
  compound_target_pred_batman,
  file = "../data/processed/tcm/07_batman_targets.RData"
)
```

# Step 2: Extract Targets from Super-Pred
```{r}
rm(list = ls())

source("utils.R")


load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry,Gene.Names.primary)

# known targets
compound_target_known_super <- read_compound_target("../data/raw/tcm/super/known") %>%
  dplyr::select(CID, Uniprot.ID = `UniProt ID`) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "SuperPred-Known")

# Predicted targets (Prob >= 60, Model Accuracy >= 85) (modify,70/90)
compound_target_pred_super <- read_compound_target("../data/raw/tcm/super") %>%
  dplyr::filter(Probability >= 60, `Model accuracy` >= 80) %>%
  dplyr::select(CID, Uniprot.ID = `UniProt ID`) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  distinct(CID, Symbol, .keep_all = TRUE)%>%
  mutate(Source = "SuperPred-Pred")

length(unique(compound_target_pred_super$CID))
length(unique(compound_target_pred_super$Symbol))
length(unique(compound_target_pred_super$Symbol)) / length(unique(compound_target_pred_super$CID))

# Save processed Super-Pred targets
save(
  compound_target_known_super,
  compound_target_pred_super,
  file = "../data/processed/tcm/08_super_targets.RData"
)
```

# Step 3: Extract Targets from SwissTargetPrediction
```{r}
rm(list = ls())

source("utils.R")

load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry,Gene.Names.primary)

# Process SwissTargetPrediction results (Probability > 0)

compound_target_pred_swiss <- read_compound_target("../data/raw/tcm/swiss") %>%
  filter(`Probability*` > 0) %>%
  separate_rows(`Common name`, `Uniprot ID`, sep = " ") %>%
  mutate(
    CID = clean_cids(CID),
    Uniprot.ID = gsub("[[:space:]]", "", `Uniprot ID`)
  ) %>%
  dplyr::select(CID, Uniprot.ID) %>%
  distinct(CID, Uniprot.ID, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  distinct(CID, Symbol, .keep_all = TRUE)%>%
  mutate(Source = "SwissTargetPrediction")

length(unique(compound_target_pred_swiss$Symbol))

length(unique(compound_target_pred_swiss$CID))

length(unique(compound_target_pred_swiss$Symbol)) / length(unique(compound_target_pred_swiss$CID))


# Save processed SwissTargetPrediction targets
save(
  compound_target_pred_swiss,
  file = "../data/processed/tcm/09_swiss_targets.RData"
)
```

---

# Step 4: Extract Targets from PharmMapper
```{r}
rm(list = ls())

source("utils.R")

load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry,Entry.Name, Gene.Names.primary)

# Process PharmMapper results (Human only, high-confidence filters) (modify,0.9)
compound_target_pred_pharm <- read_compound_target2("../data/raw/tcm/pharm",skip_row = 1) %>%
  dplyr::filter(grepl("_HUMAN", Uniplot),zscore>0, `Norm Fit`>0.5) %>%
  dplyr::select(CID, Uniplot) %>%
  distinct(CID, Uniplot, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Uniplot" = "Entry.Name")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  dplyr::select(CID,Uniprot.ID = Entry,Symbol) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "PharmMapper")

length(unique(compound_target_pred_pharm$Symbol))

length(unique(compound_target_pred_pharm$CID))

length(unique(compound_target_pred_pharm$Symbol)) / length(unique(compound_target_pred_pharm$CID))

# Save processed PharmMapper targets
save(
  compound_target_pred_pharm,
  file = "../data/processed/tcm/10_pharm_targets.RData"
)
```

---

# Step 5: Extract Targets from SEA
```{r}
rm(list = ls())

source("utils.R")

load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry, Entry.Name, Gene.Names.primary)


# Process SEA results (confidence filters)
# P < 1e-5、MaxTc >= 0.4、Z > 2 (modify)
compound_target_pred_sea <- read.csv("../data/raw/tcm/sea/sea-results.xls") %>%
  mutate(CID = as.numeric(gsub("compound","",Query.ID)), .before = 1) %>%
  arrange(CID) %>%
  mutate(CID = as.character(CID)) %>%
  dplyr::select(-Query.ID) %>%
  dplyr::filter(
    grepl("_HUMAN", Target.ID), 
    P.Value < 1e-5, 
    Max.Tc >= 0.35
  ) %>%
  distinct(CID, Target.ID, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Target.ID" = "Entry.Name")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  dplyr::select(CID,Uniprot.ID = Entry,Symbol) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "SEA")

length(unique(compound_target_pred_sea$Symbol))

length(unique(compound_target_pred_sea$CID))

length(unique(compound_target_pred_sea$Symbol)) / length(unique(compound_target_pred_sea$CID))


# Save processed SEA targets
save(
  compound_target_pred_sea,
  file = "../data/processed/tcm/13_sea_targets.RData"
)
```

---

# Step 6: Integrate All Targets and Deduplicate
```{r}
rm(list = ls())

# Load all processed target datasets
load("../data/processed/tcm/07_batman_targets.RData")
load("../data/processed/tcm/08_super_targets.RData")
load("../data/processed/tcm/09_swiss_targets.RData")
load("../data/processed/tcm/10_pharm_targets.RData")
load("../data/processed/tcm/13_sea_targets.RData")

# Merge all targets 
compound_target <- rbind(
  compound_target_known_batman,
  compound_target_known_super,
  compound_target_pred_batman, 
  compound_target_pred_super, 
  compound_target_pred_swiss,
  compound_target_pred_pharm,
  compound_target_pred_sea
) %>%
  group_by(CID, Symbol) %>%  
  summarise(
    Source = paste(unique(na.omit(Source)), collapse = "; "),
    across(everything(), ~ dplyr::first(.)),  
    .groups = "drop"
  )

# Filter high-confidence targets (known or predicted by ≥2  databases)
compound_target_final <- compound_target %>%
  mutate(
    is_known = str_detect(Source, "Known"),
    platform_count = str_split(Source, "; ") %>%
      map_dbl(~ length(unique(.)))
  ) %>%
  filter(
    is_known | platform_count >= 1
  ) %>%
  dplyr::select(CID,Symbol,Uniprot.ID,Source)

length(unique(compound_target_final$Symbol))

length(unique(compound_target_final$CID))

length(unique(compound_target_final$Symbol)) / length(unique(compound_target_final$CID))

# Save final target list
save(
  compound_target_final,
  file = "../data/processed/tcm/14_compound_target_final.RData"
)
```

---

# Summary: Target Statistics Across Prediction Platforms.
```{r}
# Generate summary table for manuscript
summary_table <- tibble(
  Database = c(
    "BATMAN-TCM (Known)",
    "SuperPred (Known)",
    "BATMAN-TCM (Predicted)",
    "SuperPred (Predicted)",
    "SwissTargetPrediction",
    "PharmMapper",
    "SEA",
    "Final Integrated Results"
  ),
  
  # Number of unique compounds with targets (CID)
  `Number of compounds with targets` = c(
    n_distinct(compound_target_known_batman$CID),
    n_distinct(compound_target_known_super$CID),
    n_distinct(compound_target_pred_batman$CID),
    n_distinct(compound_target_pred_super$CID),
    n_distinct(compound_target_pred_swiss$CID),
    n_distinct(compound_target_pred_pharm$CID),
    n_distinct(compound_target_pred_sea$CID),
    n_distinct(compound_target_final$CID)
  ),
  
  # Number of unique targets based on UniProt ID
  `Unique Targets (UniProt ID)` = c(
    n_distinct(compound_target_known_batman$Uniprot.ID),
    n_distinct(compound_target_known_super$Uniprot.ID),
    n_distinct(compound_target_pred_batman$Uniprot.ID),
    n_distinct(compound_target_pred_super$Uniprot.ID),
    n_distinct(compound_target_pred_swiss$Uniprot.ID),
    n_distinct(compound_target_pred_pharm$Uniprot.ID),
    n_distinct(compound_target_pred_sea$Uniprot.ID),
    n_distinct(compound_target_final$Uniprot.ID)
  )
)

# Save statistics

save(summary_table, file =  "../data/processed/tcm/15_target_statistics.RData")
```

# herb-compound-target
```{r}
rm(list = ls())

load("../data/processed/tcm/05_compounds_final.RData")
load("../data/processed/tcm/14_compound_target_final.RData")

herb_compound_target <- compound_target_final %>%
  left_join(
    compounds_final,
    by = "CID",
    relationship = "many-to-many"
  ) %>%
  dplyr::select(
    Herb.Name.Chinese, Herb.Name.Pinyin, Herb.Name.Latin, Herb.Name.English,
    Compound, CID, MolecularFormula, MolecularWeight, SMILES, IUPACName,
    Symbol, Uniprot.ID
  ) %>%
  distinct(Herb.Name.Chinese, CID, Symbol, .keep_all = TRUE)

length(unique(herb_compound_target$Herb.Name.Chinese))
length(unique(herb_compound_target$CID))
length(unique(herb_compound_target$Symbol))

save(herb_compound_target, file = "../data/processed/tcm/16_herb_compound_target.RData")
```

## Table S2: Potential Targets of ZWHQD.
```{r}
# Export Table S2
write.xlsx(
  list(`Potential Targets of ZWHQD` = herb_compound_target),
  "../tables/Table_S2.xlsx"
)
```