---
title: "03. Acquisition of GEO datasets for DCM‑HF"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

# Load Packages
```{r}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("data.table")) install.packages("data.table")
if (!require("openxlsx")) install.packages("openxlsx")

if (!require("GEOquery")) BiocManager::install("GEOquery")

library(tidyverse)
library(data.table)
library(openxlsx)
library(GEOquery)
```

# GSE57338: GPL11532

313 individuals with/without Heart Failure

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE57338

```{r,warning=FALSE}
rm(list=ls())

raw_dir <- "../data/raw/geo/GSE57338"
if (!dir.exists(raw_dir)) dir.create(raw_dir, recursive = TRUE)

Sys.setlocale("LC_ALL", "C")
gse57338 <- getGEO("GSE57338", destdir = raw_dir, getGPL = TRUE, AnnotGPL = F)
Sys.setlocale("LC_ALL", "")

gse57338_pheno_raw <- pData(gse57338[[1]])
gse57338_exprs_raw <- exprs(gse57338[[1]])
gse57338_gpl <- gse57338[[1]]@annotation

save(
  gse57338_pheno_raw,
  gse57338_exprs_raw,
  gse57338_gpl,
  file = "../data/raw/geo/gse57338_raw.RData"
)
```

# GSE116250: high throughput sequencing, GPL16791

64 samples from human left ventricle tissue: 14 non-failing donors, 37 dilated cardiomyopathy, and 13 ischemic cardiomyopathy

N_DCM/N_NF = 37 / 14

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE116250

```{r,warning=FALSE}
rm(list=ls())

raw_dir <- "../data/raw/geo/GSE5406"
if (!dir.exists(raw_dir)) dir.create(raw_dir, recursive = TRUE)

Sys.setlocale("LC_ALL", "C")
gse5406 <- getGEO("GSE5406", destdir = raw_dir, getGPL = TRUE, AnnotGPL = F)
Sys.setlocale("LC_ALL", "")

gse5406_pheno_raw <- pData(gse5406[[1]])
gse5406_exprs_raw <- exprs(gse5406[[1]])
gse5406_gpl <- gse5406[[1]]@annotation

save(
  gse5406_pheno_raw,
  gse5406_exprs_raw,
  gse5406_gpl,
  file = "../data/raw/geo/gse5406_raw.RData"
)
```

# GSE141910: high throughput sequencing

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE141910

left ventricular transcriptomes of non-failing healthy donors, Peripartum cardiomyopathy (PPCM), Hypertrophic cardiomyopathy (HCM) and Dilated cardiomyopathy (DCM) samples
```{r}
```

# GSE183852: scRNA-Seq

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE183852


```{r}
rm(list=ls())

raw_dir <- "../data/raw/geo/GSE183852"
if (!dir.exists(raw_dir)) dir.create(raw_dir, recursive = TRUE)

Sys.setlocale("LC_ALL", "C")
gse183852  <- getGEO("GSE183852",  destdir = raw_dir, getGPL = TRUE, AnnotGPL = F)
Sys.setlocale("LC_ALL", "")

gse183852_pheno_raw <- pData(gse183852[[1]])
gse183852_gpl <- gse183852[[1]]@annotation

save(
  gse183852_gpl,
  gse183852_pheno_raw,
  file = "../data/raw/geo/gse183852_raw.RData"
)
```

**NOTE**: **Proceed with caution!** GSE183852_DCM_Integrated.Robj.gz is 11.7GB. Long download time.
Alternative: download directly via browser from https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE183852

```{r}
library(curl)

url <- "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE183852&format=file&file=GSE183852%5FIntegrated%5FCounts%2Ecsv%2Egz"
h <- new_handle()
curl_fetch_disk(url, path = "../data/raw/geo/GSE183852/GSE183852_Integrated_Counts.csv.gz",handle = h)

# 11.7 GB
url <- "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE183852&format=file&file=GSE183852%5FDCM%5FIntegrated%2ERobj%2Egz" 

h <- new_handle()
curl_fetch_disk(url, path = "../data/raw/geo/GSE183852/GSE183852_DCM_Integrated.Robj.gz",handle = h)

rm(list=ls())
gc()
```
