---
title: "04. Differentially Expressed Genes (DEGs)  (DEG)"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

# Load Packages
```{r}
if (!require("tidyverse")) install.packages("tidyverse", dependencies = TRUE)
if (!require("data.table")) install.packages("data.table", dependencies = TRUE)
if (!require("openxlsx")) install.packages("openxlsx", dependencies = TRUE)

if (!require("RColorBrewer")) install.packages("RColorBrewer", dependencies = TRUE)
if (!require("patchwork")) install.packages("patchwork", dependencies = TRUE)
if (!require("pheatmap")) install.packages("pheatmap", dependencies = TRUE)
if (!require("ggsci")) install.packages("ggsci", dependencies = TRUE)

if (!require("AnnoProbe")) install.packages("AnnoProbe", dependencies = TRUE)
if (!require("genefilter")) BiocManager::install("genefilter", dependencies = TRUE)

if (!require("limma")) install.packages("limma", dependencies = TRUE)
if (!require("org.Hs.eg.db")) BiocManager::install("org.Hs.eg.db", dependencies = TRUE)


library(tidyverse)
library(data.table)
library(openxlsx)

library(RColorBrewer)
library(patchwork)
library(pheatmap)
library(ggsci)

library(AnnoProbe)
library(genefilter)
library(limma)
library(org.Hs.eg.db)
```

# Step 1: Loading Data
```{r}
rm(list = ls())
load("../data/raw/geo/GSE57338_raw.RData")

HC <- gse57338_pheno_raw$`disease status:ch1` == "non-failing"
HF <- gse57338_pheno_raw$`disease status:ch1` == "idiopathic dilated CMP"

keep_samples <- rownames(gse57338_pheno_raw)[HC |HF]
pheno <- gse57338_pheno_raw[keep_samples, ]

pheno$group <- ifelse(pheno$`disease status:ch1` == "non-failing", "HC", "HF") 
pheno$group <- factor(pheno$group, levels = c("HC", "HF"))

expr_keep <- gse57338_exprs_raw[,keep_samples]
```

---

# Step 2: Preprocessing

- Figure S1A: Distribution of expression density across samples.

- Figure S1B: Boxplot of gene expression levels across samples.

- Figure S1C: Principal component analysis (PCA) of sample expression profiles.

## log2 transformation
```{r}
ex <- expr_keep
qx <- as.numeric(quantile(ex, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))
LogC <- (qx[5] > 100) ||
          (qx[6]-qx[1] > 50 && qx[2] > 0)
if (LogC) { ex[which(ex <= 0)] <- NaN
  expr_keep <- log2(ex) }

# check GPL and annotate the probes to genes.

probe2gene=idmap(gse57338_gpl) 
expr_annotated <- filterEM(expr_keep, probe2gene)

# Processing of duplicated genes
expr <- as.matrix(expr_annotated)
if (any(duplicated(rownames(expr)))) {
  expr <- aggregate(expr, by = list(rownames(expr)), FUN = mean)
  rownames(expr) <- expr[,1]
  expr <- expr[,-1]
  expr <- as.matrix(expr)
}
```

## Figure S1A
```{r}
# Express the overall distribution of values and observe where the "tail" of background noise falls
df <- reshape2::melt(expr)
p <- ggplot(df, aes(x = value, color = Var2)) + 
  geom_density(show.legend = FALSE) + 
  # labs(title = "Expression density across samples") +
  xlab("log2 expression") + 
  theme_bw() +
  scale_x_continuous(breaks = seq(from = floor(min(df$value)), to = ceiling(max(df$value)), by = 1)) +
  geom_vline(xintercept = 3, linetype = "dashed", color = "blue", linewidth = 1) 

print(p)

ggsave("../figures/supplementary/Figure_S1A.pdf",p, width = 14,height = 8,units = "cm", dpi = 300)

```

# Filtering low expression genes
```{r}
threshold_log2 <- 3
ffun <- filterfun(pOverA(p = 0.5, A = threshold_log2))
keep <- genefilter(expr, ffun)

expr_filt <- expr[keep, ]
dim(expr_filt)
```

## Correlation within the group
```{r}
cor_mat <- cor(expr_filt)

within_cor <- sapply(colnames(expr_filt), function(s) {
  idx <- which(pheno$geo_accession == s)
  gr <- pheno$group[idx]
  same_group <- pheno$geo_accession[pheno$group == gr]
  same_group <- setdiff(same_group, s)
  
  if (length(same_group) == 0) return(NA)
  mean(cor_mat[s, same_group], na.rm = TRUE)
})

between_cor <- sapply(colnames(expr_filt), function(s) {
  idx <- which(pheno$geo_accession == s)
  gr <- pheno$group[idx]
  diff_group <- pheno$geo_accession[pheno$group != gr]
  if(length(diff_group)==0) return(NA)
  mean(cor_mat[s, diff_group], na.rm = TRUE)
})

mean_within = mean(within_cor, na.rm = TRUE)
mean_between = mean(between_cor, na.rm = TRUE)

cat("Average correlation coefficient within the group：", round(mean_within, 4), "\n")
cat("Average correlation coefficient between groups：", round(mean_between, 4), "\n")

if (mean_within > mean_between) {
  cat("✅ No significant batch effect\n")
} else {
  cat("⚠️ Have significant batch effect, suggest batch correction(ComBat/SVA)\n")
}

save(pheno, expr_filt, file = "../data/processed/geo/gse57338.RData")
```

## Figure S1B
```{r}
# Check normalization
pdf("../figures/supplementary/Figure_S1B.pdf", width = 14, height = 8)
boxplot(expr_filt, 
        las = 2,  
        # main = "Check normalization", 
        ylab = "log2 expression", 
        col = "lightblue")

abline(h = median(expr_filt, na.rm = TRUE), col = "red", lty = 2)
dev.off()
```

## Figure S1C
```{r}
# Check batch effect
# PCA
pca <- prcomp(t(expr_filt), scale. = TRUE)
pca_df <- data.frame(pca$x[,1:2], Group = pheno$group)
p <- ggplot(pca_df, aes(PC1, PC2, color = Group)) + 
  geom_point(size = 1) + 
  stat_ellipse() + 
  theme_minimal() + 
  scale_color_aaas()
ggsave("../figures/supplementary/Figure_S1C.pdf",p, width = 7,height = 5,units = "cm", dpi = 300)
```

# Step 3: DEG
```{r}
load("../data/processed/geo/gse57338.RData")
group <- factor(pheno$group)  
design <- model.matrix(~ 0 + group)
colnames(design) <- levels(group)
contr <- paste(levels(group)[2], levels(group)[1], sep = "-")
contr.matrix <- makeContrasts(contrasts = contr, levels = design)

fit <- lmFit(expr_filt, design)
fit2 <- contrasts.fit(fit, contr.matrix)
fit2 <- eBayes(fit2)

res <- topTable(fit2, adjust = "fdr", number = Inf)

save(res, file = "../data/processed/geo/gse57338_res.RData")
```

# Step 4: Significant DEGs
```{r}
# lfc <- log2(1.5)
lfc <- 0.585

deg_all <- as.data.frame(res) %>%
  dplyr::rename(log2FC = logFC)

deg_up <- deg_all %>% filter(adj.P.Val < 0.05, log2FC > lfc)
deg_down <- deg_all %>% filter(adj.P.Val < 0.05, log2FC < -lfc)

cat("The number of all DEGs:", nrow(deg_all), "\n")
cat("The number of up DEGs:", nrow(deg_up), "\n")
cat("The number of down DEGs:", nrow(deg_down), "\n")
```

---

##  Figure 2A: Volcano Plot.

```{r}
deg_all$significance <- ifelse(
  deg_all$`adj.P.Val` < 0.05 & abs(deg_all$log2FC) > lfc,
  ifelse(deg_all$log2FC > lfc, paste0("Up(",nrow(deg_up),")"), paste0("Down(",nrow(deg_down),")")), "NS")

p_volcano <- ggplot(deg_all, aes(x = log2FC, y = -log10(adj.P.Val), color = significance)) +
  geom_point(alpha = 0.7, size = 1) +
  scale_color_manual(values = c(aaas_colors[2], "gray", aaas_colors[1])) +
  geom_vline(xintercept = c(-lfc, lfc), linetype = "dashed", color = "gray40", linewidth = 0.4) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "gray40", linewidth = 0.4) +
  theme_bw(base_size = 10) +
  labs(
    # tag = "A",
    x = "log2(Fold Change)", 
    y = "-log10(Adjusted P-value)") +
  theme_bw() + theme(
    plot.title = element_text(hjust = 0.5)
  )

ggsave("../figures/main/Figure_2A.pdf",p_volcano, width = 7,height = 5,dpi = 300)
```

---

## Figure 2B. Heatmap
```{r}
deg_sig <- deg_all %>% filter(adj.P.Val < 0.05, abs(log2FC) > lfc)

N_top <- 50

up_top <- deg_sig %>%
  filter(log2FC > 0) %>%
  arrange(desc(log2FC)) %>%
  head(N_top)

down_top <- deg_sig %>%
  filter(log2FC < 0) %>%
  arrange(log2FC) %>%
  head(N_top)

heat_genes <- c(rownames(up_top), rownames(down_top))

sample_order <- pheno %>%
  arrange(group) %>%   
  pull(geo_accession)  

mat <- expr_filt[heat_genes, sample_order]

mat_z <- t(scale(t(mat)))

breaks <- seq(-2, 2, length.out = 100)
palette <- colorRampPalette(c(aaas_colors[2], "white", aaas_colors[1]))(99)

# 分组注释
anno_col <- data.frame(
  Group = pheno$group[match(sample_order, pheno$geo_accession)]
)
rownames(anno_col) <- sample_order

p_heat <- pheatmap(
  mat_z,
  annotation_col = anno_col,
  annotation_names_col = F,
  annotation_legend = T,
  annotation_height = 0.2,
  color = palette,
  breaks = breaks,
  scale = "none",      
  show_rownames = FALSE,
  show_colnames = FALSE,
  treeheight_row = 20,
  treeheight_col = 0,
  cluster_cols = FALSE,    
  cluster_rows = TRUE,   
  border_color = NA,
  fontsize = 10
)

pdf("../figures/main/Figure_2B.pdf", width = 7, height = 5)
print(p_heat)
dev.off()
```

# Step 7: Saving DEG Results
```{r}
save(deg_all,deg_sig,deg_up, deg_down, lfc, file = "../data/processed/geo/gse57338_degs_results.RData")
```

## Table S2: DEG Result
```{r}
load("../data/processed/geo/gse57338_degs_results.RData")

deg_all <- deg_all %>%
  arrange(significance)

write.xlsx(
  list(DEG_Result = deg_all),
  "../tables/Table_S3.xlsx"
)
```
