---
title: "06. GWAS Significant Site Adjacent Genes"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

---

# Load packages
```{r}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("data.table")) install.packages("data.table")
if (!require("openxlsx")) install.packages("openxlsx", dependencies = TRUE)

if (!require("GenomicRanges")) BiocManager::install("GenomicRanges")

if (!require("org.Hs.eg.db")) BiocManager::install("org.Hs.eg.db")

if (!require("TxDb.Hsapiens.UCSC.hg19.knownGene")) BiocManager::install("TxDb.Hsapiens.UCSC.hg19.knownGene")

if (!require("CMplot")) install.packages("CMplot")

library(tidyverse)
library(data.table)
library(openxlsx)

library(org.Hs.eg.db)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
library(GenomicRanges)
library(CMplot) # Manhattan Map
# https://ai.so.com/search?search=%E4%BF%9D%E5%A7%86%E7%BA%A7%E6%95%99%E7%A8%8B%21%E5%B0%8F%E6%9E%9C%E6%95%99%E4%BD%A0%E5%AD%A6%E4%B9%A0%E6%9B%BC%E5%93%88%E9%A1%BF%E5%9B%BE%21Manhattan%E5%9B%BE%21+&src=so_result_natural&srcg=so_result_natural&chatid=4945eaf8e25ef908
```

# Load HERMES2.0 GWAS summary 
```{r}
gwas_file <- "../data/raw/opengwas/HERMES2_GWAS_DCM_EUR/FORMAT-METAL_Pheno5-DCM_EUR.tsv.gz"

gwas <- fread(gwas_file, header = TRUE, sep = "\t", data.table = FALSE)
```

# Screening significant SNP （P < 5e-8）
```{r}
sig_snps <- gwas %>% 
  filter(pval < 5e-8) %>%
  distinct(chr, pos_b37, .keep_all = TRUE)
cat(sprintf("显著SNP数量: %d\n", nrow(sig_snps)))
```

# 4. 获取基因注释
```{r}
txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene  
genes_gr <- genes(txdb, single.strand.genes.only = FALSE) 

gene_ids <- names(genes_gr)

gene_sym <- AnnotationDbi::select(org.Hs.eg.db, keys = gene_ids, columns = "SYMBOL", keytype = "ENTREZID")

gene_sym_vec <- setNames(gene_sym$SYMBOL, gene_sym$ENTREZID)

```

# 5. 构建显著SNP的GRanges对象
```{r}
sig_gr <- GRanges(
  seqnames = paste0("chr", sig_snps$chr),
  ranges = IRanges(start = sig_snps$pos_b37 - 50000, end = sig_snps$pos_b37 + 50000),
  rsID = sig_snps$rsID,
  pval = sig_snps$pval,
  pos = sig_snps$pos_b37
)

overlaps <- findOverlaps(sig_gr, genes_gr)
neighbor_genes <- data.frame(
  SNP_rsID = sig_gr$rsID[queryHits(overlaps)],
  SNP_chr = as.character(seqnames(sig_gr)[queryHits(overlaps)]),
  SNP_pos = sig_gr$pos[queryHits(overlaps)],
  Gene_ENTREZID = names(genes_gr)[subjectHits(overlaps)],
  stringsAsFactors = FALSE
) %>% distinct()

# 添加基因符号
neighbor_genes$Gene_SYMBOL <- gene_sym_vec[neighbor_genes$Gene_ENTREZID]

gwas_genes <- unique(neighbor_genes$Gene_SYMBOL[!is.na(neighbor_genes$Gene_SYMBOL)])
writeLines(gwas_genes, "../data/processed/opengwas/GWAS_neighbor_genes.txt")
```

# 6. 已知DCM致病基因
```{r}
known_dcm_genes <- c(
  "ADRB1", "CACNA1C", "BAG3", "TTN", "LMNA", "MYH7", "MYBPC3", "TNNT2", "TNNI3",
  "ACTC1", "TPM1", "MYL2", "MYL3", "ACTN2", "DES", "DMD", "FHL2", "LAMA4", "LDB3",
  "NEXN", "PLN", "PSEN1", "PSEN2", "RBM20", "SCN5A", "SGCD", "TAZ", "TCAP", "VCL",
  "ANKRD1", "CSRP3", "CRYAB", "JPH2", "MYOZ2", "NEBL"
)
```

# 7. 合并集合B
```{r}
setB <- unique(c(gwas_genes, known_dcm_genes))

writeLines(sort(setB), "../data/processed/opengwas/SetB_DCM_genes.txt")

gene_source <- data.frame(
  Gene = setB,
  Source = ifelse(setB %in% gwas_genes, "GWAS_near_sig", "Known_DCM_gene")
)
write.xlsx(gene_source, "../data/processed/opengwas/SetB_DCM_genes_with_source.xlsx")
write.xlsx(neighbor_genes, "../data/processed/opengwas/GWAS_significant_SNPs_neighbor_genes.xlsx")
```

# 10. 输出结果
```{r}
write.xlsx(
  list(gwas_genes = data.frame(Gene = setB)), "../tables/Table_S3_gwas_genes.xlsx")
```

# 11. 曼哈顿图
```{r}
colnames(gwas)[colnames(gwas) == "rsID"] <- "SNP"
colnames(gwas)[colnames(gwas) == "chr"] <- "CHR"
colnames(gwas)[colnames(gwas) == "pos_b37"] <- "BP"
colnames(gwas)[colnames(gwas) == "pval"] <- "P"


```

# 12. QQ图
```{r}
pdf("./figures/QQ_HERMES2.0_DCM.pdf", width = 5, height = 5)
qq(gwas$pval, main = "QQ plot of HERMES2.0 DCM GWAS")
dev.off()
```

# 13. 显著位点邻近基因详情表（补充材料）
```{r}
sig_with_genes <- sig_snps %>%
  left_join(neighbor_genes_raw, by = c("rsid" = "SNP")) %>%
  group_by(rsid) %>%
  summarise(
    CHR = first(chr),
    POS = first(pos_b37),
    P = first(pval),
    Nearby_Genes = paste(unique(Gene_symbol[!is.na(Gene_symbol)]), collapse = ";")
  )
write.csv(sig_with_genes, "./supplementary/Table_S5_GWAS_significant_SNPs_genes.csv", row.names = FALSE)
```

# 14. 集合B来源表
```{r}
setB_summary <- data.frame(
  Gene = setB,
  Source = ifelse(setB %in% known_dcm_genes, "Known DCM gene", "GWAS neighboring gene")
)
write.csv(setB_summary, "./results/05_setB_gwas_genes_with_source.csv", row.names = FALSE)

cat("集合B构建完成！\n")
```


