---
title: "05. Weighted Gene Co-expression Network Analysis (WGCNA)"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

---

# Load packages
```{r}
library(WGCNA)
library(tidyverse)
library(ggvenn)
library(corrplot)
library(flashClust)
library(robustbase) 
library(dynamicTreeCut)
library(openxlsx)
```

---

# Step 1: Loading Data
```{r}
rm(list = ls())

load("../data/processed/geo/gse57338.RData")

# expression matrix
# WGCNA required: Sample (row) x  Gene(column)

dat <- expr_filt

common_samples <- intersect(colnames(dat), rownames(pheno))

dat <- dat[, common_samples, drop = FALSE]
pheno <- pheno[match(common_samples, rownames(pheno)), , drop = FALSE]

if (ncol(dat) == nrow(pheno)) {
  datExpr <- as.data.frame(t(dat))
} else if (nrow(dat) == nrow(pheno)) {
  datExpr <- as.data.frame(dat)
} else {
  stop("Error: Sample dimension mismatch between expression matrix and phenotype data.")
} 
rownames(datExpr) <- common_samples

dim(datExpr)

# trait: group (HC=0, HF=1)
datTraits <- data.frame(
  Group = factor(pheno$group)
)
rownames(datTraits) <- colnames(expr_filt)

datTraits$Disease <- ifelse(pheno$group == "HF", 1, 0)
```


# Step 2: Sample clustering (checking for outlier samples, quality control) 

- Figure S2A. 

- 
```{r}
# goodSamplesGenes
# Quality control: Remove bad genes/samples
gsg <- goodSamplesGenes(datExpr, verbose = 3)
if (!gsg$allOK) {
  if (sum(!gsg$goodGenes) > 0) {
    printFlush(paste("Removing genes:", paste(names(datExpr)[!gsg$goodGenes], collapse = ",")))
  }
  if (sum(!gsg$goodSamples) > 0) {
    printFlush(paste("Removing samples:", paste(rownames(datExpr)[!gsg$goodSamples], collapse = ",")))
  }
  datExpr <- datExpr[gsg$goodSamples, gsg$goodGenes]
}

# Screening for highly variable genes: MAD. Retain the first 5000
gene_mad <- apply(datExpr, 2, mad, na.rm = TRUE) 
N_top <- 5000
n_select <- min(N_top, ncol(datExpr))   
selected_genes <- names(sort(gene_mad, decreasing = TRUE)[1:n_select])
datExpr <- datExpr[, selected_genes]
cat("Retain the number of highly variable genes：", ncol(datExpr), "\n")

# Sample clustering: sampleTree
sampleTree <- hclust(dist(datExpr), method = "average")
group_color = ifelse(pheno$group=="HF","tomato","cyan")
color_mat <- cbind(Group = group_color)

pdf("../figures/supplementary/nFigure_S2A.pdf", width=10, height=5)
plotDendroAndColors(sampleTree, color_mat, 
                    dendroLabels = FALSE, 
                    addGuide = TRUE,       
                    guideCol = "grey"
                    # ,   
                    # main = "Sample clustering"
                    )
dev.off()


save(datTraits,datExpr,pheno,file = "../data/processed/geo/gse57338_datExpr.RData")
```

---

# Step 3: Select soft threshold

- Figure S2B. 

```{r}
# load("../data/processed/geo/gse57338_datExpr.RData")

enableWGCNAThreads()
powers <- c(c(1:10), seq(from = 12, to=30, by=2))

R2_cut <- 0.8

sft <- pickSoftThreshold(
  datExpr,
  powerVector = powers,
  RsquaredCut = R2_cut, 
  verbose = 5,
  networkType = "unsigned"
)
disableWGCNAThreads()

power <- sft$powerEstimate
if(is.na(power)) power <- 8  
cat("power = ", power, "\n")

pdf("../figures/supplementary/nFigure_S2B.pdf", width=14, height=5)
par(mfrow = c(1,2))

# Left: R2
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Power",ylab="Scale Free R^2",type="n")
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers, col="red")
abline(h=0.8, col="blue")

# Right: Connectivity
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Power",ylab="Mean Connectivity",type="n")
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, col="red")
dev.off()

save(sft,file="../data/processed/geo/gse57338_sft.RData")

```

# Step 4: Construct a co-expression network

---

```{r}
enableWGCNAThreads()
set.seed(123)

net <- blockwiseModules(
  datExpr,
  power = power,
  TOMType = "unsigned",
  minModuleSize = 30,      # Minimum number of genes in the module
  mergeCutHeight = 0.25,   # Merge similar modules
  numericLabels = TRUE,
  pamRespectsDendro = FALSE,
  verbose = 3
)
disableWGCNAThreads()

save(net,file = "../data/processed/geo/gse57338_net.RData")

```

# Step 5: Clustering

---

- Figure_2C

```{r}
load("../data/processed/geo/gse57338_net.RData")

# Convert to Color
moduleColors <- labels2colors(net$colors)
table(moduleColors) # Check the number of genes in each module

# Genetree & module color
pdf("../figures/main/nFigure_2C.pdf", width=14, height=5)
plotDendroAndColors(
  net$dendrograms[[1]],
  moduleColors[net$blockGenes[[1]]],
  "Module colors",
  dendroLabels = FALSE,
  hang = 0.03,
  addGuide = TRUE,
  guideHang = 0.05
)
dev.off()
```

# Step 6: Module correlation with diseases

---

- Figure_2D.

```{r}
load("../data/processed/geo/gse57338_datExpr.RData")

MEs <- net$MEs
MEs0 <- moduleEigengenes(datExpr, moduleColors)$eigengenes
MEs <- orderMEs(MEs0)

datTraits <- model.matrix(~0 + group, data = data.frame(group = pheno$group))

colnames(datTraits) <- levels(factor(pheno$group))  
rownames(datTraits) <- rownames(pheno)

moduleTraitCor <- cor(MEs, datTraits, use = "p")
moduleTraitPvalue <- corPvalueStudent(moduleTraitCor, nrow(datExpr))

pdf("../figures/main/nFigure_2D.pdf", width = 7, height = 10)

textMatrix <- paste0(signif(moduleTraitCor, 2), "\n(", signif(moduleTraitPvalue, 1), ")")
dim(textMatrix) <- dim(moduleTraitCor)

labeledHeatmap(
  Matrix = moduleTraitCor,
  xLabels = colnames(datTraits),        # HC + HF
  yLabels = names(MEs),
  colors = blueWhiteRed(50),
  textMatrix = textMatrix,
  setStdMargins = FALSE,
  cex.text = 0.6,
  zlim = c(-1, 1),
  main = "Module-Trait Correlation",
  xLabelsAngle = 0,        
  xLabelsAdj = 0.5          
)
dev.off()
```


# Step 7: Extract DCM_HFrEF‑related modules and candidate hub genes

---

- Figure_2E.

- Figure_2F.

```{r}
key_modules <- names(MEs)[abs(moduleTraitCor) > 0.5 & moduleTraitPvalue < 0.05]
key_modules <- gsub("ME","",key_modules)

print(key_modules)

mod_cor_HF = moduleTraitCor[,"HF"]
mod_p_HF  = moduleTraitPvalue[,"HF"]
mod_name = sub("ME","",rownames(moduleTraitCor))

mod_df = data.frame(
  Module = mod_name,
  Cor_HF = mod_cor_HF,
  Pval_HF = mod_p_HF,
  AbsCor = abs(mod_cor_HF)
) 

mod_sig <- mod_df %>%
  dplyr::filter(Pval_HF < 0.05, AbsCor >= 0.5) %>%
  arrange(desc(AbsCor))

# Calculate kME values for all genes, Module membership degree, screening the internal core of the module
kME <- signedKME(datExpr, MEs)
colnames(kME) <- paste0("kME.", gsub("kME", "", colnames(kME)))

# Calculate GS (Gene Disease Correlation)
colnames(datTraits) <- levels(factor(pheno$group))
HF <- datTraits[, "HF"] 

GS <- apply(datExpr, 2, function(x) cor(x, HF, use = "p"))
GS_p <- apply(datExpr, 2, function(x) corPvalueStudent(cor(x, HF, use = "p"), nSamples = nrow(datExpr)))

geneInfo <- data.frame(
  Gene = colnames(datExpr),
  Module = moduleColors,
  GS_HF = GS,
  GS_pval = GS_p
)
geneInfo <- cbind(geneInfo, kME)

# Screening Hub genes (Key Modules + |kME| > 0.7 + |GS| > 0.2)

mm_cut <- 0.7
gs_cut <- 0.2

# hub genes
hubGenes <- geneInfo %>%
  subset(Module %in% mod_sig$Module) %>%
  rowwise() %>%
  mutate(
    own_kME = get(paste0("kME.", Module)), 
    own_abs_kME = abs(own_kME)
  ) %>%
  subset(own_abs_kME > mm_cut & abs(GS_HF) > gs_cut)

cat("Number of screened Hub genes: ", nrow(hubGenes), "\n")
table(hubGenes$Module)


# M~M
cor_ME <- cor(MEs)
mod_names <- gsub("ME","",colnames(MEs))
colnames(cor_ME) <- mod_names
rownames(cor_ME) <- mod_names
wgcnacol <- blueWhiteRed(100)

pdf("../figures/main/nFigure_2E.pdf", width=7, height=5)
par(mar = c(0, 0, 0, 0), oma = c(0, 0, 0, 0))
corrplot(
  cor_ME,
  method = "circle",
  type = "upper",
  diag = FALSE,
  col = wgcnacol,
  tl.col = "black", 
  # tl.col = mod_names,
  tl.cex = 0.90,
  tl.srt = 45,
  mar = c(0, 0, 0, 0),   
  cl.pos = "r",
  cl.ratio = 0.15,
  addgrid.col = NA           
)
dev.off()


# |GS|~|MM|
library(cowplot)

select_mod <- mod_sig$Module

plot_list <- lapply(select_mod, function(mod) {
  df <- geneInfo %>% filter(Module == mod)
  df$MM <- abs(df[[paste0("kME.", mod)]])
  df$GS_abs <- abs(df$GS_HF)
  df$group <- ifelse(df$MM > mm_cut & df$GS_abs > gs_cut, "selected", "unselected")
  
  ggplot(df, aes(x = MM, y = GS_abs)) +
    geom_point(aes(color = group), size = 1.3, alpha = 0.6) +
    scale_color_manual(values = c(selected = "#E6550D", unselected = "gray70")) +
    geom_vline(xintercept = mm_cut, linetype = "dashed", col = "red") +
    geom_hline(yintercept = gs_cut, linetype = "dashed", col = "red")  + xlab("|MM|") + ylab("|GS|") +
    theme_bw() + theme(legend.position = "none")
})

pdf("../figures/main/nFigure_2F.pdf", width=7, height=5)
plot_grid(plotlist = plot_list, ncol=3)
dev.off()

hub_genes_wgcna <- hubGenes[,"Gene"]
save(hub_genes_wgcna, file =  "../data/processed/geo/gse57338_hub_genes_wgcna.RData")
```

# Step 8: Disease_genes: intersecting DEGs and WGCNA

- Figure 2G.

```{r}
rm(list = ls())

load("../data/processed/geo/gse57338_degs_results.RData")
load("../data/processed/geo/gse57338_hub_genes_wgcna.RData")

deg_genes <- rownames(deg_sig) 
wgcna_genes <- hub_genes_wgcna$Gene

# disease_genes: intersecting DEGs and WGCNA
disease_genes <- intersect(deg_genes,wgcna_genes)

deg_genes_df <- data.frame(Gene = deg_genes) %>%
  arrange(Gene)

wgcna_genes_df <- hub_genes_wgcna %>%
  arrange(Gene)

disease_genes_df <- data.frame(Gene = disease_genes) %>%
  arrange(Gene)

gene_list <- list(
  DEGs = deg_genes,
  `WGCNA Genes` = wgcna_genes
)

library(ggsci)
p <- ggvenn(
  gene_list,
  fill_color = pal_aaas()(2),
  stroke_color = "white",
  stroke_size = 1,
  set_name_size = 5,
  text_size = 5,
  text_color = "white",
  show_percentage = FALSE
) +
  theme(plot.title = element_text(hjust = 0.5, size=10))

# save
ggsave("../figures/main/nFigure_2G.pdf", p, width=6, height=6, dpi=300)


save(deg_genes_df, wgcna_genes_df,disease_genes_df, file = "../data/processed/geo/disease_genes.RData")

write.xlsx(
  list(
    DEGs = deg_genes_df,
    WGCNAGenes = wgcna_genes_df
  ),
  "../tables/Table_S3.xlsx"
)

deg_candidate <- deg_sig[rownames(deg_sig) %in% disease_genes, ]

up_genes <- rownames(deg_candidate[ deg_candidate$significance =="Up(151)",])

down_genes <- rownames(deg_candidate[ deg_candidate$significance =="Down(162)",])

cat("Up genes: ", length(up_genes), "\n")
print(up_genes)

cat("\nDown genes: ", length(down_genes), "\n")
print(down_genes)
```

# Figure 2
```{r}
library(magick)

files <- c(
  "../figures/main/nFigure_2A.pdf",
  "../figures/main/nFigure_2B.pdf",
  "../figures/main/nFigure_2C.pdf",
  "../figures/main/nFigure_2D.pdf",
  "../figures/main/nFigure_2E.pdf",
  "../figures/main/nFigure_2F.pdf",
  "../figures/main/nFigure_2G.pdf"
)

labels <- c("A", "B", "C", "D", "E", "F","G")

pdfs <- list()
for (i in seq_along(files)) {
  img <- image_read_pdf(files[i], density = 300)
  img <- image_annotate(img, labels[i], 
                        gravity = "northwest",  
                        location = "+10+10",    
                        size = 12,                                 weight = 700,    
                        color = "black")
  pdfs[[i]] <- img
}

row1 <- image_append(c(pdfs[[1]],pdfs[[2]]), stack = FALSE)
row2 <- image_append(c(pdfs[[3]]), stack = FALSE)
row3 <- image_append(c(pdfs[[4]],pdfs[[5]]), stack = FALSE)
row4 <- image_append(c(pdfs[[6]],pdfs[[7]]), stack = FALSE)

combined <- image_append(c(row1, row2, row3, row4), stack = TRUE)

image_write(combined, "../figures/main/nFigure_2.pdf", format = "pdf")
image_write(combined, "../figures/main/nFigure_2.png", format = "png", density = 300)
```

---