---
title: "01. Collection and characterization of bioactive ingredients in ZWHQD"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

---
# Loading Packages
```{r}
library(tidyverse)
library(data.table)
library(openxlsx)
```

# Step 1: Loading Raw Data from DCABM-TCM
```{r dcabm}
rm(list = ls())

compounds_raw <- read.xlsx("../data/raw/tcm/zwhqd_compound_dcabm.xlsx")

save(
  compounds_raw,
  file = "../data/processed/tcm/01_compounds_raw.RData"
)
```

---

# Step 2: Raw Data Preprocessing
Parse compound names, CIDs, formulas, and clean formatting.

```{r}
source("utils.R")

compounds_clean <- compounds_raw %>%
  mutate(across(where(is.character), trimws)) %>%
  separate_rows(Compounds, sep = "\\|") %>%
  mutate(
    CID = str_extract(Compounds, "(?<=Pubchem CID:)[^)]+") %>% str_trim(),
    DCABM.ID = str_extract(Compounds, "(?<=DCABM ID:)[^)]+") %>% str_trim(),
    temp = str_remove(Compounds, "\\s+\\(Pubchem CID:[^)]+\\)")
  ) %>%
  mutate(
    temp2 = str_remove(temp, "\\([^)]+\\)$") %>% str_trim(),
    Compound = str_remove(temp2, "\\([^)]+\\)$") %>% str_trim(),
    Formula = str_match(temp2, "\\(([^)]+)\\)$")[, 2] %>% str_trim()
  ) %>%
  dplyr::select(-Compounds, -temp, -temp2, -DCABM.ID) %>%
  mutate(across(where(is.character), trimws))

# Save processed data
save(
  compounds_clean,
  file = "../data/processed/tcm/02_compounds_clean.RData"
)
```

---

# Step 3: Manually Add Cycloastragenol
```{r}
compounds_add <- compounds_clean %>%
  add_row(
    Herb.Name.Chinese = "黄芪",
    Herb.Name.Pinyin = "HUANG QI",
    Herb.Name.English = "Astragali Radix",
    Herb.Name.Latin = "Astragalus membranaceus",
    Type = "Metabolite",
    CID = "13943286",
    Compound = "cycloastragenol",
    Formula = "C30H50O5"
  ) %>%
  arrange(Herb.Name.Pinyin, CID)

save(
  compounds_add,
  file = "../data/processed/tcm/03_compounds_add_cycloastragenol.RData"
)
```

---

# Step 4: Filter Invalid & Duplicate Compounds
```{r}
compounds_filtered <- compounds_add %>%
  filter(CID != "-", !is.na(CID), CID != "") %>%
  mutate(CID = clean_cids(CID)) %>%
  distinct(Herb.Name.Chinese, CID, .keep_all = TRUE) %>%
  distinct(Herb.Name.Chinese, Compound, .keep_all = TRUE)

save(
  compounds_filtered,
  file = "../data/processed/tcm/04_compounds_filtered.RData"
)
```

---

# Step 5: Load Manually Curated Properties & Final Filter

**Input**: Manual annotated compounds file `compounds_properties_manual.xlsx`

```{r}
compounds_manual <- read.xlsx("../data/raw/tcm/compounds_properties_manual.xlsx")
herbs <- compounds_filtered %>%
  dplyr::select(Herb.Name.Chinese,Herb.Name.Pinyin,Herb.Name.Latin,Herb.Name.English) %>%
  distinct(Herb.Name.Pinyin, .keep_all = TRUE)

compounds_final_filtered <- compounds_manual %>%
  dplyr::filter(Drop == "No") %>%
  dplyr::select(
    -c(Note, Compound.Toxic, Drop)
  ) %>% 
  left_join(herbs, by = c("Herb.Name.Pinyin")) %>%
  distinct(Herb.Name.Pinyin, CID, .keep_all = TRUE) %>%
  dplyr::select(Herb.Name.Chinese,Herb.Name.Pinyin,Herb.Name.Latin,Herb.Name.English,CID, Compound.Name, Compound.Category)

save(
  compounds_final_filtered,
  file = "../data/processed/tcm/05_compounds_final_filtered.RData"
)
```

---

# Step 6: Batch Retrieve PubChem Annotations
```{r}
library(future)
library(furrr)
library(progressr)

# load("../data/processed/tcm/05_compounds_final_filtered.RData")

compounds_pubchem <- compounds_final_filtered %>%
  mutate(`CID` = clean_cids(CID)) %>%
  filter(!is.na(CID)) %>%
  distinct(CID, .keep_all = TRUE)

properties <- c("IUPACName", "MolecularFormula", "MolecularWeight",
                "SMILES")

unique_cids <- unique(compounds_pubchem$CID)
cid_batches <- split(unique_cids, ceiling(seq_along(unique_cids)/20))

plan(multisession, workers = availableCores() - 2)
with_progress({
  p <- progressor(steps = length(cid_batches))
  chem_info <- future_map_dfr(cid_batches, function(b) {
    p()
    fetch_batch(properties, b)
  })
})
plan(sequential)


# Merge & Finalize
compounds_final <- compounds_final_filtered %>%
  left_join(chem_info, by = "CID") %>%
  dplyr::select(
    Herb.Name.Chinese,
    Herb.Name.Pinyin, 
    Herb.Name.Latin,
    Herb.Name.English,
    Compound.Name, 
    IUPACName, 
    Compound.Category,
    MolecularFormula, 
    MolecularWeight, 
    CID, 
    SMILES
  ) %>%
  mutate(across(where(is.character), trimws))
```

---

# Step 7: Compound Sharing Statistics (Across Herbs)
```{r}
compound_sharing <- compounds_final %>%
  distinct(Herb.Name.Pinyin, CID,.keep_all = TRUE) %>%
  group_by(CID) %>%
  summarise(
    Herb_Count = n(),
    Herbs = paste(unique(Herb.Name.Pinyin), collapse = ", ")
  ) %>%
  filter(Herb_Count >= 2) %>%
  arrange(desc(Herb_Count))  %>%
  left_join(compounds_final[,c("CID","Compound.Name")],by="CID") %>%
  distinct(CID,.keep_all = TRUE)

# Save all final results
save(
  compounds_final, compound_sharing,
  file = "../data/processed/tcm/06_compounds_final.RData"
)
```

---

# Step 8: Download SDF files
```{r}
library(httr)

unique_cids <- unique(compounds_pubchem$CID)
sdf_files <- "../data/raw/tcm/sdf/"

results <- download_all_sdf(unique_cids, sdf_files)
```

# Table S1. Bioactivate ingredients of ZWHQD.
```{r}
load("../data/processed/tcm/06_compounds_final.RData")

compounds_final <- compounds_final %>%
  dplyr::rename(
   ` Herb(Chinese)` = Herb.Name.Chinese,
    `Herb(Pinyin)` = Herb.Name.Pinyin,
    `Herb(Latin)` = Herb.Name.Latin,
   `Herb(English)` = Herb.Name.English,
   `Ingredient` = Compound.Name,
   `Ingredient(IUPAC)` = IUPACName,
   `Category` = Compound.Category,
   `Molecular Formula (g/mol)` = MolecularFormula,
   `Pubchme CID` = CID,
   `SMILES` = SMILES
  )

write.xlsx(list(Ingredients_ZWHQD = compounds_final), "../tables/Table_S1.xls")
```