---
title: "10. MR + coloc"
format: 
  html:
    toc: true
    number-sections: true
---
  
```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

# Load Packages
```{r}
library(data.table)    # Fast data reading and manipulation
library(dplyr)         # Tidyverse data manipulation framework
library(TwoSampleMR)   # Two-sample Mendelian randomization (MR) analysis
library(MRPRESSO)      # MR sensitivity analysis (pleiotropy testing)
library(coloc)         # Colocalization analysis for genetic signals
library(ggplot2)       # Statistical plotting and visualization
library(patchwork)     # Combine multiple ggplot2 figures into one layout
```

# Step 1: Define file paths and directories
```{r}
rm(list = ls())
gc()

# Outcome GWAS file (DCM)
outcome_file <- "../data/raw/opengwas/Jurgens_DCM_GWAS/DCM_GWAS/Jurgens_DCM_GWAS_META.tsv.gz"

# GTEx v8 eQTL data (Heart Left Ventricle)
eqtl_signif_file <- "../data/raw/opengwas/GTEx_Analysis_v8_eQTL/Heart_Left_Ventricle.v8.signif_variant_gene_pairs.txt.gz"
eqtl_egene_file  <- "../data/raw/opengwas/GTEx_Analysis_v8_eQTL/Heart_Left_Ventricle.v8.egenes.txt.gz"

# Hub genes of interest
hub6 <- c("STAT3", "HIF1A", "SERPINE1", "BCL2L1", "CDKN1A", "CASP1")

# Create output director
out_dir <- "../data/processed/opengwas"
if (!dir.exists(out_dir)) dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)
```

# Step 2: Load GTEx eGenes data and map gene IDs to hub genes
```{r}
egenes <- fread(eqtl_egene_file, header = TRUE, data.table = FALSE)

gene_map <- egenes %>% 
  dplyr::select(gene_id, gene_name) %>%
  distinct(gene_id, .keep_all = TRUE)

hub_genes <- gene_map %>% filter(gene_name %in% hub6)
if(nrow(hub_genes) == 0) stop("Gene_id corresponding to Hub gene not found, please check for gene name matching")

cat("hub genes and their corresponding gene IDs: \n")
print(hub_genes)
```

# Step 3: Load significant eQTL data and retain only associations with Hub genes
```{r}
# Load significant eQTL results, select key columns for downstream analysis
signif <- fread(eqtl_signif_file, 
                select = c("variant_id", "gene_id", "pval_nominal", "slope", "slope_se", 
                           "maf", "ma_samples", "ma_count"),
                data.table = FALSE)

# Rename p-value column for consistency in downstream analysis
names(signif)[names(signif) == "pval_nominal"] <- "pvalue"

# Filter the dataset to keep only eQTLs associated with hub genes
signif_hub <- signif %>% filter(gene_id %in% hub_genes$gene_id)

rm(signif); gc()

cat("Number of eQTL rows after screening:", nrow(signif_hub), "\n")
```

# Step 4: Functions 
```{r}
# Convert GTEx variant_id to GWAS ID_B38 format
gtex_to_gwas_id <- function(variant_id) {
  # Remove "chr" prefix and "_b38" suffix from input ID
  x <- gsub("^chr", "", variant_id)
  x <- gsub("_b38$", "", x)
  
  # Split the ID by underscore into components: chr_pos_ref_alt
  parts <- strsplit(x, "_")[[1]]
  
  if(length(parts) >= 4) {
    # Convert reference and alternate alleles to uppercase
    parts[3] <- toupper(parts[3])
    parts[4] <- toupper(parts[4])
    
    # Reassemble with colon separators for GWAS format
    result <- paste(parts[1], parts[2], parts[3], parts[4], sep=":")
  } else {
    # Fallback: replace underscores with colons and convert to uppercase
    result <- gsub("_", ":", x)
    result <- toupper(result)
  }
  return(result)
}

# Reverse conversion: Convert GWAS ID_B38 back to GTEx variant_id format
# Alleles are converted to lowercase to match the original GTEx format
gwas_to_gtex_id <- function(id_b38) {
  # Input example: "6:36307545:G:A" or "1:54712:TTTTC:T"
  # Output example: "chr6_36307545_g_a_b38" (lowercase alleles)
  
  # Split the ID by colon into components
  parts <- strsplit(id_b38, ":")[[1]]
  
  
  if(length(parts) >= 4) {
    # Convert reference and alternate alleles to lowercase
    parts[3] <- tolower(parts[3])
    parts[4] <- tolower(parts[4])
    
    # Reassemble with underscore separators
    result <- paste(parts[1], parts[2], parts[3], parts[4], sep="_")
  } else {
    # Fallback: replace colons with underscores and convert to lowercase
    result <- gsub(":", "_", id_b38)
    result <- tolower(result)
  }
  
  # Add "chr" prefix and "_b38" suffix to match GTEx format
  result <- paste0("chr", result, "_b38")
  return(result)
}
```

# Step 5: Load and preprocess the outcome GWAS dataset
```{r}
# Load GWAS summary statistics using fast fread function
outcome_gwas <- fread(outcome_file, header = TRUE, data.table = FALSE)

# Define mandatory columns that must exist in the GWAS file
required_columns <- c("ID_B38", "EA", "NEA", "BETA", "SE", "P", "EAFREQ", "N_cases", "N_controls", "rsID")

# Check if all required columns are present
if (!all(required_columns %in% colnames(outcome_gwas))) {
  missing_cols <- setdiff(required_columns, colnames(outcome_gwas))
  stop(
    "The GWAS file is missing required columns: ", paste(missing_cols, collapse = ", "),
    "\nAvailable columns: ", paste(colnames(outcome_gwas), collapse = ", ")
  )
}

# Format GWAS data into GTEx-style standardized columns
# ID_B38 format: "chr:position:reference_allele:alternative_allele"
gwas_clean <- outcome_gwas %>%
  transmute(
    ID_B38 = ID_B38,
    effect_allele = EA,
    other_allele = NEA,
    beta = BETA,
    se = SE,
    pval = P,
    eaf = EAFREQ,
    N_cases = N_cases,
    N_controls = N_controls,
    N_total = N_cases + N_controls
  ) %>%
  # Remove rows with missing key statistical values and alleles
  drop_na(beta, se, pval, effect_allele, other_allele)

# Print the number of variants after quality control filtering
cat("Number of variants after GWAS QC filtering:", nrow(gwas_clean), "\n")
```

# Step 6: Mendelian Randomization (MR) and Colocalization (Coloc) Analyses for Each Hub Gene
```{r}
# ----------------------
# Configuration Parameters ----
# ----------------------
# P-value threshold for selecting significant instrumental variables (IVs)
PVAL_THRESHOLD <- 1e-4
# Minimum number of overlapping SNPs required for coloc analysis
COLOC_MIN_SNPS <- 50
# Sample size for eQTL dataset (GTEx)
EQTL_SAMPLE_SIZE <- 462
# Outcome name (Dilated Cardiomyopathy)
OUTCOME_NAME <- "DCM"

# ----------------------
# Initialize Result Storage ----
# ----------------------
# Lists to store MR and coloc results for all hub genes
all_mr_results <- list()
all_coloc_results <- list()

# ----------------------
# Main Loop: Analyze Each Hub Gene ----
# ----------------------
for (i in seq_len(nrow(hub_genes))) {
  # Extract current gene information
  gene_name <- hub_genes$gene_name[i]
  gene_id <- hub_genes$gene_id[i]
  
  # Print analysis progress
  cat(sprintf("\n==== Processing Hub Gene: %s ======\n", gene_name))
  
  # ----------------------
  # Step 1: Filter eQTL data for the current gene ----
  # ----------------------
  eqtl_df <- signif_hub %>% filter(gene_id == !!gene_id)
  
  # Skip if no eQTL data exists for this gene
  if (nrow(eqtl_df) == 0) {
    cat("Skipped: No eQTL data available\n")
    next
  }
  
  # Filter significant eQTLs based on p-value threshold
  sig_eqtl <- eqtl_df %>% filter(pvalue < PVAL_THRESHOLD)
  
  # Skip if no significant IVs
  if (nrow(sig_eqtl) == 0) {
    cat(sprintf("Skipped: No significant eQTLs (p < %s)\n", PVAL_THRESHOLD))
    next
  }
  cat(sprintf("Number of significant eQTLs (p < %s): %d\n", PVAL_THRESHOLD, nrow(sig_eqtl)))
  
  # ----------------------
  # Step 2: Format Exposure Data (Gene Expression) for TwoSampleMR ----
  # ----------------------
  # Parse SNP IDs to extract effect/reference alleles
  snp_components <- strsplit(sig_eqtl$variant_id, "_")
  
  # Create raw exposure data frame
  exposure_raw <- data.frame(
    SNP = sig_eqtl$variant_id,
    effect_allele = sapply(snp_components, function(x) tolower(x[length(x)-1])),
    other_allele = sapply(snp_components, function(x) tolower(x[length(x)-2])),
    beta = sig_eqtl$slope,
    se = sig_eqtl$slope_se,
    pval = sig_eqtl$pvalue,
    eaf = sig_eqtl$maf,
    samplesize = EQTL_SAMPLE_SIZE,
    stringsAsFactors = FALSE
  )
  
  # Format data to TwoSampleMR standard format
  exposure_dat <- format_data(
    dat = exposure_raw,
    type = "exposure",
    snp_col = "SNP",
    beta_col = "beta",
    se_col = "se",
    effect_allele_col = "effect_allele",
    other_allele_col = "other_allele",
    eaf_col = "eaf",
    pval_col = "pval",
    samplesize_col = "samplesize"
  )
  exposure_dat$exposure <- gene_name
  
  # ----------------------
  # Step 3: Match and Format Outcome Data (DCM GWAS) ----
  # ----------------------
  # Convert SNP IDs to uppercase for consistent matching
  gwas_ids_upper <- sapply(exposure_dat$SNP, function(snp) toupper(gtex_to_gwas_id(snp)))
  gwas_subset <- gwas %>% filter(toupper(ID_B38) %in% gwas_ids_upper)
  
  # Skip if no matching SNPs in GWAS
  if (nrow(gwas_subset) == 0) {
    cat("Skipped: No matching SNPs found in GWAS\n")
    next
  }
  cat(sprintf("Number of matched SNPs in GWAS: %d\n", nrow(gwas_subset)))
  
  # Map back to original GTEx SNP IDs and format outcome data
  gwas_subset$SNP_original <- exposure_dat$SNP[match(toupper(gwas_subset$ID_B38), gwas_ids_upper)]
  outcome_raw <- gwas_subset %>%
    transmute(
      SNP = SNP_original,
      effect_allele = effect_allele,
      other_allele = other_allele,
      beta = beta,
      se = se,
      pval = pval,
      eaf = eaf
    )
  
  # Format outcome data
  outcome_dat <- format_data(
    dat = outcome_raw,
    type = "outcome",
    snp_col = "SNP",
    beta_col = "beta",
    se_col = "se",
    effect_allele_col = "effect_allele",
    other_allele_col = "other_allele",
    eaf_col = "eaf",
    pval_col = "pval"
  )
  outcome_dat$outcome <- OUTCOME_NAME
  
  # ----------------------
  # Step 4: Harmonize Exposure and Outcome Data ----
  # ----------------------
  harmonised_dat <- harmonise_data(exposure_dat, outcome_dat, action = 2)
  
  # Skip if no valid SNPs after harmonization
  if (nrow(harmonised_dat) == 0) {
    cat("Skipped: No valid SNPs after harmonization\n")
    next
  }
  cat(sprintf("Number of valid IVs after harmonization: %d\n", nrow(harmonised_dat)))
  
  # ----------------------
  # Step 5: Calculate F-statistic (IV Strength Evaluation) ----
  # ----------------------
  harmonised_dat <- harmonised_dat %>%
    mutate(F_statistic = (beta.exposure / se.exposure)^2)
  f_median <- median(harmonised_dat$F_statistic, na.rm = TRUE)
  cat(sprintf("Median F-statistic: %.2f\n", f_median))
  
  # ----------------------
  # Step 6: Primary Mendelian Randomization Analysis ----
  # ----------------------
  # Select appropriate MR method based on IV count
  if (nrow(harmonised_dat) >= 2) {
    mr_results <- mr(
      dat = harmonised_dat,
      method_list = c("mr_ivw", "mr_egger_regression", "mr_weighted_median")
    )
  } else {
    mr_results <- mr(dat = harmonised_dat, method_list = "mr_wald_ratio")
  }
  
  # Generate odds ratios and store results
  mr_results <- generate_odds_ratios(mr_results)
  all_mr_results[[gene_name]] <- mr_results
  print(mr_results)
  
  # ----------------------
  # Step 7: Sensitivity Analyses (if ≥3 IVs available) ----
  # ----------------------
  if (nrow(harmonised_dat) >= 3) {
    # MR-Egger intercept test for horizontal pleiotropy
    pleiotropy_test <- mr_pleiotropy_test(harmonised_dat)
    cat(sprintf("MR-Egger Intercept P-value: %s\n", pleiotropy_test$pval))
    
    # MR-PRESSO test (with error handling)
    presso_results <- tryCatch(
      mr_presso(
        BetaOutcome = "beta_outcome",
        BetaExposure = "beta_exposure",
        SdOutcome = "se_outcome",
        SdExposure = "se_exposure",
        data = harmonised_dat,
        OUTLIERtest = TRUE,
        DISTORTIONtest = TRUE
      ),
      error = function(e) NULL
    )
    if (!is.null(presso_results)) {
      cat(sprintf("MR-PRESSO Global Test P-value: %s\n", presso_results$`MR-PRESSO results`$`Global Test`$Pvalue))
    }
  }
  
  # ----------------------
  # Step 8: Generate MR Visualization Plots ----
  # ----------------------
  # Scatter plot
  scatter_plot <- mr_scatter_plot(mr_results, harmonised_dat)[[1]] +
    ggtitle(paste(gene_name, "-", OUTCOME_NAME))
  ggsave(
    filename = paste0("../figures/main/0", i, "_scatter.pdf"),
    plot = scatter_plot,
    width = 7,
    height = 5
  )
  
  # Forest plot (≥2 IVs)
  if (nrow(harmonised_dat) >= 2) {
    forest_plot <- mr_forest_plot(mr_singlesnp(harmonised_dat))[[1]] +
      ggtitle(gene_name)
    ggsave(
      filename = paste0("../figures/main/0", i, "_forest.pdf"),
      plot = forest_plot,
      width = 7,
      height = 5
    )
  }
  
  # Funnel plot & Leave-one-out plot
  funnel_plot <- mr_funnel_plot(mr_singlesnp(harmonised_dat))[[1]] + ggtitle(gene_name)
  ggsave(paste0("../figures/main/0", i, "_funnel.pdf"), funnel_plot, width=7, height=5)
  
  loo_plot <- mr_leaveoneout_plot(mr_leaveoneout(harmonised_dat))[[1]] + ggtitle(gene_name)
  ggsave(paste0("../figures/main/0", i, "_leaveoneout.pdf"), loo_plot, width=7, height=5)
  
  # ----------------------
  # Step 9: Steiger Directionality Test ----
  # ----------------------
  steiger_results <- tryCatch(directionality_test(harmonised_dat), error = function(e) NULL)
  if (!is.null(steiger_results)) {
    cat(sprintf("Steiger Test P-value: %s\n", steiger_results$steiger_pval))
  }
  
  # ----------------------
  # Step 10: Colocalization (Coloc) Analysis ----
  # ----------------------
  cat("Performing colocalization analysis...\n")
  # Prepare full eQTL dataset (no p-value filter)
  full_eqtl <- eqtl_df %>%
    select(SNP = variant_id, beta = slope, se = slope_se, MAF = maf) %>%
    mutate(varbeta = se^2) %>%
    filter(complete.cases(.), MAF > 0, MAF < 1)
  
  # Match SNPs between eQTL and GWAS for coloc
  full_eqtl$ID_B38_upper <- sapply(full_eqtl$SNP, function(snp) toupper(gtex_to_gwas_id(snp)))
  coloc_merged <- gwas %>%
    mutate(ID_B38_upper = toupper(ID_B38)) %>%
    filter(ID_B38_upper %in% full_eqtl$ID_B38_upper) %>%
    select(ID_B38_upper, beta, se, eaf, pval) %>%
    mutate(varbeta = se^2) %>%
    rename(MAF = eaf) %>%
    inner_join(full_eqtl, by = "ID_B38_upper", suffix = c(".outcome", ".exposure"))
  
  # Run coloc only if sufficient overlapping SNPs
  if (nrow(coloc_merged) >= COLOC_MIN_SNPS) {
    # Set default sample sizes if missing
    N_cases <- unique(gwas$N_cases)[1]
    N_controls <- unique(gwas$N_controls)[1]
    if (is.na(N_cases)) {
      N_cases <- 4000
      N_controls <- 16000
    }
    N_total <- N_cases + N_controls
    
    # Prepare coloc input objects
    coloc_exposure <- list(
      beta = coloc_merged$beta.exposure,
      varbeta = coloc_merged$varbeta.exposure,
      type = "quant",
      snp = coloc_merged$SNP,
      MAF = coloc_merged$MAF.exposure,
      N = EQTL_SAMPLE_SIZE
    )
    
    coloc_outcome <- list(
      beta = coloc_merged$beta.outcome,
      varbeta = coloc_merged$varbeta.outcome,
      type = "cc",
      snp = coloc_merged$SNP,
      MAF = coloc_merged$MAF.outcome,
      N = N_total,
      s = N_cases / N_total
    )
    
    # Run coloc analysis
    coloc_results <- coloc.abf(coloc_exposure, coloc_outcome)
    all_coloc_results[[gene_name]] <- coloc_results
    cat("Colocalization Posterior Probabilities:\n")
    print(coloc_results$summary)
    
    # Plot top Coloc PP.H4 values
    pp_data <- data.frame(
      SNP = coloc_merged$SNP,
      PP.H4 = coloc_results$results$SNP.PP.H4
    ) %>%
      arrange(desc(PP.H4))
    
    top_snps <- min(100, nrow(pp_data))
    coloc_plot <- ggplot(pp_data[1:top_snps, ], aes(x = seq_len(top_snps), y = PP.H4)) +
      geom_point() +
      theme_bw() +
      labs(
        title = paste(gene_name, "- Coloc PP.H4 (Top", top_snps, "SNPs)"),
        x = "SNP Index",
        y = "Posterior Probability H4"
      )
    ggsave(paste0("../figures/main/0", i, "_coloc_pph4.pdf"), coloc_plot, width=8, height=4)
    
  } else {
    cat(sprintf("Skipped Coloc: Insufficient overlapping SNPs (n < %d)\n", COLOC_MIN_SNPS))
  }
  
  # ----------------------
  # Step 11: Reverse MR Analysis (DCM → Gene Expression) ----
  # ----------------------
  cat("Performing Reverse MR (DCM → Gene Expression)...\n")
  dcm_significant_snps <- gwas %>% filter(pval < 5e-8)
  
  if (nrow(dcm_significant_snps) > 0) {
    # Convert SNP IDs between GWAS and GTEx
    dcm_significant_snps$GTEx_ID <- sapply(dcm_significant_snps$ID_B38, gwas_to_gtex_id)
    
    # Prepare reverse exposure/outcome data
    reverse_outcome <- eqtl_df %>%
      filter(variant_id %in% dcm_significant_snps$GTEx_ID) %>%
      select(SNP = variant_id, beta = slope, se = slope_se, eaf = maf, pval = pvalue) %>%
      mutate(samplesize = EQTL_SAMPLE_SIZE)
    
    if (nrow(reverse_outcome) > 0) {
      reverse_exposure <- dcm_significant_snps %>%
        filter(GTEx_ID %in% reverse_outcome$SNP) %>%
        mutate(SNP = GTEx_ID, samplesize = N_total)
      
      # Format and harmonize reverse MR data
      rev_exposure_dat <- format_data(reverse_exposure, type = "exposure")
      rev_exposure_dat$exposure <- OUTCOME_NAME
      
      rev_outcome_dat <- format_data(reverse_outcome, type = "outcome")
      rev_outcome_dat$outcome <- gene_name
      
      rev_harmonised <- harmonise_data(rev_exposure_dat, rev_outcome_dat, action = 2)
      
      # Run reverse MR if valid data exists
      if (nrow(rev_harmonised) > 0) {
        reverse_mr_res <- mr(rev_harmonised, method_list = "mr_ivw")
        if (nrow(reverse_mr_res) > 0) {
          cat(sprintf(
            "Reverse MR IVW OR = %.4f, P-value = %s\n",
            exp(reverse_mr_res$b),
            reverse_mr_res$pval
          ))
        }
      }
    }
  } else {
    cat("Skipped Reverse MR: No genome-wide significant SNPs in DCM GWAS\n")
  }
}
```

# Step 7: Save and export Results
```{r}
if(length(all_mr_results) > 0) {
  combined_mr <- bind_rows(all_mr_results, .id = "Gene")
  write.csv(combined_mr, file.path(out_dir, "all_mr_results.csv"), row.names = FALSE)
  cat("所有MR结果已保存至:", file.path(out_dir, "all_mr_results.csv"), "\n")
}

if(length(all_coloc_results) > 0) {
  coloc_summary <- lapply(names(all_coloc_results), function(g) {
    res <- all_coloc_results[[g]]$summary
    data.frame(Gene = g, t(res))
  }) %>% bind_rows()
  write.csv(coloc_summary, file.path(out_dir, "all_coloc_summary.csv"), row.names = FALSE)
  cat("Coloc结果汇总已保存\n")
}

cat("\n===== 分析完成！所有图表和结果保存在目录:", out_dir, "=====\n")
```