---
title: "01. Collection and characterization of bioactive Constituents in ZWHQD"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
# Global chunk settings for reproducible output
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

# Extended timeout for large file download
options(timeout = 36000)

# Disable automatic string-to-factor conversion
options(stringsAsFactors = FALSE)

# Download settings for stable HTTPS requests
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

# Set Bioconductor & CRAN mirrors (Tsinghua)
options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

---
# Load Packages
```{r}
library(tidyverse)  #Core toolkit for data manipulation, pipe operations, and data cleaning
library(data.table) # Fast, memory-efficient handling of large datasets
library(openxlsx)   # Read and write Excel files
library(future)     # Parallel processing framework for batch API queries
library(furrr)      # Parallel processing framework for batch API queries
library(progressr)  # Real-time progress tracking for parallel operations
library(httr)       # HTTP request handling for file downloads
```

# Step 1: Load Raw Compound Data from DCABM-TCM Database
```{r dcabm}
# Clear all objects from the working environment
rm(list = ls())

# Import raw compound data
compounds_raw <- read.xlsx("../data/raw/tcm/zwhqd_compound_dcabm.xlsx")

# Save raw data backup for reproducibility
save(
  compounds_raw,
  file = "../data/processed/tcm/01_compounds_raw.RData"
)
```

---

# Step 2: Raw Data Preprocessing & Standardization

Description: Parse compound metadata, clean formatting, extract core fields (compound name, CID, molecular formula),and standardize all character values.

```{r}
# Load custom utility functions (required for CID cleaning, PubChem data fetching, and SDF downloading)
source("utils.R")

# Data cleaning pipeline
compounds_clean <- compounds_raw %>%
  # Trim leading/trailing whitespace from all character columns
  mutate(across(where(is.character), trimws)) %>%
  
  # Split multiple compounds separated by "|" into individual rows
  separate_rows(Compounds, sep = "\\|") %>%
  
  # Extract core metadata from the unstructured Compounds string
  mutate(
    CID = str_extract(Compounds, "(?<=Pubchem CID:)[^)]+") %>% str_trim(),
    DCABM.ID = str_extract(Compounds, "(?<=DCABM ID:)[^)]+") %>% str_trim(),
    temp = str_remove(Compounds, "\\s+\\(Pubchem CID:[^)]+\\)")
  ) %>%
  
  # Extract clean compound name and molecular formula
  mutate(
    temp2 = str_remove(temp, "\\([^)]+\\)$") %>% str_trim(),
    Compound = str_remove(temp2, "\\([^)]+\\)$") %>% str_trim(),
    Formula = str_match(temp2, "\\(([^)]+)\\)$")[, 2] %>% str_trim()
  ) %>%
  
  # Remove redundant intermediate columns
  dplyr::select(-Compounds, -temp, -temp2, -DCABM.ID) %>%
  
  # Final whitespace standardization for all character columns
  mutate(across(where(is.character), trimws))

# Save cleaned dataset
save(
  compounds_clean,
  file = "../data/processed/tcm/02_compounds_clean.RData"
)
```

---

# Step 3: Manually Add Cycloastragenol Entry

Description: Add cycloastragenol (a key bioactive metabolite from *Astragali Radix*) missing from the raw dataset, then re-sort the dataset for consistent ordering.

```{r}
compounds_add <- compounds_clean %>%
  add_row(
    Herb.Name.Chinese = "黄芪",
    Herb.Name.Pinyin = "HUANG QI",
    Herb.Name.English = "Astragali Radix",
    Herb.Name.Latin = "Astragalus membranaceus",
    Type = "Metabolite",
    CID = "13943286",
    Compound = "cycloastragenol",
    Formula = "C30H50O5"
  ) %>%
  arrange(Herb.Name.Pinyin, CID)

# Save augmented dataset
save(
  compounds_add,
  file = "../data/processed/tcm/03_compounds_add_cycloastragenol.RData"
)
```

---

# Step 4: Filter Invalid & Duplicate Compounds

Description: Remove entries with invalid/empty CIDs, standardize CID format, and deduplicate entries to ensure one unique entry per herb-CID and herb-compound name pair.

```{r}
compounds_filtered <- compounds_add %>%
  # Remove entries with invalid, missing, or empty CIDs
  filter(CID != "-", !is.na(CID), CID != "") %>%
  
  # Standardize CID format using custom utility function
  mutate(CID = clean_cids(CID)) %>%
  
  # Primary deduplication: keep one unique entry per herb-CID pair
  distinct(Herb.Name.Chinese, CID, .keep_all = TRUE) %>%
  
  # Secondary deduplication: keep one unique entry per herb-compound name pair
  distinct(Herb.Name.Chinese, Compound, .keep_all = TRUE)

# Save filtered dataset
save(
  compounds_filtered,
  file = "../data/processed/tcm/04_compounds_filtered.RData"
)
```

---

# Step 5: Merge Manually Curated Properties & Final Filter

Description: Load manually annotated compound properties, filter out entries marked for removal, merge with standardized herb metadata, and generate the final curated compound list.

Input:

- Manually annotated Excel file: "../data/raw/tcm/compounds_properties_manual.xlsx"

```{r}
# Load manually curated compound properties
compounds_manual <- read.xlsx("../data/raw/tcm/compounds_properties_manual.xlsx")

# Extract unique, standardized herb metadata from the filtered dataset
herbs <- compounds_filtered %>%
  dplyr::select(Herb.Name.Chinese,Herb.Name.Pinyin,Herb.Name.Latin,Herb.Name.English) %>%
  distinct(Herb.Name.Pinyin, .keep_all = TRUE)

# Merge manual annotations with herb metadata, filter, and deduplicate
compounds_final_filtered <- compounds_manual %>%
  dplyr::filter(Drop == "No") %>%
  dplyr::select(
    -c(Note, Compound.Toxic, Drop)
  ) %>% 
  left_join(herbs, by = c("Herb.Name.Pinyin")) %>%
  distinct(Herb.Name.Pinyin, CID, .keep_all = TRUE) %>%
  dplyr::select(Herb.Name.Chinese,Herb.Name.Pinyin,Herb.Name.Latin,Herb.Name.English,CID, Compound.Name, Compound.Category)

# Save final curated dataset
save(
  compounds_final_filtered,
  file = "../data/processed/tcm/05_compounds_final_filtered.RData"
)
```

---

# Step 6: Batch Retrieve PubChem Compound Annotations

Description: Use parallel processing to batch-retrieve standardized compound annotations from PubChem, including IUPAC name, molecular formula, molecular weight, and SMILES structure.

```{r}
# load("../data/processed/tcm/05_compounds_final_filtered.RData")

# Standardize CID format and isolate unique CIDs for efficient PubChem queries
compounds_pubchem <- compounds_final_filtered %>%
  mutate(`CID` = clean_cids(CID)) %>%
  filter(!is.na(CID)) %>%
  distinct(CID, .keep_all = TRUE)

# Define PubChem properties to retrieve
properties <- c("IUPACName", "MolecularFormula", "MolecularWeight",
                "SMILES")

# Split unique CIDs into batches of 20 to comply with PubChem API rate limits
unique_cids <- unique(compounds_pubchem$CID)
cid_batches <- split(unique_cids, ceiling(seq_along(unique_cids)/20))

# Set up parallel processing: use all available cores minus 2 to avoid system overload
plan(multisession, workers = availableCores() - 2)

# Execute parallel batch queries with real-time progress tracking
with_progress({
  p <- progressor(steps = length(cid_batches))
  chem_info <- future_map_dfr(cid_batches, function(b) {
    p()
    fetch_batch(properties, b)
  })
})

# Reset to sequential processing mode
plan(sequential)

# Merge PubChem annotations with the curated compound list and standardize final output
compounds_final <- compounds_final_filtered %>%
  left_join(chem_info, by = "CID") %>%
  dplyr::select(
    Herb.Name.Chinese,
    Herb.Name.Pinyin, 
    Herb.Name.Latin,
    Herb.Name.English,
    Compound.Name, 
    IUPACName, 
    Compound.Category,
    MolecularFormula, 
    MolecularWeight, 
    CID, 
    SMILES
  ) %>%
  mutate(across(where(is.character), trimws))
```

---

# Step 7: Calculate Cross-Herb Compound Sharing Statistics

Description: Identify compounds shared across multiple herbs, count the number of herbs each compound appears in, and generate a sorted list of shared compounds for downstream analysis.

```{r}
compound_sharing <- compounds_final %>%
  distinct(Herb.Name.Pinyin, CID,.keep_all = TRUE) %>%
  group_by(CID) %>%
  summarise(
    Herb_Count = n(),
    Herbs = paste(unique(Herb.Name.Pinyin), collapse = ", ")
  ) %>%
  filter(Herb_Count >= 2) %>%
  arrange(desc(Herb_Count))  %>%
  left_join(compounds_final[,c("CID","Compound.Name")],by="CID") %>%
  distinct(CID,.keep_all = TRUE)

# Save all final results
save(
  compounds_final, compound_sharing,
  file = "../data/processed/tcm/06_compounds_final.RData"
)
```

---

# Step 8: Batch Download PubChem SDF Structure Files

Description: Download 3D or 2D structure SDF files for all unique compounds from PubChem, with automatic directory creation and download summary reporting.

```{r}
unique_cids <- unique(compounds_pubchem$CID)
sdf_files <- "../data/raw/tcm/sdf/"

results <- download_all_sdf(unique_cids, sdf_files)
```

# Table S1. Bioactive Constituents of ZWHQD.

Description: Rename columns to publication-ready format and export the final compound list as an Excel file.

```{r}
load("../data/processed/tcm/06_compounds_final.RData")

table_s1_data <- compounds_final %>%
  dplyr::rename(
   ` Herb(Chinese)` = Herb.Name.Chinese,
    `Herb(Pinyin)` = Herb.Name.Pinyin,
    `Herb(Latin)` = Herb.Name.Latin,
   `Herb(English)` = Herb.Name.English,
   `Compound` = Compound.Name,
   `Compound(IUPAC)` = IUPACName,
   `Category` = Compound.Category,
   `Molecular Formula (g/mol)` = MolecularFormula,
   `Pubchme CID` = CID,
   `SMILES` = SMILES
  )

# Export to publication-ready Excel file
write.xlsx(
  list(`Bioactive Constituents of ZWHQD` = table_s1_data), 
  "../tables/Table_S1.xlsx"
)

cat("Table S1 has been successfully generated and saved to `../tables/Table_S1.xlsx`\n")
```