---
title: "10. MR + coloc"
format: 
  html:
    toc: true
    number-sections: true
---
  
```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

# Loading Packages
```{r}
library(data.table)
library(dplyr)
library(TwoSampleMR)
library(MRPRESSO)
library(coloc)
library(ggplot2)
library(patchwork)
library(ieugwasr)
```

# MR

## Step 1: Files and folder
```{r}
rm(list = ls())
gc()

# outcome file
outcome_file <- "../data/raw/opengwas/HERMES2_GWAS_DCM_EUR/FORMAT-METAL_Pheno5-DCM_EUR.tsv.gz"

# GTEx v8 heart left ventricle data
eqtl_signif_file <- "../data/raw/opengwas/GTEx_Analysis_v8_eQTL/Heart_Left_Ventricle.v8.signif_variant_gene_pairs.txt.gz"
eqtl_egene_file  <- "../data/raw/opengwas/GTEx_Analysis_v8_eQTL/Heart_Left_Ventricle.v8.egenes.txt.gz"

# Hub genes
hub6 <- c("STAT3", "HIF1A", "SERPINE1", "BCL2L1", "CDKN1A", "CASP1")

# output folder
out_dir <- "../data/processed/opengwas"
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)
```

## Step 2: Loading GTEx genes data and matching the gene_id for Hub gene
```{r}
egenes <- fread(eqtl_egene_file, header = TRUE, data.table = FALSE)

gene_map <- egenes %>% 
  dplyr::select(gene_id, gene_name) %>%
  distinct(gene_id, .keep_all = TRUE)

hub_genes <- gene_map %>% filter(gene_name %in% hub6)
if(nrow(hub_genes) == 0) stop("Gene_id corresponding to Hub gene not found, please check for gene name matching")

cat("Find the following Hub genes and their gene_id:\n")
print(hub_genes)
```

## Step 3: Loading significant eQTL data, retaining only the association of Hub genes
```{r}
signif <- fread(eqtl_signif_file, 
                select = c("variant_id", "gene_id", "pvalue", "slope", "slope_se", 
                           "maf", "ma_samples", "ma_count"),
                data.table = FALSE)

# Selecting of eQTL for Hub genes
signif_hub <- signif %>% filter(gene_id %in% hub_genes$gene_id)
rm(signif); gc()

cat("Number of eQTL rows after screening:", nrow(signif_hub), "\n")
```

## Step 4: function: Extract IVs (p<5e-8, club) for each gene
```{r}
get_ivs_for_gene <- function(gene_id, pval_thresh = 5e-8, r2 = 0.001, kb = 10000) {
  sub <- signif_hub %>% filter(gene_id == !!gene_id)
  if(nrow(sub) == 0) return(NULL)
  
  sub <- sub %>% filter(pvalue < pval_thresh)
  if(nrow(sub) == 0) return(NULL)
  
  sub <- sub %>%
    mutate(
      chr_pos = gsub("_b38$", "", variant_id),
      chr = as.numeric(gsub("_.*", "", chr_pos)),
      pos = as.numeric(gsub(".*_", "", gsub("_.*_.*", "", chr_pos)))
    ) %>%
    filter(!is.na(chr), !is.na(pos))
  
  if(nrow(sub) == 0) return(NULL)
  
  # LD clumping  IEU OpenGWAS API
  clumped <- tryCatch({
    ieugwasr::ld_clump(
      dat = data.frame(
        SNP = sub$variant_id,
        CHR = sub$chr,
        POS = sub$pos,
        P = sub$pvalue
      ),
      clump_r2 = r2,
      clump_kb = kb,
      plink_bin = NULL,   
      pop = "EUR"
    )
  }, error = function(e) NULL)
  
  if(is.null(clumped) || nrow(clumped) == 0) return(NULL)
  
  ivs <- sub %>% filter(variant_id %in% clumped$SNP)
  return(ivs)
}
```

## Step 5: Loading outcome GWAS data
```{r}
outcome_gwas <- fread(outcome_file, header = TRUE, data.table = FALSE)

outcome_gwas <- outcome_gwas %>%
  rename(
    SNP = SNP,
    chr_name = CHR,
    pos = POS,
    effect_allele = EA,
    other_allele = OA,
    beta = BETA,
    se = SE,
    pval = P,
    eaf = EAF
  )
cat("结局GWAS行数:", nrow(outcome_gwas), "\n")
```