---
title: "02. Prediction and Identification of Potential Targets of ZWHQD"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
# Global chunk settings for reproducible output
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

# Extended timeout for large file download
options(timeout = 36000)

# Disable automatic string-to-factor conversion
options(stringsAsFactors = FALSE)

# Download settings for stable HTTPS requests
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

# Set Bioconductor & CRAN mirrors (Tsinghua)
options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

---

# Step 1: Extract Targets from BATMAN-TCM
```{r}
# Clear environment to avoid variable conflicts
rm(list = ls())

load("../data/raw/uniprot/uniprot_human_reviewed.RData") 
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry,Gene.Names.primary)

# Load input datasets
load("../data/processed/tcm/06_compounds_final.RData")
load("../data/raw/tcm/batman/batman_target_known.RData")
load("../data/raw/tcm/batman/batman_target_pred.RData")

# Known targets from BATMAN-TCM
compound_target_known_batman <- batman_target_known %>%
  dplyr::filter(CID %in% unique(compounds_final$CID)) %>%
  dplyr::select(CID, Uniprot.ID) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "BATMAN-Known")

# Predicted targets from BATMAN-TCM (Score ≥ 0.84)
compound_target_pred_batman <- batman_target_pred %>%
  dplyr::filter(CID %in% unique(compounds_final$CID)) %>%
  dplyr::select(CID, Uniprot.ID, Score) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol), Score >= 0.84) %>%
  dplyr::select(-Score) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "BATMAN-Pred")

# Save processed BATMAN-TCM targets
save(
  compound_target_known_batman,
  compound_target_pred_batman,
  file = "../data/processed/tcm/07_batman_targets.RData"
)
```

# Step 2: Extract Targets from Super-Pred
```{r}
rm(list = ls())

load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry,Gene.Names.primary)

source("utils.R")

# known targets
compound_target_known_super <- read_compound_target("../data/raw/tcm/super/known") %>%
  dplyr::select(CID, Uniprot.ID = `UniProt ID`) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "SuperPred-Known")

# Predicted targets (Prob >= 60, Model Accuracy >= 85) (modify,70/90)
compound_target_pred_super <- read_compound_target("../data/raw/tcm/super") %>%
  dplyr::filter(Probability >= 60, `Model accuracy` >= 85) %>%
  dplyr::select(CID, Uniprot.ID = `UniProt ID`) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  distinct(CID, Symbol, .keep_all = TRUE)%>%
  mutate(Source = "SuperPred-Pred")

# Save processed Super-Pred targets
save(
  compound_target_known_super,
  compound_target_pred_super,
  file = "../data/processed/tcm/08_super_targets.RData"
)
```

# Step 3: Extract Targets from SwissTargetPrediction
```{r}
rm(list = ls())

source("utils.R")

load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry,Gene.Names.primary)

# Process SwissTargetPrediction results (Probability > 0)

compound_target_pred_swiss <- read_compound_target("../data/raw/tcm/swiss") %>%
  filter(`Probability*` > 0) %>%
  separate_rows(`Common name`, `Uniprot ID`, sep = " ") %>%
  mutate(
    CID = clean_cids(CID),
    Uniprot.ID = gsub("[[:space:]]", "", `Uniprot ID`)
  ) %>%
  dplyr::select(CID, Uniprot.ID) %>%
  distinct(CID, Uniprot.ID, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  distinct(CID, Symbol, .keep_all = TRUE)%>%
  mutate(Source = "SwissTargetPrediction")


# Save processed SwissTargetPrediction targets
save(
  compound_target_pred_swiss,
  file = "../data/processed/tcm/09_swiss_targets.RData"
)
```

---

# Step 4: Extract Targets from PharmMapper
```{r}
rm(list = ls())

source("utils.R")

load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry,Entry.Name, Gene.Names.primary)

# Process PharmMapper results (Human only, high-confidence filters) (modify,0.9)
compound_target_pred_pharm <- read_compound_target2("../data/raw/tcm/pharm",skip_row = 1) %>%
  dplyr::filter(grepl("_HUMAN", Uniplot),zscore>0, `Norm Fit`>0.5) %>%
  dplyr::select(CID, Uniplot) %>%
  distinct(CID, Uniplot, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Uniplot" = "Entry.Name")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  dplyr::select(CID,Uniprot.ID = Entry,Symbol) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "PharmMapper")

# Save processed PharmMapper targets
save(
  compound_target_pred_pharm,
  file = "../data/processed/tcm/10_pharm_targets.RData"
)
```

---

# Step 5: Extract Targets from TargetNet
```{r}
rm(list = ls())

source("utils.R")

load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry, Gene.Names.primary)

# Read query and result files
query <- read.xlsx("../data/raw/tcm/targetnet/query.xlsx")
cids<- query$CID
col_names <- c("Uniprot.ID","Protein",cids)

# Process TargetNet results (Probability > 0.5)
result <- read.xlsx("../data/raw/tcm/targetnet/results.xlsx") 
compound_target_pred_targetnet  <- result %>%
  dplyr::select(-Details) %>%
  setNames(col_names) %>%
  pivot_longer(
    cols = -c(Uniprot.ID, Protein),   
    names_to = "CID",                 
    values_to = "Probability"     
  ) %>%
  dplyr::filter(!is.na(Probability), Probability > 0.5) %>%
  dplyr::select(CID, Uniprot.ID) %>% 
  distinct(CID, Uniprot.ID, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "TargetNet")

# Save processed TargetNet targets
save(
  compound_target_pred_targetnet,
  file = "../data/processed/tcm/11_targetnet_targets.RData"
)
```

---

# Step 6: Extract Targets from PPB3
```{r}
rm(list = ls())

source("utils.R")

load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry, Gene.Names.primary)

# Filter human single-protein targets
result <- read_compound_target("../data/raw/tcm/ppb3") %>%
  dplyr::filter( 
    `Target Type` == 'SINGLE PROTEIN', 
    `Target Organism` == "Homo sapiens",
    `Confidence Score` > 0.2
  )

# Convert ChEMBL IDs to UniProt IDs via API
library(httr)
library(jsonlite)

chembl_ids <- unique(result$`Target ID`)
uniprot_list <- sapply(chembl_ids, chembl_to_uniprot)

# Save API mapping result to avoid repeated requests
save(uniprot_list,file = "../data/raw/tcm/ppb3/uniprot_list.RData")

# load("../data/raw/tcm/ppb3/uniprot_list.RData")

# Create mapping dataframe
df <- data.frame(
  Chembl.ID = names(uniprot_list),
  Uniprot.ID = sapply(uniprot_list, `[`, 1),  
  row.names = NULL
)

# Annotate and clean targets
compound_target_pred_ppb3 <- result %>%
  left_join(df, by=c("Target ID" = "Chembl.ID")) %>% 
  distinct(CID, Uniprot.ID, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  dplyr::select(CID,Uniprot.ID,Symbol) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "PPB3")

# Save processed PPB3 targets
save(
  compound_target_pred_ppb3,
  file = "../data/processed/tcm/12_ppb3_targets.RData"
)
```

---

# Step 7: Extract Targets from SEA
```{r}
rm(list = ls())

source("utils.R")

load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry, Entry.Name, Gene.Names.primary)


# Process SEA results (stringent high-confidence filters)
# P < 1e-3、MaxTc > 0.3、Z > 2 (modify)
compound_target_pred_sea <- read.csv("../data/raw/tcm/sea/sea-results.xls") %>%
  mutate(CID = as.numeric(gsub("compound","",Query.ID)), .before = 1) %>%
  arrange(CID) %>%
  mutate(CID = as.character(CID)) %>%
  dplyr::select(-Query.ID) %>%
  dplyr::filter(
    grepl("_HUMAN", Target.ID), 
    P.Value < 1e-3, 
    Max.Tc > 0.3, 
    Z.Score > 2
  ) %>%
  distinct(CID, Target.ID, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Target.ID" = "Entry.Name")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  dplyr::select(CID,Uniprot.ID = Entry,Symbol) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "SEA")

# Save processed SEA targets
save(
  compound_target_pred_sea,
  file = "../data/processed/tcm/13_sea_targets.RData"
)
```

---

# Step 8: Integrate All Targets and Deduplicate
```{r}
rm(list = ls())

# Standardized column names for consistent merging
# cnames <- c(
#   "Herb.Name.Chinese",
#   "Herb.Name.Pinyin", 
#   "Herb.Name.English",
#   "Herb.Name.Latin",
#   "Compound.Name", "IUPACName", "Compound.Category",
#   "MolecularFormula", "MolecularWeight", "CID",
#   "SMILES", "Symbol", "Uniprot.ID"
# )


# Load all processed target datasets
load("../data/processed/tcm/07_batman_targets.RData")
load("../data/processed/tcm/08_super_targets.RData")
load("../data/processed/tcm/09_swiss_targets.RData")
load("../data/processed/tcm/10_pharm_targets.RData")
load("../data/processed/tcm/11_targetnet_targets.RData")
load("../data/processed/tcm/12_ppb3_targets.RData")
load("../data/processed/tcm/13_sea_targets.RData")

# Merge all targets 
compound_target <- rbind(
  compound_target_known_batman,
  compound_target_known_super,
  compound_target_pred_batman, 
  compound_target_pred_super, 
  compound_target_pred_swiss,
  compound_target_pred_pharm,
  compound_target_pred_targetnet,
  compound_target_pred_ppb3,
  compound_target_pred_sea
) %>%
  group_by(CID, Symbol) %>%  
  summarise(
    Source = paste(unique(na.omit(Source)), collapse = "; "),
    across(everything(), ~ dplyr::first(.)),  
    .groups = "drop"
  )

# Filter high-confidence targets (known or predicted by ≥2  databases)
compound_target_final <- compound_target %>%
  mutate(
    is_known = str_detect(Source, "Known"),
    platform_count = str_split(Source, "; ") %>%
      map_dbl(~ length(unique(.)))
  ) %>%
  filter(
    is_known | platform_count >= 2
  ) %>%
  dplyr::select(CID,Symbol,Uniprot.ID,Source)

# Save final target list
save(
  compound_target_final,
  file = "../data/processed/tcm/14_compound_target_final.RData"
)
```

---

# Summary: Target Statistics Across Prediction Platforms.
```{r}
# Generate summary table for manuscript
summary_table <- tibble(
  Database = c(
    "BATMAN-TCM (Known)",
    "SuperPred (Known)",
    "BATMAN-TCM (Predicted)",
    "SuperPred (Predicted)",
    "SwissTargetPrediction",
    "PharmMapper",
    "TargetNet",
    "PPB3",
    "SEA",
    "Final Integrated Results"
  ),
  
  # Number of unique compounds with targets (CID)
  `Number of compounds with targets` = c(
    n_distinct(compound_target_known_batman$CID),
    n_distinct(compound_target_known_super$CID),
    n_distinct(compound_target_pred_batman$CID),
    n_distinct(compound_target_pred_super$CID),
    n_distinct(compound_target_pred_swiss$CID),
    n_distinct(compound_target_pred_pharm$CID),
    n_distinct(compound_target_pred_targetnet$CID),
    n_distinct(compound_target_pred_ppb3$CID),
    n_distinct(compound_target_pred_sea$CID),
    n_distinct(compound_target_final$CID)
  ),
  
  # Number of unique targets based on UniProt ID
  `Unique Targets (UniProt ID)` = c(
    n_distinct(compound_target_known_batman$Uniprot.ID),
    n_distinct(compound_target_known_super$Uniprot.ID),
    n_distinct(compound_target_pred_batman$Uniprot.ID),
    n_distinct(compound_target_pred_super$Uniprot.ID),
    n_distinct(compound_target_pred_swiss$Uniprot.ID),
    n_distinct(compound_target_pred_pharm$Uniprot.ID),
    n_distinct(compound_target_pred_targetnet$Uniprot.ID),
    n_distinct(compound_target_pred_ppb3$Uniprot.ID),
    n_distinct(compound_target_pred_sea$Uniprot.ID),
    n_distinct(compound_target_final$Uniprot.ID)
  )
)

# Save statistics

save(summary_table, file =  "../data/processed/tcm/15_target_statistics.RData")
```

# herb-compound-target
```{r}
rm(list = ls())

load("../data/processed/tcm/06_compounds_final.RData")
load("../data/processed/tcm/14_compound_target_final.RData")

herb_compound_target <- compound_target_final %>%
  left_join(
    compounds_final,
    by = "CID",
    relationship = "many-to-many"
  ) %>%
  dplyr::select(
    Herb.Name.Chinese, Herb.Name.Pinyin, Herb.Name.Latin, Herb.Name.English,
    Compound.Name, CID, MolecularFormula, MolecularWeight, SMILES, IUPACName,
    Symbol, Uniprot.ID
  ) %>%
  distinct(Herb.Name.Chinese, CID, Symbol, .keep_all = TRUE)

length(unique(herb_compound_target$Herb.Name.Chinese))
length(unique(herb_compound_target$CID))
length(unique(herb_compound_target$Symbol))

save(herb_compound_target, file = "../data/processed/tcm/16_herb_compound_target.RData")
```

# Table S2: Potential Targets of ZWHQD.
```{r}
# Export Table S2
write.xlsx(
  list(`Potential Targets of ZWHQD` = herb_compound_target),
  "../tables/Table_S2.xlsx"
)
```