---
title: "Part 3: Mechanism exploration of ZWHQD against HFrEF"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

---

# Identification of Candidate Targets

## Load Packages
```{r}
library(ggvenn)
library(tidyverse)
library(openxlsx)
library(ggsci)
```

## Candidate Genes
```{r}
rm(list = ls())

load("../data/processed/geo/GSE141910/degs_results.RData")

deg_sig <- deg_all %>% filter(padj < 0.05, abs(log2FoldChange) > lfc)

deg_genes <- rownames(deg_sig) 

load("../data/processed/tcm/14_compound_target_final.RData")

herb_targets <- unique(compound_target_final$Symbol)

load("../data/processed/geo/GSE141910/wgcna_candidates_genes.RData")

# disease_genes: intersecting DEGs and WGCNA
disease_genes <- intersect(deg_genes,wgcna_candidates_genes$Gene)

# candidate_genes: intersecting herb_targets and disease_genes
candidate_genes <-intersect(herb_targets, disease_genes)

herb_targets_df <- data.frame(Gene = herb_targets) %>%
  arrange(Gene)

deg_genes_df <- data.frame(Gene = deg_genes) %>%
  arrange(Gene)

wgcna_genes_df <- data.frame(Gene = wgcna_candidates_genes$Gene) %>%
  arrange(Gene)

disease_genes_df <- data.frame(Gene = disease_genes) %>%
  arrange(Gene)

candidate_genes_df <- data.frame(Gene = candidate_genes) %>%
  arrange(Gene)

gene_list <- list(
  DEGs = deg_genes,
  `Herb targets` = herb_targets,
  `WGCNA Genes` = wgcna_candidates_genes$Gene
)

p <- ggvenn(
  gene_list,
  fill_color = pal_aaas()(4),
  stroke_color = "white",
  stroke_size = 1,
  set_name_size = 5,
  text_size = 5,
  text_color = "white",
  show_percentage = FALSE
) +
  theme(plot.title = element_text(hjust = 0.5, size=10))

# save
ggsave("../figures/main/Figure_3A.pdf", p, width=6, height=6, dpi=300)
ggsave("../figures/main/Figure_3A.png", p, width=6, height=6, dpi=300)


save(herb_targets_df, deg_genes_df, wgcna_genes_df,disease_genes_df, candidate_genes_df, file = "../data/processed/network/01_herb_disease_genes.RData")

write.xlsx(
  list(
    HerbTargets = herb_targets_df,
    DEGs = deg_genes_df,
    WGCNAGenes = wgcna_genes_df,
    DiseaseGenes = disease_genes_df,
    CandidateGenes = candidate_genes_df
  ),
  "../tables/candidate_genes.xlsx"
)

deg_candidate <- deg_all[rownames(deg_all) %in% candidate_genes, ]
 
deg_candidate$regulation <- ifelse(
  deg_candidate$log2FoldChange >= lfc, "Up",
  ifelse(deg_candidate$log2FoldChange <= -lfc, "Down", "NS")
)

deg_candidate$Gene <- rownames(deg_candidate)

up_genes <- deg_candidate$Gene[deg_candidate$regulation == "Up"]
down_genes <- deg_candidate$Gene[deg_candidate$regulation == "Down"]

cat("Up genes: ", length(up_genes), "\n")
print(up_genes)

cat("\nDown genes: ", length(down_genes), "\n")
print(down_genes)
```

---

# Herb-Ingredient-Target (HIT) Network
## Load Packages

```{r}
library(tidyverse)
library(openxlsx)
library(data.table)

library(clusterProfiler)
library(igraph)
library(centiserve)
library(RCy3)
```

## Step 1: Loading data
```{r}
rm(list = ls())

load("../data/processed/network/01_herb_disease_genes.RData")
load("../data/processed/tcm/14_compound_target_final.RData")

herb_ingredient_target <- candidate_genes_df %>%
  left_join(compound_target_final, by = c("Gene" = "Symbol"),relationship = "many-to-many") %>%
  dplyr::select(Herb.Name.Pinyin, Compound.Name, Gene) %>%
  dplyr:: rename(Herb = Herb.Name.Pinyin, Ingredient = Compound.Name, Target = Gene)  %>%
  distinct() %>%
  arrange(Herb,Target)

save(herb_ingredient_target,file = "../data/processed/network/02_herb_ingredient_target.RData")

```

## Step 2: Build herb-compound-target network
```{r}
herb_ingredient <- herb_ingredient_target %>%
  dplyr::select(Herb, Ingredient) %>%
  distinct()
  
ingredient_target <- herb_ingredient_target %>%
  dplyr::select(Ingredient, Target) %>%
  distinct()
  
# edges
edges <- bind_rows(
  herb_ingredient %>% 
    dplyr::rename(source = Herb, target = Ingredient) %>% 
    mutate(interaction = "herb_ingredient"),
  ingredient_target %>%
    dplyr::rename(source = Ingredient, target = Target) %>%
    mutate(interaction = "ingredient_target")
)

# nodes
nodes <- data.frame(id = unique(c(edges$source, edges$target))) %>%
  mutate(type = case_when(
    id %in% herb_ingredient$Herb        ~ "Herb",
    id %in% ingredient_target$Ingredient  ~ "Ingredient",
    id %in% ingredient_target$Target    ~ "Target",
    TRUE                              ~ "Other"
  ))

save(nodes,edges,file = "../data/processed/network/03_HIT_network.RData")
```

## Step 3: HIT network visualization
```{r}
rm(list = ls())

load("../data/processed/network/03_HIT_network.RData")

cytoscapePing()  

deleteAllNetworks() 

createNetworkFromDataFrames(
  nodes = nodes,
  edges = edges,
  title = "HFrEF_HIT_Network",
  collection = "HIT_Network"
)

# layout
layoutNetwork("force-directed")

# style
style_name <- "HIT_Style_By_Nodes_Type"
try(deleteVisualStyle(style_name), silent = TRUE)

# default style
defaults <- list(
  NODE_SHAPE = "RECTANGLE",
  NODE_SIZE = 30,
  NODE_FILL_COLOR = "#CCCCCC",
  NODE_LABEL_FONT_FACE = "Arial,plain,12",
  EDGE_WIDTH = 0.5,
  EDGE_STROKE_COLOR = "#E0E0E0",
  EDGE_TRANSPARENCY = 200         
)

# mapping style
mappings <- list(
  # node shape
  mapVisualProperty(
    visual.prop = 'node shape', 
    table.column = 'type', 
    mapping.type = 'd',
    table.column.values = c("Herb", "Ingredient", "Target"),
    visual.prop.values = c("ROUND RECTANGLE","ELLIPSE", "DIAMOND")
  ),
  
  # node color
  mapVisualProperty(
    visual.prop = 'node fill color', 
    table.column = 'type', 
    mapping.type = 'd',
    table.column.values = c("Herb", "Ingredient", "Target"),
    visual.prop.values = c("#fe9acc", "#fe0000", "#0401b9")
  ),
  
  # node size
  mapVisualProperty(
    visual.prop = 'node size',
    table.column = 'type',
    mapping.type = 'd',
    table.column.values = c("Herb", "Ingredient", "Target"),
    visual.prop.values = c(30, 20, 30)
  ),
  
  # node label
  mapVisualProperty(
    visual.prop = 'node label', 
    table.column = 'name', 
    mapping.type = 'p'
  ),
  
  mapVisualProperty(
    visual.prop = 'node label color', 
    table.column = 'type', 
    mapping.type = 'd',
    table.column.values = c("Herb", "Ingredient", "Target"),
    visual.prop.values = c("#000000", "#000000", "#FFFFFF")
  )
)

# apply style
createVisualStyle(
  style.name = style_name, 
  defaults = defaults, 
  mappings = mappings
)

setVisualStyle(style_name)

# Figure_3B.pdf / Figure_3B.png
```

---

# Functional Enrichment Analysis

## Load packages
```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)
library(ggplot2)
library(STRINGdb)
library(dplyr)
library(ggvenn)
```

## Step 1: SYMBOL to ENTREZID

```{r}
rm(list = ls())
load("../data/processed/network/01_herb_disease_genes.RData")

# candidate_genes
gene_id_map <- bitr(
  geneID = candidate_genes_df$Gene,
  fromType = "SYMBOL",
  toType = "ENTREZID",
  OrgDb = org.Hs.eg.db
)
gene_entrez <- gene_id_map$ENTREZID
```

## Step 2: GO
```{r}
go_enrich <- enrichGO(
  gene = gene_entrez,
  OrgDb = org.Hs.eg.db,
  keyType = "ENTREZID",
  ont = "ALL",
  pAdjustMethod = "fdr",
  qvalueCutoff = 0.05,
  readable = TRUE
)

```

## Step 3: KEGG
```{r}
kegg_enrich <- enrichKEGG(
  gene = gene_entrez,
  organism = "hsa",
  pAdjustMethod = "fdr",
  qvalueCutoff = 0.05
)

save(kegg_enrich,go_enrich, file = "../data/processed/functional/enrich.RData")
```

## Step 4: visualization
```{r}
load("../data/processed/functional/enrich.RData")

go_top <- go_enrich %>%
  as.data.frame() %>%
  group_by(ONTOLOGY) %>%
  slice_max(-log10(p.adjust), n = 5, with_ties = FALSE) %>% 
  ungroup() %>%
  mutate(
    Description_wrapped = str_wrap(Description, width = 30),
    Description_wrapped = as.factor(Description_wrapped),
    Description_wrapped = fct_reorder(Description_wrapped, p.adjust, .desc = FALSE),
    ONTOLOGY = factor(ONTOLOGY, levels = c("BP", "CC", "MF"))
  )

go_plot <- ggplot(go_top, aes(x = Count, y = Description_wrapped)) +
  geom_col(aes(fill = -log10(p.adjust)), width = 0.7) +
  facet_grid(ONTOLOGY ~ ., scales = "free_y", space = "free_y") +
  scale_fill_gradient(low = "#fde725", high = "#21918c", name = "-log10(p.adjust)") +
  labs(x = "Gene Count", y = "") +
  theme_bw(base_size = 12) +
  theme(
    strip.text.y = element_text(angle = 0, face = "bold", size = 11),
    axis.text.y = element_text(size = 8),          
    axis.title = element_text(size = 10),
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  ) +
  guides(fill = guide_colorbar(title.position = "top", barwidth = 10, barheight = 0.6))

ggsave("../figures/main/Figure_3C.pdf", go_plot, width = 7, height = 10, dpi = 300)
ggsave("../figures/main/Figure_3C.png", go_plot, width = 7, height = 10, dpi = 300)


# KEGG plot
kegg_plot <- dotplot(kegg_enrich, showCategory = 15) + theme_bw()
ggsave("../figures/main/Figure_3D.pdf", kegg_plot, width = 7, height = 10, dpi = 300)
ggsave("../figures/main/Figure_3D.png", kegg_plot, width = 7, height = 10, dpi = 300)
```

---

# Figure 3
```{r}
library(magick)

total_width_cm <- 14        
dpi <- 300                    
px_per_cm <- dpi / 2.54     
total_width_px <- round(total_width_cm * px_per_cm) 

ratios_row1 <- c(1, 2)
ratios_row2 <- c(1, 1)   # õ┐«µö╣õĖ║ 1:1

width_row1_left  <- round(total_width_px * ratios_row1[1] / sum(ratios_row1))
width_row1_right <- round(total_width_px * ratios_row1[2] / sum(ratios_row1))
width_row2_left  <- round(total_width_px * ratios_row2[1] / sum(ratios_row2))
width_row2_right <- round(total_width_px * ratios_row2[2] / sum(ratios_row2))

ratio <- c(width_row1_left,width_row1_right,width_row2_left,width_row2_right)

files <- c(
  "../figures/main/Figure_3A.pdf",
  "../figures/main/Figure_3B.pdf",
  "../figures/main/Figure_3C.pdf",
  "../figures/main/Figure_3D.pdf"
)

labels <- c("A", "B", "C", "D")

label_size <- 12   

pdfs <- list()
for (i in seq_along(files)) {
  img <- image_read_pdf(files[i], density = 300)
  img <- image_resize(img,ratio[i])
  img <- image_annotate(img, labels[i], 
                        gravity = "northwest",  
                        location = "+10+10",    
                        size = label_size,     
                        weight = 700,    
                        color = "black")
  pdfs[[i]] <- img
}

row1 <- image_append(c(pdfs[[1]], pdfs[[2]]), stack = FALSE)
row2 <- image_append(c(pdfs[[3]], pdfs[[4]]), stack = FALSE)

combined <- image_append(c(row1, row2), stack = TRUE)


image_write(combined, "../figures/main/Figure_3.pdf", format = "pdf")
image_write(combined, "../figures/main/Figure_3.png", format = "png", density = dpi)
```

---


# PPI Network

## Step 1: Obtain the PPI network from the STRING database
```{r}
rm(list = ls())

load("../data/processed/network/01_herb_disease_genes.RData")

gene_list <- candidate_genes_df$Gene
ppi_list <- clusterProfiler::getPPI(gene_list, taxID = "9606",required_score = 400)

save(ppi_list,file = "../data/processed/network/04_ppi_list.RData")
```

## Step 2: PPI network construction data preparation
```{r}
nodes <- data.frame(id = igraph::V(ppi_list)$name)
nodes$degree <- igraph::degree(ppi_list)

edges <- igraph::as_data_frame(ppi_list, what = "edges")
colnames(edges)[colnames(edges) == "from"] <- "source"
colnames(edges)[colnames(edges) == "to"]   <- "target"
# add an "interaction" column to edges
if(!"interaction" %in% colnames(edges)){
    edges$interaction <- "interacts with"
}

save(nodes,edges,file = "../data/processed/network/05_ppi_network.RData")
```

## Step 3: PPI network visualization
```{r}
# Ping Cytoscape 
cytoscapePing() 

try(deleteNetwork("HFrEF_PPI_Network"), silent = TRUE)
deleteAllNetworks() 

createNetworkFromDataFrames(
  nodes = nodes,  
  edges = edges,  
  title = "HFrEF_PPI_Network", 
  collection = "PPI_Network" 
)

# Create based on degree values, node attributes must be synchronized into Cytoscape first

loadTableData(nodes, data.key.column = "id", table = "node")

# New style
try(deleteVisualStyle("PPI_Style_By_Nodes_Degree"), silent = TRUE)
style_name <- "PPI_Style_By_Nodes_Degree"

# default style
default_lists <- list(
  NODE_SHAPE = "ellipse",           
  NODE_SIZE = 40,                   
  NODE_FILL_COLOR = "#88CCEE",      
  # NODE_LABEL_FONT_SIZE = 12,
  NODE_LABEL_FONT_FACE = "Arial,plain,12",  
  EDGE_WIDTH = 0.5,
  EDGE_STROKE_COLOR = "#E0E0E0",
  EDGE_TRANSPARENCY = 200          
)

mapping_lists <- list(
  # Node size: small degree ŌåÆ 20, large degree ŌåÆ 100
  mapVisualProperty(
    visual.prop = 'node size', 
    table.column = 'degree', 
    mapping.type = 'c',  
    table.column.values = c(min(nodes$degree), max(nodes$degree)), 
    visual.prop.values = c(20, 100)
  ),
  # Node fill color mapping: small degree ŌåÆ light yellow, large degree ŌåÆ red
  mapVisualProperty(
    visual.prop = 'node fill color', 
    table.column = 'degree', 
    mapping.type = 'c',
    table.column.values = c(min(nodes$degree), max(nodes$degree)),
    visual.prop.values = c("#FFFFCC", "#CC0000")
  )
)

createVisualStyle(
  style.name = style_name,
  defaults = default_lists,
  mappings = mapping_lists
)

setVisualStyle(style_name)

# Node Label
setNodeLabelMapping(table.column = 'id', style.name = style_name)

layoutNetwork("force-directed")
```

---

# Identification of Hub Genes
## Step 1: Cytohubba
```{r}
cytohubba_res <- read_csv("../data/processed/network/06_cytohubba_res.csv") %>%
  column_to_rownames(var = colnames(.)[1])

cytohubba_res <- read_csv("../tables/cytohubba.csv") %>%
  column_to_rownames(var = colnames(.)[1])

# cytohubba
n <- min(20, nrow(cytohubba_res))

top_mcc <- rownames(cytohubba_res[order(cytohubba_res$MCC, decreasing = TRUE), ])[1:n]

top_degree <- rownames(cytohubba_res[order(cytohubba_res$Degree,decreasing = T),])[1:n]

top_mnc <- rownames(cytohubba_res[order(cytohubba_res$MNC,decreasing = T),])[1:n]

cytohubba_hub <- Reduce(intersect, list(top_mcc, top_degree, top_mnc))

cytohubba_hub

# Figure 4B_d
gene_list <- list(
  MCC = top_mcc,
  MNC = top_mnc,
  Degree = top_degree
)

library(ggsci)
library(ggvenn)

p <- ggvenn(
  gene_list,
  fill_color = pal_aaas()(3),
  stroke_color = "white",
  stroke_size = 0.5,
  set_name_size = 2.5,
  text_size = 2.5,
  text_color = "white",
  show_percentage = FALSE
) +
  theme(plot.title = element_text(hjust = 0.5, size=4))

ggsave("../figures/main/Figure_4B_d.pdf", p, width=1.38, height=1.38, units="in", dpi=300)
ggsave("../figures/main/Figure_4B_d.png", p, width=2.76, height=1.38, units="in", dpi=300)


# Figure 4B

library(magick)

total_width_cm <- 7        
dpi <- 300                    
px_per_cm <- dpi / 2.54  

files <- c(
  "../figures/main/Figure_4B_a.pdf",
  "../figures/main/Figure_4B_b.pdf",
  "../figures/main/Figure_4B_c.pdf",
  "../figures/main/Figure_4B_d.pdf"
)

labels <- c("a", "b", "c", "d")

label_size <- 8   

pdfs <- list()
for (i in seq_along(files)) {
  img <- image_read_pdf(files[i], density = 300)
  img <- image_annotate(img, labels[i], 
                        gravity = "northwest",  
                        location = "+5+5",    
                        size = label_size,     
                        weight = 700,    
                        color = "black")
  pdfs[[i]] <- img
}

row1 <- image_append(c(pdfs[[1]], pdfs[[2]]), stack = FALSE)
row2 <- image_append(c(pdfs[[3]], pdfs[[4]]), stack = FALSE)

combined <- image_append(c(row1, row2), stack = TRUE)


image_write(combined, "../figures/main/Figure_4B.pdf", format = "pdf")
image_write(combined, "../figures/main/Figure_4B.png", format = "png", density = dpi)
```

## Step 2: MCODE
```{r}
mcode_hub <- data.table::fread("../data/processed/network/07_mcode_res.txt") %>%
  dplyr::slice(1) %>%                   
  dplyr::select(last_col()) %>%          
  dplyr::pull(1) %>%                     
  strsplit(",\\s*") %>%         
  unlist() 

# intersect
hub_genes <- Reduce(intersect, list(top_mcc, top_degree, top_mnc,mcode_hub))

hub_genes_tbl <- cytohubba_res[rownames(cytohubba_res) %in% hub_genes, c("MCC","MNC","Degree")] %>%
  mutate(MCC = as.numeric(MCC)) %>%
  arrange(desc(MCC)) %>%
    rownames_to_column("Gene")

hub_genes_tbl

hub_genes <- hub_genes_tbl$Gene

hub_genes

save(hub_genes_tbl,hub_genes,top_mcc,top_degree,top_mnc, 
     file = "../data/processed/network/08_hub_genes.RData")

write.xlsx(
  hub_genes_tbl,
  "../tables/Table_S5.xlsx"
)
```

## Step 3: Visualization
```{r}
gene_list <- list(
  MCC = top_mcc,
  MNC = top_mnc,
  Degree = top_degree,
  MCODE = mcode_hub
)

library(ggsci)
library(ggvenn)

p <- ggvenn(
  gene_list,
  fill_color = pal_aaas()(4),
  stroke_color = "white",
  stroke_size = 1,
  set_name_size = 5,
  text_size = 5,
  text_color = "white",
  show_percentage = FALSE
) +
  theme(plot.title = element_text(hjust = 0.5, size=10))

# save
ggsave("../figures/main/Figure_4D.pdf", p, width=2.76, height=2.76, units="in", dpi=300)
ggsave("../figures/main/Figure_4D.png", p, width=2.76, height=2.76, units="in", dpi=300)
```

# Figure 4
```{r}
library(magick)

total_width_cm <- 14        
dpi <- 300                    
px_per_cm <- dpi / 2.54     
total_width_px <- round(total_width_cm * px_per_cm) 

ratios_row1 <- c(1, 1)
ratios_row2 <- c(1, 1)  

width_row1_left  <- round(total_width_px * ratios_row1[1] / sum(ratios_row1))
width_row1_right <- round(total_width_px * ratios_row1[2] / sum(ratios_row1))
width_row2_left  <- round(total_width_px * ratios_row2[1] / sum(ratios_row2))
width_row2_right <- round(total_width_px * ratios_row2[2] / sum(ratios_row2))

ratio <- c(width_row1_left,width_row1_right,width_row2_left,width_row2_right)

files <- c(
  "../figures/main/Figure_4A.pdf",
  "../figures/main/Figure_4B.pdf",
  "../figures/main/Figure_4C.pdf",
  "../figures/main/Figure_4D.pdf"
)

labels <- c("A", "B", "C", "D")

label_size <- 12   

pdfs <- list()
for (i in seq_along(files)) {
  img <- image_read_pdf(files[i], density = 300)
  img <- image_resize(img,ratio[i])
  img <- image_annotate(img, labels[i], 
                        gravity = "northwest",  
                        location = "+10+10",    
                        size = label_size,     
                        weight = 700,    
                        color = "black")
  pdfs[[i]] <- img
}

row1 <- image_append(c(pdfs[[1]], pdfs[[2]]), stack = FALSE)
row2 <- image_append(c(pdfs[[3]], pdfs[[4]]), stack = FALSE)

combined <- image_append(c(row1, row2), stack = TRUE)


image_write(combined, "../figures/main/Figure_4.pdf", format = "pdf")
image_write(combined, "../figures/main/Figure_4.png", format = "png", density = dpi)
```


# Immune Infiltration Profiling

## Load Package
```{r}
library(GSVA)
library(GSEABase)
library(msigdbr)
library(pheatmap) 
library(tidyverse) 
library(rstatix) 
library(corrplot) 
library(ggpubr) 
library(ggsci)
library(IOBR)
```

https://www.medrxiv.org/content/10.1101/2024.04.26.24306471v1

We first determined the levels of immune cells in failing myocardium using five immune infiltration algorithms: Cibersort, xCell, ssGSEA, MCPcounter, and QuanTIseq.

## Step 1: FPKM & group

```{r}
rm(list = ls())
load("../data/processed/geo/GSE141910/GSE141910_preprocessed_final.RData")

group <- pheno_final %>% 
  rownames_to_column("Sample") %>%
  dplyr::select(Sample,group) %>%
  dplyr::rename(Group = group)


fpkm <- 2^expr_fpkm_corr - 1
fpkm[fpkm < 0] <- 0
```

## Step 2: deconvo_tme
```{r}
# MCPcounter
res_mcp   <- deconvo_tme(eset = fpkm, method = "mcpcounter")

# xCell
res_xcell <- deconvo_tme(eset = fpkm, method = "xcell")

# quanTIseq
res_quant <- deconvo_tme(eset = fpkm, method = "quantiseq")

# EPIC
res_epic  <- deconvo_tme(eset = fpkm, method = "epic")

# ssGSEA
# signature_score_calculation_methods
res_ssg <- calculate_sig_score(
  pdata = NULL,
  eset = fpkm,
  signature = load_data("go_bp"),
  method = "ssgsea",
  mini_gene_count = 2
)

save(res_mcp, res_xcell, res_quant, res_epic, res_ssg, file = "../data/processed/ime/res.RData")
```

## Step 3: Visualization function
```{r}
plot_tme <- function(df, title=NULL){
  if(!"ID" %in% colnames(df)) df <- rownames_to_column(df, "ID")
  df_long <- df %>%
    column_to_rownames("ID") %>%
    t() %>%
    as.data.frame() %>%
    rownames_to_column("cell") %>%
    reshape2::melt(id.vars="cell", variable.name="Sample", value.name="Score") %>%
    left_join(group, by = "Sample")
  
  df_long <- df_long %>%
    group_by(Sample) %>%
    mutate(Percent = Score / sum(Score) * 100) %>%
    ungroup()
  
  ggplot(df_long, aes(x = Sample, y = Percent, fill = cell)) +
    geom_col(position = "stack", width = 0.9) +
    facet_wrap(~Group, scales = "free_x") +
    theme_bw() + 
    RotatedAxis() +
    ggtitle(title) +
    ylab("Proportion (%)") +
    theme(legend.position = "bottom",axis.text.x = element_blank(),axis.ticks.x = element_blank())
}
```

## Step 4: Visualization
```{r}
p5A <- plot_tme(res_mcp)
pS3A <- plot_tme(res_xcell)
pS3B <- plot_tme(res_quant)
pS3C <- plot_tme(res_epic)
pS3D <- plot_tme(res_ssg)

ggsave("../figures/Figure_5A.pdf", pA, width = 10, height = 5, units = "cm", dpi = 300)
ggsave("../figures/Figure_5B.pdf", pB, width = 10, height = 5, units = "cm", dpi = 300)
ggsave("../figures/Figure_5C.pdf", pC, width = 10, height = 5, units = "cm", dpi = 300)
ggsave("../figures/Figure_5D.pdf", pD, width = 10, height = 5, units = "cm", dpi = 300)
ggsave("../figures/Figure_5E.pdf", pE, width = 10, height = 5, units = "cm", dpi = 300)

# Immune cell correlation heatmap
mcp_mat <- res_mcp %>% column_to_rownames("ID")
pdf("../figures/Figure_5F.pdf", width = 8, height = 7)
pheatmap(cor(mcp_mat, method = "spearman"),
         color = colorRampPalette(c("blue","white","red"))(100),
         treeheight_col = 15, treeheight_row = 15)
dev.off()

# Differential immune cell box plot
df_g <- mcp_mat %>% rownames_to_column("Sample") %>% left_join(group)
pvals <- apply(df_g[,-1:-2], 2, function(x) wilcox.test(x~df_g$Group)$p.value)
sig_cell <- names(pvals[pvals < 0.05])

plot_g <- df_g[,c("Sample","Group",sig_cell)] %>% reshape2::melt(id.vars = c("Sample","Group"))
pdf("../figures/Figure_5G.pdf", width = 11, height = 5.5)
ggboxplot(plot_g, x = "variable", y = "value", color = "Group") +
  stat_compare_means(method = "wilcox.test", label = "p.signif") +
  theme_bw() + rotate_x_text(45)
dev.off()

# Hub genes-immune correlation heatmap
load("../data/processed/network/08_hub_genes.RData")

hub_expr <- t(expr_fpkm_corr[hub_genes,])
imm_sig_expr <- mcp_mat[,sig_cell]

cor_mat <- cor(hub_expr, imm_sig_expr, method = "spearman")
p_mat <- corrplot::cor.mtest(hub_expr, imm_sig_expr)$p

pdf("../figures/Figure_5H.pdf", width = 9, height = 5)
corrplot(cor_mat, p.mat = p_mat, sig.level = 0.05, insig = "blank",
         col = colorRampPalette(c("royalblue","white","firebrick"))(200))
dev.off()
```


## Step 3: ÕģŹń¢½ń╗åĶā×õĖ░Õ║”ÕĀåÕÅĀÕøŠ’╝łµĀĘµ£¼µ»öõŠŗÕłåÕĖā’╝ē

```{r}
cibersort_data <- cibersort_result %>% 
  dplyr::select(where(~ any(. != 0))) %>%
  tibble::rownames_to_column('sample') %>%  
  tidyr::pivot_longer(
    cols = -c(group, sample), 
    names_to = 'cell', 
    values_to = 'score'
)

p1 <- ggplot(cibersort_data, aes(x = sample, y = score, fill = cell)) +
  geom_col(position = 'fill', alpha = 0.7) +  
  labs(x = '', 
       y = 'Estimated proportion', 
       fill = 'Cell type'
  ) + 
  guides(fill = guide_legend(ncol = 1)) + 
  theme(
    axis.text.x = element_blank(),  
    axis.text.y = element_text(size = 10),  
    axis.title = element_text(size = 10), 
    axis.ticks.x = element_blank(),  
    legend.title = element_text(size = 10),  
    legend.text = element_text(size = 10),  
    legend.key.size = unit(2, 'mm'),  
    strip.text = element_text(size = 10),  
    strip.background = element_rect(fill = 'gray', colour = 'black'), 
    panel.grid = element_blank(), 
    panel.background = element_rect(fill = NA, color = NA)  
  ) +
  facet_grid(. ~ group, scales = "free")  

print(p1)  
```

## Step 4: ÕģŹń¢½ń╗åĶā×ÕĘ«Õ╝éÕłåµ×ÉÕÅŖń«▒ń║┐ÕøŠ
```{r}
state_result <- cibersort_data %>% 
  group_by(cell) %>% 
  wilcox_test(score ~ group, p.adjust.method = 'BH')
state_result$p <- signif(state_result$p, 3)
state_result$p.label <- ifelse(state_result$p < 0.05, 
                               ifelse(state_result$p < 0.01, 
                                      ifelse(state_result$p < 0.001, 
                                             ifelse(state_result$p < 0.0001, "****", "***"), "**"), "*"), "ns")

change_all <- cibersort_data %>%
  group_by(cell) %>%
  summarise(
    case_median = median(score[group == "DCM_HFrEF"]),
    control_median = median(score[group == "Normal"]),
    signif = ifelse(case_median > control_median, "sig_high", "sig_low")
  )

change_all$signif <- ifelse(state_result$p.label[match(change_all$cell, state_result$cell)] == "ns", 
                            "not significant", change_all$signif)
cibersort_data$change <- change_all$signif[match(cibersort_data$cell, change_all$cell)]

cibersort_data <- cibersort_data %>% arrange(cell)
cibersort_data$cell <- factor(cibersort_data$cell, levels = unique(cibersort_data$cell))

state_result <- state_result %>%
  left_join(
    cibersort_data %>% distinct(cell, group),
    by = c("cell" = "cell")
  )

p2 <- ggplot(cibersort_data, aes(x = cell, y = score, fill = group)) +
  geom_boxplot() +  
  scale_fill_manual(values = c(Normal = "#4E4E56", DCM_HFrEF = "#DA635D")) +  
  stat_pvalue_manual(state_result,
                     x = "cell",
                     y.position = max(cibersort_data$score) * 1.1,
                     size = 3,
                     color = "blue",
                     label = "p.label") +
  labs(x = "", y = "Estimated Proportion", fill = "Group") + 
  theme_minimal() + 
  theme(
    axis.title.y = element_text(size = 10), 
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10, color = "black"),  
    axis.text = element_text(size = 10),  
    legend.text = element_text(size = 10), 
    legend.position = 'top',  
    legend.title = element_text(size = 10),  
    panel.grid.major.x = element_blank(),  
    panel.grid.minor = element_blank() 
  )

print(p2)  
```

## Step 5: ÕģŹń¢½ń╗åĶā×ķŚ┤ńøĖÕģ│µĆ¦Õłåµ×É
```{r}
# ńøĖÕģ│µĆ¦Õłåµ×É’╝Üõ╗ģõ┐ØńĢÖµĢ░ÕĆ╝ÕłŚ’╝łµÄÆķÖżÕłåń╗äńŁēķØ×µĢ░ÕĆ╝õ┐Īµü»’╝ē
numeric_cols <- sapply(cibersort_result, is.numeric)
# Ķ«Īń«ŚSpearmanńøĖÕģ│ń│╗µĢ░’╝łķĆéÕÉłķØ×µŁŻµĆüÕłåÕĖāµĢ░µŹ«’╝ē
cor_sorce <- cor(cibersort_result[, numeric_cols], method = 'spearman')
# Ķ«Īń«ŚńøĖÕģ│ń│╗µĢ░ńÜäPÕĆ╝’╝łõĮ┐ńö©WGCNAÕīģńÜäÕćĮµĢ░’╝ē
cor_sorce_p <- WGCNA::corPvalueStudent(cor_sorce, nrow(cibersort_result))
# Ķ«ŠńĮ«ķó£Ķē▓µĖÉÕÅś’╝Üõ╗ÄµĘ▒ĶōØ’╝łĶ┤¤ńøĖÕģ│’╝ēÕł░µĄģķ╗äÕåŹÕł░µĘ▒ń║ó’╝łµŁŻńøĖÕģ│’╝ē
col2 <- colorRampPalette(c("#053061", "#2166AC", "#4393C3", "#92C5DE", "#D1E5F0",
                           "#FDDBC7", "#F4A582", "#D6604D", "#B2182B", "#67001F"))(50)
# Õ«Üõ╣ēń╗śÕłČńøĖÕģ│µĆ¦ńāŁÕøŠńÜäÕćĮµĢ░’╝Ü
# õĖŖõĖēĶ¦ÆµśŠńż║ńøĖÕģ│ń│╗µĢ░µĢ░ÕĆ╝’╝īõĖŗõĖēĶ¦ÆµśŠńż║Õ£åÕ£ł’╝łÕż¦Õ░ÅĶĪ©ńż║ńøĖÕģ│µĆ¦Õ╝║Õ╝▒’╝ēÕÆīµśŠĶæŚµĆ¦µĀćńŁŠ
corr_plot <- function() {
  corrplot(cor_sorce,
           na.label = " ",  # ń╝║Õż▒ÕĆ╝ńö©ń®║µĀ╝ĶĪ©ńż║
           diag = T,  # µśŠńż║Õ»╣Ķ¦Æń║┐’╝łĶć¬Ķ║½ńøĖÕģ│µĆ¦’╝ē
           method = "number",  # õĖŖõĖēĶ¦Æńö©µĢ░ÕŁŚĶĪ©ńż║ńøĖÕģ│ń│╗µĢ░
           type = "upper",  # õ╗ģµśŠńż║õĖŖõĖēĶ¦Æ
           tl.pos = "lt",  # µĀćńŁŠõĮŹńĮ«’╝łÕĘ”ÕÆīõĖŖ’╝ē
           tl.col = "black",  # µĀćńŁŠķó£Ķē▓
           tl.srt = 90,  # µĀćńŁŠµŚŗĶĮ¼90Õ║”
           tl.cex = 0.9,  # µĀćńŁŠÕŁŚõĮōÕż¦Õ░Å
           cl.length = 8,  # ķó£Ķē▓ÕøŠõŠŗńÜäÕłåµ«ĄµĢ░
           cl.cex = 0.8,  # ķó£Ķē▓ÕøŠõŠŗÕŁŚõĮōÕż¦Õ░Å
           col = col2,  # õĮ┐ńö©Ķć¬Õ«Üõ╣ēķó£Ķē▓
           number.cex = 0.7)  # ńøĖÕģ│ń│╗µĢ░µĢ░ÕĆ╝Õż¦Õ░Å

  # õĖŗõĖēĶ¦ÆµĘ╗ÕŖĀµśŠĶæŚµĆ¦µĀćńŁŠ
  corrplot(cor_sorce,
           p.mat = cor_sorce_p,  # PÕĆ╝ń¤®ķśĄ
           insig = "label_sig",  # Õ»╣õĖŹµśŠĶæŚńÜäńøĖÕģ│µĆ¦µĀćĶ«░
           add = TRUE,  # Õ£©ÕĘ▓µ£ēÕøŠõĖŖµĘ╗ÕŖĀ
           type = "lower",  # õ╗ģµśŠńż║õĖŗõĖēĶ¦Æ
           method = "circle",  # ńö©Õ£åÕ£łĶĪ©ńż║ńøĖÕģ│µĆ¦
           col = col2,  # ķó£Ķē▓õĖÄõĖŖõĖēĶ¦ÆõĖĆĶć┤
           na.label = " ",  # ń╝║Õż▒ÕĆ╝ÕżäńÉå
           diag = FALSE,  # õĖŹµśŠńż║Õ»╣Ķ¦Æń║┐
           tl.pos = "n",  # õĖŹµśŠńż║µĀćńŁŠ
           cl.pos = "n",  # õĖŹµśŠńż║ķó£Ķē▓ÕøŠõŠŗ
           sig.level = c(.0001, .001, .01, .05),  # µśŠĶæŚµĆ¦µ░┤Õ╣│Õłåń║¦
           pch.cex = 0.9,  # µśŠĶæŚµĆ¦µĀćńŁŠÕż¦Õ░Å
           pch.col = "black")  # µśŠĶæŚµĆ¦µĀćńŁŠķó£Ķē▓
}
corr_plot()  # ń╗śÕłČÕ╣ČµśŠńż║ńøĖÕģ│µĆ¦ńāŁÕøŠ
```

## Step 6: ńø«µĀćÕ¤║ÕøĀõĖÄÕģŹń¢½ń╗åĶā×ńÜäńøĖÕģ│µĆ¦Õłåµ×É
```{r}
# Õ«Üõ╣ēµé©µä¤Õģ┤ĶČŻńÜäÕ¤║ÕøĀ’╝łĶ»ĘµĀ╣µŹ«Õ«×ķÖģńĀöń®ČńÜäÕ¤║ÕøĀõ┐«µö╣’╝ē’╝ÜĶ┐Öķćīõ╗źACEÕÆīCASP1õĖ║õŠŗ
interest_genes <- c("ACE","CASP1")  
gene_expr <- expr_fpkm_corr[rownames(expr_fpkm_corr) %in% interest_genes, , drop = FALSE]

gene_expr_t <- as.data.frame(t(gene_expr))
colnames(gene_expr_t) <- rownames(gene_expr)  
common_samples <- intersect(rownames(cibersort_result), rownames(gene_expr_t))
cibersort_sub <- cibersort_result[common_samples, cell_cols, drop = FALSE] %>%
  dplyr::select(where(~ any(. != 0)))

gene_expr_sub <- gene_expr_t[common_samples, , drop = FALSE] 
combined_data <- cbind(cibersort_sub, gene_expr_sub)

correlation_matrix <- cor(combined_data, method = "spearman")

n_immune <- ncol(cibersort_sub)  
n_genes <- length(interest_genes)  
gene_cell_cor <- correlation_matrix[(n_immune + 1):(n_immune + n_genes), 1:n_immune, drop = FALSE]

p_matrix <- matrix(0, nrow = n_genes, ncol = n_immune)
rownames(p_matrix) <- interest_genes
colnames(p_matrix) <- colnames(cibersort_sub)
for (gene in interest_genes) {
  for (cell in colnames(cibersort_sub)) {
    test <- cor.test(combined_data[[gene]], combined_data[[cell]], method = "spearman")
    p_matrix[gene, cell] <- test$p.value  
  }
}

sig_matrix <- ifelse(p_matrix < 0.001, "***", 
                     ifelse(p_matrix < 0.01, "**", 
                            ifelse(p_matrix < 0.05, "*", "")))
pheatmap(gene_cell_cor,
         color = colorRampPalette(c("#2166AC", "white", "#B2182B"))(100), 
         display_numbers = sig_matrix,  
         number_color = "black",
         number_format = "%s",  
         main = "Correlation between Genes and Immune Cells",  # ÕøŠµĀćķóś
         treeheight_row = 20,  
         treeheight_col = 20, 
         fontsize = 10,  
         fontsize_row = 12,  
         fontsize_col = 10,  
         cellwidth = 25,  
         cellheight = 30)  
dev.off()  
```

# Figure 5

Immune cell infiltration analysis between HF and control. (A) Result of Cibersort algorithm. (B) Result of MCPcounter algorithm. (C) Result of QuanTIseq algorithm. (D) Result of ssGSEA algorithm. (E) Result of xCell algorithm.

```{r}

```

---

<table class="nav-table" width="100%">
  <tr>
    <td align="left">
      [Home](../index.qmd) 
    </td>
    <td align="right">
      [Previous](Part_2_Screening_of_HFrEF_related_genes.qmd) | [Next](Part_4_Mechanism_validation_via_bioinformatics_and_docking.qmd)
    </td>
  </tr>
</table>

# References {-}