---
title: "Part 3: Mechanism exploration of ZWHQD against HFrEF"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

---

# Identification of Candidate Targets

## Load Packages
```{r}
library(ggvenn)
library(tidyverse)
library(openxlsx)
library(ggsci)
```

## Candidate Genes
```{r}
rm(list = ls())

load("../data/processed/geo/GSE141910/degs_results.RData")

deg_sig <- deg_all %>% filter(padj < 0.05, abs(log2FoldChange) > lfc)

deg_genes <- rownames(deg_sig) 

load("../data/processed/tcm/14_compound_target_final.RData")

herb_targets <- unique(compound_target_final$Symbol)

load("../data/processed/geo/GSE141910/wgcna_candidates_genes.RData")

# disease_genes: intersecting DEGs and WGCNA
disease_genes <- intersect(deg_genes,wgcna_candidates_genes$Gene)

# candidate_genes: intersecting herb_targets and disease_genes
candidate_genes <-intersect(herb_targets, disease_genes)

herb_targets_df <- data.frame(Gene = herb_targets) %>%
  arrange(Gene)

deg_genes_df <- data.frame(Gene = deg_genes) %>%
  arrange(Gene)

wgcna_genes_df <- data.frame(Gene = wgcna_candidates_genes$Gene) %>%
  arrange(Gene)

disease_genes_df <- data.frame(Gene = disease_genes) %>%
  arrange(Gene)

candidate_genes_df <- data.frame(Gene = candidate_genes) %>%
  arrange(Gene)

gene_list <- list(
  DEGs = deg_genes,
  `Herb targets` = herb_targets,
  `WGCNA Genes` = wgcna_candidates_genes$Gene
)

p <- ggvenn(
  gene_list,
  fill_color = pal_aaas()(4),
  stroke_color = "white",
  stroke_size = 1,
  set_name_size = 5,
  text_size = 5,
  text_color = "white",
  show_percentage = FALSE
) +
  theme(plot.title = element_text(hjust = 0.5, size=10))

# save
ggsave("../figures/main/Figure_3A.pdf", p, width=6, height=6, dpi=300)
ggsave("../figures/main/Figure_3A.png", p, width=6, height=6, dpi=300)


save(herb_targets_df, deg_genes_df, wgcna_genes_df,disease_genes_df, candidate_genes_df, file = "../data/processed/network/01_herb_disease_genes.RData")

write.xlsx(
  list(
    HerbTargets = herb_targets_df,
    DEGs = deg_genes_df,
    WGCNAGenes = wgcna_genes_df,
    DiseaseGenes = disease_genes_df,
    CandidateGenes = candidate_genes_df
  ),
  "../tables/candidate_genes.xlsx"
)

deg_candidate <- deg_all[rownames(deg_all) %in% candidate_genes, ]
 
deg_candidate$regulation <- ifelse(
  deg_candidate$log2FoldChange >= lfc, "Up",
  ifelse(deg_candidate$log2FoldChange <= -lfc, "Down", "NS")
)

deg_candidate$Gene <- rownames(deg_candidate)

up_genes <- deg_candidate$Gene[deg_candidate$regulation == "Up"]
down_genes <- deg_candidate$Gene[deg_candidate$regulation == "Down"]

cat("Up genes: ", length(up_genes), "\n")
print(up_genes)

cat("\nDown genes: ", length(down_genes), "\n")
print(down_genes)
```

---

# Herb-Ingredient-Target (HIT) Network
## Load Packages

```{r}
library(tidyverse)
library(openxlsx)
library(data.table)

library(clusterProfiler)
library(igraph)
library(centiserve)
library(RCy3)
```

## Step 1: Loading data
```{r}
rm(list = ls())

load("../data/processed/network/01_herb_disease_genes.RData")
load("../data/processed/tcm/14_compound_target_final.RData")

herb_ingredient_target <- candidate_genes_df %>%
  left_join(compound_target_final, by = c("Gene" = "Symbol"),relationship = "many-to-many") %>%
  dplyr::select(Herb.Name.Pinyin, Compound.Name, Gene) %>%
  dplyr:: rename(Herb = Herb.Name.Pinyin, Ingredient = Compound.Name, Target = Gene)  %>%
  distinct() %>%
  arrange(Herb,Target)

save(herb_ingredient_target,file = "../data/processed/network/02_herb_ingredient_target.RData")

```

## Step 2: Build herb-compound-target network
```{r}
herb_ingredient <- herb_ingredient_target %>%
  dplyr::select(Herb, Ingredient) %>%
  distinct()
  
ingredient_target <- herb_ingredient_target %>%
  dplyr::select(Ingredient, Target) %>%
  distinct()
  
# edges
edges <- bind_rows(
  herb_ingredient %>% 
    dplyr::rename(source = Herb, target = Ingredient) %>% 
    mutate(interaction = "herb_ingredient"),
  ingredient_target %>%
    dplyr::rename(source = Ingredient, target = Target) %>%
    mutate(interaction = "ingredient_target")
)

# nodes
nodes <- data.frame(id = unique(c(edges$source, edges$target))) %>%
  mutate(type = case_when(
    id %in% herb_ingredient$Herb        ~ "Herb",
    id %in% ingredient_target$Ingredient  ~ "Ingredient",
    id %in% ingredient_target$Target    ~ "Target",
    TRUE                              ~ "Other"
  ))

save(nodes,edges,file = "../data/processed/network/03_HIT_network.RData")
```

## Step 3: HIT network visualization
```{r}
rm(list = ls())

load("../data/processed/network/03_HIT_network.RData")

cytoscapePing()  

deleteAllNetworks() 

createNetworkFromDataFrames(
  nodes = nodes,
  edges = edges,
  title = "HFrEF_HIT_Network",
  collection = "HIT_Network"
)

# layout
layoutNetwork("force-directed")

# style
style_name <- "HIT_Style_By_Nodes_Type"
try(deleteVisualStyle(style_name), silent = TRUE)

# default style
defaults <- list(
  NODE_SHAPE = "RECTANGLE",
  NODE_SIZE = 30,
  NODE_FILL_COLOR = "#CCCCCC",
  NODE_LABEL_FONT_FACE = "Arial,plain,12",
  EDGE_WIDTH = 0.5,
  EDGE_STROKE_COLOR = "#E0E0E0",
  EDGE_TRANSPARENCY = 200         
)

# mapping style
mappings <- list(
  # node shape
  mapVisualProperty(
    visual.prop = 'node shape', 
    table.column = 'type', 
    mapping.type = 'd',
    table.column.values = c("Herb", "Ingredient", "Target"),
    visual.prop.values = c("ROUND RECTANGLE","ELLIPSE", "DIAMOND")
  ),
  
  # node color
  mapVisualProperty(
    visual.prop = 'node fill color', 
    table.column = 'type', 
    mapping.type = 'd',
    table.column.values = c("Herb", "Ingredient", "Target"),
    visual.prop.values = c("#fe9acc", "#fe0000", "#0401b9")
  ),
  
  # node size
  mapVisualProperty(
    visual.prop = 'node size',
    table.column = 'type',
    mapping.type = 'd',
    table.column.values = c("Herb", "Ingredient", "Target"),
    visual.prop.values = c(30, 20, 30)
  ),
  
  # node label
  mapVisualProperty(
    visual.prop = 'node label', 
    table.column = 'name', 
    mapping.type = 'p'
  ),
  
  mapVisualProperty(
    visual.prop = 'node label color', 
    table.column = 'type', 
    mapping.type = 'd',
    table.column.values = c("Herb", "Ingredient", "Target"),
    visual.prop.values = c("#000000", "#000000", "#FFFFFF")
  )
)

# apply style
createVisualStyle(
  style.name = style_name, 
  defaults = defaults, 
  mappings = mappings
)

setVisualStyle(style_name)

# Figure_3B.pdf / Figure_3B.png
```

---

# Functional Enrichment Analysis

## Load packages
```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)
library(ggplot2)
library(STRINGdb)
library(dplyr)
library(ggvenn)
```

## Step 1: SYMBOL to ENTREZID

```{r}
rm(list = ls())
load("../data/processed/network/01_herb_disease_genes.RData")

# candidate_genes
gene_id_map <- bitr(
  geneID = candidate_genes_df$Gene,
  fromType = "SYMBOL",
  toType = "ENTREZID",
  OrgDb = org.Hs.eg.db
)
gene_entrez <- gene_id_map$ENTREZID
```

## Step 2: GO
```{r}
go_enrich <- enrichGO(
  gene = gene_entrez,
  OrgDb = org.Hs.eg.db,
  keyType = "ENTREZID",
  ont = "ALL",
  pAdjustMethod = "fdr",
  qvalueCutoff = 0.05,
  readable = TRUE
)

```

## Step 3: KEGG
```{r}
kegg_enrich <- enrichKEGG(
  gene = gene_entrez,
  organism = "hsa",
  pAdjustMethod = "fdr",
  qvalueCutoff = 0.05
)

save(kegg_enrich,go_enrich, file = "../data/processed/functional/enrich.RData")
```

## Step 4: visualization
```{r}
load("../data/processed/functional/enrich.RData")

go_top <- go_enrich %>%
  as.data.frame() %>%
  group_by(ONTOLOGY) %>%
  slice_max(-log10(p.adjust), n = 5, with_ties = FALSE) %>% 
  ungroup() %>%
  mutate(
    Description_wrapped = str_wrap(Description, width = 30),
    Description_wrapped = as.factor(Description_wrapped),
    Description_wrapped = fct_reorder(Description_wrapped, p.adjust, .desc = FALSE),
    ONTOLOGY = factor(ONTOLOGY, levels = c("BP", "CC", "MF"))
  )

go_plot <- ggplot(go_top, aes(x = Count, y = Description_wrapped)) +
  geom_col(aes(fill = -log10(p.adjust)), width = 0.7) +
  facet_grid(ONTOLOGY ~ ., scales = "free_y", space = "free_y") +
  scale_fill_gradient(low = "#fde725", high = "#21918c", name = "-log10(p.adjust)") +
  labs(x = "Gene Count", y = "") +
  theme_bw(base_size = 12) +
  theme(
    strip.text.y = element_text(angle = 0, face = "bold", size = 11),
    axis.text.y = element_text(size = 8),          
    axis.title = element_text(size = 10),
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  ) +
  guides(fill = guide_colorbar(title.position = "top", barwidth = 10, barheight = 0.6))

ggsave("../figures/main/Figure_3C.pdf", go_plot, width = 7, height = 10, dpi = 300)
ggsave("../figures/main/Figure_3C.png", go_plot, width = 7, height = 10, dpi = 300)


# KEGG plot
kegg_plot <- dotplot(kegg_enrich, showCategory = 15) + theme_bw()
ggsave("../figures/main/Figure_3D.pdf", kegg_plot, width = 7, height = 10, dpi = 300)
ggsave("../figures/main/Figure_3D.png", kegg_plot, width = 7, height = 10, dpi = 300)
```

---

# Figure 3
```{r}
library(magick)

total_width_cm <- 14        
dpi <- 300                    
px_per_cm <- dpi / 2.54     
total_width_px <- round(total_width_cm * px_per_cm) 

ratios_row1 <- c(1, 2)
ratios_row2 <- c(1, 1)   # 修改为 1:1

width_row1_left  <- round(total_width_px * ratios_row1[1] / sum(ratios_row1))
width_row1_right <- round(total_width_px * ratios_row1[2] / sum(ratios_row1))
width_row2_left  <- round(total_width_px * ratios_row2[1] / sum(ratios_row2))
width_row2_right <- round(total_width_px * ratios_row2[2] / sum(ratios_row2))

ratio <- c(width_row1_left,width_row1_right,width_row2_left,width_row2_right)

files <- c(
  "../figures/main/Figure_3A.pdf",
  "../figures/main/Figure_3B.pdf",
  "../figures/main/Figure_3C.pdf",
  "../figures/main/Figure_3D.pdf"
)

labels <- c("A", "B", "C", "D")

label_size <- 12   

pdfs <- list()
for (i in seq_along(files)) {
  img <- image_read_pdf(files[i], density = 300)
  img <- image_resize(img,ratio[i])
  img <- image_annotate(img, labels[i], 
                        gravity = "northwest",  
                        location = "+10+10",    
                        size = label_size,     
                        weight = 700,    
                        color = "black")
  pdfs[[i]] <- img
}

row1 <- image_append(c(pdfs[[1]], pdfs[[2]]), stack = FALSE)
row2 <- image_append(c(pdfs[[3]], pdfs[[4]]), stack = FALSE)

combined <- image_append(c(row1, row2), stack = TRUE)


image_write(combined, "../figures/main/Figure_3.pdf", format = "pdf")
image_write(combined, "../figures/main/Figure_3.png", format = "png", density = dpi)
```

---


# PPI Network

## Step 1: Obtain the PPI network from the STRING database
```{r}
rm(list = ls())

load("../data/processed/network/01_herb_disease_genes.RData")

gene_list <- candidate_genes_df$Gene
ppi_list <- clusterProfiler::getPPI(gene_list, taxID = "9606",required_score = 400)

save(ppi_list,file = "../data/processed/network/04_ppi_list.RData")
```

## Step 2: PPI network construction data preparation
```{r}
nodes <- data.frame(id = igraph::V(ppi_list)$name)
nodes$degree <- igraph::degree(ppi_list)

edges <- igraph::as_data_frame(ppi_list, what = "edges")
colnames(edges)[colnames(edges) == "from"] <- "source"
colnames(edges)[colnames(edges) == "to"]   <- "target"
# add an "interaction" column to edges
if(!"interaction" %in% colnames(edges)){
    edges$interaction <- "interacts with"
}

save(nodes,edges,file = "../data/processed/network/05_ppi_network.RData")
```

## Step 3: PPI network visualization
```{r}
# Ping Cytoscape 
cytoscapePing() 

try(deleteNetwork("HFrEF_PPI_Network"), silent = TRUE)
deleteAllNetworks() 

createNetworkFromDataFrames(
  nodes = nodes,  
  edges = edges,  
  title = "HFrEF_PPI_Network", 
  collection = "PPI_Network" 
)

# Create based on degree values, node attributes must be synchronized into Cytoscape first

loadTableData(nodes, data.key.column = "id", table = "node")

# New style
try(deleteVisualStyle("PPI_Style_By_Nodes_Degree"), silent = TRUE)
style_name <- "PPI_Style_By_Nodes_Degree"

# default style
default_lists <- list(
  NODE_SHAPE = "ellipse",           
  NODE_SIZE = 40,                   
  NODE_FILL_COLOR = "#88CCEE",      
  # NODE_LABEL_FONT_SIZE = 12,
  NODE_LABEL_FONT_FACE = "Arial,plain,12",  
  EDGE_WIDTH = 0.5,
  EDGE_STROKE_COLOR = "#E0E0E0",
  EDGE_TRANSPARENCY = 200          
)

mapping_lists <- list(
  # Node size: small degree → 20, large degree → 100
  mapVisualProperty(
    visual.prop = 'node size', 
    table.column = 'degree', 
    mapping.type = 'c',  
    table.column.values = c(min(nodes$degree), max(nodes$degree)), 
    visual.prop.values = c(20, 100)
  ),
  # Node fill color mapping: small degree → light yellow, large degree → red
  mapVisualProperty(
    visual.prop = 'node fill color', 
    table.column = 'degree', 
    mapping.type = 'c',
    table.column.values = c(min(nodes$degree), max(nodes$degree)),
    visual.prop.values = c("#FFFFCC", "#CC0000")
  )
)

createVisualStyle(
  style.name = style_name,
  defaults = default_lists,
  mappings = mapping_lists
)

setVisualStyle(style_name)

# Node Label
setNodeLabelMapping(table.column = 'id', style.name = style_name)

layoutNetwork("force-directed")
```

---

# Identification of Hub Genes
## Step 1: Cytohubba
```{r}
cytohubba_res <- read_csv("../data/processed/network/06_cytohubba_res.csv") %>%
  column_to_rownames(var = colnames(.)[1])

cytohubba_res <- read_csv("../tables/cytohubba.csv") %>%
  column_to_rownames(var = colnames(.)[1])

# cytohubba
n <- min(20, nrow(cytohubba_res))

top_mcc <- rownames(cytohubba_res[order(cytohubba_res$MCC, decreasing = TRUE), ])[1:n]

top_degree <- rownames(cytohubba_res[order(cytohubba_res$Degree,decreasing = T),])[1:n]

top_mnc <- rownames(cytohubba_res[order(cytohubba_res$MNC,decreasing = T),])[1:n]

cytohubba_hub <- Reduce(intersect, list(top_mcc, top_degree, top_mnc))

cytohubba_hub

# Figure 4B_d
gene_list <- list(
  MCC = top_mcc,
  MNC = top_mnc,
  Degree = top_degree
)

library(ggsci)
library(ggvenn)

p <- ggvenn(
  gene_list,
  fill_color = pal_aaas()(3),
  stroke_color = "white",
  stroke_size = 0.5,
  set_name_size = 2.5,
  text_size = 2.5,
  text_color = "white",
  show_percentage = FALSE
) +
  theme(plot.title = element_text(hjust = 0.5, size=4))

ggsave("../figures/main/Figure_4B_d.pdf", p, width=1.38, height=1.38, units="in", dpi=300)
ggsave("../figures/main/Figure_4B_d.png", p, width=2.76, height=1.38, units="in", dpi=300)


# Figure 4B

library(magick)

total_width_cm <- 7        
dpi <- 300                    
px_per_cm <- dpi / 2.54  

files <- c(
  "../figures/main/Figure_4B_a.pdf",
  "../figures/main/Figure_4B_b.pdf",
  "../figures/main/Figure_4B_c.pdf",
  "../figures/main/Figure_4B_d.pdf"
)

labels <- c("a", "b", "c", "d")

label_size <- 8   

pdfs <- list()
for (i in seq_along(files)) {
  img <- image_read_pdf(files[i], density = 300)
  img <- image_annotate(img, labels[i], 
                        gravity = "northwest",  
                        location = "+5+5",    
                        size = label_size,     
                        weight = 700,    
                        color = "black")
  pdfs[[i]] <- img
}

row1 <- image_append(c(pdfs[[1]], pdfs[[2]]), stack = FALSE)
row2 <- image_append(c(pdfs[[3]], pdfs[[4]]), stack = FALSE)

combined <- image_append(c(row1, row2), stack = TRUE)


image_write(combined, "../figures/main/Figure_4B.pdf", format = "pdf")
image_write(combined, "../figures/main/Figure_4B.png", format = "png", density = dpi)
```

## Step 2: MCODE
```{r}
mcode_hub <- data.table::fread("../data/processed/network/07_mcode_res.txt") %>%
  dplyr::slice(1) %>%                   
  dplyr::select(last_col()) %>%          
  dplyr::pull(1) %>%                     
  strsplit(",\\s*") %>%         
  unlist() 

# intersect
hub_genes <- Reduce(intersect, list(top_mcc, top_degree, top_mnc,mcode_hub))

hub_genes_tbl <- cytohubba_res[rownames(cytohubba_res) %in% hub_genes, c("MCC","MNC","Degree")] %>%
  mutate(MCC = as.numeric(MCC)) %>%
  arrange(desc(MCC)) %>%
    rownames_to_column("Gene")

hub_genes_tbl

hub_genes <- hub_genes_tbl$Gene

hub_genes

save(hub_genes_tbl,hub_genes,top_mcc,top_degree,top_mnc, 
     file = "../data/processed/network/08_hub_genes.RData")

write.xlsx(
  hub_genes_tbl,
  "../tables/Table_S5.xlsx"
)
```

## Step 3: Visualization
```{r}
gene_list <- list(
  MCC = top_mcc,
  MNC = top_mnc,
  Degree = top_degree,
  MCODE = mcode_hub
)

library(ggsci)
library(ggvenn)

p <- ggvenn(
  gene_list,
  fill_color = pal_aaas()(4),
  stroke_color = "white",
  stroke_size = 1,
  set_name_size = 5,
  text_size = 5,
  text_color = "white",
  show_percentage = FALSE
) +
  theme(plot.title = element_text(hjust = 0.5, size=10))

# save
ggsave("../figures/main/Figure_4D.pdf", p, width=2.76, height=2.76, units="in", dpi=300)
ggsave("../figures/main/Figure_4D.png", p, width=2.76, height=2.76, units="in", dpi=300)
```

# Figure 4
```{r}
library(magick)

total_width_cm <- 14        
dpi <- 300                    
px_per_cm <- dpi / 2.54     
total_width_px <- round(total_width_cm * px_per_cm) 

ratios_row1 <- c(1, 1)
ratios_row2 <- c(1, 1)  

width_row1_left  <- round(total_width_px * ratios_row1[1] / sum(ratios_row1))
width_row1_right <- round(total_width_px * ratios_row1[2] / sum(ratios_row1))
width_row2_left  <- round(total_width_px * ratios_row2[1] / sum(ratios_row2))
width_row2_right <- round(total_width_px * ratios_row2[2] / sum(ratios_row2))

ratio <- c(width_row1_left,width_row1_right,width_row2_left,width_row2_right)

files <- c(
  "../figures/main/Figure_4A.pdf",
  "../figures/main/Figure_4B.pdf",
  "../figures/main/Figure_4C.pdf",
  "../figures/main/Figure_4D.pdf"
)

labels <- c("A", "B", "C", "D")

label_size <- 12   

pdfs <- list()
for (i in seq_along(files)) {
  img <- image_read_pdf(files[i], density = 300)
  img <- image_resize(img,ratio[i])
  img <- image_annotate(img, labels[i], 
                        gravity = "northwest",  
                        location = "+10+10",    
                        size = label_size,     
                        weight = 700,    
                        color = "black")
  pdfs[[i]] <- img
}

row1 <- image_append(c(pdfs[[1]], pdfs[[2]]), stack = FALSE)
row2 <- image_append(c(pdfs[[3]], pdfs[[4]]), stack = FALSE)

combined <- image_append(c(row1, row2), stack = TRUE)


image_write(combined, "../figures/main/Figure_4.pdf", format = "pdf")
image_write(combined, "../figures/main/Figure_4.png", format = "png", density = dpi)
```


# Immune Infiltration Profiling

## Load Package
```{r}
library(GSVA)
library(GSEABase)
library(msigdbr)
library(pheatmap) 
library(tidyverse) 
library(rstatix) 
library(corrplot) 
library(ggpubr) 
library(ggsci)
```

## Step 1: ssGSEA
```{r}
rm(list = ls())
load("../data/processed/geo/GSE141910/GSE141910_preprocessed_final.RData")

data(LM22)

results <- cibersort(
  sig_matrix = LM22, 
  mixture_file = expr_fpkm_corr, 
  perm = 1000, 
  QN = TRUE
)

save(results,file = "../data/processed/tme/cibersort_res.RData")
write.table(
  rownames_to_column(as.data.frame(results), 'Sample'), 
  '../data/processed/tme/cibersort_res.xlsx',  
  sep = '\t', 
  quote = F, 
  row.names = F)
```

## Step 2: 提取22种免疫细胞的比例数据
```{r}
cell_cols <- intersect(colnames(results), colnames(LM22)) 
cibersort_result <- results[, cell_cols] %>% as.data.frame()
```

## Step 3: Adding group infos
```{r}
group_info <- data.frame(
  samples = rownames(pheno_final), 
  group =pheno_final$group)

group_info <- group_info[order(match(group_info$sample, rownames(results))), ]

cibersort_result$group <- group_info$group[match(rownames(cibersort_result), group_info$sample)]
```


## Step 3: 免疫细胞丰度堆叠图（样本比例分布）

```{r}
cibersort_data <- cibersort_result %>% 
  dplyr::select(where(~ any(. != 0))) %>%
  tibble::rownames_to_column('sample') %>%  
  tidyr::pivot_longer(
    cols = -c(group, sample), 
    names_to = 'cell', 
    values_to = 'score'
)

p1 <- ggplot(cibersort_data, aes(x = sample, y = score, fill = cell)) +
  geom_col(position = 'fill', alpha = 0.7) +  
  labs(x = '', 
       y = 'Estimated proportion', 
       fill = 'Cell type'
  ) + 
  guides(fill = guide_legend(ncol = 1)) + 
  theme(
    axis.text.x = element_blank(),  
    axis.text.y = element_text(size = 10),  
    axis.title = element_text(size = 10), 
    axis.ticks.x = element_blank(),  
    legend.title = element_text(size = 10),  
    legend.text = element_text(size = 10),  
    legend.key.size = unit(2, 'mm'),  
    strip.text = element_text(size = 10),  
    strip.background = element_rect(fill = 'gray', colour = 'black'), 
    panel.grid = element_blank(), 
    panel.background = element_rect(fill = NA, color = NA)  
  ) +
  facet_grid(. ~ group, scales = "free")  

print(p1)  
```

## Step 4: 免疫细胞差异分析及箱线图
```{r}
state_result <- cibersort_data %>% 
  group_by(cell) %>% 
  wilcox_test(score ~ group, p.adjust.method = 'BH')
state_result$p <- signif(state_result$p, 3)
state_result$p.label <- ifelse(state_result$p < 0.05, 
                               ifelse(state_result$p < 0.01, 
                                      ifelse(state_result$p < 0.001, 
                                             ifelse(state_result$p < 0.0001, "****", "***"), "**"), "*"), "ns")

change_all <- cibersort_data %>%
  group_by(cell) %>%
  summarise(
    case_median = median(score[group == "DCM_HFrEF"]),
    control_median = median(score[group == "Normal"]),
    signif = ifelse(case_median > control_median, "sig_high", "sig_low")
  )

change_all$signif <- ifelse(state_result$p.label[match(change_all$cell, state_result$cell)] == "ns", 
                            "not significant", change_all$signif)
cibersort_data$change <- change_all$signif[match(cibersort_data$cell, change_all$cell)]

cibersort_data <- cibersort_data %>% arrange(cell)
cibersort_data$cell <- factor(cibersort_data$cell, levels = unique(cibersort_data$cell))

state_result <- state_result %>%
  left_join(
    cibersort_data %>% distinct(cell, group),
    by = c("cell" = "cell")
  )

p2 <- ggplot(cibersort_data, aes(x = cell, y = score, fill = group)) +
  geom_boxplot() +  
  scale_fill_manual(values = c(Normal = "#4E4E56", DCM_HFrEF = "#DA635D")) +  
  stat_pvalue_manual(state_result,
                     x = "cell",
                     y.position = max(cibersort_data$score) * 1.1,
                     size = 3,
                     color = "blue",
                     label = "p.label") +
  labs(x = "", y = "Estimated Proportion", fill = "Group") + 
  theme_minimal() + 
  theme(
    axis.title.y = element_text(size = 10), 
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10, color = "black"),  
    axis.text = element_text(size = 10),  
    legend.text = element_text(size = 10), 
    legend.position = 'top',  
    legend.title = element_text(size = 10),  
    panel.grid.major.x = element_blank(),  
    panel.grid.minor = element_blank() 
  )

print(p2)  
```

## Step 5: 免疫细胞间相关性分析
```{r}
# 相关性分析：仅保留数值列（排除分组等非数值信息）
numeric_cols <- sapply(cibersort_result, is.numeric)
# 计算Spearman相关系数（适合非正态分布数据）
cor_sorce <- cor(cibersort_result[, numeric_cols], method = 'spearman')
# 计算相关系数的P值（使用WGCNA包的函数）
cor_sorce_p <- WGCNA::corPvalueStudent(cor_sorce, nrow(cibersort_result))
# 设置颜色渐变：从深蓝（负相关）到浅黄再到深红（正相关）
col2 <- colorRampPalette(c("#053061", "#2166AC", "#4393C3", "#92C5DE", "#D1E5F0",
                           "#FDDBC7", "#F4A582", "#D6604D", "#B2182B", "#67001F"))(50)
# 定义绘制相关性热图的函数：
# 上三角显示相关系数数值，下三角显示圆圈（大小表示相关性强弱）和显著性标签
corr_plot <- function() {
  corrplot(cor_sorce,
           na.label = " ",  # 缺失值用空格表示
           diag = T,  # 显示对角线（自身相关性）
           method = "number",  # 上三角用数字表示相关系数
           type = "upper",  # 仅显示上三角
           tl.pos = "lt",  # 标签位置（左和上）
           tl.col = "black",  # 标签颜色
           tl.srt = 90,  # 标签旋转90度
           tl.cex = 0.9,  # 标签字体大小
           cl.length = 8,  # 颜色图例的分段数
           cl.cex = 0.8,  # 颜色图例字体大小
           col = col2,  # 使用自定义颜色
           number.cex = 0.7)  # 相关系数数值大小

  # 下三角添加显著性标签
  corrplot(cor_sorce,
           p.mat = cor_sorce_p,  # P值矩阵
           insig = "label_sig",  # 对不显著的相关性标记
           add = TRUE,  # 在已有图上添加
           type = "lower",  # 仅显示下三角
           method = "circle",  # 用圆圈表示相关性
           col = col2,  # 颜色与上三角一致
           na.label = " ",  # 缺失值处理
           diag = FALSE,  # 不显示对角线
           tl.pos = "n",  # 不显示标签
           cl.pos = "n",  # 不显示颜色图例
           sig.level = c(.0001, .001, .01, .05),  # 显著性水平分级
           pch.cex = 0.9,  # 显著性标签大小
           pch.col = "black")  # 显著性标签颜色
}
corr_plot()  # 绘制并显示相关性热图
```

## Step 6: 目标基因与免疫细胞的相关性分析
```{r}
# 定义您感兴趣的基因（请根据实际研究的基因修改）：这里以ACE和CASP1为例
interest_genes <- c("ACE","CASP1")  
gene_expr <- expr_fpkm_corr[rownames(expr_fpkm_corr) %in% interest_genes, , drop = FALSE]

gene_expr_t <- as.data.frame(t(gene_expr))
colnames(gene_expr_t) <- rownames(gene_expr)  
common_samples <- intersect(rownames(cibersort_result), rownames(gene_expr_t))
cibersort_sub <- cibersort_result[common_samples, cell_cols, drop = FALSE] %>%
  dplyr::select(where(~ any(. != 0)))

gene_expr_sub <- gene_expr_t[common_samples, , drop = FALSE] 
combined_data <- cbind(cibersort_sub, gene_expr_sub)

correlation_matrix <- cor(combined_data, method = "spearman")

n_immune <- ncol(cibersort_sub)  
n_genes <- length(interest_genes)  
gene_cell_cor <- correlation_matrix[(n_immune + 1):(n_immune + n_genes), 1:n_immune, drop = FALSE]

p_matrix <- matrix(0, nrow = n_genes, ncol = n_immune)
rownames(p_matrix) <- interest_genes
colnames(p_matrix) <- colnames(cibersort_sub)
for (gene in interest_genes) {
  for (cell in colnames(cibersort_sub)) {
    test <- cor.test(combined_data[[gene]], combined_data[[cell]], method = "spearman")
    p_matrix[gene, cell] <- test$p.value  
  }
}

sig_matrix <- ifelse(p_matrix < 0.001, "***", 
                     ifelse(p_matrix < 0.01, "**", 
                            ifelse(p_matrix < 0.05, "*", "")))
pheatmap(gene_cell_cor,
         color = colorRampPalette(c("#2166AC", "white", "#B2182B"))(100), 
         display_numbers = sig_matrix,  
         number_color = "black",
         number_format = "%s",  
         main = "Correlation between Genes and Immune Cells",  # 图标题
         treeheight_row = 20,  
         treeheight_col = 20, 
         fontsize = 10,  
         fontsize_row = 12,  
         fontsize_col = 10,  
         cellwidth = 25,  
         cellheight = 30)  
dev.off()  
```

# Figure 5
```{r}
```

---

<table class="nav-table" width="100%">
  <tr>
    <td align="left">
      [Home](../index.qmd) 
    </td>
    <td align="right">
      [Previous](Part_2_Screening_of_HFrEF_related_genes.qmd) | [Next](Part_4_Mechanism_validation_via_bioinformatics_and_docking.qmd)
    </td>
  </tr>
</table>

# References {-}