---
title: "Part 4: Mechanism validation via bioinformatics and docking"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

---

# Identification and Validation of Biomarker Genes by Machine Learning

## Load Packages
```{r}
library(tidyverse)
library(data.table)
library(caret)
library(glmnet)
library(e1071)
library(randomForest)
library(gbm)
library(pROC)
library(ggplot2)
library(ggsci)
library(corrplot)
library(openxlsx)
library(ggvenn)
```

## Step 1: Loading data
```{r}
rm(list = ls())

load("../data/processed/geo/GSE141910/GSE141910_preprocessed_final.RData")

load("../data/processed/network/08_hub_genes.RData")

expr_matrix <- t(expr_fpkm_corr[rownames(expr_fpkm_corr) %in% hub_genes,])

expr_matrix <- expr_matrix[,hub_genes]

group <- pheno_final[rownames(expr_matrix),"group"] %>% as.numeric()-1

expr_data <- data.frame(expr_matrix, group = group)
```

## Step 2: Stratified sampling
```{r}
set.seed(42)

train_index <- caret::createDataPartition(
  expr_data$group, 
  p = 0.7, 
  list = FALSE
)

train_data <- expr_data[train_index, ]
test_data  <- expr_data[-train_index, ]

x_train <- as.matrix(train_data[, -ncol(train_data)])
y_train <- train_data$group
x_test  <- as.matrix(test_data[, -ncol(test_data)])
y_test  <- test_data$group

save(train_data, test_data, x_train,y_train,x_test,y_test, file = "../data/processed/ml/01_model_data.RData")
```

## Step 4: Hub genes correlation analysis
```{r}
cor_matrix <- cor(x_train)

off_diag_abs <- abs(cor_matrix[lower.tri(cor_matrix)])

max_abs <- max(off_diag_abs, na.rm = TRUE)
min_abs <- min(off_diag_abs, na.rm = TRUE)

cor_df <- cor_matrix %>%
  as.data.frame() %>%
  rownames_to_column("Gene1") %>%  
  tidyr::pivot_longer(
    cols = -Gene1,                
    names_to = "Gene2",            
    values_to = "cor"              
  )

cor_df <- cor_df[as.character(cor_df$Gene1) != as.character(cor_df$Gene2), ]

cor_df$row_id <- as.numeric(factor(cor_df$Gene1, levels = hub_genes))

cor_df$col_id <- as.numeric(factor(cor_df$Gene2, levels = hub_genes))

cor_df <- cor_df[cor_df$row_id > cor_df$col_id, ]

p <- ggplot(cor_df, aes(x = col_id, y = row_id)) +
  geom_point(aes(fill = cor), shape = 21, size = 10, stroke = 0) +
  geom_text(aes(label = sprintf("%.2f", cor)), size = 2.5, color = "black") +  
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", 
                       midpoint = 0, limits = c(-1, 1),
                       name = "Correlation") +
  scale_y_continuous(breaks = 1:length(hub_genes), labels = hub_genes, trans = "reverse") +
  scale_x_continuous(breaks = 1:length(hub_genes), labels = hub_genes) +
  labs(x="",y="") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.2, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = c(0.85,0.55)) +
  coord_fixed()

ggsave("../figures/supplementary/Figure_S4A.pdf", p, width = 7, height = 7, dpi = 300)
```

## Step 4: ROC-AUC for each gene
```{r}
compute_auc <- function(gene, x, y) {
  roc_obj <- roc(response = y, predictor = x[, gene], quiet = TRUE, direction = "auto")
  auc_val <- round(auc(roc_obj), 3)
  return(auc_val)
}

gene_auc_df <- data.frame(
  Gene = hub_genes,
  Train_AUC = sapply(hub_genes, compute_auc, x = x_train, y = y_train),
  Test_AUC  = sapply(hub_genes, compute_auc, x = x_test,  y = y_test)
) 

hub_genes_tbl2 <- hub_genes_tbl %>%
  left_join(gene_auc_df,by="Gene")

save(gene_auc_df, hub_genes_tbl2, 
     file = "../data/processed/ml/02_gene_auc.RData")

write.xlsx(
  hub_genes_tbl2,
  "../tables/Table2_Hub_Genes_AUC.xlsx"
)
```

## Step 5: Feature selection

### z-scores
```{r}
rm(list = ls())

load("../data/processed/ml/01_model_data.RData")

train_data$group <- factor(train_data$group, levels = c(0,1), labels = c("Normal","DCM_HFrEF"))

test_data$group <- factor(test_data$group, levels = c(0,1), labels = c("Normal","DCM_HFrEF"))

df <- train_data
preProc <- caret::preProcess(df[, -which(colnames(df)=="group")], method = c("center", "scale"))

df_scaled <- df
df_scaled[, -which(colnames(df)=="group")] <- predict(preProc, df[, -which(colnames(df)=="group")])

set.seed(42)
train_control <- caret::trainControl(
  method = "cv",
  number = 10,
  classProbs = TRUE,
  summaryFunction = twoClassSummary
)
```

### LASSO
```{r}
set.seed(42)

lasso_model <- caret::train(
  group ~ .,
  data = df_scaled,
  method = "glmnet",
  metric = "ROC",
  trControl = train_control,
  tuneGrid = expand.grid(alpha=1, lambda=seq(0.001,0.1,0.001))
)

lasso_coef <- coef(lasso_model$finalModel, lasso_model$bestTune$lambda)
lasso_selected <- rownames(lasso_coef)[which(lasso_coef[,1]!=0)][-1]

cat("LASSO selected:", length(lasso_selected), "genes\n")
cat(lasso_selected, "\n")

save(lasso_model,lasso_selected, file = "../data/processed/ml/02_lasso.RData")
```

### SVM-RFE
```{r}
set.seed(42)

svm_model <- caret::train(
  group ~ .,
  data = df_scaled,
  method = "svmLinear",   
  metric = "ROC",
  trControl = train_control  
)

imp_svm <- varImp(svm_model)$importance
svm_score <- imp_svm$DCM_HFrEF   
names(svm_score) <- rownames(imp_svm)
svm_selected <- names(sort(svm_score, decreasing=TRUE))[svm_score>0]

cat("SVM-RFE selected:", length(svm_selected), "genes\n")
cat(svm_selected, "\n")

save(svm_model,svm_selected, file = "../data/processed/ml/03_svm_rfe.RData")
```

### Random Forest
```{r}
set.seed(42)
rf_model <- caret::train(
  group ~ .,
  data = df_scaled,
  method = "rf",          
  metric = "ROC",
  trControl = train_control,  
  ntree = 500
)
imp_rf <- varImp(rf_model)$importance
imp_rf <- imp_rf[order(imp_rf$Overall, decreasing = TRUE), , drop=FALSE] 
rf_selected <- rownames(imp_rf)[imp_rf$Overall > 0] 

cat("RF selected:", length(rf_selected), "genes\n")

cat(rf_selected, "\n")

save(rf_model, rf_selected, file = "../data/processed/ml/04_rf.RData")
```

### GBM
```{r}
set.seed(42)
gbm_grid <- expand.grid(
  n.trees = 100,
  interaction.depth = 2,
  shrinkage = 0.1,
  n.minobsinnode = 5
)

gbm_model <- caret::train(
  group ~ .,
  data = df_scaled,
  method = "gbm",
  metric = "ROC",
  trControl = train_control,
  tuneGrid = gbm_grid,
  verbose = FALSE 
)

imp_gbm <- varImp(gbm_model)$importance
gbm_selected <- rownames(imp_gbm)[imp_gbm$Overall > 0]

cat("GBM selected:", length(gbm_selected), "features\n")
cat(gbm_selected, "\n")

save(gbm_model,gbm_selected, file = "../data/processed/ml/05_gbm.RData")
```

### Consensus genes
```{r}
consensus_genes <- Reduce(intersect, list(lasso_selected, svm_selected, rf_selected, gbm_selected))
cat("Consensus genes:", length(consensus_genes), "\n")
print(consensus_genes)

consensus_genes <- data.frame(Gene = consensus_genes)

save(consensus_genes, file =  "../data/processed/ml/06_consensus_genes.RData")

write.xlsx(consensus_genes, "../tables/consensus_genes.xlsx")
```

## Step 6: Feature selection Visualization
```{r}
# Figure S5A
lasso_coef <- coef(lasso_model$finalModel, lasso_model$bestTune$lambda)
lasso_df <- data.frame(
  Gene = rownames(lasso_coef)[-1],
  Coef = as.numeric(lasso_coef)[-1]
) %>% filter(Coef != 0)

pS5A <- ggplot(lasso_df, aes(x = reorder(Gene, Coef), y = Coef, fill = Coef > 0)) +
  geom_col() +
  coord_flip() +
  labs( x = "", y = "Coefficient") +
  theme_bw() +
  theme(
    text = element_text( size = 10),
    axis.text = element_text(size=7),
    axis.title = element_text(size = 8, face = "bold"),
    legend.position = "none",
    panel.grid = element_blank(),
    # axis.ticks.y = element_blank()
  )
ggsave("../figures/supplementary/Figure_S5A.pdf",pS5A, width = 7, height = 5,units = "cm",dpi = 300)


# Figure S5B
imp_svm <- varImp(svm_model)$importance
svm_df <- data.frame(
  Gene = rownames(imp_svm),
  Score = imp_svm$DCM_HFrEF
) %>% arrange(desc(Score)) %>% head(15)

pS5B <- ggplot(svm_df, aes(x = reorder(Gene, Score), y = Score, fill = Score)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "", y = "Importance") +
  theme_bw() +
  theme(
    text = element_text( size = 10),
    axis.text = element_text(size=7),
    axis.title = element_text(size = 8, face = "bold"),
    legend.position = "none",
    panel.grid = element_blank(),
    axis.ticks.y = element_blank()
  )

ggsave("../figures/supplementary/Figure_S5B.pdf",pS5B, width = 7, height = 5,units = "cm",dpi = 300)


# Figure S5C
imp_rf <- varImp(rf_model)$importance
rf_df <- data.frame(
  Gene = rownames(imp_rf),
  Score = imp_rf$Overall
) %>% arrange(desc(Score))

pS5C <- ggplot(rf_df, aes(x = reorder(Gene, Score), y = Score, fill = Score)) +
  geom_col() +
  coord_flip() +
  labs(x = "", y = "Importance score") +
  scale_y_continuous(expand = c(0, 0)) +
  theme_bw() +
  theme(
    text = element_text( size = 10),
    axis.text = element_text(size=7),
    axis.title = element_text(size = 8, face = "bold"),
    legend.position = "none",
    panel.grid = element_blank(),
    axis.ticks.y = element_blank()
  )
ggsave("../figures/supplementary/Figure_S5C.pdf", pS5C, width = 7, height = 5)

## Figure S5D
imp_gbm <- varImp(gbm_model)$importance
gbm_df <- data.frame(
  Gene = rownames(imp_gbm),
  Score = imp_gbm$Overall
) %>% arrange(desc(Score)) %>% head(15)

pS5D <- ggplot(gbm_df, aes(x = reorder(Gene, Score), y = Score, fill = Score)) +
  geom_col() +
  coord_flip() +
  labs(x = "", y = "Importance") +
  scale_y_continuous(expand = c(0, 0)) +
  theme_bw() +
  theme(
    text = element_text( size = 10),
    axis.text = element_text(size=7),
    axis.title = element_text(size = 8, face = "bold"),
    legend.position = "none",
    panel.grid = element_blank(),
    axis.ticks.y = element_blank()
  )

ggsave("../figures/supplementary/Figure_S5D.pdf", pS5D, width = 7, height = 5)
```

### Venn Plot: Consensus genes 
```{r}
# Venn
gene_list <- list(
  LASSO = lasso_selected,
  `SVM-RFE` = svm_selected,
  RF = rf_selected,
  GBM = gbm_selected
)

p <- ggvenn(
  gene_list,
  fill_color = pal_aaas()(4),
  stroke_color = "white",
  stroke_size = 1,
  set_name_size = 5,
  text_size = 5,
  text_color = "white",
  show_percentage = FALSE
) +
  theme(plot.title = element_text(hjust = 0.5, size=7))

# save
ggsave("../figures/main/Figure_6A.pdf", p, width=7, height=5, dpi=300, units="cm")
ggsave("../figures/main/Figure_6A.png", p, width=7, height=5, dpi=300, units="cm")
```

## Step 7: Logistic regression model
```{r}
load("../data/processed/ml/06_consensus_genes.RData")
load("../data/processed/ml/01_model_data.RData")

formula_lr <- as.formula(paste("group ~", paste(consensus_genes$Gene, collapse = "+")))
lr_model <- glm(formula_lr, data = train_data, family = binomial())

save(lr_model, file = "../data/processed/ml/07_lr_model.RData")
```

## Step 8: Nomogram
```{r}
library(rms)
# rms
dd <- datadist(train_data)
options(datadist = dd)

lrm_mod <- lrm(formula_lr, data = train_data, x = TRUE, y = TRUE)

nom <- nomogram(lrm_mod,
                fun = plogis,          
                funlabel = "Predicted Probability",
                fun.at = c(0.05, 0.5,0.95),
                lp = FALSE)     

pdf("../figures/main/Figure_6B.pdf", width = 7, height = 5)
par(mar = c(1, 1, 1, 1))  
plot(nom,
     xfrac = 0.35,   
     cex.axis = 0.6, 
     cex.lab = 0.7)  
dev.off()

```

## Step 8: External validation

### GSE116250: External validation set
```{r}
library(org.Hs.eg.db)

rm(list = ls())

load("../data/processed/ml/06_consensus_genes.RData")
load("../data/processed/ml/07_lr_model.RData")
load("../data/raw/geo/gse116250_raw.RData")

gse116250_expr_trans <- gse116250_fpkm_raw %>%
  column_to_rownames("GeneID") %>%
  mutate(across(everything(), ~ .x + 1)) %>%  
  log2() %>%
  t() %>% 
  as.data.frame()

entrez_ids <- colnames(gse116250_expr_trans)  
symbols <- mapIds(org.Hs.eg.db, keys = entrez_ids, keytype = "ENTREZID", column = "SYMBOL", multiVals = "first")

keep <- !is.na(symbols)
gse116250_expr_annot <- t(gse116250_expr_trans[, keep]) %>%
  as.data.frame() %>%
  mutate(Gene = symbols[keep], .before = 1)

common_genes <- intersect(consensus_genes$Gene, (gse116250_expr_annot$Gene))

if (all(common_genes %in% consensus_genes$Gene)) {
  cat("Good dataset\n")
} else {
  cat("Bad dataset\n")
}

gse116250_expr_merged <- gse116250_expr_annot %>%
  as.data.frame() %>%
  dplyr::filter(Gene %in% common_genes) %>%
  group_by(Gene) %>%
  summarise(across(everything(), mean, na.rm = TRUE)) %>%
  column_to_rownames("Gene") %>%
  t() %>%                          
  as.data.frame()

gse116250_expr_merged <- gse116250_expr_merged[,consensus_genes$Gene]

# pheno info
Normal <- trimws(gse116250_pheno_raw$characteristics_ch1.1) == "disease: non-failing"
DCM_HFrEF <- trimws(gse116250_pheno_raw$characteristics_ch1.1) == "disease: dilated cardiomyopathy"

gse116250_expr_filtered <- rbind(gse116250_expr_merged[Normal,],gse116250_expr_merged[DCM_HFrEF,]) %>%
  mutate(group = c(rep(0,sum(Normal)), rep(1,sum(DCM_HFrEF))))

save(gse116250_expr_filtered, file = "../data/processed/ml/gse116250.RData")
```

### GSE19303: External validation set  
```{r}
rm(list = ls())

library(AnnoProbe)

load("../data/processed/ml/06_consensus_genes.RData")
load("../data/processed/ml/07_lr_model.RData")
load("../data/raw/geo/gse19303_raw.RData")

ex <- gse19303_exprs_raw
qx <- as.numeric(quantile(ex, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))

LogC <- (qx[5] > 100) ||
          (qx[6]-qx[1] > 50 && qx[2] > 0)
if (LogC) { ex[which(ex <= 0)] <- NaN
  ex <- log2(ex + 1) }

probe2gene <- idmap("GPL570")
gse19303_expr_annot <-  as.data.frame(ex) %>% 
  rownames_to_column("probe_id") %>%
  inner_join(probe2gene, by = "probe_id") %>%
  dplyr::select(-probe_id) %>%
  mutate(Gene = symbol, .before = 1) %>%
  dplyr::select(-symbol)

common_genes <- intersect(consensus_genes$Gene, (gse19303_expr_annot$Gene))

if (all(common_genes %in% consensus_genes$Gene)) {
  cat("Good dataset\n")
} else {
  cat("Bad dataset\n")
}

gse19303_expr_merged <- gse19303_expr_annot %>%
  as.data.frame() %>%
  dplyr::filter(Gene %in% common_genes) %>%
  group_by(Gene) %>%
  summarise(across(everything(), mean, na.rm = TRUE)) %>%
  column_to_rownames("Gene") %>%
  t() %>%                          
  as.data.frame()

gse19303_expr_merged <- gse19303_expr_merged[,consensus_genes$Gene]

# pheno info
Normal <- !grepl("^DCM_", gse19303_pheno_raw$title)
DCM_HFrEF <- grepl("^DCM_", gse19303_pheno_raw$title) & 
                  as.numeric(gse19303_pheno_raw$`lvef:ch1`) <= 40

# grepl("_baseline", gse19303_pheno_raw$title) &

gse19303_expr_filtered <- rbind(gse19303_expr_merged[Normal,],gse19303_expr_merged[DCM_HFrEF,]) %>%
  mutate(group = c(rep(0,sum(Normal)), rep(1,sum(DCM_HFrEF))))

save(gse19303_expr_filtered, file = "../data/processed/ml/gse19303.RData")
```


## Step 9: ROC 


```{r}
rm(list = ls())

source("external_roc_cm.R")

load("../data/processed/ml/01_model_data.RData")
load("../data/processed/ml/gse116250.RData")
load("../data/processed/ml/gse19303.RData")
load("../data/processed/ml/07_lr_model.RData")

dataset_list <- list(
  Train = list(data=train_data, name="GSE141910_Train"),
  Test = list(data=test_data, name="GSE141910_Test"),
  GSE116250 = list(data=gse116250_expr_filtered, name="GSE116250")
  # ,
  # GSE19303 = list(data=gse19303_expr_filtered, name="GSE19303")
)


roc_all <- map_dfr(dataset_list, ~calc_roc(lr_model, .x$data, .x$name)$df)

p6C <- ggplot(
  roc_all, 
  aes(x=Specificity, y=Sensitivity, color=Group)
  ) +
  geom_line(linewidth = 1) +
  geom_abline(slope=1, intercept=0, linetype="dashed", color="gray50") +
  scale_color_aaas() +
  labs(x="1 - Specificity", y="Sensitivity") +
  theme_bw() + 
  theme(legend.position = c(0.7,0.2), 
        legend.background = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        axis.title = element_text(size = 10),
        panel.grid = element_blank(),
      )

ggsave("../figures/main/Figure_6C.pdf", p6C, width=7, height=7, dpi=300)

```

## Step 10: calibration curve
```{r}
dd <- datadist(train_data)
options(datadist = "dd")
form <- as.formula(paste0("group ~ ", paste(consensus_genes$Gene, collapse = "+")))
fit <- lrm(form,data=train_data,x=T,y=T)
cal_train <- calibrate(fit,method="boot",B=1000)

cal_manual <- function(testdata,model,name){
  pred <- predict(model,newdata = testdata,type="fitted")
  df <- data.frame(pred=pred,obs=testdata$group)
  fit_loess <- loess(obs~pred,data=df,span=0.75)
  seq_x <- seq(0,1,length.out=100)
  seq_y <- predict(fit_loess,newdata=data.frame(pred=seq_x))
  seq_y[seq_y>1] <-1;seq_y[seq_y<0]<-0
  return(data.frame(Pred=seq_x,Obs=seq_y,Dataset=name))
}

df_train <- data.frame(
  Pred = cal_train[, "predy"],
  Obs  = cal_train[, "calibrated.corrected"],
  Dataset = "GSE141910_Train"
)

cal_manual <- function(newdat, mod, name){
  pred <- predict(mod, newdata = newdat, type = "fitted")
  obs <- newdat$group
  df <- data.frame(pred = pred, obs = obs)
  lo <- loess(obs ~ pred, data = df, span = 0.75)
  xseq <- seq(0, 1, length.out = 100)
  yseq <- predict(lo, newdata = data.frame(pred = xseq))
  yseq[yseq > 1] <- 1
  yseq[yseq < 0] <- 0
  return(data.frame(Pred = xseq, Obs = yseq, Dataset = name))
}

df_test <- cal_manual(test_data[, c(consensus_genes$Gene, "group")], fit, "GSE141910_Test")
df_gse116250  <- cal_manual(gse116250_expr_filtered[, c(consensus_genes$Gene, "group")],  fit, "GSE116250")

all_cal <- rbind(df_train, df_test, df_gse116250)

p6D <- ggplot(all_cal, aes(x=Pred, y=Obs, color=Dataset)) +
  geom_smooth(method="loess", se=F, span=0.8, size=1) +
  geom_abline(slope=1, intercept=0, linetype="dashed", color="black") +
  scale_color_brewer(palette = "Set2") +
  labs(x="Predicted Probability", y="Actual Probability") +
  xlim(0,1) + ylim(0,1) +
  theme_bw()+
  theme(panel.grid=element_blank())

ggsave("../figures/main/Figure_6D.pdf", p6D, width=7, height=7, dpi=300)
```

## herb-ingredient-target
```{r}
load("../data/processed/tcm/14_compound_target_final.RData")
load("../data/processed/ml/06_consensus_genes.RData")

hit_core <- compound_target_final %>%
    dplyr::filter(Symbol %in% consensus_genes$Gene) %>%
  arrange(Herb.Name.Pinyin,Symbol)

unique(hit_core$Herb.Name.Pinyin)

key_compounds <- hit_core[,c("Herb.Name.Pinyin", "Compound.Name","Symbol")]
```

# Figure 6
```{r}

```


# Single‑Cell Transcriptome Expression Mapping
## Load Packages
```{r}
library(Seurat)
library(ggplot2)
library(patchwork)
library(dplyr)
library(Matrix)
library(harmony)
```

## Step 1: Loading data

```{r}
rm(list=ls())
gc()
load("../data/processed/ml/06_consensus_genes.RData")

load("../data/raw/geo/GSE183852/GSE183852_DCM_Cells/GSE183852_DCM_Cells.Robj")

new.cluster.ids <- c(
  "Fibroblasts",
  "Endothelium",
  "Pericytes",
  "Macrophages",
  "T_Cells",
  "Smooth_Muscle",
  "Monocytes",
  "NK_Cells",
  "Neurons",
  "B_Cells",
  "Lymphatic"
)

```

## Step 2: 细胞注释UMAP
```{r}
p1 <- DimPlot(HDCM, label = TRUE, repel = TRUE, pt.size = 0.1, label.size = 2.5) +
  theme_bw(base_family = "sans") +  
  theme(
    panel.border = element_rect(fill = NA, color = "black", size = 1.0),
    legend.text = element_text(size = 7),
    legend.title = element_text(size = 8),
    legend.spacing.y = unit(0.02, "cm"),  
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 8),  
    axis.title = element_text(size = 8)  
  )

ggsave("../results/figures/main/Figure_6A_UMAP.pdf", p1, width = 10, height = 8, units = "cm",dpi = 300)

ggsave("../results/figures/main/Figure_6A_UMAP.png", p1, width = 10, height = 8, units = "cm",dpi = 300)
```



---




# Bayesian co-localization analysis

## Load Packages
```{r}
library(data.table)   
library(ggplot2)      
library(coloc)         
library(tidyverse) 
library(arrow)
library(rtracklayer)
library(GenomicRanges)
library(future)
library(future.apply)
```

## Step 1: Set paths
```{r}
# download.file("http://hgdownload.cse.ucsc.edu/goldenPath/hg19/liftOver/hg19ToHg38.over.chain.gz", destfile = "../data/raw/opengwas/hg19ToHg38.over.chain.gz")

data_dir <- "../data/raw/opengwas"          
out_dir  <- "../data/processed/opengwas" 

chain_file <- file.path(data_dir, "hg19ToHg38.over.chain") 

if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)

# Core genes
load("../data/processed/ml/06_consensus_genes.RData")

```

## Step 2: reading and cleaning GTEx eQTL data
```{r}
cat("=== Reading GTEx eQTL (left ventricle) ===\n")

eqtl_file <- file.path(data_dir, "GTEx_Analysis_v8_eQTL/Heart_Left_Ventricle.v8.signif_variant_gene_pairs.txt.gz")

eqtl_raw <- fread(eqtl_file, header = TRUE, sep = "\t")
save(eqtl_raw, file = file.path(out_dir,"eqtl_heart_LV_raw.RData"))

eqtl_clean <- eqtl_raw %>%
  dplyr::mutate(
    gene_id = stringr::str_remove(gene_id, "\\..*$"),
    pval = pval_nominal,
    beta = slope,
    varbeta = slope_se^2,
    MAF = maf
  ) %>%
  dplyr::filter(pval < 1e-5) %>%
  dplyr::mutate(
    id_temp = stringr::str_remove(variant_id, "_b38$"),
    chr = stringr::str_extract(id_temp, "^chr[0-9XYM]+"),
    pos = as.numeric(stringr::str_extract(id_temp, "(?<=_)[0-9]+(?=_)")),
    ref = stringr::str_extract(id_temp, "(?<=_)[ATCG]+(?=_[ATCG]+$)"),
    alt = stringr::str_extract(id_temp, "(?<=_)[ATCG]+$")
  ) %>%
  dplyr::filter(
    !is.na(chr), 
    !is.na(pos), 
    !is.na(ref), 
    !is.na(alt)
  ) %>%
  dplyr::select(
    gene_id, 
    chr, 
    pos, 
    ref, 
    alt, 
    beta, 
    varbeta, 
    MAF, 
    pval, 
    variant_id
  )

cat("eQTL valid records (P<1e-5):", nrow(eqtl_clean), "\n")

save(eqtl_clean, file= file.path(out_dir, "eqtl_heart_LV_clean.RData"))

rm(eqtl_raw, eqtl_clean); gc()


```

## Step 3: reading and lifting over HERMES GWAS (hg19 → hg38)

- [HERMES dilated cardiomyopathy 2024 GWAS: European ancestry](https://www.kp4cd.org/node/1482)

- 

```{r}
# cat("\n=== Reading HERMES GWAS ===\n")
# hermes_raw <- fread(file.path(data_dir, "HERMES2_GWAS_DCM_EUR/FORMAT-METAL_Pheno5-DCM_EUR.tsv.gz"))
# 
# save(hermes_raw, file = file.path(out_dir, "hermes_DCM_GWAS_raw.RData"))

load(file.path(out_dir, "hermes_DCM_GWAS_raw.RData"))

# sample sizes
# https://doi.org/10.1038/s41588-024-01952-y
N_cases_hermes   <- 6001
N_controls_hermes <- 449382   # from the paper
N_total_hermes    <- N_cases_hermes + N_controls_hermes   # 455383
s_hermes          <- N_cases_hermes / N_total_hermes

cat("HERMES sample sizes: cases =", N_cases_hermes, ", controls =", N_controls_hermes,
    ", total =", N_total_hermes, "\n")

# Prepare GRanges for liftOver
hermes_gr <- GenomicRanges::GRanges(
  seqnames = paste0("chr", hermes_raw$chr),
  ranges = IRanges::IRanges(start = hermes_raw$pos_b37, end = hermes_raw$pos_b37)
)
chain <- rtracklayer::import.chain(chain_file)
hermes_gr_hg38 <- rtracklayer::liftOver(hermes_gr, chain)
success <- which(lengths(hermes_gr_hg38) > 0)
hermes_hg38 <- hermes_raw[success, ]
hermes_hg38$chr <- as.character(GenomicRanges::seqnames(unlist(hermes_gr_hg38[success])))
hermes_hg38$pos <- IRanges::start(unlist(hermes_gr_hg38[success]))

# Build HERMES GWAS dataset
hermes_gwas <- hermes_hg38 %>%
  dplyr::mutate(
    ref = A2,
    alt = A1,
    beta = A1_beta,
    varbeta = se^2,
    pval = pval,
    variant_id = rsID,
    MAF_raw = A1_freq,
    MAF = ifelse(MAF_raw > 0.5, 1 - MAF_raw, MAF_raw)
  ) %>%
  dplyr::filter(!is.na(beta), !is.na(varbeta), !is.na(pval)) %>%
  dplyr::select(chr, pos, ref, alt, beta, varbeta, pval, variant_id, MAF)

cat("HERMES successfully converted SNPs:", nrow(hermes_gwas),
    sprintf("(%.2f%%)", 100 * nrow(hermes_gwas) / nrow(hermes_raw)), "\n")


save(hermes_gwas, N_cases_hermes, N_total_hermes,  file = file.path(out_dir, "hermes_DCM_GWAS_hg38.RData"))

rm(hermes_raw, hermes_gr, hermes_gr_hg38, chain, success); gc()
```

## Step 4: reading Jurgens GWAS (hg38)

- [Dilated cardiomyopathy 2024 GWAS: European ancestry](https://www.kp4cd.org/node/1558)

```{r}
# Jurgens GWAS summary statistics
# jurgens_raw <- fread("../data/raw/opengwas/Jurgens_DCM_GWAS/DCM_GWAS/Jurgens_DCM_GWAS_META.tsv.gz", header = TRUE)

# save(jurgens_raw, file = file.path(out_dir, "jurgens_DCM_GWAS_raw.RData"))

load(file.path(out_dir, "jurgens_DCM_GWAS_raw.RData"))

# names(jurgens_raw)

# https://www.nature.com/articles/s41588-024-01975-5
# N_cases_jurgens <- max(jurgens_raw$N_cases)
# N_controls_jurgens <- max(jurgens_raw$N_controls)
N_cases_jurgens <- 9365
N_controls_jurgens <- 946368  # from paper
N_total_jurgens <- N_cases_jurgens + N_controls_jurgens

s_jurgens <- N_cases_jurgens / N_total_jurgens

jurgens_gwas <- jurgens_raw %>%
  dplyr::mutate(
    chr = paste0("chr", CHR),
    pos = POS,
    ref = NEA,
    alt = EA,
    beta = BETA,
    varbeta = SE^2,
    pval = P,
    variant_id = ifelse(is.na(rsID) | rsID == "", ID_B38, rsID),
    MAF_raw = EAFREQ,
    MAF = ifelse(MAF_raw > 0.5, 1 - MAF_raw, MAF_raw)
  ) %>%
  dplyr::filter(!is.na(pval)) %>%
  dplyr::select(chr, pos, ref, alt, beta, varbeta, pval, variant_id, MAF)

cat("Jurgens 有效 SNP 数:", nrow(jurgens_gwas), "\n")
rm(jurgens_raw); gc()


save(jurgens_gwas, N_cases_jurgens, N_controls_jurgens, N_total_jurgens, s_jurgens, file = "../data/processed/opengwas/jurgens_DCM_GWAS_hg38.RData")

rm(jurgens_raw, jurgens_gwas, jurgens_gwas); gc()
```

## Step 5: matching SNPs (only identical alleles, no flipping)
```{r}
match_identical <- function(eqtl_df, gwas_df) {
  eqtl_match <- eqtl_df %>%
    dplyr::select(gene_id, chr, pos, ref_eqtl = ref, alt_eqtl = alt,
                  beta_eqtl = beta, varbeta_eqtl = varbeta, MAF_eqtl = MAF,
                  pval_eqtl = pval, variant_id_eqtl = variant_id)
  gwas_match <- gwas_df %>%
    dplyr::select(chr, pos, ref_gwas = ref, alt_gwas = alt,
                  beta_gwas = beta, varbeta_gwas = varbeta, MAF_gwas = MAF,
                  pval_gwas = pval, variant_id_gwas = variant_id)
  matched_pos <- dplyr::inner_join(eqtl_match, gwas_match, by = c("chr", "pos"))
  matched_identical <- matched_pos %>%
    dplyr::filter(ref_eqtl == ref_gwas & alt_eqtl == alt_gwas) %>%
    dplyr::select(-ref_eqtl, -alt_eqtl, -ref_gwas, -alt_gwas)
  return(matched_identical)
}

load(file.path(out_dir, "eqtl_heart_LV_clean.RData"))
load(file.path(out_dir, "hermes_DCM_GWAS_hg38.RData"))
load(file.path(out_dir, "jurgens_DCM_GWAS_hg38.RData"))

matched_hermes <- match_identical(eqtl_clean, hermes_gwas)
matched_jurgens <- match_identical(eqtl_clean, jurgens_gwas)

cat("\n=== Matching results ===\n")
cat("HERMES matched records (identical alleles):", nrow(matched_hermes), "\n")
cat("Jurgens matched records (identical alleles):", nrow(matched_jurgens), "\n")

rm(hermes_gwas, jurgens_gwas); gc()
```

## Step 6: evaluating match quality
```{r}
evaluate_matching <- function(matched_df, original_eqtl, name) {
  cat("\n========== ", name, " matching quality ==========\n")
  total_eqtl <- nrow(original_eqtl)
  matched_n <- nrow(matched_df)
  cat("Match rate:", matched_n, "/", total_eqtl, sprintf("(%.2f%%)", 100 * matched_n / total_eqtl), "\n")
  
  snp_per_gene <- matched_df %>% dplyr::group_by(gene_id) %>% dplyr::summarise(n = dplyr::n())
  cat("SNPs per gene: median =", median(snp_per_gene$n), 
      ", mean =", round(mean(snp_per_gene$n),1),
      ", range =", min(snp_per_gene$n), "-", max(snp_per_gene$n), "\n")
  cat("Genes with ≥10 SNPs:", sum(snp_per_gene$n >= 10), "/", nrow(snp_per_gene), "\n")
  
  if (all(c("MAF_eqtl", "MAF_gwas") %in% names(matched_df))) {
    maf_cor <- cor(matched_df$MAF_eqtl, matched_df$MAF_gwas, method = "spearman", use = "complete.obs")
    cat("MAF Spearman correlation:", round(maf_cor, 4), "\n")
  }
  
  if (all(c("pval_eqtl", "pval_gwas") %in% names(matched_df))) {
    logp_eqtl <- -log10(matched_df$pval_eqtl)
    logp_gwas <- -log10(matched_df$pval_gwas)
    p_cor <- cor(logp_eqtl, logp_gwas, method = "spearman", use = "complete.obs")
    cat("-log10(P) Spearman correlation:", round(p_cor, 4), "\n")
  }
  
  original_genes <- unique(original_eqtl$gene_id)
  matched_genes <- unique(matched_df$gene_id)
  cat("Gene coverage:", length(matched_genes), "/", length(original_genes),
      sprintf("(%.2f%%)", 100 * length(matched_genes) / length(original_genes)), "\n")
}

# Save original eqtl_clean for evaluation
eqtl_original <- eqtl_clean
evaluate_matching(matched_hermes, eqtl_original, "HERMES")
evaluate_matching(matched_jurgens, eqtl_original, "Jurgens")
```

## Step 7: filtering genes with ≥10 SNPs
```{r}
filter_genes <- function(matched_df) {
  genes_keep <- matched_df %>%
    dplyr::group_by(gene_id) %>%
    dplyr::summarise(n = dplyr::n()) %>%
    dplyr::filter(n >= 10) %>%
    dplyr::pull(gene_id)
  filtered <- matched_df %>% dplyr::filter(gene_id %in% genes_keep)
  cat("Genes retained (≥10 SNPs):", length(genes_keep), "\n")
  cat("Total SNP‑gene pairs after filtering:", nrow(filtered), "\n")
  return(filtered)
}

matched_hermes_final <- filter_genes(matched_hermes)
matched_jurgens_final <- filter_genes(matched_jurgens)

rm(eqtl_clean, eqtl_original, matched_hermes, matched_jurgens); gc()
```

---

## Step 8: running P‑value based colocalisation in parallel
```{r}
run_coloc_pval_gene <- function(sub_df, N_gwas, s_gwas, N_eqtl = 386,
                                p1 = 1e-4, p2 = 1e-4, p12 = 1e-5) {
  if (nrow(sub_df) < 10) return(NULL)
  d1 <- list(pvalues = sub_df$pval_eqtl,
             snp = sub_df$variant_id_eqtl,
             N = N_eqtl,
             type = "quant")
  d2 <- list(pvalues = sub_df$pval_gwas,
             snp = sub_df$variant_id_gwas,
             N = N_gwas,
             type = "cc",
             s = s_gwas)
  res <- tryCatch(coloc.abf(d1, d2, p1 = p1, p2 = p2, p12 = p12),
                  error = function(e) NULL)
  if (is.null(res)) return(NULL)
  data.frame(gene = unique(sub_df$gene_id),
             n_snps = res$summary["n snps"],
             PP.H0 = res$summary["PP.H0"],
             PP.H1 = res$summary["PP.H1"],
             PP.H2 = res$summary["PP.H2"],
             PP.H3 = res$summary["PP.H3"],
             PP.H4 = res$summary["PP.H4"])
}
```

## Step 9: Parallel wrapper
```{r}
run_parallel_pval <- function(matched_df, N_gwas, s_gwas, ncores = 8) {
  gene_list <- split(matched_df, matched_df$gene_id)
  cat("Total genes to analyse:", length(gene_list), "\n")
  plan(multisession, workers = ncores)
  results <- future_lapply(gene_list, function(x) {
    run_coloc_pval_gene(x, N_gwas, s_gwas)
  }, future.seed = TRUE)
  plan(sequential)
  results <- do.call(rbind, results)
  return(results)
}
```


## Step 10: Run colocalisation
```{r}
cat("\n=== 4. Running HERMES colocalisation (P‑value based) ===\n")

ncores <- parallel::detectCores() - 1
res_hermes <- run_parallel_pval(matched_hermes_final, N_total_hermes, s_hermes, ncores = ncores)

cat("\n=== 5. Running Jurgens colocalisation (P‑value based) ===\n")
res_jurgens <- run_parallel_pval(matched_jurgens_final, N_total_jurgens, s_jurgens, ncores = ncores)
```

## Step: 11. Identify high‑confidence colocalisation (PP.H4 > 0.8) 
```{r}
hits_hermes <- res_hermes %>% dplyr::filter(PP.H4 > 0.8)
hits_jurgens <- res_jurgens %>% dplyr::filter(PP.H4 > 0.8)
```

## Step 12. Save results
```{r}
fwrite(res_hermes, file.path(out_dir, "coloc_results_HERMES_pval.csv"))
fwrite(hits_hermes, file.path(out_dir, "coloc_hits_HERMES_pval_PPH4_0.8.csv"))
fwrite(res_jurgens, file.path(out_dir, "coloc_results_Jurgens_pval.csv"))
fwrite(hits_jurgens, file.path(out_dir, "coloc_hits_Jurgens_pval_PPH4_0.8.csv"))
```

## Step 13. Summary 
```{r}
cat("\n========== FINAL SUMMARY ==========\n")
cat("HERMES: analysed", nrow(res_hermes), "genes; high‑confidence hits (PP.H4>0.8):", nrow(hits_hermes), "\n")
if (nrow(hits_hermes) > 0) print(hits_hermes[, c("gene", "PP.H4")])
cat("\nJurgens: analysed", nrow(res_jurgens), "genes; high‑confidence hits (PP.H4>0.8):", nrow(hits_jurgens), "\n")
if (nrow(hits_jurgens) > 0) print(hits_jurgens[, c("gene", "PP.H4")])

overlap <- intersect(hits_hermes$gene, hits_jurgens$gene)
cat("\nOverlap between GWAS hits:", length(overlap), "genes\n")
if (length(overlap) > 0) {
  cat("Overlapping genes:", paste(overlap, collapse = ", "), "\n")
  writeLines(overlap, file.path(out_dir, "overlap_genes.txt"))
}

cat("\nAll results saved in:", out_dir, "\n")
cat("Analysis completed.\n")
```

# GSEA Pathway Analysis of Causal Genes

# Figure 7
```{r}
```
---

# Docking of Active Ingredients and Core Targets


# Figure 8
```{r}
```
---

<table class="nav-table" width="100%">
  <tr>
    <td align="left">
      [Home](../index.qmd) 
    </td>
    <td align="right">
      [Previous](Part_3_Mechanism_exploration_of_ZWHQD_against_HFrEF.qmd) | [Next](Part_5_TF_miRNA_regulatory_network.qmd)
    </td>
  </tr>
</table>

# References {-}