---
title: "Part 1. Data acquisition and preprocessing"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

---

# Identification of Potential Targets for ZWHQD

## Active Ingredients Acquisition of ZWHQD
In the present study, the [DCABM-TCM database](http://bionet.ncpsb.org.cn/dcabm-tcm/) [@liu2023DCABM] was used as the primary source to systematically retrieve blood‑absorbed compounds from each medicinal herb in ZWHQD, along with their chemical structural information. These candidate components were screened according to established criteria, and standardized chemical data were obtained from the [PubChem database] (https://pubchem.ncbi.nlm.nih.gov/) [@kim2025pubchem].

Specifically, we identified and collected blood-absorbed compounds from six medicinal herbs of ZWHQD (see @tbl-1) by performing Chinese Pinyin-based herb queries in the [DCABM-TCM database](http://bionet.ncpsb.org.cn/dcabm-tcm/) [@liu2023DCABM]. To refine the candidate list, we applied multiple stringent filters: (i) compounds without a valid PubChem Compound Identifier (CID) were discarded; (ii) highly toxic diester diterpenoid alkaloids (e.g., aconitine) derived from *Aconiti Radix Lateralis Preparata* (Fu Zi) were excluded, because during the standard decoction process of ZWHQD, these compounds undergo substantial hydrolysis, converting into less toxic monoester alkaloids (e.g., benzoylaconine) and further into non‑toxic amine alcohol alkaloids (e.g., aconine), resulting in negligible or undetectable levels in the final decoction. [@gao2022Aconitine; @huang2007Aconitine] (iii) ubiquitous endogenous substances, including common saccharides, amino acids, organic acids and their metabolites, as well as other nutritional components, were excluded because they lack target‑specific pharmacological functions and contribute minimally to the therapeutic properties of ZWHQD.

Given the notable pharmacological bioactivity of compounds from *Astragalus Membranaceus*, we manually supplemented cycloastragenol, an in vivo bioactive metabolite of astragaloside IV with well-ocumented effects [@zhao2025cycloastragenol], into the active compound dataset. After completing systematic filtration and manual supplementation, we defined the remaining compounds as the final candidate active ingredients of ZWHQD.

Finally, to characterize the molecular properties of these candidate active ingredients, we queried each qualified compound in [PubChem](https://pubchem.ncbi.nlm.nih.gov/) [@kim2025pubchem].  For each compound, we extracted core chemical data, including IUPAC names, molecular formulas, molecular weights, SMILES strings, InChI, and InChIKey. Additionally, we downloaded the corresponding SDF files. The SMILES strings (or SDF files) were used for target prediction, and the SDF files also provided the essential 3D conformers for subsequent molecular docking.

---

### Loading Packages
```{r}
library(tidyverse)
library(data.table)
library(openxlsx)
```

### Step 1: Loading Raw Data from DCABM-TCM
```{r dcabm}
rm(list = ls())

compounds_raw <- read.xlsx("../data/raw/tcm/zwhqd_compound_dcabm.xlsx")

save(
  compounds_raw,
  file = "../data/processed/tcm/01_compounds_raw.RData"
)
```

---

### Step 2: Raw Data Preprocessing
Parse compound names, CIDs, formulas, and clean formatting.

```{r}
source("utils.R")

compounds_clean <- compounds_raw %>%
  mutate(across(where(is.character), trimws)) %>%
  separate_rows(Compounds, sep = "\\|") %>%
  mutate(
    CID = str_extract(Compounds, "(?<=Pubchem CID:)[^)]+") %>% str_trim(),
    DCABM.ID = str_extract(Compounds, "(?<=DCABM ID:)[^)]+") %>% str_trim(),
    temp = str_remove(Compounds, "\\s+\\(Pubchem CID:[^)]+\\)")
  ) %>%
  mutate(
    temp2 = str_remove(temp, "\\([^)]+\\)$") %>% str_trim(),
    Compound = str_remove(temp2, "\\([^)]+\\)$") %>% str_trim(),
    Formula = str_match(temp2, "\\(([^)]+)\\)$")[, 2] %>% str_trim()
  ) %>%
  dplyr::select(-Compounds, -temp, -temp2, -DCABM.ID) %>%
  mutate(across(where(is.character), trimws))

# Save processed data
save(
  compounds_clean,
  file = "../data/processed/tcm/02_compounds_clean.RData"
)
```

---

### Step 3: Manually Add Cycloastragenol
```{r}
compounds_add <- compounds_clean %>%
  add_row(
    Herb.Name.Chinese = "黄芪",
    Herb.Name.Pinyin = "HUANG QI",
    Herb.Name.English = "Astragali Radix",
    Herb.Name.Latin = "Astragalus membranaceus",
    Type = "Metabolite",
    CID = "13943286",
    Compound = "cycloastragenol",
    Formula = "C30H50O5"
  ) %>%
  arrange(Herb.Name.Pinyin, CID)

save(
  compounds_add,
  file = "../data/processed/tcm/03_compounds_add_cycloastragenol.RData"
)
```

---

### Step 4: Filter Invalid & Duplicate Compounds
```{r}
compounds_filtered <- compounds_add %>%
  filter(CID != "-", !is.na(CID), CID != "") %>%
  mutate(CID = clean_cids(CID)) %>%
  distinct(Herb.Name.Chinese, CID, .keep_all = TRUE) %>%
  distinct(Herb.Name.Chinese, Compound, .keep_all = TRUE)

save(
  compounds_filtered,
  file = "../data/processed/tcm/04_compounds_filtered.RData"
)
```

---

### Step 5: Load Manually Curated Properties & Final Filter

**Input**: Manual annotated compounds file `compounds_properties_manual.xlsx`

```{r}
compounds_manual <- read.xlsx("../data/raw/tcm/compounds_properties_manual.xlsx")

compounds_final_filtered <- compounds_manual %>%
  dplyr::filter(Drop == "No") %>%
  dplyr::select(
    -c(Note, Compound.Toxic, Drop)
  ) %>%
  distinct(Herb.Name.Pinyin, CID, .keep_all = TRUE)

save(
  compounds_final_filtered,
  file = "../data/processed/tcm/05_compounds_final_filtered.RData"
)
```

---

### Step 6: Batch Retrieve PubChem Annotations
```{r}
library(future)
library(furrr)
library(progressr)

compounds_pubchem <- compounds_final_filtered %>%
  mutate(`CID` = clean_cids(CID)) %>%
  filter(!is.na(CID)) %>%
  distinct(CID, .keep_all = TRUE)

properties <- c("IUPACName", "MolecularFormula", "MolecularWeight",
                "SMILES")

unique_cids <- unique(compounds_pubchem$CID)
cid_batches <- split(unique_cids, ceiling(seq_along(unique_cids)/20))

plan(multisession, workers = availableCores() - 2)
with_progress({
  p <- progressor(steps = length(cid_batches))
  chem_info <- future_map_dfr(cid_batches, function(b) {
    p()
    fetch_batch(properties, b)
  })
})
plan(sequential)


# Merge & Finalize
compounds_final <- compounds_pubchem %>%
  left_join(chem_info, by = "CID") %>%
  select(
    # Herb.Name.Chinese, 
    Herb.Name.Pinyin, 
    # Herb.Name.English,
    Compound.Name, IUPACName, Compound.Category,
    MolecularFormula, MolecularWeight, CID, SMILES
  ) %>%
  mutate(across(where(is.character), trimws))
```

---

### Step 7: Compound Sharing Statistics (Across Herbs)
```{r}
compound_sharing <- compounds_final %>%
  distinct(Herb.Name.Pinyin, CID) %>%
  group_by(CID) %>%
  summarise(
    Herb_Count = n(),
    Herbs = paste(unique(Herb.Name.Pinyin), collapse = ", ")
  ) %>%
  filter(Herb_Count >= 2) %>%
  arrange(desc(Herb_Count))

# Save all final results
save(
  compounds_final, compound_sharing,
  file = "../data/processed/tcm/06_compounds_final.RData"
)
```

---

### Step 8: Download SDF files
```{r}
library(httr)

unique_cids <- unique(compounds_pubchem$CID)
sdf_files <- "../data/raw/tcm/sdf/"

results <- download_all_sdf(unique_cids, sdf_files)
```

---

## Targets Prediction for ZWHQD

we predicted potential targets of the blood-absorbed compounds using seven authoritative online platforms, including three classic mainstream databases and four additional professional target fishing servers: **BATMAN-TCM 2.0** @kong2023batman, **SuperPred 3.0** @gallo2022superpred, **SwissTargetPrediction** @daina2019swiss, **PharmMapper** @liu2010pharm,@Wang2017pharm,  **TargetNet** @yao2016targetnet, **PPB3** @Darsaraee2026ppb3, and **SEA** @keiser2007sea. 

Subsequently, we retained high‑confidence target candidates using the following thresholds: score ≥0.84 for BATMAN‑TCM 2.0; probability ≥70% and accuracy ≥90% for SuperPred; probability ≥0.1 for SwissTargetPrediction; z‑score >0 and Norm Fit >0.9 for PharmMapper; probability ≥0.5 for TargetNet; single human protein for PPB3; and P Value < 1e<sup>-10</sup>, Z.Score > 5 and  Max.Tc >0.5 for SEA, all predicted targets restricted to human origin only.

To ensure analytical uniformity and interoperability, we then mapped and standardized all retained candidates to official human gene symbols using the UniProt knowledgebase (https://www.uniprot.org/). We confirmed non‑standard names or aliases, excluded targets that could not be standardized, and removed redundant entries.

Finally, we refined the target list through multi‑tool voting combined with literature evidence: we preferentially retained targets appearing in at least two tools, and we also considered single‑tool high‑confidence targets when supported by published literature.

### Step 1: BATMAN-TCM Targets

We retrieved target predictions for each blood‑absorbed compound from the BATMAN‑TCM 2.0 @kong2023batman using the compound PubChem CID strings as input.

**output**:

- batman_target_known `../data/raw/tcm/batman/batman_target_known.RData` 

- batman_target_pred `../data/raw/tcm/batman/batman_target_pred.RData`

```{r}
rm(list = ls())

source("utils.R")

load("../data/processed/tcm/06_compounds_final.RData")
load("../data/raw/tcm/batman/batman_target_known.RData")
load("../data/raw/tcm/batman/batman_target_pred.RData")

# Known Targets
compound_target_known_batman <- compounds_final %>%
  left_join(batman_target_known, by = "CID", relationship = "many-to-many") %>%
  select(-ends_with(".x")) %>%
  rename_with(~ str_replace(., "\\.y$", ""), ends_with(".y")) %>%
  distinct(Herb.Name.Pinyin, CID, Uniprot.ID, .keep_all = TRUE) %>%
  filter(!is.na(Symbol))

# Predicted Targets (Score ≥ 0.84)
compound_target_pred_batman <- compounds_final %>%
  left_join(batman_target_pred, by = "CID", relationship = "many-to-many") %>%
  select(-ends_with(".x")) %>%
  rename_with(~ str_replace(., "\\.y$", ""), ends_with(".y")) %>%
  distinct(Herb.Name.Pinyin, CID, Uniprot.ID, .keep_all = TRUE) %>%
  filter(!is.na(Symbol), Score >= 0.84)

save(
  compound_target_known_batman,
  compound_target_pred_batman,
  file = "../data/processed/tcm/07_batman_targets.RData"
)
```

### Step 2: Super-Pred Targets
```{r}
rm(list = ls())

source("utils.R")

load("../data/processed/tcm/06_compounds_final.RData")
load("../data/raw/uniprot/uniprot_human_reviewed.RData")

# Known targets
known <- read_compound_target("../data/raw/tcm/super/known") %>%
  select(CID, Protein.name = `Target Name`, Uniprot.ID = `UniProt ID`)

# Predicted targets (Prob > 70, Model Accuracy > 90)
pred <- read_compound_target("../data/raw/tcm/super") %>%
  filter(Probability >= 70, `Model accuracy` >= 90) %>%
  select(CID, Protein.name = `Target Name`, Uniprot.ID = `UniProt ID`)

# Merge & Annotate
process_super <- function(df) {
  df %>%
    left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
    filter(!is.na(Gene.Names.primary)) %>%
    mutate(Symbol = Gene.Names.primary) %>%
    distinct(CID, Uniprot.ID, .keep_all = TRUE)
}

known_clean <- process_super(known)
pred_clean <- process_super(pred)

# Merge with compounds
compound_target_known_super <- compounds_final %>%
  left_join(known_clean, by = "CID", relationship = "many-to-many") %>%
  distinct(Herb.Name.Pinyin, CID, Uniprot.ID, .keep_all = TRUE) %>%
  filter(!is.na(Symbol))

compound_target_pred_super <- compounds_final %>%
  left_join(pred_clean, by = "CID", relationship = "many-to-many") %>%
  distinct(Herb.Name.Pinyin, CID, Uniprot.ID, .keep_all = TRUE) %>%
  filter(!is.na(Symbol))

# Save
save(
  compound_target_known_super,
  compound_target_pred_super,
  file = "../data/processed/tcm/08_super_targets.RData"
)
```

### Step 3: SwissTargetPrediction Targets
```{r}
rm(list = ls())

source("utils.R")

load("../data/processed/tcm/06_compounds_final.RData")
load("../data/raw/uniprot/uniprot_human_reviewed.RData")

result <- read_compound_target("../data/raw/tcm/swiss") %>%
  filter(`Probability*` > 0.10) %>%
  separate_rows(`Common name`, `Uniprot ID`, sep = " ") %>%
  mutate(
    CID = clean_cids(CID),
    Uniprot.ID = gsub("[[:space:]]", "", `Uniprot ID`),
    Symbol = gsub("[[:space:]]", "", `Common name`)
  ) %>%
  select(CID, Uniprot.ID) %>%
  distinct(CID, Uniprot.ID, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  filter(!is.na(Gene.Names.primary))

compound_target_pred_swiss <- compounds_final %>%
  left_join(result, by = "CID", relationship = "many-to-many") %>%
  filter(!is.na(Gene.Names.primary)) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  distinct(Herb.Name.Pinyin, CID, Symbol, .keep_all = TRUE)

# Save
save(
  compound_target_pred_swiss,
  file = "../data/processed/tcm/09_swiss_targets.RData"
)
```

---

### Step 4: PharmMapper Targets
```{r}
rm(list = ls())

source("utils.R")

load("../data/processed/tcm/06_compounds_final.RData")
load("../data/raw/uniprot/uniprot_human_reviewed.RData")

result <- read_compound_target2("../data/raw/tcm/pharm",skip_row = 1) 

result <- result %>%
  dplyr::filter(grepl("_HUMAN", Uniplot),zscore>0, `Norm Fit`>0.9) %>%
  dplyr::select(CID, Uniplot) %>%
  distinct(CID, Uniplot, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Uniplot" = "Entry.Name")) %>%
  filter(!is.na(Gene.Names.primary))

compound_target_pred_pharm <- compounds_final %>%
  left_join(result, by = "CID", relationship = "many-to-many") %>%
  filter(!is.na(Gene.Names.primary)) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  distinct(Herb.Name.Pinyin, CID, Symbol, .keep_all = TRUE) %>%
  dplyr::rename(Uniprot.ID = Entry)

# save
save(
  compound_target_pred_pharm,
  file = "../data/processed/tcm/10_pharm_targets.RData"
)
```

---

### Step 5: TtargetNet Targets
```{r}
rm(list = ls())

source("utils.R")

load("../data/processed/tcm/06_compounds_final.RData")
load("../data/raw/uniprot/uniprot_human_reviewed.RData")


query <- read.xlsx("../data/raw/tcm/targetnet/query.xlsx")

cids<- query$CID
col_names <- c("Uniprot.ID","Protein",cids)

result <- read.xlsx("../data/raw/tcm/targetnet/results.xlsx") 
result <- result %>%
  dplyr::select(-Details) %>%
  setNames(col_names) %>%
  pivot_longer(
    cols = -c(Uniprot.ID, Protein),   
    names_to = "CID",                 
    values_to = "Probability"     
  ) %>%
  dplyr::filter(!is.na(Probability), Probability > 0.5) %>%
  mutate(CID = as.numeric(CID)) %>%
  arrange(CID,Probability) %>% 
  mutate(CID = as.character(CID)) %>%
  dplyr::select(CID,Uniprot.ID,Probability) %>% 
  distinct(CID, Uniprot.ID, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  filter(!is.na(Gene.Names.primary))

compound_target_pred_targetnet <- compounds_final %>%
  left_join(result, by = "CID", relationship = "many-to-many") %>%
  filter(!is.na(Gene.Names.primary)) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  distinct(Herb.Name.Pinyin, CID, Symbol, .keep_all = TRUE)

# Save
save(
  compound_target_pred_targetnet,
  file = "../data/processed/tcm/11_targetnet_targets.RData"
)
```

---

### Step 6: PPB3 Targets
```{r}
rm(list = ls())

source("utils.R")

load("../data/processed/tcm/06_compounds_final.RData")
load("../data/raw/uniprot/uniprot_human_reviewed.RData")

result <- read_compound_target("../data/raw/tcm/ppb3") 

result <- result %>%
  dplyr::filter( `Target Type` == 'SINGLE PROTEIN', `Target Organism` == "Homo sapiens")


library(httr)
library(jsonlite)

chembl_ids <- unique(result$`Target ID`)

uniprot_list <- sapply(chembl_ids, chembl_to_uniprot)

save(uniprot_list,file = "../data/raw/tcm/ppb3/uniprot_list.RData")

# load("../data/raw/tcm/ppb3/uniprot_list.RData")

df <- data.frame(
  Chembl.ID = names(uniprot_list),
  Uniprot.ID = sapply(uniprot_list, `[`, 1),  
  row.names = NULL
)

result <- result %>%
  left_join(df, by=c("Target ID" = "Chembl.ID")) %>% 
  distinct(CID, Uniprot.ID, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  filter(!is.na(Gene.Names.primary)) 

compound_target_pred_ppb3 <- compounds_final %>%
  left_join(result, by = "CID", relationship = "many-to-many") %>%
  filter(!is.na(Gene.Names.primary)) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  distinct(Herb.Name.Pinyin, CID, Symbol, .keep_all = TRUE) %>%
  dplyr::select(-c(Rank,`Nearest Neighbors`))

# Save
save(
  compound_target_pred_ppb3,
  file = "../data/processed/tcm/12_ppb3_targets.RData"
)
```

---

### Step 7: SEA Targets
```{r}
rm(list = ls())

source("utils.R")

load("../data/processed/tcm/06_compounds_final.RData")
load("../data/raw/uniprot/uniprot_human_reviewed.RData")

result <- read.csv("../data/raw/tcm/sea/sea-results.xls") %>%
  mutate(CID = as.numeric(gsub("compound","",Query.ID)), .before = 1) %>%
  arrange(CID) %>%
  mutate(CID = as.character(CID)) %>%
  dplyr::select(-Query.ID) %>%
  dplyr::filter(
    grepl("_HUMAN", Target.ID), 
    P.Value < 1e-10, 
    Max.Tc > 0.5, 
    Z.Score > 5
  ) %>%
  distinct(CID, Target.ID, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Target.ID" = "Entry.Name")) %>%
  filter(!is.na(Gene.Names.primary))

compound_target_pred_sea <- compounds_final %>%
  left_join(result, by = "CID", relationship = "many-to-many") %>%
  filter(!is.na(Gene.Names.primary)) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  distinct(Herb.Name.Pinyin, CID, Symbol, .keep_all = TRUE) %>% 
  dplyr::rename(Uniprot.ID = Entry)

# save
save(
  compound_target_pred_sea,
  file = "../data/processed/tcm/13_sea_targets.RData"
)
```

---

### Step 8: Merge All Targets & Deduplicate
```{r}
rm(list = ls())

# Common Columns
cnames <- c(
  # "Herb.Name.Chinese", 
  "Herb.Name.Pinyin", 
  # "Herb.Name.English", 
  # "Herb.Name.Latin",
  "Compound.Name", "IUPACName", "Compound.Category",
  "MolecularFormula", "MolecularWeight", "CID",
  "SMILES", "Symbol", "Uniprot.ID"
)


load("../data/processed/tcm/07_batman_targets.RData")

load("../data/processed/tcm/08_super_targets.RData")

load("../data/processed/tcm/09_swiss_targets.RData")

load("../data/processed/tcm/10_pharm_targets.RData")

load("../data/processed/tcm/11_targetnet_targets.RData")

load("../data/processed/tcm/12_ppb3_targets.RData")

load("../data/processed/tcm/13_sea_targets.RData")

# Known Targets
compound_target_known <- rbind(
  compound_target_known_batman[, cnames],
  compound_target_known_super[, cnames]
) %>%
  distinct(Herb.Name.Pinyin, CID, Symbol, .keep_all = TRUE) %>%
  filter(!is.na(Symbol))

# Predicted Targets
pred_batman <- compound_target_pred_batman[, cnames] %>% mutate(database = "BATMAN")

pred_super <- compound_target_pred_super[, cnames] %>% mutate(database = "SuperPred")

pred_swiss <- compound_target_pred_swiss[, cnames] %>% mutate(database = "Swiss")

pred_pharm <- compound_target_pred_pharm[, cnames] %>% mutate(database = "PharmMapper")

pred_targetnet <- compound_target_pred_targetnet[, cnames] %>% mutate(database = "TargetNet")

pred_ppb3 <- compound_target_pred_ppb3[, cnames] %>% mutate(database = "PPB3")

pred_sea <- compound_target_pred_swiss[, cnames] %>% mutate(database = "SEA")


# all targets
compound_target_pred_all <- rbind(
  pred_batman, 
  pred_super, 
  pred_swiss,
  pred_pharm,
  pred_targetnet,
  pred_ppb3,
  pred_sea
) %>%
  distinct(Herb.Name.Pinyin, CID, Symbol, database, .keep_all = TRUE) %>%
  filter(!is.na(Symbol))

# counts
compound_target_pred_count <- compound_target_pred_all %>%
  group_by(Herb.Name.Pinyin, CID, Symbol) %>%
  summarise(
    database_count = n_distinct(database),
    Uniprot.ID = first(Uniprot.ID),
    across(-c(database, Uniprot.ID), first),
    .groups = "drop"
  )

# retain targets predicted by ≥2 databases
compound_target_pred_filtered <- compound_target_pred_count %>%
  dplyr::filter(database_count >= 2) %>%
  dplyr::select(all_of(cnames)) %>%
  distinct()

# Final predicted target
compound_target_pred <- compound_target_pred_filtered

load("../data/raw/uniprot/uniprot_human_reviewed.RData")

# All Targets
compound_target_final <- rbind(compound_target_known, compound_target_pred) %>%
  distinct(Herb.Name.Pinyin, CID, Symbol, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::select(-Reviewed, -Entry.Name)

save(
  compound_target_final,
  file = "../data/processed/tcm/14_compound_target_final.RData"
)

## Export Final Supplementary Table
compound_target_export <- compound_target_final %>%
  dplyr::select(-Sequence)
```

---

### Step 9: Targets Statistics
```{r}
# Create summary table with statistics for each database and final results
summary_table <- tibble(
  # Database name (Known / Predicted)
  Database = c(
    "BATMAN-TCM (Known)",
    "SuperPred (Known)",
    "BATMAN-TCM (Predicted)",
    "SuperPred (Predicted)",
    "SwissTargetPrediction",
    "PharmMapper",
    "TargetNet",
    "PPB3",
    "SEA",
    "Final Integrated Results"
  ),
  
  # Number of unique compounds with targets (CID)
  'Number of compounds with targets' = c(
    n_distinct(compound_target_known_batman$CID),
    n_distinct(compound_target_known_super$CID),
    n_distinct(compound_target_pred_batman$CID),
    n_distinct(compound_target_pred_super$CID),
    n_distinct(compound_target_pred_swiss$CID),
    n_distinct(compound_target_pred_pharm$CID),
    n_distinct(compound_target_pred_targetnet$CID),
    n_distinct(compound_target_pred_ppb3$CID),
    n_distinct(compound_target_pred_sea$CID),
    n_distinct(compound_target_final$CID)
  ),
  
  # Number of unique targets based on UniProt ID
  `Targets Count` = c(
    n_distinct(compound_target_known_batman$Uniprot.ID),
    n_distinct(compound_target_known_super$Uniprot.ID),
    n_distinct(compound_target_pred_batman$Uniprot.ID),
    n_distinct(compound_target_pred_super$Uniprot.ID),
    n_distinct(compound_target_pred_swiss$Uniprot.ID),
    n_distinct(compound_target_pred_pharm$Uniprot.ID),
    n_distinct(compound_target_pred_targetnet$Uniprot.ID),
    n_distinct(compound_target_pred_ppb3$Uniprot.ID),
    n_distinct(compound_target_pred_sea$Uniprot.ID),
    n_distinct(compound_target_final$Uniprot.ID)
  )
)

# Export table to Excel for manuscript supplementary materials

save(summary_table, file =  "../data/processed/tcm/15_target_statistics.RData")

write.xlsx(
  list(
    BATMAN_Known_Tagets = compound_target_known_batman,
    BATMAN_Pred_Tagets = compound_target_pred_batman,
    Super_Known_Tagets = compound_target_known_super,
    Super_Pred_Tagets = compound_target_pred_super,
    Swiss_Pred_Tagets = compound_target_pred_swiss,
    Pharm_Pred_Tagets = compound_target_pred_pharm,
    TargetNet_Pred_Tagets = compound_target_pred_targetnet,
    PPB3_Pred_Tagets = compound_target_pred_ppb3,
    SEA_Pred_Tagets = compound_target_pred_sea,
    Final_Known = compound_target_known,
    Final_Predicted = compound_target_pred,
    Final_All = compound_target_export,
    Target_Statistics = summary_table
  ),
  "../tables/Table_S2.xlsx"
)
```

```{r}
add_platform <- function(df, platform_name) {
  df %>%
    mutate(
      Prediction_platform = platform_name,
      Target_gene_name = as.character(!!sym("Symbol")), 
      UniProt_ID = as.character(!!sym("Uniprot.ID"))      
    ) %>%
    select(Target_gene_name, UniProt_ID, Prediction_platform)
}

platform_list <- list(
  BATMAN_Known_Tagets = compound_target_known_batman,
  BATMAN_Pred_Tagets  = compound_target_pred_batman,
  Super_Known_Tagets  = compound_target_known_super,
  Super_Pred_Tagets   = compound_target_pred_super,
  Swiss_Pred_Tagets   = compound_target_pred_swiss,
  Pharm_Pred_Tagets   = compound_target_pred_pharm,
  TargetNet_Pred_Tagets = compound_target_pred_targetnet,
  PPB3_Pred_Tagets    = compound_target_pred_ppb3,
  SEA_Pred_Tagets     = compound_target_pred_sea
)

all_platforms <- bind_rows(
  lapply(names(platform_list), function(name) {
    df <- platform_list[[name]]
    if(is.null(df) || nrow(df) == 0) return(NULL)
    add_platform(df, name)   
  })
)

final_intersection <- compound_target_known %>% 
  distinct(Symbol, Uniprot.ID) %>%
  mutate(In_intersection = "Yes")

table_S2_grouped <- all_platforms %>%
  group_by(Target_gene_name, UniProt_ID) %>%
  summarise(
    Prediction_platform = paste(unique(Prediction_platform), collapse = "; "),
    .groups = "drop"
  )

table_S2 <- table_S2_grouped %>%
  left_join(final_intersection, by = c( "Target_gene_name" ="Symbol" , "UniProt_ID" = "Uniprot.ID" )) %>%
  mutate(`Included in intersection` = ifelse(is.na(In_intersection), "No", "Yes")) %>%
  select(-In_intersection)

write.xlsx(
  list(`Table S2` = table_S2),
  "../tables/Table_S2.xlsx"
)
```

---

# Acquisition of DCM_HFrEF-Related Gene Expression Data
We integrated multiple transcriptome datasets from the GEO database and systematically carried out the screening and validation of target genes related to heart failure with reduced ejection fraction (DCM_HFrEF). [GSE141910]() was selected as data sources for the bulk RNA-seq expression spectrum of DCM_HFrEF in this study, and [GSE116250](), [GSE161472](), [GSE135055](), [GSE120895](), and [GSE19303]() were selected as a validation set to perform the diagnostic value assessment with the sample heart tissues. [GSE183852]() was selected for subsequent 10X single-cell RNA sequencing (scRNA-seq) analysis. ([Supplementary Table S3]()).


**Notes:**
All these data have been downloaded and stored in the "../data/rwa/geo". Please run the following code with caution, as it is time-consuming due to the large volume of data, especially the single-cell transcriptome dataset.

## Load Packages
```{r}
library(tidyverse)
library(data.table)
library(openxlsx)
library(patchwork)
library(ggrepel)
library(vegan)
library(GEOquery)
library(limma)
library(DESeq2)
library(edgeR)
library(sva)
library(org.Hs.eg.db)
if (!require("TxDb.Hsapiens.UCSC.hg38.knownGene")) {
  BiocManager::install("TxDb.Hsapiens.UCSC.hg38.knownGene")
}
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
```

## Step 1: RNA‑seq Datasets
### GSE141910: Main training set
```{r,warning=FALSE}
rm(list=ls())

raw_dir <- "../data/raw/geo/GSE141910"
if (!dir.exists(raw_dir)) dir.create(raw_dir, recursive = TRUE)

Sys.setlocale("LC_ALL", "C")
gse141910 <- getGEO("GSE141910", destdir = raw_dir, getGPL = TRUE, AnnotGPL = F)
Sys.setlocale("LC_ALL", "")

gse141910_pheno_raw <- pData(gse141910[[1]])

# url <- "https://www.dropbox.com/scl/fi/2leazyjscrvn3ahderxt9/phenoData.csv?rlkey=vo0gpy1q503ywz18mro0d5nai&e=1&dl=1"
# gse141910_pheno_raw_with_lvef <- read_csv(url)
gse141910_pheno_raw_with_lvef <- read.csv("../data/raw/geo/GSE141910/gse141910_pheno_raw_with_lvef.csv")

url <- "https://www.ncbi.nlm.nih.gov/geo/download/?type=rnaseq_counts&acc=GSE141910&format=file&file=GSE141910_raw_counts_GRCh38.p13_NCBI.tsv.gz"
gse141910_counts_raw <- read_tsv(url)

url <- "https://www.ncbi.nlm.nih.gov/geo/download/?type=rnaseq_counts&acc=GSE141910&format=file&file=GSE141910_norm_counts_FPKM_GRCh38.p13_NCBI.tsv.gz"
gse141910_fpkm_raw <- read_tsv(url)

url <- "https://www.ncbi.nlm.nih.gov/geo/download/?type=rnaseq_counts&acc=GSE141910&format=file&file=GSE141910_norm_counts_TPM_GRCh38.p13_NCBI.tsv.gz"
gse141910_tpm_raw <- read_tsv(url)

save(
  gse141910_pheno_raw,
  gse141910_pheno_raw_with_lvef,
  gse141910_counts_raw,
  gse141910_fpkm_raw,
  gse141910_tpm_raw,
  file = "../data/raw/geo/gse141910_raw.RData"
)
```

### GSE116250: External validation set
```{r,warning=FALSE}
rm(list=ls())

raw_dir <- "../data/raw/geo/GSE116250"
if (!dir.exists(raw_dir)) dir.create(raw_dir, recursive = TRUE)

Sys.setlocale("LC_ALL", "C")
gse116250 <- getGEO("GSE116250", destdir = raw_dir, getGPL = TRUE, AnnotGPL = F)
Sys.setlocale("LC_ALL", "")

gse116250_pheno_raw <- pData(gse116250[[1]])

url <- "https://www.ncbi.nlm.nih.gov/geo/download/?type=rnaseq_counts&acc=GSE116250&format=file&file=GSE116250_raw_counts_GRCh38.p13_NCBI.tsv.gz"
gse116250_counts_raw <- read_tsv(url)

url <- "https://www.ncbi.nlm.nih.gov/geo/download/?type=rnaseq_counts&acc=GSE116250&format=file&file=GSE116250_norm_counts_FPKM_GRCh38.p13_NCBI.tsv.gz"
gse116250_fpkm_raw <- read_tsv(url)

url <- "https://www.ncbi.nlm.nih.gov/geo/download/?type=rnaseq_counts&acc=GSE116250&format=file&file=GSE116250_norm_counts_TPM_GRCh38.p13_NCBI.tsv.gz"
gse116250_tpm_raw <- read_tsv(url)

save(
  gse116250_pheno_raw,
  gse116250_counts_raw,
  gse116250_fpkm_raw,
  gse116250_tpm_raw,
  file = "../data/raw/geo/gse116250_raw.RData"
)
```

### GSE161472: External validation set
```{r,warning=FALSE}
rm(list=ls())

raw_dir <- "../data/raw/geo/GSE161472"
if (!dir.exists(raw_dir)) dir.create(raw_dir, recursive = TRUE)

Sys.setlocale("LC_ALL", "C")
gse161472 <- getGEO("GSE161472", destdir = raw_dir, getGPL = TRUE, AnnotGPL = F)
Sys.setlocale("LC_ALL", "")

gse161472_pheno_raw <- pData(gse161472[[1]])

url <- "https://www.ncbi.nlm.nih.gov/geo/download/?type=rnaseq_counts&acc=GSE161472&format=file&file=GSE161472_raw_counts_GRCh38.p13_NCBI.tsv.gz"
gse161472_counts_raw <- read_tsv(url)

url <- "https://www.ncbi.nlm.nih.gov/geo/download/?type=rnaseq_counts&acc=GSE161472&format=file&file=GSE161472_norm_counts_FPKM_GRCh38.p13_NCBI.tsv.gz"
gse161472_fpkm_raw <- read_tsv(url)

url <- "https://www.ncbi.nlm.nih.gov/geo/download/?type=rnaseq_counts&acc=GSE161472&format=file&file=GSE161472_norm_counts_TPM_GRCh38.p13_NCBI.tsv.gz"
gse161472_tpm_raw <- read_tsv(url)

save(
  gse161472_pheno_raw,
  gse161472_counts_raw,
  gse161472_fpkm_raw,
  gse161472_tpm_raw,
  file = "../data/raw/geo/gse161472_raw.RData"
)
```

### GSE135055: External validation set
```{r,warning=FALSE}
rm(list=ls())

raw_dir <- "../data/raw/geo/GSE135055"
if (!dir.exists(raw_dir)) dir.create(raw_dir, recursive = TRUE)

Sys.setlocale("LC_ALL", "C")
gse135055 <- getGEO("GSE135055", destdir = raw_dir, getGPL = TRUE, AnnotGPL = F)
Sys.setlocale("LC_ALL", "")

gse135055_pheno_raw <- pData(gse135055[[1]])

url <- "https://www.ncbi.nlm.nih.gov/geo/download/?type=rnaseq_counts&acc=GSE135055&format=file&file=GSE135055_raw_counts_GRCh38.p13_NCBI.tsv.gz"
gse135055_counts_raw <- read_tsv(url)

url <- "https://www.ncbi.nlm.nih.gov/geo/download/?type=rnaseq_counts&acc=GSE135055&format=file&file=GSE135055_norm_counts_FPKM_GRCh38.p13_NCBI.tsv.gz"
gse135055_fpkm_raw <- read_tsv(url)

url <- "https://www.ncbi.nlm.nih.gov/geo/download/?type=rnaseq_counts&acc=GSE135055&format=file&file=GSE135055_norm_counts_TPM_GRCh38.p13_NCBI.tsv.gz"
gse135055_tpm_raw <- read_tsv(url)

save(
  gse135055_pheno_raw,
  gse135055_counts_raw,
  gse135055_fpkm_raw,
  gse135055_tpm_raw,
  file = "../data/raw/geo/gse135055_raw.RData"
)
```

## Step 2: Microarray Datasets
### GSE120895: External validation set 
```{r,warning=FALSE}
rm(list=ls())

raw_dir <- "../data/raw/geo/GSE120895"
if (!dir.exists(raw_dir)) dir.create(raw_dir, recursive = TRUE)

Sys.setlocale("LC_ALL", "C")
gse120895 <- getGEO("GSE120895", destdir = raw_dir, getGPL = TRUE, AnnotGPL = F)
Sys.setlocale("LC_ALL", "")

gse120895_pheno_raw <- pData(gse120895[[1]])
gse120895_exprs_raw <- exprs(gse120895[[1]])

save(
  gse120895_pheno_raw,
  gse120895_exprs_raw,
  file = "../data/raw/geo/gse120895_raw.RData"
)
```

### GSE19303: External validation set 
```{r,warning=FALSE}
rm(list=ls())

raw_dir <- "../data/raw/geo/GSE19303"
if (!dir.exists(raw_dir)) dir.create(raw_dir, recursive = TRUE)

Sys.setlocale("LC_ALL", "C")
gse19303  <- getGEO("GSE19303",  destdir = raw_dir, getGPL = TRUE, AnnotGPL = F)
Sys.setlocale("LC_ALL", "")

gse19303_pheno_raw <- pData(gse19303[[1]])
gse19303_exprs_raw <- exprs(gse19303[[1]])

save(
  gse19303_pheno_raw,
  gse19303_exprs_raw,
  file = "../data/raw/geo/gse19303_raw.RData"
)
```

## Step 3: Single‑cell RNA‑seq Dataset

```{r}
rm(list=ls())

raw_dir <- "../data/raw/geo/GSE183852"
if (!dir.exists(raw_dir)) dir.create(raw_dir, recursive = TRUE)

Sys.setlocale("LC_ALL", "C")
gse183852  <- getGEO("GSE183852",  destdir = raw_dir, getGPL = TRUE, AnnotGPL = F)
Sys.setlocale("LC_ALL", "")

gse183852_pheno_raw <- pData(gse183852[[1]])
gse183852_gpl <- gse183852[[1]]@annotation

save(
  gse183852_gpl,
  gse183852_pheno_raw,
  file = "../data/raw/geo/gse183852_raw.RData"
)
```

**NOTE**: **Proceed with caution!** GSE183852_DCM_Integrated.Robj.gz is 11.7GB. Long download time.
Alternative: download directly via browser from https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE183852

```{r}
library(curl)

url <- "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE183852&format=file&file=GSE183852%5FIntegrated%5FCounts%2Ecsv%2Egz"
h <- new_handle()
curl_fetch_disk(url, path = "../data/raw/geo/GSE183852/GSE183852_Integrated_Counts.csv.gz",handle = h)

# 11.7 GB
url <- "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE183852&format=file&file=GSE183852%5FDCM%5FIntegrated%2ERobj%2Egz" 

h <- new_handle()
curl_fetch_disk(url, path = "../data/raw/geo/GSE183852/GSE183852_DCM_Integrated.Robj.gz",handle = h)

rm(list=ls())
gc()
```

---

# Preprocessing of the GSE141910 Dataset (for DEG & WGCNA & ML)

We corrected for batch effects on the normalized count matrix using the ComBat‑seq function from the sva package [@zhang2020SVA].


## Step 1: Loading Raw Counts & FPKM
```{r}
rm(list=ls())

# Loading raw data
load("../data/raw/geo/gse141910_raw.RData")

# Extracting expression matrix and phenotype
counts <- gse141910_counts_raw %>% 
  column_to_rownames("GeneID")
pheno <- gse141910_pheno_raw

# Adding LVEF information
pheno_lvef <- gse141910_pheno_raw_with_lvef
pheno <- pheno %>% 
  left_join(pheno_lvef, by = c("title" = "sample_name")) %>%
  column_to_rownames("geo_accession")
```

## Step 2: Filtering Normal / DCM_HFrEF Samples
```{r}
# Filtering Pheno Data of Good Samples 
pheno <- pheno[rownames(pheno) %in% colnames(counts),]

# Normal <- pheno$etiology == "NF" & !is.na(pheno$LVEF) & pheno$LVEF >= 0.5 & pheno$LVEF <= 0.7
Normal <- pheno$etiology == "NF" 
DCM_HFrEF  <- pheno$etiology == "DCM" & !is.na(pheno$LVEF) & pheno$LVEF <= 0.4

keep_samples <- rownames(pheno)[Normal | DCM_HFrEF]
pheno <- pheno[keep_samples, ]
counts_keep <- counts[, keep_samples]

pheno$group <- factor(ifelse(pheno$etiology == "NF", "Normal", "DCM_HFrEF"), levels = c("Normal","DCM_HFrEF"))
```

## Step 3: Gene Annotation(Probe ID converted to gene symbol) 
```{r}
# Obtaining gene length from TxDb
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
trans_len <- transcriptLengths(txdb, with.cds_len = TRUE)

gene_len_vec_txdb <- trans_len %>%
  filter(!is.na(tx_len)) %>%
  group_by(gene_id) %>%
  summarise(Length_bp = max(tx_len, na.rm = TRUE)) %>%
  ungroup() %>%
  pull(Length_bp, name = gene_id)   

# Convert Entrez IDs to official HGNC gene symbols
entrez <- rownames(counts_keep)

# Obtain EntrezID with length
has_len <- entrez %in% names(gene_len_vec_txdb)

# Checking the matching rate
cat("TxDb matching rate：", round(100 * sum(has_len) / nrow(counts_keep), 2), "%\n")

# Retain genes with length
entrez_filt <- entrez[has_len]

sym <- mapIds(
  org.Hs.eg.db,
  keys = entrez_filt,
  keytype = "ENTREZID",
  column = "SYMBOL",
  multiVals = "first"
)

# Create annotation data frame
gene_anno_filt <- data.frame(
  EntrezID = entrez_filt,
  Symbol = sym,
  row.names = entrez_filt,
  stringsAsFactors = FALSE
) %>% 
  drop_na(Symbol)


# Subset count matrix to keep only annotated genes
counts_anno_filt <- counts_keep[gene_anno_filt$EntrezID, ]
```

## Step 4: Rename the row name from EntrezID to Symbol
```{r}
counts_anno_mat <- as.matrix(counts_anno_filt)
rownames(counts_anno_mat) <- gene_anno_filt$Symbol[match(rownames(counts_anno_mat), gene_anno_filt$EntrezID)]
```

## Step 5: Processing duplicate genes (take the average)
```{r}
counts_dup_filt <- limma::avereps(counts_anno_mat) %>%
  round() %>% 
  as.data.frame()
```


## Step 6: Filter low expression genes based on CPM

```{r}
dge <- DGEList(counts_dup_filt)
cpms <- cpm(dge)
keep <- rowSums(cpms > 0) >= 0.8 * ncol(cpms)  
counts_low_filt <- counts_dup_filt[keep, ]
```

## Step 7: compute_fpkm_from_counts
```{r}
compute_fpkm_from_counts <- function(count_mat, gene_anno, gene_len_vec, lib_size_mil) {
  anno_len <- gene_anno %>%
    mutate(Length = gene_len_vec[as.character(EntrezID)]) %>%
    group_by(Symbol) %>%
    summarise(Length_bp = mean(Length, na.rm = TRUE))
  
  anno_len <- anno_len[match(rownames(count_mat), anno_len$Symbol), ]
  if(any(is.na(anno_len$Length_bp))) {
    warning("Some Symbols have no length info, set to median")
    anno_len$Length_bp[is.na(anno_len$Length_bp)] <- median(anno_len$Length_bp, na.rm = TRUE)
  }
  
  gene_len_kb <- anno_len$Length_bp / 1000
  names(gene_len_kb) <- anno_len$Symbol
  
  fpkm <- sweep(count_mat, 2, lib_size_mil, "/")
  fpkm <- sweep(fpkm, 1, gene_len_kb, "/")
  return(fpkm)
}


counts_final <- counts_low_filt[rownames(counts_low_filt) %in% gene_anno_filt$Symbol, ]

lib_size_mil_before_filter <- colSums(counts_dup_filt) / 1e6

# fpkm_final
fpkm_final <- compute_fpkm_from_counts(
  count_mat = counts_final,
  gene_anno = gene_anno_filt,
  gene_len_vec = gene_len_vec_txdb,
  lib_size_mil = lib_size_mil_before_filter[colnames(counts_final)]
) %>% as.data.frame()
```

## Step 8: SVA Batch Correction for Counts
```{r}
# log2 (CPM+1) matrix 
dge_norm <- calcNormFactors(DGEList(counts_final))
logcpm <- cpm(dge_norm, log = TRUE, prior.count = 1)   # log2(CPM+1)

# Design Matrix
mod <- model.matrix(~ group + age + race + gender, pheno)
mod0 <- model.matrix(~ 1, pheno)

# Estimate the number of potential factors
n.sv <- num.sv(as.matrix(logcpm), mod, method="leek")

if (n.sv > 0) {
  # SVA
  svobj <-sva(as.matrix(logcpm), mod, mod0, n.sv = n.sv)
  
  # Correct logCPM 
  expr_counts_corr <- removeBatchEffect(logcpm, covariates = svobj$sv, design = mod)
  
  # Correct logFPKM
  logfpkm <- log2(fpkm_final + 1)
  expr_fpkm_corr <- removeBatchEffect(logfpkm, covariates = svobj$sv, design = mod)
  
  # 将 SVA 因子添加到表型数据中
  svdf <- as.data.frame(svobj$sv)
  colnames(svdf) <- paste0("SVA", 1:ncol(svdf))
  pheno_final <- cbind(pheno, svdf)
} else {
  expr_counts_corr <- logcpm
  expr_fpkm_corr <- log2(fpkm_final + 1)
  pheno_final <- pheno
  warning("No significant surrogate variables found. Proceeding without batch correction.")
}

```

## Step 8: Saving data for DEG & WGCNA & Immune infiltration analysis & ML)
```{r}
proc_dir <- "../data/processed/geo/GSE141910"
if (!dir.exists(proc_dir)) dir.create(proc_dir, recursive = TRUE)

# Save
save(
  expr_counts_corr,
  expr_fpkm_corr,
  counts_final,
  pheno_final,
  fpkm_final,
  file = file.path(proc_dir, "GSE141910_preprocessed_final.RData")
)
```

## Step 9: PCA Before / After Correction
```{r}
load("../data/processed/geo/GSE141910/GSE141910_preprocessed_final.RData")
dds <- DESeqDataSetFromMatrix(
  countData = counts_final, 
  colData = pheno_final, 
  design = ~ group
)
vsd <- vst(dds, blind = FALSE)

# PCA plot BEFORE batch correction
pca_before <- plotPCA(vsd, intgroup = "group", returnData = TRUE)
var_before <- round(100 * attr(pca_before, "percentVar"))

p_before <- ggplot(pca_before, aes(x = PC1, y = PC2, color = group, fill = group)) +
  geom_point(size = 2.5, alpha = 0.7) +
  stat_ellipse(type = "t", level = 0.95, geom = "polygon", alpha = 0.15) +
  geom_text_repel(aes(label = rownames(pca_before)), size = 3, show.legend = FALSE) + 
  xlab(paste0("PC1: ", var_before[1], "% variance")) +
  ylab(paste0("PC2: ", var_before[2], "% variance")) +
  ggtitle("Before Batch Correction") +
  theme_bw(base_size = 10) +
  labs(tag = "A") + 
  theme(
    plot.tag.position = c(0.10, 0.90),
    plot.tag = element_text(size = 12, face = "bold"),
    plot.title = element_text(hjust = 0.5),
    legend.position = "none"
  )

# PCA plot AFTER batch correction
assay(vsd) <- expr_counts_corr

pca_after <- plotPCA(vsd, intgroup = "group", returnData = TRUE)
var_after <- round(100 * attr(pca_after, "percentVar"))

p_after <- ggplot(pca_after, aes(x = PC1, y = PC2, color = group, fill = group)) +
  geom_point(size = 2.5, alpha = 0.7) +
  stat_ellipse(type = "t", level = 0.95, geom = "polygon", alpha = 0.15) +
  geom_text_repel(aes(label = rownames(pca_after)), size = 3, show.legend = FALSE) + 
  xlab(paste0("PC1: ", var_after[1], "% variance")) +
  ylab(paste0("PC2: ", var_after[2], "% variance")) +
  ggtitle("After SVA Batch Correction") +
  theme_bw(base_size = 10) +
  labs(tag = "B") +
  theme(
    plot.tag.position = c(0.10, 0.90),
    plot.tag = element_text(size = 12, face = "bold"),
    plot.title = element_text(hjust = 0.5),
    legend.position = "right"
  )

# Combine plots with shared legend and add internal A/B tags
p_combined <- (p_before + p_after) +
  plot_annotation(tag_levels = "A") +
  theme(
    plot.tag.position = c(0.10, 0.90),
    plot.tag = element_text(size = 12, face = "bold")
  )

ggsave("Figure_S1_A.pdf", p_before, path = "../figures/supplementary", width=7, height=5)
ggsave("Figure_S1_B.pdf",  p_after,  path = "../figures/supplementary", width=7, height=5)


# Save combined PCA plot
ggsave(
  "Figure_S1.pdf", 
  plot = p_combined, 
  path = "../figures/supplementary", 
  width = 14, 
  height = 6, 
  dpi = 300
)

ggsave(
  filename = "Figure_S1.png",
  plot = p_combined,
  path = "../figures/supplementary",
  width = 14,
  height = 6,
  dpi = 300
)
```

---

<table class="nav-table" width="100%">
  <tr>
    <td align="left">
      [Home](../index.qmd) 
    </td>
    <td align="right">
      [Next](Part_2_Screening_of_HFrEF_related_genes.qmd)
    </td>
  </tr>
</table>

# References {-}