---
title: "Part 3: Mechanism exploration of ZWHQD against HFrEF"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

---

# Identification of Candidate Targets

## Load Packages
```{r}
library(ggvenn)
library(tidyverse)
library(openxlsx)
library(ggsci)
```

## Candidate Genes
```{r}
rm(list = ls())

load("../data/processed/geo/GSE141910/degs_results.RData")

deg_sig <- deg_all %>% filter(padj < 0.05, abs(log2FoldChange) > lfc)

deg_genes <- rownames(deg_sig) 

load("../data/processed/tcm/14_compound_target_final.RData")

herb_targets <- unique(compound_target_final$Symbol)

load("../data/processed/geo/GSE141910/wgcna_candidates_genes.RData")

# disease_genes: intersecting DEGs and WGCNA
disease_genes <- intersect(deg_genes,wgcna_candidates_genes$Gene)

# candidate_genes: intersecting herb_targets and disease_genes
candidate_genes <-intersect(herb_targets, disease_genes)

herb_targets_df <- data.frame(Gene = herb_targets) %>%
  arrange(Gene)

deg_genes_df <- data.frame(Gene = deg_genes) %>%
  arrange(Gene)

wgcna_genes_df <- data.frame(Gene = wgcna_candidates_genes$Gene) %>%
  arrange(Gene)

disease_genes_df <- data.frame(Gene = disease_genes) %>%
  arrange(Gene)

candidate_genes_df <- data.frame(Gene = candidate_genes) %>%
  arrange(Gene)

gene_list <- list(
  DEGs = deg_genes,
  `Herb targets` = herb_targets,
  `WGCNA Genes` = wgcna_candidates_genes$Gene
)

p <- ggvenn(
  gene_list,
  fill_color = pal_aaas()(4),
  stroke_color = "white",
  stroke_size = 1,
  set_name_size = 5,
  text_size = 5,
  text_color = "white",
  show_percentage = FALSE
) +
  theme(plot.title = element_text(hjust = 0.5, size=10))

# save
ggsave("../figures/main/Figure_3A.pdf", p, width=6, height=6, dpi=300)
ggsave("../figures/main/Figure_3A.png", p, width=6, height=6, dpi=300)


save(herb_targets_df, deg_genes_df, wgcna_genes_df,disease_genes_df, candidate_genes_df, file = "../data/processed/network/01_herb_disease_genes.RData")

write.xlsx(
  list(
    HerbTargets = herb_targets_df,
    DEGs = deg_genes_df,
    WGCNAGenes = wgcna_genes_df,
    DiseaseGenes = disease_genes_df,
    CandidateGenes = candidate_genes_df
  ),
  "../tables/candidate_genes.xlsx"
)

deg_candidate <- deg_all[rownames(deg_all) %in% candidate_genes, ]
 
deg_candidate$regulation <- ifelse(
  deg_candidate$log2FoldChange >= lfc, "Up",
  ifelse(deg_candidate$log2FoldChange <= -lfc, "Down", "NS")
)

deg_candidate$Gene <- rownames(deg_candidate)

up_genes <- deg_candidate$Gene[deg_candidate$regulation == "Up"]
down_genes <- deg_candidate$Gene[deg_candidate$regulation == "Down"]

cat("Up genes: ", length(up_genes), "\n")
print(up_genes)

cat("\nDown genes: ", length(down_genes), "\n")
print(down_genes)
```

---

# Herb-Ingredient-Target (HIT) Network
## Load Packages

```{r}
library(tidyverse)
library(openxlsx)
library(data.table)

library(clusterProfiler)
library(igraph)
library(centiserve)
library(RCy3)
```

## Step 1: Loading data
```{r}
rm(list = ls())

load("../data/processed/network/01_herb_disease_genes.RData")
load("../data/processed/tcm/14_compound_target_final.RData")

herb_ingredient_target <- candidate_genes_df %>%
  left_join(compound_target_final, by = c("Gene" = "Symbol"),relationship = "many-to-many") %>%
  dplyr::select(Herb.Name.Pinyin, Compound.Name, Gene) %>%
  dplyr:: rename(Herb = Herb.Name.Pinyin, Ingredient = Compound.Name, Target = Gene)  %>%
  distinct() %>%
  arrange(Herb,Target)

save(herb_ingredient_target,file = "../data/processed/network/02_herb_ingredient_target.RData")

```

## Step 2: Build herb-compound-target network
```{r}
herb_ingredient <- herb_ingredient_target %>%
  dplyr::select(Herb, Ingredient) %>%
  distinct()
  
ingredient_target <- herb_ingredient_target %>%
  dplyr::select(Ingredient, Target) %>%
  distinct()
  
# edges
edges <- bind_rows(
  herb_ingredient %>% 
    dplyr::rename(source = Herb, target = Ingredient) %>% 
    mutate(interaction = "herb_ingredient"),
  ingredient_target %>%
    dplyr::rename(source = Ingredient, target = Target) %>%
    mutate(interaction = "ingredient_target")
)

# nodes
nodes <- data.frame(id = unique(c(edges$source, edges$target))) %>%
  mutate(type = case_when(
    id %in% herb_ingredient$Herb        ~ "Herb",
    id %in% ingredient_target$Ingredient  ~ "Ingredient",
    id %in% ingredient_target$Target    ~ "Target",
    TRUE                              ~ "Other"
  ))

save(nodes,edges,file = "../data/processed/network/03_HIT_network.RData")
```

## Step 3: HIT network visualization
```{r}
rm(list = ls())

load("../data/processed/network/03_HIT_network.RData")

cytoscapePing()  

deleteAllNetworks() 

createNetworkFromDataFrames(
  nodes = nodes,
  edges = edges,
  title = "HFrEF_HIT_Network",
  collection = "HIT_Network"
)

# layout
layoutNetwork("force-directed")

# style
style_name <- "HIT_Style_By_Nodes_Type"
try(deleteVisualStyle(style_name), silent = TRUE)

# default style
defaults <- list(
  NODE_SHAPE = "RECTANGLE",
  NODE_SIZE = 30,
  NODE_FILL_COLOR = "#CCCCCC",
  NODE_LABEL_FONT_FACE = "Arial,plain,12",
  EDGE_WIDTH = 0.5,
  EDGE_STROKE_COLOR = "#E0E0E0",
  EDGE_TRANSPARENCY = 200         
)

# mapping style
mappings <- list(
  # node shape
  mapVisualProperty(
    visual.prop = 'node shape', 
    table.column = 'type', 
    mapping.type = 'd',
    table.column.values = c("Herb", "Ingredient", "Target"),
    visual.prop.values = c("ROUND RECTANGLE","ELLIPSE", "DIAMOND")
  ),
  
  # node color
  mapVisualProperty(
    visual.prop = 'node fill color', 
    table.column = 'type', 
    mapping.type = 'd',
    table.column.values = c("Herb", "Ingredient", "Target"),
    visual.prop.values = c("#fe9acc", "#fe0000", "#0401b9")
  ),
  
  # node size
  mapVisualProperty(
    visual.prop = 'node size',
    table.column = 'type',
    mapping.type = 'd',
    table.column.values = c("Herb", "Ingredient", "Target"),
    visual.prop.values = c(30, 20, 30)
  ),
  
  # node label
  mapVisualProperty(
    visual.prop = 'node label', 
    table.column = 'name', 
    mapping.type = 'p'
  ),
  
  mapVisualProperty(
    visual.prop = 'node label color', 
    table.column = 'type', 
    mapping.type = 'd',
    table.column.values = c("Herb", "Ingredient", "Target"),
    visual.prop.values = c("#000000", "#000000", "#FFFFFF")
  )
)

# apply style
createVisualStyle(
  style.name = style_name, 
  defaults = defaults, 
  mappings = mappings
)

setVisualStyle(style_name)

# Figure_3B.pdf / Figure_3B.png
```

---

# Functional Enrichment Analysis

## Load packages
```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)
library(ggplot2)
library(STRINGdb)
library(dplyr)
library(ggvenn)
```

## Step 1: SYMBOL to ENTREZID

```{r}
rm(list = ls())
load("../data/processed/network/01_herb_disease_genes.RData")

# candidate_genes
gene_id_map <- bitr(
  geneID = candidate_genes_df$Gene,
  fromType = "SYMBOL",
  toType = "ENTREZID",
  OrgDb = org.Hs.eg.db
)
gene_entrez <- gene_id_map$ENTREZID
```

## Step 2: GO
```{r}
go_enrich <- enrichGO(
  gene = gene_entrez,
  OrgDb = org.Hs.eg.db,
  keyType = "ENTREZID",
  ont = "ALL",
  pAdjustMethod = "fdr",
  qvalueCutoff = 0.05,
  readable = TRUE
)

```

## Step 3: KEGG
```{r}
kegg_enrich <- enrichKEGG(
  gene = gene_entrez,
  organism = "hsa",
  pAdjustMethod = "fdr",
  qvalueCutoff = 0.05
)

save(kegg_enrich,go_enrich, file = "../data/processed/functional/enrich.RData")
```

## Step 4: visualization
```{r}
load("../data/processed/functional/enrich.RData")

go_top <- go_enrich %>%
  as.data.frame() %>%
  group_by(ONTOLOGY) %>%
  slice_max(-log10(p.adjust), n = 5, with_ties = FALSE) %>% 
  ungroup() %>%
  mutate(
    Description_wrapped = str_wrap(Description, width = 30),
    Description_wrapped = as.factor(Description_wrapped),
    Description_wrapped = fct_reorder(Description_wrapped, p.adjust, .desc = FALSE),
    ONTOLOGY = factor(ONTOLOGY, levels = c("BP", "CC", "MF"))
  )

go_plot <- ggplot(go_top, aes(x = Count, y = Description_wrapped)) +
  geom_col(aes(fill = -log10(p.adjust)), width = 0.7) +
  facet_grid(ONTOLOGY ~ ., scales = "free_y", space = "free_y") +
  scale_fill_gradient(low = "#fde725", high = "#21918c", name = "-log10(p.adjust)") +
  labs(x = "Gene Count", y = "") +
  theme_bw(base_size = 12) +
  theme(
    strip.text.y = element_text(angle = 0, face = "bold", size = 11),
    axis.text.y = element_text(size = 8),          
    axis.title = element_text(size = 10),
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  ) +
  guides(fill = guide_colorbar(title.position = "top", barwidth = 10, barheight = 0.6))

ggsave("../figures/main/Figure_3C.pdf", go_plot, width = 7, height = 10, dpi = 300)
ggsave("../figures/main/Figure_3C.png", go_plot, width = 7, height = 10, dpi = 300)


# KEGG plot
kegg_plot <- dotplot(kegg_enrich, showCategory = 15) + theme_bw()
ggsave("../figures/main/Figure_3D.pdf", kegg_plot, width = 7, height = 10, dpi = 300)
ggsave("../figures/main/Figure_3D.png", kegg_plot, width = 7, height = 10, dpi = 300)
```

---

# Figure 3
```{r}
library(magick)

total_width_cm <- 14        
dpi <- 300                    
px_per_cm <- dpi / 2.54     
total_width_px <- round(total_width_cm * px_per_cm) 

ratios_row1 <- c(1, 2)
ratios_row2 <- c(1, 1)   # 修改为 1:1

width_row1_left  <- round(total_width_px * ratios_row1[1] / sum(ratios_row1))
width_row1_right <- round(total_width_px * ratios_row1[2] / sum(ratios_row1))
width_row2_left  <- round(total_width_px * ratios_row2[1] / sum(ratios_row2))
width_row2_right <- round(total_width_px * ratios_row2[2] / sum(ratios_row2))

ratio <- c(width_row1_left,width_row1_right,width_row2_left,width_row2_right)

files <- c(
  "../figures/main/Figure_3A.pdf",
  "../figures/main/Figure_3B.pdf",
  "../figures/main/Figure_3C.pdf",
  "../figures/main/Figure_3D.pdf"
)

labels <- c("A", "B", "C", "D")

label_size <- 12   

pdfs <- list()
for (i in seq_along(files)) {
  img <- image_read_pdf(files[i], density = 300)
  img <- image_resize(img,ratio[i])
  img <- image_annotate(img, labels[i], 
                        gravity = "northwest",  
                        location = "+10+10",    
                        size = label_size,     
                        weight = 700,    
                        color = "black")
  pdfs[[i]] <- img
}

row1 <- image_append(c(pdfs[[1]], pdfs[[2]]), stack = FALSE)
row2 <- image_append(c(pdfs[[3]], pdfs[[4]]), stack = FALSE)

combined <- image_append(c(row1, row2), stack = TRUE)


image_write(combined, "../figures/main/Figure_3.pdf", format = "pdf")
image_write(combined, "../figures/main/Figure_3.png", format = "png", density = dpi)
```

---


# PPI Network

## Step 1: Obtain the PPI network from the STRING database
```{r}
rm(list = ls())

load("../data/processed/network/01_herb_disease_genes.RData")

gene_list <- candidate_genes_df$Gene
ppi_list <- clusterProfiler::getPPI(gene_list, taxID = "9606",required_score = 400)

save(ppi_list,file = "../data/processed/network/04_ppi_list.RData")
```

## Step 2: PPI network construction data preparation
```{r}
nodes <- data.frame(id = igraph::V(ppi_list)$name)
nodes$degree <- igraph::degree(ppi_list)

edges <- igraph::as_data_frame(ppi_list, what = "edges")
colnames(edges)[colnames(edges) == "from"] <- "source"
colnames(edges)[colnames(edges) == "to"]   <- "target"
# add an "interaction" column to edges
if(!"interaction" %in% colnames(edges)){
    edges$interaction <- "interacts with"
}

save(nodes,edges,file = "../data/processed/network/05_ppi_network.RData")
```

## Step 3: PPI network visualization
```{r}
# Ping Cytoscape 
cytoscapePing() 

try(deleteNetwork("HFrEF_PPI_Network"), silent = TRUE)
deleteAllNetworks() 

createNetworkFromDataFrames(
  nodes = nodes,  
  edges = edges,  
  title = "HFrEF_PPI_Network", 
  collection = "PPI_Network" 
)

# Create based on degree values, node attributes must be synchronized into Cytoscape first

loadTableData(nodes, data.key.column = "id", table = "node")

# New style
try(deleteVisualStyle("PPI_Style_By_Nodes_Degree"), silent = TRUE)
style_name <- "PPI_Style_By_Nodes_Degree"

# default style
default_lists <- list(
  NODE_SHAPE = "ellipse",           
  NODE_SIZE = 40,                   
  NODE_FILL_COLOR = "#88CCEE",      
  # NODE_LABEL_FONT_SIZE = 12,
  NODE_LABEL_FONT_FACE = "Arial,plain,12",  
  EDGE_WIDTH = 0.5,
  EDGE_STROKE_COLOR = "#E0E0E0",
  EDGE_TRANSPARENCY = 200          
)

mapping_lists <- list(
  # Node size: small degree → 20, large degree → 100
  mapVisualProperty(
    visual.prop = 'node size', 
    table.column = 'degree', 
    mapping.type = 'c',  
    table.column.values = c(min(nodes$degree), max(nodes$degree)), 
    visual.prop.values = c(20, 100)
  ),
  # Node fill color mapping: small degree → light yellow, large degree → red
  mapVisualProperty(
    visual.prop = 'node fill color', 
    table.column = 'degree', 
    mapping.type = 'c',
    table.column.values = c(min(nodes$degree), max(nodes$degree)),
    visual.prop.values = c("#FFFFCC", "#CC0000")
  )
)

createVisualStyle(
  style.name = style_name,
  defaults = default_lists,
  mappings = mapping_lists
)

setVisualStyle(style_name)

# Node Label
setNodeLabelMapping(table.column = 'id', style.name = style_name)

layoutNetwork("force-directed")
```

---

# Identification of Hub Genes
## Step 1: Cytohubba
```{r}
cytohubba_res <- read_csv("../data/processed/network/06_cytohubba_res.csv") %>%
  column_to_rownames(var = colnames(.)[1])

cytohubba_res <- read_csv("../tables/cytohubba.csv") %>%
  column_to_rownames(var = colnames(.)[1])

# cytohubba
n <- min(20, nrow(cytohubba_res))

top_mcc <- rownames(cytohubba_res[order(cytohubba_res$MCC, decreasing = TRUE), ])[1:n]

top_degree <- rownames(cytohubba_res[order(cytohubba_res$Degree,decreasing = T),])[1:n]

top_mnc <- rownames(cytohubba_res[order(cytohubba_res$MNC,decreasing = T),])[1:n]

cytohubba_hub <- Reduce(intersect, list(top_mcc, top_degree, top_mnc))

cytohubba_hub

# Figure 4B_d
gene_list <- list(
  MCC = top_mcc,
  MNC = top_mnc,
  Degree = top_degree
)

library(ggsci)
library(ggvenn)

p <- ggvenn(
  gene_list,
  fill_color = pal_aaas()(3),
  stroke_color = "white",
  stroke_size = 0.5,
  set_name_size = 2.5,
  text_size = 2.5,
  text_color = "white",
  show_percentage = FALSE
) +
  theme(plot.title = element_text(hjust = 0.5, size=4))

ggsave("../figures/main/Figure_4B_d.pdf", p, width=1.38, height=1.38, units="in", dpi=300)
ggsave("../figures/main/Figure_4B_d.png", p, width=2.76, height=1.38, units="in", dpi=300)


# Figure 4B

library(magick)

total_width_cm <- 7        
dpi <- 300                    
px_per_cm <- dpi / 2.54  

files <- c(
  "../figures/main/Figure_4B_a.pdf",
  "../figures/main/Figure_4B_b.pdf",
  "../figures/main/Figure_4B_c.pdf",
  "../figures/main/Figure_4B_d.pdf"
)

labels <- c("a", "b", "c", "d")

label_size <- 8   

pdfs <- list()
for (i in seq_along(files)) {
  img <- image_read_pdf(files[i], density = 300)
  img <- image_annotate(img, labels[i], 
                        gravity = "northwest",  
                        location = "+5+5",    
                        size = label_size,     
                        weight = 700,    
                        color = "black")
  pdfs[[i]] <- img
}

row1 <- image_append(c(pdfs[[1]], pdfs[[2]]), stack = FALSE)
row2 <- image_append(c(pdfs[[3]], pdfs[[4]]), stack = FALSE)

combined <- image_append(c(row1, row2), stack = TRUE)


image_write(combined, "../figures/main/Figure_4B.pdf", format = "pdf")
image_write(combined, "../figures/main/Figure_4B.png", format = "png", density = dpi)
```

## Step 2: MCODE
```{r}
mcode_hub <- data.table::fread("../data/processed/network/07_mcode_res.txt") %>%
  dplyr::slice(1) %>%                   
  dplyr::select(last_col()) %>%          
  dplyr::pull(1) %>%                     
  strsplit(",\\s*") %>%         
  unlist() 

# intersect
hub_genes <- Reduce(intersect, list(top_mcc, top_degree, top_mnc,mcode_hub))

hub_genes_tbl <- cytohubba_res[rownames(cytohubba_res) %in% hub_genes, c("MCC","MNC","Degree")] %>%
  mutate(MCC = as.numeric(MCC)) %>%
  arrange(desc(MCC)) %>%
    rownames_to_column("Gene")

hub_genes_tbl

hub_genes <- hub_genes_tbl$Gene

hub_genes

save(hub_genes_tbl,hub_genes,top_mcc,top_degree,top_mnc, 
     file = "../data/processed/network/08_hub_genes.RData")

write.xlsx(
  hub_genes_tbl,
  "../tables/Table_S5.xlsx"
)
```

## Step 3: Visualization
```{r}
gene_list <- list(
  MCC = top_mcc,
  MNC = top_mnc,
  Degree = top_degree,
  MCODE = mcode_hub
)

library(ggsci)
library(ggvenn)

p <- ggvenn(
  gene_list,
  fill_color = pal_aaas()(4),
  stroke_color = "white",
  stroke_size = 1,
  set_name_size = 5,
  text_size = 5,
  text_color = "white",
  show_percentage = FALSE
) +
  theme(
    text = element_text(family = "serif", size = 10)
    plot.title = element_text(hjust = 0.5, size=10)
    )

# save
ggsave("../figures/main/Figure_4D.pdf", p, width=2.76, height=2.76, units="in", dpi=300)
ggsave("../figures/main/Figure_4D.png", p, width=2.76, height=2.76, units="in", dpi=300)
```

# Figure 4
```{r}
library(magick)

total_width_cm <- 14        
dpi <- 300                    
px_per_cm <- dpi / 2.54     
total_width_px <- round(total_width_cm * px_per_cm) 

ratios_row1 <- c(1, 1)
ratios_row2 <- c(1, 1)  

width_row1_left  <- round(total_width_px * ratios_row1[1] / sum(ratios_row1))
width_row1_right <- round(total_width_px * ratios_row1[2] / sum(ratios_row1))
width_row2_left  <- round(total_width_px * ratios_row2[1] / sum(ratios_row2))
width_row2_right <- round(total_width_px * ratios_row2[2] / sum(ratios_row2))

ratio <- c(width_row1_left,width_row1_right,width_row2_left,width_row2_right)

files <- c(
  "../figures/main/Figure_4A.pdf",
  "../figures/main/Figure_4B.pdf",
  "../figures/main/Figure_4C.pdf",
  "../figures/main/Figure_4D.pdf"
)

labels <- c("A", "B", "C", "D")

label_size <- 12   

pdfs <- list()
for (i in seq_along(files)) {
  img <- image_read_pdf(files[i], density = 300)
  img <- image_resize(img,ratio[i])
  img <- image_annotate(img, labels[i], 
                        gravity = "northwest",  
                        location = "+10+10",    
                        size = label_size,     
                        weight = 700,    
                        color = "black")
  pdfs[[i]] <- img
}

row1 <- image_append(c(pdfs[[1]], pdfs[[2]]), stack = FALSE)
row2 <- image_append(c(pdfs[[3]], pdfs[[4]]), stack = FALSE)

combined <- image_append(c(row1, row2), stack = TRUE)


image_write(combined, "../figures/main/Figure_4.pdf", format = "pdf")
image_write(combined, "../figures/main/Figure_4.png", format = "png", density = dpi)
```


# Immune Infiltration Profiling

## Load Package
```{r}
library(GSVA)
library(GSEABase)
library(msigdbr)
library(pheatmap) 
library(tidyverse) 
library(rstatix) 
library(corrplot) 
library(ggpubr) 
library(ggsci)
library(ggsignif)
library(ggplot2)
library(IOBR)
library(openxlsx)
library(Seurat)
```

https://www.medrxiv.org/content/10.1101/2024.04.26.24306471v1

We first determined the levels of immune cells in failing myocardium using five immune infiltration algorithms: Cibersort, xCell, ssGSEA, MCPcounter, and QuanTIseq.

## Step 1: FPKM & group

```{r}
rm(list = ls())
load("../data/processed/geo/GSE141910/GSE141910_preprocessed_final.RData")

group <- pheno_final %>% 
  rownames_to_column("Sample") %>%
  dplyr::select(Sample,group) %>%
  dplyr::rename(Group = group)


fpkm <- 2^expr_fpkm_corr - 1
fpkm[fpkm < 0] <- 0
```

## Step 2: deconvo_tme
```{r}
# MCPcounter
res_mcp   <- deconvo_tme(eset = fpkm, method = "mcpcounter")

# xCell
res_xcell <- deconvo_tme(eset = fpkm, method = "xcell")

# quanTIseq
res_quant <- deconvo_tme(eset = fpkm, method = "quantiseq")

# EPIC
res_epic  <- deconvo_tme(eset = fpkm, method = "epic")

save(res_mcp, res_xcell, res_quant, res_epic, file = "../data/processed/ime/res.RData")
```

## Step 3: Extracting data
```{r}
export_table <- function(res, name){
  out <- res %>% rename(Sample=ID) %>% left_join(group,by="Sample")
}

xCell <- export_table(res_xcell,  "xCell")
MCPcounter <- export_table(res_mcp,    "MCPcounter")
quanTIseq <- export_table(res_quant,  "quanTIseq")
EPIC <- export_table(res_epic,   "EPIC")

mcp_mat <- res_mcp %>% column_to_rownames("ID")
colnames(mcp_mat) <- gsub("_MCPcounter", "", names(mcp_mat))

df_all <- mcp_mat %>% 
  rownames_to_column("Sample") %>%
  left_join(group,by="Sample")

# Stacking data
df_stack <- df_all %>%
  pivot_longer(cols=-c(Sample,Group), names_to="Cell", values_to="Abundance") %>% 
  group_by(Group, Cell) %>% 
  summarise(Mean = mean(Abundance), .groups = "drop")

pvals <- apply(df_all[,2:(ncol(df_all)-1)], 2, function(x) {
  wilcox.test(x ~ df_all$Group)$p.value
})

sig_cells <- names(pvals)[pvals < 0.05]

df_plot <- df_all[,c("Sample","Group",sig_cells)] %>%
  pivot_longer(cols=-c(Sample,Group), names_to="Cell", values_to="Abundance")

group_levels <- unique(df_plot$Group)
comparison_list <- list(c(group_levels[1], group_levels[2]))
```

## Step 4: Figure 5A~E
```{r}
# Figure A: Differential immune cell box plot
df_clean <- df_plot %>%
  group_by(Cell) %>%
  mutate(
    Q1 = quantile(Abundance, 0.25),
    Q3 = quantile(Abundance, 0.75),
    IQR = Q3 - Q1,
    upper_limit = Q3 + 1.5 * IQR
  ) %>%
  filter(Abundance <= upper_limit) %>% 
  ungroup()

pA <- ggplot(df_clean, aes(x = Group, y = Abundance, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.6) +
  stat_compare_means(
    method = "wilcox.test",
    aes(label = paste0("p = ", ..p.format..)),
    # parse = TRUE,  
    size = 2.5,                 
    vjust = 0.6,               
    hjust = 0.5,
    label.x.npc = 0.5,          
    hide.ns = FALSE      
  ) +
  
  facet_wrap(~Cell, ncol = 4, scales = "free_y") +

  scale_fill_jco() +
  
  theme_bw(base_size = 10, base_family = "serif") +
  theme(
    axis.text.x = element_text(size = 7),
    strip.background = element_rect(fill = "gray90"),
    strip.text = element_text(face = "bold", size = 7),
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 6),
    legend.position = "bottom"
  ) +
  labs(x = NULL, y = "Abundance")

ggsave("../figures/main/Figure_5A.pdf",pA,width=14, height=10, units="cm",dpi=300)

# Figure B: Stacking
pB <- ggplot(df_stack, 
       aes(x = Group,                
           y = Mean, 
           fill = Cell)) +
  geom_col(position = "fill", width = 0.6) +
  scale_y_continuous(labels = scales::percent_format()) +
  ggsci::scale_fill_jco() +
  labs(x = NULL, y = "Proportion", fill = "Cell") +
  theme_bw(base_size = 8, base_family = "serif") +
  theme(
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    axis.title = element_text(size = 8),
    legend.title = element_text(size = 7),  
    legend.text  = element_text(size = 6), 
    legend.key.size = unit(0.25, "cm"), 
    legend.spacing.x = unit(0.15, "cm"),
    legend.spacing.y = unit(0.05, "cm"),
    legend.margin = margin(2,2,2,2), 
    legend.box.margin = margin(0,0,0,0),
    legend.position = "right",  
    legend.direction = "vertical"
  )

ggsave("../figures/main/Figure_5B.pdf", pB, width=7, height=5, units="cm", dpi=300)

# Figure C: Bar chart of differential immune cell sorting
df_mean <- df_plot %>% 
  group_by(Cell, Group) %>%
  summarise(Mean = mean(Abundance), .groups = "drop") %>%
  pivot_wider(names_from = Group, values_from = Mean)

# FC = DCM-HFrEF / Normal
df_mean$FoldChange <- df_mean[[2]] / df_mean[[3]]
df_mean <- df_mean %>% arrange(FoldChange)  

pC <- ggplot(df_mean, 
             aes(x = FoldChange, 
                 y = reorder(Cell, FoldChange), 
                 fill = Cell)) + 
  geom_col(width = 0.65) +
  geom_vline(xintercept = 1, color = "red", linetype = "dashed", linewidth = 0.5) +
  ggsci::scale_fill_aaas() +   
  labs(x = "Fold Change (DCM-HFrEF/Normal)", y = "") +
  theme_bw(base_size = 7, base_family = "serif") + 
  scale_x_continuous(
    breaks = seq(0, 1.2, 0.2),
    expand = c(0, 0) 
  ) + 
  theme(
    axis.text.y = element_text(size = 7),
    axis.text.x = element_text(size = 7),
    axis.title.x = element_text(size = 7),
    axis.ticks.y = element_blank(),
    plot.margin = margin(t = 1, r = 1, b = 1, l = -2, unit = "mm"),
    legend.position = "none"
  )

ggsave("../figures/main/Figure_5C.pdf", pC, width=7, height=5, units="cm", dpi=300)

# Figure D: Heat map of immune cell correlation
library(corrplot)
imm_exp  <- mcp_mat[, sig_cells] %>%
  rownames_to_column("Sample") %>%
  left_join(group,by="Sample") %>%
  column_to_rownames("Sample") %>%
  dplyr::filter(Group=="DCM_HFrEF") %>%
  dplyr::select(-Group)

cor_mat_cell <- cor(imm_exp, method = "spearman")

pdf("../figures/main/Figure_5D.pdf", width=7, height=5)
corrplot(
  cor_mat_cell,
  method = "color",
  type = "upper",   
  diag = FALSE,    
  order = "original",
  col = colorRampPalette(c("navy","white","red3"))(100),
  tl.cex = 0.7,      
  cl.cex = 0.6,
  tl.srt = 45,
  tl.col = "black",
  mar = c(1, 0, 1, 0),    
  cl.offset = 0.2,        
  t.pos = "left",
  addCoef.col = "black",
  number.cex = 0.6,
)
dev.off()

# Figure E: Hub genes-immune cell correlation heatmap
load("../data/processed/network/08_hub_genes.Rdata")

hub_exp  <- expr_fpkm_corr[hub_genes, ] %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("Sample") %>%
  left_join(group,by="Sample") %>%
  column_to_rownames("Sample") %>%
  dplyr::filter(Group=="DCM_HFrEF") %>%
  dplyr::select(-Group)

cor_mat <- cor(imm_exp, hub_exp, method="spearman")
p_mat <- matrix(NA, nrow = ncol(imm_exp), ncol = ncol(hub_exp),
                dimnames = list(colnames(imm_exp), colnames(hub_exp)))

for(i in 1:ncol(imm_exp)){
  for(j in 1:ncol(hub_exp)){
    p_mat[i,j] <- cor.test(imm_exp[,i], hub_exp[,j], method = "spearman")$p.value
  }
}

library(ggplot2)
library(ggsci)
library(reshape2)

cor_long <- melt(cor_mat)
p_long <- melt(p_mat)
cor_long$p <- p_long$value
colnames(cor_long) <- c("Cell","HubGene","Corr","Pvalue")

pE <- ggplot(cor_long, aes(x = HubGene, y = Cell)) +
  geom_tile(aes(fill = Corr), color = "gray90", size = 0.1) +
  scale_fill_gradient2(low = "#2166AC", mid = "white", high = "#B2182B",midpoint = 0, limits = c(-1,1)) +
  theme_minimal() +
  theme(
    text = element_text(family = "serif", size = 7),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 6,margin = margin(t = 0.5, unit = "mm")),
    axis.text.y = element_text(size = 6,margin = margin(r = 0.5, unit = "mm")),
    axis.title = element_blank(),
    legend.position = "right",
    legend.key.size = unit(0.15, "cm"),  
    legend.title = element_text(size = 5),
    legend.text  = element_text(size = 4), 
    legend.box.spacing = unit(0.01, "cm"), 
    legend.margin = margin(0,0,0,0),       
    plot.margin = margin(0.5,0.5,0.5,0.5, "mm"),
    panel.grid = element_blank()
  ) +
  labs(fill = "Correlation")

ggsave("../figures/main/Figure_5E.pdf", pE, width = 7, height = 5, units = "cm", dpi = 300)
```

## Figure 5

```{r}
library(magick)

total_width_cm <- 14        
dpi <- 300                    
px_per_cm <- dpi / 2.54     
total_width_px <- round(total_width_cm * px_per_cm) 

ratios_row1 <- c(1)
ratios_row2 <- c(1, 1) 
ratios_row3 <- c(1, 1) 

width_row1_left  <- round(total_width_px * ratios_row1[1] / sum(ratios_row1))
width_row2_left  <- round(total_width_px * ratios_row2[1] / sum(ratios_row2))
width_row2_right <- round(total_width_px * ratios_row2[2] / sum(ratios_row2))

width_row3_left  <- round(total_width_px * ratios_row3[1] / sum(ratios_row3))
width_row3_right <- round(total_width_px * ratios_row3[2] / sum(ratios_row3))

ratio <- c(width_row1_left,width_row2_left,width_row2_right,width_row3_left,width_row3_right)

files <- c(
  "../figures/main/Figure_5A.pdf",
  "../figures/main/Figure_5B.pdf",
  "../figures/main/Figure_5C.pdf",
  "../figures/main/Figure_5D.pdf",
  "../figures/main/Figure_5E.pdf"
)

labels <- c("A", "B", "C", "D", "E")

label_size <- 12   

pdfs <- list()
for (i in seq_along(files)) {
  img <- image_read_pdf(files[i], density = 300)
  img <- image_resize(img,ratio[i])
  img <- image_annotate(img, labels[i], 
                        gravity = "northwest",  
                        location = "+10+10",    
                        size = label_size,     
                        weight = 700,    
                        color = "black")
  pdfs[[i]] <- img
}

row1 <- image_append(c(pdfs[[1]]), stack = FALSE)
row2 <- image_append(c(pdfs[[2]], pdfs[[3]]), stack = FALSE)
row3 <- image_append(c(pdfs[[4]], pdfs[[5]]), stack = FALSE)

combined <- image_append(c(row1, row2, row3), stack = TRUE)


image_write(combined, "../figures/main/Figure_5.pdf", format = "pdf")
image_write(combined, "../figures/main/Figure_5.png", format = "png", density = dpi)

```

---

<table class="nav-table" width="100%">
  <tr>
    <td align="left">
      [Home](../index.qmd) 
    </td>
    <td align="right">
      [Previous](Part_2_Screening_of_HFrEF_related_genes.qmd) | [Next](Part_4_Mechanism_validation_via_bioinformatics_and_docking.qmd)
    </td>
  </tr>
</table>

# References {-}