---
title: "Part 4: Mechanism validation via bioinformatics and docking"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

---

# Identification and Validation of Biomarker Genes by Machine Learning

## Load Packages
```{r}
library(tidyverse)
library(caret)
library(glmnet)
library(e1071)
library(Boruta)
library(xgboost)
library(pROC)
library(ggplot2)
library(ggsci)
```

## Step 1: Loading data
```{r}
rm(list = ls())

load("../data/processed/geo/GSE141910GSE141910_preprocessed_final.RData")

load("../data/processed/network/08_hub_genes.Rdata")

expr_matrix <- t(expr_fpkm_corr[rownames(expr_fpkm_corr) %in% hub_genes,])

group <- pheno_final[rownames(expr_matrix),"group"] %>% as.numeric()-1

expr_data <- data.frame(expr_matrix, group = group)
```

## Step 2: Stratified sampling
```{r}
set.seed(42)

train_index <- createDataPartition(
  expr_data$group, 
  p = 0.7, 
  list = FALSE
)

train_data <- expr_data[train_index, ]
test_data  <- expr_data[-train_index, ]

x_train <- as.matrix(train_data[, -ncol(train_data)])
y_train <- train_data$group
x_test  <- as.matrix(test_data[, -ncol(test_data)])
y_test  <- test_data$group

save(x_train,y_train,x_test,y_test, file = "../data/processed/ml/01_model_data.Rdata")
```

## Step 3: Feature selection

### LASSO
```{r}
set.seed(42)
cv_lasso <- cv.glmnet(
  x_train, 
  y_train, 
  family = "binomial", 
  alpha = 1, 
  nfolds = 10
)

lasso_coef <- coef(cv_lasso, s = "lambda.min")

lasso_selected <- rownames(lasso_coef)[which(lasso_coef[,1] != 0)][-1]

cat("LASSO selected:", length(lasso_selected), "genes\n")
cat(lasso_selected, "\n")

save(cv_lasso,lasso_selected, file = "../data/processed/ml/02_lasso.Rdata")
```

### SVM-RFE
```{r}
set.seed(42)

svmFuncs <- caretFuncs
svmFuncs$summary <- function(data, lev = NULL, model = NULL) defaultSummary(data, lev, model)
rfe_ctrl <- rfeControl(functions = svmFuncs, method = "cv", number = 10, verbose = FALSE)
svm_rfe <- rfe(x_train, as.factor(y_train), sizes = c(1:18), rfeControl = rfe_ctrl)
svm_selected <- predictors(svm_rfe)
cat("SVM-RFE selected:", length(svm_selected), "genes\n")
cat(svm_selected, "\n")

save(svm_rfe,svm_selected, file = "../data/processed/ml/03_svm_rfe.Rdata")
```

### Boruta
```{r}
set.seed(42)
folds <- createFolds(y_train, k = 5, list = TRUE)

importance_list <- list() 
decision_list <- list()   

for (i in seq_along(folds)) {
  train_idx <- folds[[i]]
  x_sub <- x_train[train_idx, ]
  y_sub <- y_train[train_idx]
  
  boruta_sub <- Boruta(x_sub, as.factor(y_sub), 
                       maxRuns = 100, pValue = 0.05, doTrace = 0)
  
  imp_stats <- attStats(boruta_sub)
  importance_list[[i]] <- imp_stats$medianImp
  names(importance_list[[i]]) <- rownames(imp_stats)
  
  decision_list[[i]] <- boruta_sub$finalDecision
  names(decision_list[[i]]) <- rownames(imp_stats)
}

all_features <- unique(unlist(lapply(importance_list, names)))
imp_matrix <- matrix(NA, nrow = length(all_features), ncol = 5,
                     dimnames = list(all_features, paste0("Fold", 1:5)))

for (i in 1:5) {
  imp_matrix[names(importance_list[[i]]), i] <- importance_list[[i]]
}

confirm_matrix <- sapply(decision_list, function(d) {
    confirmed <- names(d[d == "Confirmed"])
    return(all_features %in% confirmed)
})
confirm_freq <- rowSums(confirm_matrix)   # 频次 0-5

boruta_selected <- all_features[confirm_freq == 5]
cat("Boruta selected:", length(boruta_selected), "genes\n")

cat("Boruta selected:", length(boruta_selected), "genes\n")
cat(boruta_selected, "\n")

save(boruta_result,boruta_selected, file = "../data/processed/ml/04_boruta.Rdata")
```

### XGBoost
```{r}
set.seed(42)

dtrain <- xgb.DMatrix(data = x_train, label = y_train)

params <- list(
  objective = "binary:logistic",
  max_depth = 3,
  eta = 0.1,
  eval_metric = "logloss"
)

xgb_model <- xgb.train(
  params = params,
  data = dtrain,
  nrounds = 100,
  verbose = 0
)

importance <- xgb.importance(
  feature_names = colnames(x_train), 
  model = xgb_model)

xgb_selected <- head(importance$Feature, 15)
cat("XGBoost selected:", length(xgb_selected), "features\n")
cat(xgb_selected, "\n")

save(xgb_model,xgb_selected, file = "../data/processed/ml/05_xgb.Rdata")
```

### Consensus genes
```{r}
consensus_genes <- Reduce(intersect, list(lasso_selected, svm_selected, boruta_selected, xgb_selected))
cat("Consensus genes:", length(consensus_genes), "\n")
print(consensus_genes)

write.csv(data.frame(Gene = consensus_genes), "../data/processed/ml/05_consensus_genes.csv", row.names = FALSE)
```

## Step 4: Feature selection Visualization
### LASSO
```{r}
# Figure S3A
pdf("../results/figures/supplementary/Figure_S3_A_LASSO_cv_curve.pdf", width = 7, height = 5)

plot(cv_lasso, 
     # main = "LASSO Cross-Validation Error Curve",
     xlab = expression(-log(lambda)), 
     ylab = "Partial Likelihood Deviance"
)
abline(v = -log(cv_lasso$lambda.min), col = "red", lty = 2)
abline(v = -log(cv_lasso$lambda.1se), col = "blue", lty = 2)
legend("topright", 
       legend = c("lambda.min", "lambda.1se"), 
       col = c("red", "blue"), 
       lty = 2, 
       bty = "n")
dev.off()

# Figure S3B

pdf("../results/figures/supplementary/Figure_S3_B_LASSO_coef_path.pdf", width = 7, height = 5)

plot(cv_lasso$glmnet.fit, 
     # main = "LASSO Coefficient Path", 
     xvar = "lambda", 
     label = FALSE)
abline(v = -log(cv_lasso$lambda.min), col = "red", lty = 2)
abline(v = -log(cv_lasso$lambda.1se), col = "blue", lty = 2)
legend("topleft", 
       legend = c("lambda.min", "lambda.1se"), 
       col = c("red", "blue"), 
       lty = 2, 
       bty = "n")
dev.off()

  # Figure S3C
if (length(lasso_selected) <= 30 & length(lasso_selected) > 0) {
  coef_vals <- as.matrix(lasso_coef)[-1, , drop = FALSE]
  coef_df <- data.frame(Gene = rownames(coef_vals), Coefficient = coef_vals[,1])
  coef_df <- coef_df[coef_df$Coefficient != 0, ]
  coef_df <- coef_df[order(abs(coef_df$Coefficient), decreasing = TRUE), ]
  p <- ggplot(coef_df, aes(x = reorder(Gene, Coefficient), y = Coefficient, fill = Coefficient > 0)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    labs(
      title = "LASSO Selected Features Coefficients",
      x = "Gene", 
      y = "Coefficient") +
    scale_fill_manual(values = c("red", "green"), labels = c("Negative", "Positive")) +
    theme_minimal()
  ggsave("../results/figures/supplementary/Figure_S3_C_LASSO_coef_bar.pdf", p, width = 7, height = max(5, length(lasso_selected)*0.3))
}

```

### SVM-RFE
```{r}
# Figure S4A
plot_data <- svm_rfe$results
best_idx <- which.max(plot_data$Accuracy)
best_point <- plot_data[best_idx, ]

p <- ggplot(plot_data, aes(x = Variables, y = Accuracy)) +
    geom_line() +
    geom_point(shape = 8, size = 2) + 
    geom_point(data = best_point, aes(x = Variables, y = Accuracy), 
               color = "red", size = 4, shape = 16) +
    labs(x = "Number of Features", y = "Accuracy (Cross-Validation)") +
    theme_bw()
ggsave("../results/figures/supplementary/Figure_S4_A_SVMRFE_performance.pdf", p, width = 7, height = 5)

# Figure S4B
if (length(svm_rfe$optVariables) > 0) {
    top_vars <- svm_rfe$optVariables[1:min(20, length(svm_rfe$optVariables))]
    importance_score <- rev(seq_along(top_vars))  
    var_imp_df <- data.frame(
      Gene = factor(top_vars, levels = rev(top_vars)),
      Importance = importance_score)
    
    p <- ggplot(var_imp_df, aes(x = Gene, y = Importance, fill = Importance)) +
        geom_bar(stat = "identity") +
        coord_flip() +
        scale_fill_gradient(low = "lightblue", high = "darkblue", guide = "none") +
        labs(
          # title = "SVM-RFE Feature Ranking", 
          x = "Gene", 
          y = "Importance Score (higher = more important)") +
        theme_minimal() +
        theme(axis.text.y = element_text(size = 8))
  ggsave("../results/figures/supplementary/Figure_S4_B_SVMRFE_ranking.pdf", p, width = 7, height = max(5, length(top_vars)*0.3))
}
```

### Boruta
```{r}
# Figure S5A
freq_df <- data.frame(Feature = all_features, Freq = confirm_freq)
freq_df <- freq_df[order(freq_df$Freq, decreasing = TRUE), ]

top_n <- min(30, nrow(freq_df))
freq_top <- freq_df[1:top_n, ]

p1 <- ggplot(freq_top, aes(x = reorder(Feature, Freq), y = Freq, fill = Freq)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_gradient(low = "lightblue", high = "darkblue", breaks = 0:5) +
  labs(
    # title = "Boruta Feature Selection Stability Across 5 CV Folds",
       x = "Feature", y = "Number of Folds where Confirmed") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8))
ggsave("../results/figures/supplementary/Figrue_S5_A_Boruta_CV_Frequency.pdf", p1, width = 7, height = max(5, top_n*0.25))

# # Figure S5B
# pdf("../results/figures/supplementary/Figure_S5_B_Boruta_decision.pdf", width = 8, height = 6)
# plotImpHistory(boruta_result, main = "Boruta Importance History")
# dev.off()

# Figure S5B
library(reshape2)

imp_mean <- rowMeans(imp_matrix, na.rm = TRUE)
ordered_features <- names(sort(imp_mean, decreasing = TRUE))
imp_melt <- melt(imp_matrix[ordered_features, ], 
                 varnames = c("Feature", "Fold"), 
                 value.name = "Importance")

p2 <- ggplot(imp_melt, aes(x = Fold, y = Feature, fill = Importance)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red", na.value = "grey90") +
  labs(
    # title = "Boruta Median Importance Across CV Folds",
    
       x = "Cross-Validation Fold", y = "Feature", fill = "Median Imp.") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(size = 8))
ggsave("../results/figures/supplementary/Figure_S5_B_Boruta_CV_Heatmap.pdf", p2, width = 7, height = max(6, nrow(imp_matrix)*0.1))

# Figure S5C
top_features <- names(sort(imp_mean, decreasing = TRUE))
imp_top <- imp_matrix[top_features, ]
imp_top_melt <- melt(imp_top, varnames = c("Feature", "Fold"), value.name = "Importance")

p3 <- ggplot(imp_top_melt, aes(x = reorder(Feature, -Importance, FUN = median, na.rm = TRUE), 
                               y = Importance)) +
  geom_boxplot(fill = "lightblue", outlier.color = "red", outlier.size = 1) +
  coord_flip() +
  labs(
    # title = paste("Top", top_n_box, "Features: Importance Distribution Across Folds"),
       x = "Feature", y = "Median Importance") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10))
ggsave("../results/figures/supplementary/Figure_S5_C_Boruta_CV_TopBoxplot.pdf", p3, width = 7, height = max(5, top_n_box*0.3))

# Figure S5D
fold_counts <- data.frame(
  Fold = paste0("Fold", 1:5),
  Confirmed = sapply(decision_list, function(d) sum(d == "Confirmed"))
)

p4 <- ggplot(fold_counts, aes(x = Fold, y = Confirmed, fill = Fold)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Confirmed), vjust = -0.5) +
  labs(title = "Number of Boruta-Confirmed Genes per CV Fold",
       x = "CV Fold", y = "Confirmed Genes") +
  theme_minimal() +
  theme(legend.position = "none")
ggsave("../results/figures/supplementary/Figure_S5_D_Boruta_CV_FoldCounts.pdf", p4, width = 7, height = 5)

```

### XGBoost
```{r}
importance_matrix <- xgb.importance(feature_names = colnames(x_train), model = xgb_model)

# Figure S6A
if (nrow(importance_matrix) > 0) {
  imp_top <- head(importance_matrix, 20)
  p <- ggplot(imp_top, aes(x = reorder(Feature, Gain), y = Gain, fill = Gain)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    scale_fill_gradient(low = "lightblue", high = "darkblue") +
    labs(
      # title = "XGBoost Feature Importance (Gain)", 
      x = "Feature", 
      y = "Gain") +
    theme_minimal()
  ggsave("../results/figures/supplementary/Figure_S6_A_XGBoost_importance.pdf", p, width = 8, height = max(5, nrow(imp_top)*0.3))
}

# Figure S6B
pdf("../results/figures/supplementary/Figure_S6_B_XGBoost_tree.pdf", width = 12, height = 8)
xgb.plot.tree(model = xgb_model, trees = 0:4, render = FALSE)  
dev.off()
```

### Consensus genes
```{r}
# Venn
gene_list <- list(
  LASSO = lasso_selected,
  `SVM-RFE` = svm_selected,
  Bruta = boruta_selected,
  XGBoost = xgb_selected
)

p <- ggvenn(
  gene_list,
  fill_color = pal_aaas()(4),
  stroke_color = "white",
  stroke_size = 1,
  set_name_size = 5,
  text_size = 5,
  text_color = "white",
  show_percentage = FALSE
) +
  theme(plot.title = element_text(hjust = 0.5, size=10))

# save
ggsave("../results/figures/main/Figure_5A_Venn_ML_Core_Genes.pdf", p, width=6, height=6, dpi=300)
ggsave("../results/figures/main/Figure_5A_Venn_ML_Core_Genes.png", p, width=6, height=6, dpi=300)


# Upset
library(ComplexUpset)

all_features <- unique(unlist(gene_list))
presence <- sapply(gene_list, function(x) all_features %in% x)
rownames(presence) <- all_features
upset_data <- as.data.frame(presence)
upset_data$feature <- rownames(upset_data)
upset_plot <- upset(
    upset_data,
    intersect = colnames(upset_data),  
    name = "Fold",
    width_ratio = 0.1,
    min_degree = 1,
    sort_intersections = "ascending"
)
ggsave("../results/figures/supplementary/Figure_7A_UpSet_ML_Consensus.pdf", upset_plot, width = 10, height = 6)

```

## Step 5: Logistic regression model
```{r}
library(ggplot2)
library(ggsci)
library(pROC)

formula_lr <- as.formula(paste("group ~", paste(consensus_genes, collapse = "+")))
lr_model <- glm(formula_lr, data = train_data, family = binomial())

# Train set ROC
train_pred <- predict(lr_model, train_data, type = "response")
train_roc <- roc(train_data$group, train_pred)

# Test set ROC
test_pred <- predict(lr_model, test_data, type = "response")
test_roc <- roc(test_data$group, test_pred)

# 构建数据框
# 计算 AUC 值
train_auc <- round(auc(train_roc), 3)
test_auc <- round(auc(test_roc), 3)

roc_df <- rbind(
  data.frame(
    Specificity = 1 - train_coords$specificity,  # 1 - specificity 作为 x 轴
    Sensitivity = train_coords$sensitivity,
    Group = paste0("Train AUC = ", train_auc)
  ),
  data.frame(
    Specificity = 1 - test_coords$specificity,
    Sensitivity = test_coords$sensitivity,
    Group = paste0("Test AUC = ", test_auc)
  )
)

# ROC curve
p <- ggplot(roc_df, aes(x = Specificity, y = Sensitivity, color = Group)) +
  geom_line(size = 1.2) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray50") + 
  scale_color_aaas() +  
  labs(
    # title = "ROC Curves",
    x = "1 - Specificity",
    y = "Sensitivity"
  ) +
  theme_bw() +
  theme(
    text = element_text(family = "sans", size = 10),
    legend.position = c(0.8, 0.2),
    legend.background = element_blank(),  
    legend.title = element_blank()         
  )

ggsave("../results/figures/main/Figure_5B_ROC_curves.pdf", p, width = 6, height = 6)
print(p)
```

---


# Single‑Cell Transcriptome Expression Mapping
## Step 1: Lo

```{r}
```

---




# Identification of Causal Genes by Mendelian Randomization

## Step 1:
```{r}
```

---

# GSEA Pathway Analysis of Causal Genes

---

# Docking of Active Ingredients and Core Targets

---

<table class="nav-table" width="100%">
  <tr>
    <td align="left">
      [Home](../index.qmd) | [About](../about.qmd) | [Methods](../methods.qmd) | [Results](../results.qmd)
    </td>
    <td align="right">
      [Previous](Part_3_Mechanism_exploration_of_ZWHQD_against_HFrEF.qmd) | [Next](Part_5_TF_miRNA_regulatory_network.qmd)
    </td>
  </tr>
</table>

# References {-}