---
title: "Part 3: Mechanism exploration of ZWHQD against HFrEF"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

---

# Identification of Candidate Targets

## Load Packages
```{r}
library(ggvenn)
library(tidyverse)
library(openxlsx)
```

## Candidate Genes
```{r}
rm(list = ls())

load("../data/processed/geo/GSE141910/degs_results.Rdata")

deg_sig <- deg_all %>% filter(padj < 0.05, abs(log2FoldChange) > lfc)

deg_genes <- rownames(deg_sig) 

load("../data/processed/tcm/14_compound_target_final.RData")

herb_targets <- unique(compound_target_final$Symbol)

load("../data/processed/geo/GSE141910/wgcna_candidates_genes.Rdata")

# disease_genes: intersecting DEGs and WGCNA
disease_genes <- intersect(deg_genes,wgcna_candidates_genes$Gene)

# candidate_genes: intersecting herb_targets and disease_genes
candidate_genes <-intersect(herb_targets, disease_genes)

herb_targets_df <- data.frame(Gene = herb_targets) %>%
  arrange(Gene)

deg_genes_df <- data.frame(Gene = deg_genes) %>%
  arrange(Gene)

wgcna_genes_df <- data.frame(Gene = wgcna_candidates_genes$Gene) %>%
  arrange(Gene)

disease_genes_df <- data.frame(Gene = disease_genes) %>%
  arrange(Gene)

candidate_genes_df <- data.frame(Gene = candidate_genes) %>%
  arrange(Gene)

gene_list <- list(
  DEGs = deg_genes,
  `Herb targets` = herb_targets,
  `WGCNA Genes` = wgcna_candidates_genes$Gene
)

p <- ggvenn(
  gene_list,
  fill_color = c("#F1948A", "blue","#58D68D"),
  stroke_color = "white",
  stroke_size = 1,
  set_name_size = 5,
  text_size = 5,
  show_percentage = FALSE
) +
  theme(plot.title = element_text(hjust = 0.5, size=10))

# save
ggsave("../results/figures/main/Figure_3C_Venn_Herb_DEG_WGCNA.pdf", p, width=6, height=6, dpi=300)
ggsave("../results/figures/main/Figure_3C_Venn_Herb_DEG_WGCNA.png", p, width=6, height=6, dpi=300)


save(herb_targets_df, deg_genes_df, wgcna_genes_df,disease_genes_df, candidate_genes_df, file = "../data/processed/network/herb_disease_genes.Rdata")

write.xlsx(
  list(
    HerbTargets = herb_targets_df,
    DEGs = deg_genes_df,
    WGCNAGenes = wgcna_genes_df,
    DiseaseGenes = disease_genes_df,
    CandidateGenes = candidate_genes_df
  ),
  "../results/tables/supplementary/Table_S5_herb_disease_targets.xlsx"
)

deg_candidate <- deg_all[rownames(deg_all) %in% candidate_genes, ]
 
deg_candidate$regulation <- ifelse(
  deg_candidate$log2FoldChange >= lfc, "Up",
  ifelse(deg_candidate$log2FoldChange <= -lfc, "Down", "NS")
)

deg_candidate$Gene <- rownames(deg_candidate)

up_genes <- deg_candidate$Gene[deg_candidate$regulation == "Up"]
down_genes <- deg_candidate$Gene[deg_candidate$regulation == "Down"]

cat("Up genes: ", length(up_genes), "\n")
print(up_genes)

cat("\nDown genes: ", length(down_genes), "\n")
print(down_genes)
```

---

# Functional Enrichment Analysis

## Load packages
```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)
library(ggplot2)
library(STRINGdb)
library(dplyr)
library(ggvenn)
```

## Step 1: SYMBOL to ENTREZID

```{r}
rm(list = ls())
load("../data/processed/network/herb_disease_genes.Rdata")

# candidate_genes
gene_id_map <- bitr(
  geneID = candidate_genes_df$Gene,
  fromType = "SYMBOL",
  toType = "ENTREZID",
  OrgDb = org.Hs.eg.db
)
gene_entrez <- gene_id_map$ENTREZID
```

## Step 2: GO
```{r}
go_enrich <- enrichGO(
  gene = gene_entrez,
  OrgDb = org.Hs.eg.db,
  keyType = "ENTREZID",
  ont = "ALL",
  pAdjustMethod = "fdr",
  qvalueCutoff = 0.05,
  readable = TRUE
)

go_top <- go_enrich %>%
  as.data.frame() %>%
  group_by(ONTOLOGY) %>%
  slice_max(-log10(p.adjust), n = 5) %>% 
  ungroup() %>%
  mutate(
    Description = fct_reorder(Description, p.adjust), 
    ONTOLOGY = factor(ONTOLOGY, levels = c("BP", "CC", "MF"))  
  )

go_plot <- ggplot(go_top, aes(x = Count, y = Description)) +
  geom_col(aes(fill = -log10(p.adjust)), width = 0.7) + 
  facet_grid(ONTOLOGY ~ ., scales = "free_y", space = "free_y") +
  scale_fill_gradient(low = "#fde725", high = "#21918c", name = "-log10(p.adjust)") +
  
  # 标签
  labs(
    x = "Gene Count",
    y = "GO Term"
  ) +
  theme_bw(base_size = 12) +
  theme(
    strip.text.y = element_text(angle = 0, face = "bold", size = 11),
    axis.text.y = element_text(size = 10),
    panel.grid = element_blank(),
    legend.position = "bottom"
  ) +
  guides(fill = guide_colorbar(title.position = "top", barwidth = 15, barheight = 0.8))

ggsave("../results/figures/main/Figure_3D_GO_Plot.pdf", go_plot, width = 9, height = 7, dpi = 300)

```

## Step 3: KEGG
```{r}
kegg_enrich <- enrichKEGG(
  gene = gene_entrez,
  organism = "hsa",
  pAdjustMethod = "fdr",
  qvalueCutoff = 0.05
)

# KEGG plot
kegg_plot <- dotplot(kegg_enrich, showCategory = 12) + theme_bw()
ggsave("../results/figures/main/Figure_3E_KEGG_dot.pdf", kegg_plot, width = 9, height = 7, dpi = 300)

save(kegg_enrich,go_enrich, file = "../data/processed/functional/enrich.Rdata")

# write.xlsx(
#   list(
#     GO <- 
#   )
# )
```

---

# Immune Infiltration Profiling
```{r}

```

---

# Herb-Ingredient-Target (HIT) Network
```{r}
library(RCy3)
```

---

# PPI Network


---

# Identification of Hub Genes



---

<table class="nav-table" width="100%">
  <tr>
    <td align="left">
      [Home](../index.qmd) | [About](../about.qmd) | [Methods](../methods.qmd) | [Results](../results.qmd)
    </td>
    <td align="right">
      [Previous](Part_2_Screening_of_HFrEF_related_genes.qmd) | [Next](Part_4_Mechanism_validation_via_bioinformatics_and_docking.qmd)
    </td>
  </tr>
</table>

# References {-}