---
title: "Part 4: Mechanism validation via bioinformatics and docking"
format: 
  html:
    toc: true
    number-sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.dpi = 300,
  fig.align = "center"
)

options(timeout = 36000)
options(stringsAsFactors = FALSE)
options(download.file.method = "curl")
options(download.file.extra = "-k -L")

options(BioC_mirror = "https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
```

---

# Identification and Validation of Biomarker Genes by Machine Learning

## Load Packages
```{r}
library(tidyverse)
library(data.table)
library(reshape2)
library(caret)
library(glmnet)
library(e1071)
library(Boruta)
library(xgboost)
library(pROC)
library(ggplot2)
library(ggsci)
library(corrplot)
library(openxlsx)
```

## Step 1: Loading data
```{r}
rm(list = ls())

load("../data/processed/geo/GSE141910/GSE141910_preprocessed_final.RData")

load("../data/processed/network/08_hub_genes.RData")

expr_matrix <- t(expr_fpkm_corr[rownames(expr_fpkm_corr) %in% hub_genes,])

expr_matrix <- expr_matrix[,hub_genes]

group <- pheno_final[rownames(expr_matrix),"group"] %>% as.numeric()-1

expr_data <- data.frame(expr_matrix, group = group)
```

## Step 2: Stratified sampling
```{r}
set.seed(42)

train_index <- createDataPartition(
  expr_data$group, 
  p = 0.7, 
  list = FALSE
)

train_data <- expr_data[train_index, ]
test_data  <- expr_data[-train_index, ]

x_train <- as.matrix(train_data[, -ncol(train_data)])
y_train <- train_data$group
x_test  <- as.matrix(test_data[, -ncol(test_data)])
y_test  <- test_data$group

save(train_data, test_data, x_train,y_train,x_test,y_test, file = "../data/processed/ml/01_model_data.RData")
```

## Step 4: Hub genes correlation analysis
```{r}
cor_matrix <- cor(x_train)

off_diag_abs <- abs(cor_matrix[lower.tri(cor_matrix)])

max_abs <- max(off_diag_abs, na.rm = TRUE)
min_abs <- min(off_diag_abs, na.rm = TRUE)

cor_df <- melt(cor_matrix, varnames = c("Gene1", "Gene2"), value.name = "cor")

cor_df <- cor_df[as.character(cor_df$Gene1) != as.character(cor_df$Gene2), ]

cor_df$row_id <- as.numeric(factor(cor_df$Gene1, levels = hub_genes))

cor_df$col_id <- as.numeric(factor(cor_df$Gene2, levels = hub_genes))

cor_df <- cor_df[cor_df$row_id > cor_df$col_id, ]

p <- ggplot(cor_df, aes(x = col_id, y = row_id)) +
  geom_point(aes(fill = cor), shape = 21, size = 10, stroke = 0) +
  geom_text(aes(label = sprintf("%.2f", cor)), size = 2.5, color = "black") +  
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", 
                       midpoint = 0, limits = c(-1, 1),
                       name = "Correlation") +
  scale_y_continuous(breaks = 1:length(hub_genes), labels = hub_genes, trans = "reverse") +
  scale_x_continuous(breaks = 1:length(hub_genes), labels = hub_genes) +
  labs(x="",y="") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.2, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = c(0.85,0.55)) +
  coord_fixed()

ggsave("../results/figures/supplementary/Figure_S3_A_correlation_plot.pdf", p, width = 7, height = 7, dpi = 300)
```

## Step 4: ROC-AUC for each gene
```{r}
compute_auc <- function(gene, x, y) {
  roc_obj <- roc(response = y, predictor = x[, gene], quiet = TRUE, direction = "auto")
  auc_val <- round(auc(roc_obj), 3)
  return(auc_val)
}

gene_auc_df <- data.frame(
  Gene = hub_genes,
  Train_AUC = sapply(hub_genes, compute_auc, x = x_train, y = y_train),
  Test_AUC  = sapply(hub_genes, compute_auc, x = x_test,  y = y_test)
) 

hub_genes_tbl2 <- hub_genes_tbl %>%
  left_join(gene_auc_df,by="Gene")

save(gene_auc_df, hub_genes_tbl2, 
     file = "../data/processed/ml/02_gene_auc.RData")

write.xlsx(
  hub_genes_tbl2,
  "../results/tables/main/Table2_Hub_Genes_AUC.xlsx"
)


```

## Step 5: Feature selection

### LASSO
```{r}
set.seed(42)
cv_lasso <- cv.glmnet(
  x_train, 
  y_train, 
  family = "binomial", 
  alpha = 1, 
  nfolds = 10
)

lasso_coef <- coef(cv_lasso, s = "lambda.min")

lasso_selected <- rownames(lasso_coef)[which(lasso_coef[,1] != 0)][-1]

cat("LASSO selected:", length(lasso_selected), "genes\n")
cat(lasso_selected, "\n")

save(cv_lasso,lasso_selected, file = "../data/processed/ml/02_lasso.RData")
```

### SVM-RFE
```{r}
set.seed(42)

svmFuncs <- caretFuncs
svmFuncs$summary <- function(data, lev = NULL, model = NULL) defaultSummary(data, lev, model)
rfe_ctrl <- rfeControl(functions = svmFuncs, method = "cv", number = 10, verbose = FALSE)
svm_rfe <- rfe(x_train, as.factor(y_train), sizes = c(1:18), rfeControl = rfe_ctrl)
svm_selected <- predictors(svm_rfe)
cat("SVM-RFE selected:", length(svm_selected), "genes\n")
cat(svm_selected, "\n")

save(svm_rfe,svm_selected, file = "../data/processed/ml/03_svm_rfe.RData")
```

### Boruta
```{r}
set.seed(42)
folds <- createFolds(y_train, k = 5, list = TRUE)

importance_list <- list() 
decision_list <- list()   

for (i in seq_along(folds)) {
  train_idx <- folds[[i]]
  x_sub <- x_train[train_idx, ]
  y_sub <- y_train[train_idx]
  
  boruta_sub <- Boruta(x_sub, as.factor(y_sub), 
                       maxRuns = 100, pValue = 0.05, doTrace = 0)
  
  imp_stats <- attStats(boruta_sub)
  importance_list[[i]] <- imp_stats$medianImp
  names(importance_list[[i]]) <- rownames(imp_stats)
  
  decision_list[[i]] <- boruta_sub$finalDecision
  names(decision_list[[i]]) <- rownames(imp_stats)
}

all_features <- unique(unlist(lapply(importance_list, names)))
imp_matrix <- matrix(NA, nrow = length(all_features), ncol = 5,
                     dimnames = list(all_features, paste0("Fold", 1:5)))

for (i in 1:5) {
  imp_matrix[names(importance_list[[i]]), i] <- importance_list[[i]]
}

confirm_matrix <- sapply(decision_list, function(d) {
    confirmed <- names(d[d == "Confirmed"])
    return(all_features %in% confirmed)
})
confirm_freq <- rowSums(confirm_matrix)   # 频次 0-5

boruta_selected <- all_features[confirm_freq == 5]
cat("Boruta selected:", length(boruta_selected), "genes\n")

cat("Boruta selected:", length(boruta_selected), "genes\n")
cat(boruta_selected, "\n")

save(boruta_result,boruta_selected, file = "../data/processed/ml/04_boruta.RData")
```

### XGBoost
```{r}
set.seed(42)

dtrain <- xgb.DMatrix(data = x_train, label = y_train)

params <- list(
  objective = "binary:logistic",
  max_depth = 3,
  eta = 0.1,
  eval_metric = "logloss"
)

xgb_model <- xgb.train(
  params = params,
  data = dtrain,
  nrounds = 100,
  verbose = 0
)

importance <- xgb.importance(
  feature_names = colnames(x_train), 
  model = xgb_model)

xgb_selected <- head(importance$Feature, 15)
cat("XGBoost selected:", length(xgb_selected), "features\n")
cat(xgb_selected, "\n")

save(xgb_model,xgb_selected, file = "../data/processed/ml/05_xgb.RData")
```

### Consensus genes
```{r}
consensus_genes <- Reduce(intersect, list(lasso_selected, svm_selected, boruta_selected, xgb_selected))
cat("Consensus genes:", length(consensus_genes), "\n")
print(consensus_genes)

save(consensus_genes, file =  "../data/processed/ml/06_consensus_genes.RData")

write.csv(data.frame(Gene = consensus_genes), "../data/processed/ml/06_consensus_genes.csv", row.names = FALSE)
```

## Step 6: Feature selection Visualization
### LASSO
```{r}
# Figure S3A
pdf("../results/figures/supplementary/Figure_S3_A_LASSO_cv_curve.pdf", width = 7, height = 5)

plot(cv_lasso, 
     # main = "LASSO Cross-Validation Error Curve",
     xlab = expression(-log(lambda)), 
     ylab = "Partial Likelihood Deviance"
)
abline(v = -log(cv_lasso$lambda.min), col = "red", lty = 2)
abline(v = -log(cv_lasso$lambda.1se), col = "blue", lty = 2)
legend("topright", 
       legend = c("lambda.min", "lambda.1se"), 
       col = c("red", "blue"), 
       lty = 2, 
       bty = "n")
dev.off()

# Figure S3B

pdf("../results/figures/supplementary/Figure_S3_B_LASSO_coef_path.pdf", width = 7, height = 5)

plot(cv_lasso$glmnet.fit, 
     # main = "LASSO Coefficient Path", 
     xvar = "lambda", 
     label = FALSE)
abline(v = -log(cv_lasso$lambda.min), col = "red", lty = 2)
abline(v = -log(cv_lasso$lambda.1se), col = "blue", lty = 2)
legend("topleft", 
       legend = c("lambda.min", "lambda.1se"), 
       col = c("red", "blue"), 
       lty = 2, 
       bty = "n")
dev.off()

  # Figure S3C
if (length(lasso_selected) <= 30 & length(lasso_selected) > 0) {
  coef_vals <- as.matrix(lasso_coef)[-1, , drop = FALSE]
  coef_df <- data.frame(Gene = rownames(coef_vals), Coefficient = coef_vals[,1])
  coef_df <- coef_df[coef_df$Coefficient != 0, ]
  coef_df <- coef_df[order(abs(coef_df$Coefficient), decreasing = TRUE), ]
  p <- ggplot(coef_df, aes(x = reorder(Gene, Coefficient), y = Coefficient, fill = Coefficient > 0)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    labs(
      title = "LASSO Selected Features Coefficients",
      x = "Gene", 
      y = "Coefficient") +
    scale_fill_manual(values = c("red", "green"), labels = c("Negative", "Positive")) +
    theme_minimal()
  ggsave("../results/figures/supplementary/Figure_S3_C_LASSO_coef_bar.pdf", p, width = 7, height = max(5, length(lasso_selected)*0.3))
}

```

### SVM-RFE
```{r}
# Figure S4A
plot_data <- svm_rfe$results
best_idx <- which.max(plot_data$Accuracy)
best_point <- plot_data[best_idx, ]

p <- ggplot(plot_data, aes(x = Variables, y = Accuracy)) +
    geom_line() +
    geom_point(shape = 8, size = 2) + 
    geom_point(data = best_point, aes(x = Variables, y = Accuracy), 
               color = "red", size = 4, shape = 16) +
    labs(x = "Number of Features", y = "Accuracy (Cross-Validation)") +
    theme_bw()
ggsave("../results/figures/supplementary/Figure_S4_A_SVMRFE_performance.pdf", p, width = 7, height = 5)

# Figure S4B
if (length(svm_rfe$optVariables) > 0) {
    top_vars <- svm_rfe$optVariables[1:min(20, length(svm_rfe$optVariables))]
    importance_score <- rev(seq_along(top_vars))  
    var_imp_df <- data.frame(
      Gene = factor(top_vars, levels = rev(top_vars)),
      Importance = importance_score)
    
    p <- ggplot(var_imp_df, aes(x = Gene, y = Importance, fill = Importance)) +
        geom_bar(stat = "identity") +
        coord_flip() +
        scale_fill_gradient(low = "lightblue", high = "darkblue", guide = "none") +
        labs(
          # title = "SVM-RFE Feature Ranking", 
          x = "Gene", 
          y = "Importance Score (higher = more important)") +
        theme_minimal() +
        theme(axis.text.y = element_text(size = 8))
  ggsave("../results/figures/supplementary/Figure_S4_B_SVMRFE_ranking.pdf", p, width = 7, height = max(5, length(top_vars)*0.3))
}
```

### Boruta
```{r}
# Figure S5A
freq_df <- data.frame(Feature = all_features, Freq = confirm_freq)
freq_df <- freq_df[order(freq_df$Freq, decreasing = TRUE), ]

top_n <- min(30, nrow(freq_df))
freq_top <- freq_df[1:top_n, ]

p1 <- ggplot(freq_top, aes(x = reorder(Feature, Freq), y = Freq, fill = Freq)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_gradient(low = "lightblue", high = "darkblue", breaks = 0:5) +
  labs(
    # title = "Boruta Feature Selection Stability Across 5 CV Folds",
       x = "Feature", y = "Number of Folds where Confirmed") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8))
ggsave("../results/figures/supplementary/Figrue_S5_A_Boruta_CV_Frequency.pdf", p1, width = 7, height = max(5, top_n*0.25))

# # Figure S5B
# pdf("../results/figures/supplementary/Figure_S5_B_Boruta_decision.pdf", width = 8, height = 6)
# plotImpHistory(boruta_result, main = "Boruta Importance History")
# dev.off()

# Figure S5B
library(reshape2)

imp_mean <- rowMeans(imp_matrix, na.rm = TRUE)
ordered_features <- names(sort(imp_mean, decreasing = TRUE))
imp_melt <- melt(imp_matrix[ordered_features, ], 
                 varnames = c("Feature", "Fold"), 
                 value.name = "Importance")

p2 <- ggplot(imp_melt, aes(x = Fold, y = Feature, fill = Importance)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red", na.value = "grey90") +
  labs(
    # title = "Boruta Median Importance Across CV Folds",
    
       x = "Cross-Validation Fold", y = "Feature", fill = "Median Imp.") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(size = 8))
ggsave("../results/figures/supplementary/Figure_S5_B_Boruta_CV_Heatmap.pdf", p2, width = 7, height = max(6, nrow(imp_matrix)*0.1))

# Figure S5C
top_features <- names(sort(imp_mean, decreasing = TRUE))
imp_top <- imp_matrix[top_features, ]
imp_top_melt <- melt(imp_top, varnames = c("Feature", "Fold"), value.name = "Importance")

p3 <- ggplot(imp_top_melt, aes(x = reorder(Feature, -Importance, FUN = median, na.rm = TRUE), 
                               y = Importance)) +
  geom_boxplot(fill = "lightblue", outlier.color = "red", outlier.size = 1) +
  coord_flip() +
  labs(
    # title = paste("Top", top_n_box, "Features: Importance Distribution Across Folds"),
       x = "Feature", y = "Median Importance") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10))
ggsave("../results/figures/supplementary/Figure_S5_C_Boruta_CV_TopBoxplot.pdf", p3, width = 7, height = max(5, top_n_box*0.3))

# Figure S5D
fold_counts <- data.frame(
  Fold = paste0("Fold", 1:5),
  Confirmed = sapply(decision_list, function(d) sum(d == "Confirmed"))
)

p4 <- ggplot(fold_counts, aes(x = Fold, y = Confirmed, fill = Fold)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Confirmed), vjust = -0.5) +
  labs(title = "Number of Boruta-Confirmed Genes per CV Fold",
       x = "CV Fold", y = "Confirmed Genes") +
  theme_minimal() +
  theme(legend.position = "none")
ggsave("../results/figures/supplementary/Figure_S5_D_Boruta_CV_FoldCounts.pdf", p4, width = 7, height = 5)

```

### XGBoost
```{r}
importance_matrix <- xgb.importance(feature_names = colnames(x_train), model = xgb_model)

# Figure S6A
if (nrow(importance_matrix) > 0) {
  imp_top <- head(importance_matrix, 20)
  p <- ggplot(imp_top, aes(x = reorder(Feature, Gain), y = Gain, fill = Gain)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    scale_fill_gradient(low = "lightblue", high = "darkblue") +
    labs(
      # title = "XGBoost Feature Importance (Gain)", 
      x = "Feature", 
      y = "Gain") +
    theme_minimal()
  ggsave("../results/figures/supplementary/Figure_S6_A_XGBoost_importance.pdf", p, width = 8, height = max(5, nrow(imp_top)*0.3))
}

# Figure S6B
pdf("../results/figures/supplementary/Figure_S6_B_XGBoost_tree.pdf", width = 12, height = 8)
xgb.plot.tree(model = xgb_model, trees = 0:4, render = FALSE)  
dev.off()
```

### Consensus genes
```{r}
# Venn
gene_list <- list(
  LASSO = lasso_selected,
  `SVM-RFE` = svm_selected,
  Bruta = boruta_selected,
  XGBoost = xgb_selected
)

p <- ggvenn(
  gene_list,
  fill_color = pal_aaas()(4),
  stroke_color = "white",
  stroke_size = 1,
  set_name_size = 5,
  text_size = 5,
  text_color = "white",
  show_percentage = FALSE
) +
  theme(plot.title = element_text(hjust = 0.5, size=10))

# save
ggsave("../results/figures/main/Figure_5A_Venn_ML_Core_Genes.pdf", p, width=6, height=6, dpi=300)
ggsave("../results/figures/main/Figure_5A_Venn_ML_Core_Genes.png", p, width=6, height=6, dpi=300)


# Upset
library(ComplexUpset)

all_features <- unique(unlist(gene_list))
presence <- sapply(gene_list, function(x) all_features %in% x)
rownames(presence) <- all_features
upset_data <- as.data.frame(presence)
upset_data$feature <- rownames(upset_data)
upset_plot <- upset(
    upset_data,
    intersect = colnames(upset_data),  
    name = "Fold",
    width_ratio = 0.1,
    min_degree = 1,
    sort_intersections = "ascending"
)
ggsave("../results/figures/supplementary/Figure_7A_UpSet_ML_Consensus.pdf", upset_plot, width = 10, height = 6)

```

## Step 7: Logistic regression model
```{r}
load("../data/processed/ml/06_consensus_genes.RData")
load("../data/processed/ml/01_model_data.RData")

formula_lr <- as.formula(paste("group ~", paste(consensus_genes$Gene, collapse = "+")))
lr_model <- glm(formula_lr, data = train_data, family = binomial())

save(lr_model, file = "../data/processed/ml/07_lr_model.RData")

# Train set ROC
train_pred <- predict(lr_model, train_data, type = "response")
train_roc <- roc(train_data$group, train_pred)

# Test set ROC
test_pred <- predict(lr_model, test_data, type = "response")
test_roc <- roc(test_data$group, test_pred)

# 构建数据框
# 计算 AUC 值
train_auc <- round(auc(train_roc), 3)
test_auc <- round(auc(test_roc), 3)

train_roc_smooth <- smooth(train_roc, method = "binormal")
test_roc_smooth  <- smooth(test_roc,  method = "binormal")

roc_df <- rbind(
  data.frame(
    Specificity = 1 - train_roc_smooth$specificities, 
    Sensitivity = train_roc_smooth$sensitivities,
    Group = paste0("Train AUC = ", train_auc)
  ),
  data.frame(
    Specificity = 1 - test_roc_smooth$specificities,
    Sensitivity = test_roc_smooth$sensitivities,
    Group = paste0("Test AUC = ", test_auc)
  )
)

# ROC curve
p <- ggplot(roc_df, aes(x = Specificity, y = Sensitivity, color = Group)) +
  geom_line(size = 1.2) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray50") + 
  scale_color_aaas() +  
  labs(
    # title = "ROC Curves",
    x = "1 - Specificity",
    y = "Sensitivity"
  ) +
  theme_bw() +
  theme(
    text = element_text(family = "sans", size = 10),
    legend.position = c(0.8, 0.2),
    legend.background = element_blank(),  
    legend.title = element_blank()         
  )

ggsave("../results/figures/main/Figure_5B_ROC_curves.pdf", p, width = 6, height = 6)

source("../scripts/external_roc_cm.R")

res_cm_train <- plot_external_cm(
  lr_model, 
  train_data, 
  train_data$group, 
  dataset_name = "train_data"
)

res_cm_test <- plot_external_cm(
  lr_model, 
  test_data, 
  test_data$group, 
  dataset_name = "test_data"
)
save(roc_df,res_cm_train,res_cm_test, file = "../data/processed/ml/gse141910_res.RData")
```

## Step 8: External validation

### GSE116250: External validation set
```{r}
library(org.Hs.eg.db)

rm(list = ls())

load("../data/processed/ml/06_consensus_genes.RData")
load("../data/processed/ml/07_lr_model.RData")
load("../data/raw/geo/gse116250_raw.RData")

gse116250_expr_trans <- gse116250_fpkm_raw %>%
  column_to_rownames("GeneID") %>%
  mutate(across(everything(), ~ .x + 1)) %>%  
  log2() %>%
  t() %>% 
  as.data.frame()

entrez_ids <- colnames(gse116250_expr_trans)  
symbols <- mapIds(org.Hs.eg.db, keys = entrez_ids, keytype = "ENTREZID", column = "SYMBOL", multiVals = "first")

keep <- !is.na(symbols)
gse116250_expr_annot <- t(gse116250_expr_trans[, keep]) %>%
  as.data.frame() %>%
  mutate(Gene = symbols[keep], .before = 1)

common_genes <- intersect(consensus_genes$Gene, (gse116250_expr_annot$Gene))

if (all(common_genes %in% consensus_genes$Gene)) {
  cat("Good dataset\n")
} else {
  cat("Bad dataset\n")
}

gse116250_expr_merged <- gse116250_expr_annot %>%
  as.data.frame() %>%
  dplyr::filter(Gene %in% common_genes) %>%
  group_by(Gene) %>%
  summarise(across(everything(), mean, na.rm = TRUE)) %>%
  column_to_rownames("Gene") %>%
  t() %>%                          
  as.data.frame()

gse116250_expr_merged <- gse116250_expr_merged[,consensus_genes$Gene]

# pheno info
Normal <- trimws(gse116250_pheno_raw$characteristics_ch1.1) == "disease: non-failing"
DCM_HFrEF <- trimws(gse116250_pheno_raw$characteristics_ch1.1) == "disease: dilated cardiomyopathy"

gse116250_expr_filtered <- rbind(gse116250_expr_merged[Normal,],gse116250_expr_merged[DCM_HFrEF,]) %>%
  mutate(group = c(rep(0,sum(Normal)), rep(1,sum(DCM_HFrEF))))

source("../scripts/external_roc_cm.R")

res_roc <- plot_external_roc(
  lr_model, 
  gse116250_expr_filtered, 
  gse116250_expr_filtered$group,
  dataset_name = "GSE116250"
)

res_cm <- plot_external_cm(
  lr_model, 
  gse116250_expr_filtered, 
  gse116250_expr_filtered$group, 
  dataset_name = "GSE116250"
)

if (res_roc$auc >= 0.80) {
  cat("Good dataset\n")
} else {
  cat("Bad dataset\n")
}

save(res_cm,res_roc, file = "../data/processed/ml/gse116250_res.RData")
```

### GSE161472: External validation set
```{r}
rm(list = ls())

load("../data/processed/ml/06_consensus_genes.RData")
load("../data/processed/ml/07_lr_model.RData")
load("../data/raw/geo/gse161472_raw.RData")

gse161472_expr_trans <- gse161472_fpkm_raw %>%
  column_to_rownames("GeneID") %>%
  mutate(across(everything(), ~ .x + 1)) %>%  
  log2() %>%
  t() %>% 
  as.data.frame()

entrez_ids <- colnames(gse161472_expr_trans)  
symbols <- mapIds(org.Hs.eg.db, keys = entrez_ids, keytype = "ENTREZID", column = "SYMBOL", multiVals = "first")

keep <- !is.na(symbols)
gse161472_expr_annot <- t(gse161472_expr_trans[, keep]) %>%
  as.data.frame() %>%
  mutate(Gene = symbols[keep], .before = 1)

common_genes <- intersect(consensus_genes$Gene, (gse161472_expr_annot$Gene))

if (all(common_genes %in% consensus_genes$Gene)) {
  cat("Good dataset\n")
} else {
  cat("Bad dataset\n")
}

gse161472_expr_merged <- gse161472_expr_annot %>%
  as.data.frame() %>%
  dplyr::filter(Gene %in% common_genes) %>%
  group_by(Gene) %>%
  summarise(across(everything(), mean, na.rm = TRUE)) %>%
  column_to_rownames("Gene") %>%
  t() %>%                          
  as.data.frame()

gse161472_expr_merged <- gse161472_expr_merged[,consensus_genes$Gene]

# pheno info
Normal <- trimws(gse161472_pheno_raw$`disease state:ch1`) == "NF"
DCM_HFrEF <- trimws(gse161472_pheno_raw$`disease state:ch1`) == "HFrEF"

gse161472_expr_filtered <- rbind(gse161472_expr_merged[Normal,],gse161472_expr_merged[DCM_HFrEF,]) %>%
  mutate(group = c(rep(0,sum(Normal)), rep(1,sum(DCM_HFrEF))))

source("../scripts/external_roc_cm.R")

res_roc <- plot_external_roc(
  lr_model, 
  gse161472_expr_filtered, 
  gse161472_expr_filtered$group,
  dataset_name = "GSE161472"
)

res_cm <- plot_external_cm(
  lr_model, 
  gse161472_expr_filtered, 
  gse161472_expr_filtered$group, 
  dataset_name = "GSE161472"
)

if (res_roc$auc >= 0.80) {
  cat("Good dataset\n")
} else {
  cat("Bad dataset\n")
}

save(res_cm,res_roc, file = "../data/processed/ml/gse161472_res.RData")
```

### GSE135055: External validation set
```{r}
rm(list = ls())

load("../data/processed/ml/06_consensus_genes.RData")
load("../data/processed/ml/07_lr_model.RData")
load("../data/raw/geo/gse135055_raw.RData")

gse135055_expr_trans <- gse135055_fpkm_raw %>%
  column_to_rownames("GeneID") %>%
  mutate(across(everything(), ~ .x + 1)) %>%  
  log2() %>%
  t() %>% 
  as.data.frame()

entrez_ids <- colnames(gse135055_expr_trans)  
symbols <- mapIds(org.Hs.eg.db, keys = entrez_ids, keytype = "ENTREZID", column = "SYMBOL", multiVals = "first")

keep <- !is.na(symbols)
gse135055_expr_annot <- t(gse135055_expr_trans[, keep]) %>%
  as.data.frame() %>%
  mutate(Gene = symbols[keep], .before = 1)

common_genes <- intersect(consensus_genes$Gene, (gse135055_expr_annot$Gene))

if (all(common_genes %in% consensus_genes$Gene)) {
  cat("Good dataset\n")
} else {
  cat("Bad dataset\n")
}

gse135055_expr_merged <- gse135055_expr_annot %>%
  as.data.frame() %>%
  dplyr::filter(Gene %in% common_genes) %>%
  group_by(Gene) %>%
  summarise(across(everything(), mean, na.rm = TRUE)) %>%
  column_to_rownames("Gene") %>%
  t() %>%                          
  as.data.frame()

gse135055_expr_merged <- gse135055_expr_merged[,consensus_genes$Gene]

gse135055_pheno_filtered <- gse135055_pheno_raw[rownames(gse135055_expr_merged),]

# pheno info
Normal <- trimws(gse135055_pheno_filtered$`diagnosis:ch1`) == "Not available"

lvef_col <- trimws(gse135055_pheno_filtered$`lvef (%):ch1`)
lvef_numeric <- ifelse(grepl("^[0-9.]+$", lvef_col), as.numeric(lvef_col), NA)

DCM_HFrEF <- (trimws(gse135055_pheno_filtered$`diagnosis:ch1`) != "Not available") &
             !is.na(lvef_numeric) &
             lvef_numeric <= 40

gse135055_expr_filtered <- rbind(gse135055_expr_merged[Normal,],gse135055_expr_merged[DCM_HFrEF,]) %>%
  mutate(group = c(rep(0,sum(Normal)), rep(1,sum(DCM_HFrEF))))

source("../scripts/external_roc_cm.R")

res_roc <- plot_external_roc(
  lr_model, 
  gse135055_expr_filtered, 
  gse135055_expr_filtered$group,
  dataset_name = "GSE135055"
)

res_cm <- plot_external_cm(
  lr_model, 
  gse135055_expr_filtered, 
  gse135055_expr_filtered$group, 
  dataset_name = "GSE135055"
)

if (res_roc$auc >= 0.80) {
  cat("Good dataset\n")
} else {
  cat("Bad dataset\n")
}

save(res_cm,res_roc, file = "../data/processed/ml/gse135055_res.RData")
```

---

### GSE120895: External validation set(array)
```{r}
library(AnnoProbe)

rm(list = ls())

load("../data/processed/ml/06_consensus_genes.RData")
load("../data/processed/ml/07_lr_model.RData")
load("../data/raw/geo/gse120895_raw.RData")

ex <- gse120895_exprs_raw
qx <- as.numeric(quantile(ex, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))

LogC <- (qx[5] > 100) ||
          (qx[6]-qx[1] > 50 && qx[2] > 0)
if (LogC) { ex[which(ex <= 0)] <- NaN
  ex <- log2(ex + 1) }

probe2gene <- idmap("GPL570")
gse120895_expr_annot <-  as.data.frame(ex) %>% 
  rownames_to_column("probe_id") %>%
  inner_join(probe2gene, by = "probe_id") %>%
  dplyr::select(-probe_id) %>%
  mutate(Gene = symbol, .before = 1) %>%
  dplyr::select(-symbol)

common_genes <- intersect(consensus_genes$Gene, (gse120895_expr_annot$Gene))

if (all(common_genes %in% consensus_genes$Gene)) {
  cat("Good dataset\n")
} else {
  cat("Bad dataset\n")
}

gse120895_expr_merged <- gse120895_expr_annot %>%
  as.data.frame() %>%
  dplyr::filter(Gene %in% common_genes) %>%
  group_by(Gene) %>%
  summarise(across(everything(), mean, na.rm = TRUE)) %>%
  column_to_rownames("Gene") %>%
  t() %>%                          
  as.data.frame()

gse120895_expr_merged <- gse120895_expr_merged[,consensus_genes$Gene]

# pheno info
Normal <- trimws(gse120895_pheno_raw$`condition:ch1`) == "control"
DCM_HFrEF <- trimws(gse120895_pheno_raw$`condition:ch1`) != "control"

gse120895_expr_filtered <- rbind(gse120895_expr_merged[Normal,],gse120895_expr_merged[DCM_HFrEF,]) %>%
  mutate(group = c(rep(0,sum(Normal)), rep(1,sum(DCM_HFrEF))))

source("../scripts/external_roc_cm.R")

res_roc <- plot_external_roc(
  lr_model, 
  gse120895_expr_filtered, 
  gse120895_expr_filtered$group,
  dataset_name = "GSE120895"
)

res_cm <- plot_external_cm(
  lr_model, 
  gse120895_expr_filtered, 
  gse120895_expr_filtered$group, 
  dataset_name = "GSE120895"
)

if (res_roc$auc >= 0.80) {
  cat("Good dataset\n")
} else {
  cat("Bad dataset\n")
}

save(res_cm,res_roc, file = "../data/processed/ml/gse120895_res.RData")
```

### GSE19303: External validation set  
```{r}
rm(list = ls())

load("../data/processed/ml/06_consensus_genes.RData")
load("../data/processed/ml/07_lr_model.RData")
load("../data/raw/geo/gse19303_raw.RData")

ex <- gse19303_exprs_raw
qx <- as.numeric(quantile(ex, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))

LogC <- (qx[5] > 100) ||
          (qx[6]-qx[1] > 50 && qx[2] > 0)
if (LogC) { ex[which(ex <= 0)] <- NaN
  ex <- log2(ex + 1) }

probe2gene <- idmap("GPL570")
gse19303_expr_annot <-  as.data.frame(ex) %>% 
  rownames_to_column("probe_id") %>%
  inner_join(probe2gene, by = "probe_id") %>%
  dplyr::select(-probe_id) %>%
  mutate(Gene = symbol, .before = 1) %>%
  dplyr::select(-symbol)

common_genes <- intersect(consensus_genes$Gene, (gse19303_expr_annot$Gene))

if (all(common_genes %in% consensus_genes$Gene)) {
  cat("Good dataset\n")
} else {
  cat("Bad dataset\n")
}

gse19303_expr_merged <- gse19303_expr_annot %>%
  as.data.frame() %>%
  dplyr::filter(Gene %in% common_genes) %>%
  group_by(Gene) %>%
  summarise(across(everything(), mean, na.rm = TRUE)) %>%
  column_to_rownames("Gene") %>%
  t() %>%                          
  as.data.frame()

gse19303_expr_merged <- gse19303_expr_merged[,consensus_genes$Gene]

# pheno info
Normal <- !grepl("^DCM_", gse19303_pheno_raw$title)
DCM_HFrEF <- grepl("^DCM_", gse19303_pheno_raw$title) & 
                  as.numeric(gse19303_pheno_raw$`lvef:ch1`) <= 40

# grepl("_baseline", gse19303_pheno_raw$title) &

gse19303_expr_filtered <- rbind(gse19303_expr_merged[Normal,],gse19303_expr_merged[DCM_HFrEF,]) %>%
  mutate(group = c(rep(0,sum(Normal)), rep(1,sum(DCM_HFrEF))))

source("../scripts/external_roc_cm.R")

res_roc <- plot_external_roc(
  lr_model, 
  gse19303_expr_filtered, 
  gse19303_expr_filtered$group,
  dataset_name = "GSE19303"
)

res_cm <- plot_external_cm(
  lr_model, 
  gse19303_expr_filtered, 
  gse19303_expr_filtered$group, 
  dataset_name = "GSE19303"
)

if (res_roc$auc >= 0.80) {
  cat("Good dataset\n")
} else {
  cat("Bad dataset\n")
}

save(res_cm,res_roc, file = "../data/processed/ml/gse19303_res.RData")
```

## herb-ingredient-target
```{r}
load("../data/processed/tcm/14_compound_target_final.RData")

hit_core <- compound_target_final %>%
    dplyr::filter(Symbol %in% consensus_genes$Gene) %>%
  arrange(Herb.Name.Pinyin,Symbol)

unique(hit_core$Herb.Name.Pinyin)

hit_core[,c("Herb.Name.Pinyin", "Compound.Name","Symbol")]
```
# Single‑Cell Transcriptome Expression Mapping
## Load Packages
```{r}
library(Seurat)
library(ggplot2)
library(patchwork)
library(dplyr)
library(Matrix)
library(harmony)
```

## Step 1: Loading data

```{r}
rm(list=ls())
gc()
load("../data/processed/ml/06_consensus_genes.RData")

load("../data/raw/geo/GSE183852/GSE183852_DCM_Cells/GSE183852_DCM_Cells.Robj")

new.cluster.ids <- c(
  "Fibroblasts",
  "Endothelium",
  "Pericytes",
  "Macrophages",
  "T_Cells",
  "Smooth_Muscle",
  "Monocytes",
  "NK_Cells",
  "Neurons",
  "B_Cells",
  "Lymphatic"
)

```

## Step 2: 细胞注释UMAP
```{r}
p1 <- DimPlot(HDCM, label = TRUE, repel = TRUE, pt.size = 0.1, label.size = 2.5) +
  theme_bw(base_family = "sans") +  
  theme(
    panel.border = element_rect(fill = NA, color = "black", size = 1.0),
    legend.text = element_text(size = 7),
    legend.title = element_text(size = 8),
    legend.spacing.y = unit(0.02, "cm"),  
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 8),  
    axis.title = element_text(size = 8)  
  )

ggsave("../results/figures/main/Figure_6A_UMAP.pdf", p1, width = 10, height = 8, units = "cm",dpi = 300)

ggsave("../results/figures/main/Figure_6A_UMAP.png", p1, width = 10, height = 8, units = "cm",dpi = 300)
```



---




# Bayesian co-localization analysis

## Load Packages
```{r}
library(data.table)   
library(ggplot2)      
library(coloc)         
library(tidyverse) 
library(arrow)
library(rtracklayer)
library(GenomicRanges)
```


## Step 1: Loading data
```{r}
# Loading GTEx v11 eQTLs Heart_Left_Ventricle data
eqtl_data <- read_parquet("../data/raw/opengwas/GTEx_Analysis_v11_eQTL/Heart_Left_Ventricle.v11.eQTLs.signif_pairs.parquet")

eqtl_dt <- as.data.table(eqtl_data)
eqtl_dt[, phenotype_id_clean := sub("\\..*", "", phenotype_id)]

# Loading HERMES2_GWAS_DCM_EUR GWAS data
gwas_raw <- data.table::fread("../data/raw/opengwas/HERMES2_GWAS_DCM_EUR/FORMAT-METAL_Pheno5-DCM_EUR.tsv.gz")
colnames(gwas_raw)
head(gwas_raw, 2)

# Core genes
load("../data/processed/ml/06_consensus_genes.RData")
phenotype_ids <- c(
  "ENSG00000162434",  # JAK1
  "ENSG00000118777",  # ABCG2
  "ENSG00000113578",  # FGF1
  "ENSG00000136160",  # EDNRB
  "ENSG00000100644",  # HIF1A
  "ENSG00000166825",  # ANPEP
  "ENSG00000159640"   # ACE
)
```

## Step 2: Function of single gene co-localization analysis
```{r}
#' Run coloc.abf for a given gene using GTEx v11 eQTL and DCM GWAS
#'
#' @param phenotype_id Ensembl gene ID (e.g. "ENSG00000155657")
#' @param eqtl_dt data.table containing GTEx v11 cis-eQTL significant pairs.
#'                Must have columns: phenotype_id, variant_id, slope, slope_se, af, ma_count.
#' @param gwas_dt data.table containing GWAS summary stats.
#'                Must have columns: SNP, BETA, SE, EAF, N (optional, if missing will be estimated).
#' @param gwas_n_cases number of cases in GWAS (for type="cc" s parameter)
#' @param gwas_n_controls number of controls in GWAS (for total N)
#' @param p1 prior probability a SNP is associated with trait 1 (eQTL), default 1e-4
#' @param p2 prior probability a SNP is associated with trait 2 (GWAS), default 1e-4
#' @param p12 prior probability a SNP is associated with both traits, default 1e-5
#' @param eqtl_window_kb cis-window in kb around gene TSS (not used if eqtl_dt is already filtered), default 1000
#' @param gene_tss optional integer: TSS position in bp. If NULL, function will try to infer from eqtl_dt or use 0 (skip filtering).
#' @param return_details if TRUE, return full coloc result object; otherwise return summary data.frame.
#'
#' @return A data.frame with gene, PP.H0, PP.H1, PP.H2, PP.H3, PP.H4, n_snps, status
#'         or if return_details=TRUE, a list with results and summary.
#'
#' @import data.table coloc
#' @export
#'
coloc_for_gene <- function(
  phenotype_id,
  eqtl_dt,
  gwas_dt,
  gwas_n_cases,
  gwas_n_controls,
  p1 = 1e-4,
  p2 = 1e-4,
  p12 = 1e-5,
  eqtl_window_kb = 1000,
  gene_tss = NULL,
  return_details = FALSE) {
  
  # ---------- (1) Extracting eQTL of gene ----------
  gene_eqtl <- eqtl_dt[phenotype_id == phenotype_id]
  if (nrow(gene_eqtl) == 0) {
    warning("No eQTL found for gene ", phenotype_id)
    return(data.frame(gene = phenotype_id, PP.H4 = NA, n_snps = 0, status = "no_eqtl"))
  }
  
  # ---------- (2) Analyze variant_id and calculate MAF, N_eqtl ----------
  # variant_id: "chr_pos_ref_alt_build"
  gene_eqtl[, c("chr", "pos", "ref", "alt") := tstrsplit(variant_id, "_", keep = 1:4)]
  gene_eqtl[, pos := as.numeric(pos)]
  gene_eqtl[, maf := pmin(af, 1 - af)]
  # Estimate the sample size N=ma_comunt/(2 * maf), round it up and filter out outliers
  gene_eqtl[, N_eqtl := round(ma_count / (2 * maf))]
  # Reasonable range: Typically, GTEx v11 heart samples range from 900-1000, with possible errors below 50 or above 2000. Use median instead
  med_n <- median(gene_eqtl$N_eqtl, na.rm = TRUE)
  gene_eqtl[is.na(N_eqtl) | N_eqtl < 50 | N_eqtl > 2000, N_eqtl := med_n]
  
  # ---------- (3) Cis window filtering (if eqtl-d t has not been filtered by position) ----------
  if (!is.null(gene_tss)) {
    gene_eqtl <- gene_eqtl[abs(pos - gene_tss) <= eqtl_window_kb * 1000]
    if (nrow(gene_eqtl) == 0) {
      warning("No eQTL within cis-window for gene ", phenotype_id)
      return(data.frame(gene = phenotype_id, PP.H4 = NA, n_snps = 0, status = "no_cis_eqtl"))
    }
  }
  
  # ---------- (4) Build Dataset 1 (eQTL) ----------
  d1 <- list(
    beta = gene_eqtl$slope,
    varbeta = gene_eqtl$slope_se^2,
    type = "quant",
    N = med_n,   #Use the median sample size of all eQTLs for this gene
    sdY = 1,     #Conservatively set, as the expression level of GTEx has been normalized
    snp = gene_eqtl$variant_id,
    MAF = gene_eqtl$maf
  )
  
  # Validation of Dataset 1
  if (any(is.na(d1$beta)) || any(is.na(d1$varbeta)) || any(d1$varbeta <= 0)) {
    warning("Invalid eQTL data for gene ", phenotype_id)
    return(data.frame(gene = phenotype_id, PP.H4 = NA, n_snps = nrow(gene_eqtl), status = "invalid_eqtl"))
  }
  
  # ---------- (5) Prepare GWAS data ----------
  # Ensure that GWAS has the necessary columns: SNP, BETA, SE, EAF (MAF), and N (optional)
  if (!"N_total" %in% names(gwas_dt)) {
    stop("GWAS data must contain 'N_total' column")
  }
  gwas_clean <- gwas_dt[!is.na(beta) & !is.na(varbeta) & !is.na(MAF) & varbeta > 0]
  
  total_n <- unique(gwas_clean$N_total)
  if (length(total_n) > 1) total_n <- median(gwas_clean$N_total, na.rm = TRUE)
  case_n <- unique(gwas_clean$N_case)
  if (length(case_n) > 1) case_n <- median(gwas_clean$N_case, na.rm = TRUE)
  s <- case_n / total_n
  
  # Match SNP
  common_snps <- intersect(d1$snp, gwas_clean$SNP)
  if (length(common_snps) < 5) {
    warning("Less than 5 common SNPs for gene ", phenotype_id)
    return(data.frame(gene = phenotype_id, PP.H4 = NA, n_snps = length(common_snps), status = "low_overlap"))
  }
  
  # Align by common SNP subset
  idx1 <- match(common_snps, d1$snp)
  idx2 <- match(common_snps, gwas_clean$SNP)
  
  d1_sub <- list(
    beta = d1$beta[idx1],
    varbeta = d1$varbeta[idx1],
    type = "quant",
    N = d1$N,
    sdY = d1$sdY,
    snp = d1$snp[idx1],
    MAF = d1$MAF[idx1]
  )
  
  d2_sub <- list(
    beta = gwas_clean$BETA[idx2],
    varbeta = gwas_clean$SE[idx2]^2,
    type = "cc",
    N = total_n,
    s = s,
    snp = gwas_clean$SNP[idx2],
    MAF = gwas_clean$EAF[idx2]
  )
  
  # ---------- (6) run coloc.abf ----------
  coloc_result <- tryCatch({
    coloc::coloc.abf(dataset1 = d1_sub, dataset2 = d2_sub,
                     p1 = p1, p2 = p2, p12 = p12)
  }, error = function(e) {
    warning("coloc.abf failed for gene ", phenotype_id, ": ", e$message)
    return(NULL)
  })
  
  if (is.null(coloc_result)) {
    return(data.frame(gene = phenotype_id, PP.H4 = NA, n_snps = length(common_snps), status = "coloc_error"))
  }
  
  # ---------- (7) Extraction results ----------
  summary_stats <- coloc_result$summary
  res_df <- data.frame(
    gene = phenotype_id,
    PP.H0 = summary_stats["PP.H0.abf"],
    PP.H1 = summary_stats["PP.H1.abf"],
    PP.H2 = summary_stats["PP.H2.abf"],
    PP.H3 = summary_stats["PP.H3.abf"],
    PP.H4 = summary_stats["PP.H4.abf"],
    n_snps = length(common_snps),
    status = "success",
    stringsAsFactors = FALSE
  )
  
  if (return_details) {
    return(list(summary = res_df, full_result = coloc_result))
  } else {
    return(res_df)
  }
}
```

## Step 3: Building the dataset2 object for colon
```{r}
gwas_data <- gwas_raw[, .(
    snp     = rsID,
    beta    = A1_beta,
    varbeta = se^2,
    MAF     = A1_freq,   
    N_total = N_total,
    N_case  = N_case
)] %>%
  na.omit()

# Calculate the total sample size and case ratio (using median as each row may be the same)
total_n <- unique(gwas_data$N)  
if(length(total_n) > 1) total_n <- median(gwas_data$N, na.rm=TRUE)
s <- unique(gwas_data$N_case) / total_n
if(length(s) > 1) s <- median(gwas_data$N_case, na.rm=TRUE) / total_n

d2 <- list(
    beta    = gwas_data$beta,
    varbeta = gwas_data$varbeta,
    type    = "cc",
    N       = total_n,
    s       = s,
    snp     = gwas_data$snp,
    MAF     = gwas_data$MAF
)
```

## JAK1
```{r}
jak1_result <- coloc_for_gene(
  phenotype_id = "ENSG00000162434",
  eqtl_dt = eqtl_dt,
  gwas_dt = gwas_data,
  gwas_n_cases = 6001,
  gwas_n_controls = 1199156,
  p1 = 1e-4, p2 = 1e-4, p12 = 1e-5,
  eqtl_window_kb = 1000,
  gene_tss = NULL,
  return_details = FALSE
)

print(jak1_result)
```

---

# GSEA Pathway Analysis of Causal Genes

---

# Docking of Active Ingredients and Core Targets

---

<table class="nav-table" width="100%">
  <tr>
    <td align="left">
      [Home](../index.qmd) | [About](../about.qmd) | [Methods](../methods.qmd) | [Results](../results.qmd)
    </td>
    <td align="right">
      [Previous](Part_3_Mechanism_exploration_of_ZWHQD_against_HFrEF.qmd) | [Next](Part_5_TF_miRNA_regulatory_network.qmd)
    </td>
  </tr>
</table>

# References {-}